---
title: "1_raw_data_course_catalog_validation_MNPS"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: "July 30 2024"
review:
last update: "May 19 2025"
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
# Connect to SharePoint where the data files live
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
orgfiles_site     <- get_sharepoint_site(site_url = orgfiles_site_url)
internal_drive    <- orgfiles_site$get_drive("Internal")

# Read the district config — tells us where to find files, how to rename columns, etc.
config <- yaml::read_yaml(here::here("config", "districts", "tr_nashville.yaml"))
raw_data_folder_path <- config$raw_data_folder_path

```

# Part 1. Set Variables & Read Data

## 1.1 Update Folder Path, File Names, Year and Bring in Data

```{r}
# Load the raw course catalog
course_catalog_raw <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = config$raw_files$course_catalog_raw
)

# Clean column names to remove spaces and special characters
# course_catalog_raw <- course_catalog_raw %>% clean_names()

# Rename to the template name
course_catalog_all_years <- course_catalog_raw

```



## Part 2: Validate Data

### 2.1 Run ers_column_info on your data

**ACTION: Replace "raw_data" with the name of your data frame**

```{r}
ers_column_info(course_catalog_all_years)
```

### 2.2 Manual Checks

-   **What:** A handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input the result of the check in a OneNote table, checks table, or other tool you are using to track validation checks

-   Do we have right year(s) of data?

    -   Check: Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

-   What fields do you have, and what do they mean?

    -   Check: Districts often send many fields in the course catalog. Review each field to better understand what they have sent you. If you aren't sure what a field means, ask ChatGPT, and search for them in the state's student data system. Here is the link to TX's as an example: <https://tealprod.tea.state.tx.us/TWEDS/103/0/0/0/Introduction/List/878> In Lubbock, the team discovered that RFS_P_POP_SERVED_CODE indicates the population of students served by the course (I.e., bilingual students, special education students, etc.).

**Project Team Action:** Track validation checks with the result of your manual checks. Make a list of fields that you need to ask the district's data analyst about

### 2.3 Understand File Structure

#### a. Number of Total Rows

-   **What:** Understand the structure of your file by loading the number of rows. You would expect one row per unique course name and course ID. You may see a larger number of courses than you are expecting based on district naming conventions, like if courses are named by S1 and S2 conventions. This could be anywhere from 500-2000, based on the size of the district and number of unique course offerings included.

-   **Check:** None

-   **Project Team Action:** Gut check/track. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY high school students in the district"

```{r}
course_catalog_all_years %>% nrow() # View output of check to know what to write for your status column
```

#### b. Resolution Remove Rows

-   **What:** Transform your data if necessary. You can also leave a note about district follow-up required if the team did not receive the full data set

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before

-   **Project Team Action:** Maybe change (see guidance in code block)

```{r}
# #remove rows
# # Remove rows where a specific condition is met
# # For example, remove rows where the subject detail is OTHER
# course_catalog_all_years <- course_catalog_all_years %>%
#   filter(subject_detail = "OTHER") #this is an example, you should replace this with the actual condition you want to filter on
# 
# # Print the result
# print(head(course_catalog_all_years))
```

### 2.2 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present

-   **Project Team Action:** Definitely change (see guidance in code block)

```{r}
#if the ERS Standard column is present in your data then type your district's column on the right hand side (it is okay if the names are identical)
#if the ERS Standard column is not present in your data then type "MISSING"
#you can also exclude the columns you do not need for your analysis by simply not renaming those columns in this step (the mutate function with the .keep = none argument only retains columns that follow the ERS column tagging convention)

course_catalog_all_years <- course_catalog_all_years %>%
  mutate(
    D_course_id = `mnps_course_code`,
    D_course_name = `mnps_course_name`,
    D_course_type = `state_designation`,
    D_pathway = `program_of_study`,
    D_cluster = `cluster`,
    D_course_level = `level`,
    D_state_course_code = `course_code`,
    D_state_course_name = `state_course_name`,
    .keep = "none" #"Keep-None" here will ONLY keep the columns listed above, so if you do not list a column, it will be deleted. Make sure to keep any district specific columns or other columns you need for your analysis.
  )


```

#### b. Review columns

-   **What:** Manually check that you have the same number of columns + the number of missing columns. Review the columns you are missing

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them.

### 2.3 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting

-   **Project Action:** Gut check/track

```{r}
# View duplicates in course catalog file
 course_catalog_all_years_dups <- course_catalog_all_years %>%
  group_by_all() %>%
 mutate(n = n()) %>%
    filter(n>1) %>% 
view(course_catalog_all_years_dups) %>% 
  ungroup()

# Count the number of total duplicates
  course_catalog_all_years_dups %>%
  summarise(distinct_course_ids = n_distinct(D_course_id))

```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates

-   **Check:** That the number of duplicates is now 0

-   **Project Action:** None

```{r}
# Remove duplicate rows
course_catalog_all_years <- distinct(course_catalog_all_years)

```

### 2.4 Missing & Unique Values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action**: None

```{r}
# Update blanks or NULL to NA
course_catalog_all_years <- course_catalog_all_years %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates. Gut check/track

```{r}
# Create new data frame with column information
col_info <-  data.frame(
  Column_Name = names(course_catalog_all_years),
  
  Count_Rows = sapply(course_catalog_all_years, function(x) length(na.omit(x))),
  NA_Rows = sapply(course_catalog_all_years, function(x) sum(is.na(x))),
  Unique_Rows = sapply(course_catalog_all_years, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)

# Calc % missing
col_info <- col_info %>% 
  mutate(Percent_Missing = round(NA_Rows / Count_Rows * 100,2),
         Percent_Missing = Percent_Missing, nsmall = 2) %>% 
  print()


```

#### c. Check for duplicate course names

-   **What:** Check for changing course names, and/or inconsistency in duplicate course names and or course Ids.

-   **Check:** Are there duplicate course names with different course IDs? What is the difference between these courses?

-   **Project Team Action:** Review output and note any key courses with identical names/IDs. Are they two different courses? Discuss any issues with project team.

-   **Note** Remember, course names may often look duplicated due to the course occurring twice for a student over two semesters, ie. "Algebra 1" in the fall and "Algebra 1" in the spring. This is normal and expected.

```{r}
 course_catalog_all_years %>%
  group_by(D_course_id) %>%
  mutate(count = n_distinct(D_course_name)) %>%
  ungroup() %>%
  filter(count > 1) %>%
 arrange(D_course_id)# Keep only rows where there are multiple names

# how many instances of courses having the exact same name?
course_catalog_all_years %>%
  group_by(D_course_name) %>%
  summarise(count = n_distinct(D_course_id)) %>%
  filter(count > 1) %>%
  ungroup() %>%
  arrange(desc(count)) # Keep only rows where there are multiple course IDs

# how many instances of course code being exactly the same
course_catalog_all_years %>%
  group_by(D_course_id) %>%
  summarise(count = n_distinct(D_course_name)) %>%
  filter(count > 1) %>%
  ungroup() %>%
  arrange(desc(count)) # Keep only rows where there are multiple course names
```

### 2.5 Distinct Count of Courses

#### a. Return the Distinct Count of Courses

-   **What:** Distinct count of courses

-   **Check:** Count should be slightly higher than the number of rows in each course catalog file (assuming you have multiple for different years) as we would expect that in some year, courses were offered that were not previously offered.

-   **Project Action:** Gut check/track

```{r}
#count the number of distinct courses by course ID
course_catalog_all_years %>%
  summarize(count_courses = n_distinct(D_course_id)) %>% 
  print()

#count the number of distinct courses by course name
course_catalog_all_years %>%
  summarize(count_courses = n_distinct(D_course_name)) %>% 
  print()



```

#### b. Distinct Count of Courses by ERS-Coded Subject Area- don't have this for MNPS

-   **What:** Count of courses by ERS-coded subject area. NOTE- this requires having completed course coding in excel before you bring in the file. See the transcript analysis landing page on Tettra for more information- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>

-   **Check:** Review the counts of courses by district-coded subject area. You would expect a relatively even split across core courses like ELA, Math, Science, and Social Studies, with a potentially larger number of courses in other areas like CTE, Fine Arts, and World Languages.

-   **Project Action:** Gut check/track

```{r}
# #subject area counts
# course_catalog_all_years %>% 
#   group_by(D_course_subj) %>% 
#   summarise(count_courses = length(unique(D_course_id))) %>% 
#   ungroup()

```

### 2.6 Distinct Counts of Course IDs by Flags

#### a. Return counts of Pathway courses

-   **What:** Count of courses that are Pathways

-   **Check:** You are looking for wonky things. Does the percentage of CTE courses seem reasonable considering what you know about the number of CTE pathways in the district?

-   **Project Action:** Gut check/track

```{r}
#First lets create a new column called D_pathway_flag that is a Y for every row except those where the pathway is Cambridge / IB Diploma Program
course_catalog_all_years <- course_catalog_all_years %>%
  mutate(D_pathway_flag = ifelse(D_pathway == "Cambridge / IB Diploma Program", "N", "Y"))

# # Count Pathway courses
# course_catalog_all_years%>% 
#   group_by(D_pathway_flag) %>% 
#   summarise(count_courses = n_distinct(D_course_id)) %>% 
#   ungroup()


# #calc % Pathway courses
# percent_pathway <- (sum(course_catalog_all_years_unique$D_pathway_flag =="Y") / length(course_catalog_all_years_unique$D_pathway_flag)) *100 %>% 
#   print()

```

#### b. Review courses with pathway flags

-   **What:** Return courses with pathway flag

-   **Check:** Review the rows to see if you can identify any patterns around course naming, term, and credit-earning potential for these courses versus non-pathway

-   **Project Action:** Gut check/track

```{r}
# #Filter to pathway courses, Review the resulting data frame
# courses_with_pathway_flag <- course_catalog_all_years_unique[course_catalog_all_years_unique$D_pathway_flag == "Y", ] %>% 
#   print()

```

#### c. Return counts of dual credit courses

-   **What:** Count of courses that are dual credit

-   **Check:** You are looking for wonky things. Does the percentage of dual credit courses seem reasonable considering what you know about the number of dual enrollment partnerships in the district?

-   **Project Action:** Gut check/track

```{r}
#First create a dual credit flag if D_course type is equal to "	Statewide Dual Credit" or "Dual Enrollment it gets a "Y", otherwise "N"
course_catalog_all_years <- course_catalog_all_years %>%
  mutate(D_dual_credit_flag = ifelse(D_course_type %in% c("Statewide Dual Credit", "Dual Enrollment"), "Y", "N"))


# Count dual credit courses
course_catalog_all_years %>% 
  group_by(D_dual_credit_flag) %>% 
  summarise(count_courses = n_distinct(D_course_id)) %>% 
  ungroup()

#what about if I count using state course ID
course_catalog_all_years %>% 
  group_by(D_dual_credit_flag) %>% 
  summarise(count_courses = n_distinct(D_state_course_code)) %>% 
  ungroup()

#calc % dual credit
percent_dual_credit <- (sum(course_catalog_all_years$D_dual_credit_flag =="Y") / length(course_catalog_all_years$D_dual_credit_flag)) *100

print(percent_dual_credit)
```

#### d. Review courses with dual credit flags

-   **What:** Return courses with dual credit flag

-   **Check:** Review the rows to see if you can identify any patterns around course naming, term, and credit-earning potential for these courses versus non-dual credit

-   **Project Action:** gut check/track

```{r}
# Filter to dual credit courses, Review the resulting data frame
courses_with_dual_credit_flag <- course_catalog_all_years %>% filter(D_dual_credit_flag == "Y") %>% 
  print()
```

## 2.7 Pathways Companion Coding- This was only "Pathways" Essentially for MNPS, so we don't need the below

#### a. Review and Code the Pathways Companion Sheet

-   **What:** Review and Code the Pathways Companion Sheet to ensure that all courses are coded to the correct pathway, and that there are no duplicates or N/A values in course Ids. Also standardizes pathway names. NOTE- if you are not doing a pathways analysis, you can skip straight to uploading to sharepoint.

-   **Check:** Review rows/data for any missing data.

-   **Project Action:** Gut check/track

```{r}
# #First, add ERS standard names onto the pathways fields. 
# #if the ERS Standard column is present in your data then type your district's column on the right hand side (it is okay if the names are identical)
# #if the ERS Standard column is not present in your data then type "MISSING"
# #you can also exclude the columns you do not need for your analysis by simply not renaming those columns in this step (the mutate function with the .keep = none argument only retains columns that follow the ERS column tagging convention)
# pathways_companion <- pathways %>%
#   select(
#     D_course_id = `Course.ID:`,
#     D_course_name = `Course.Name.(PDF)`,
#     D_course_level = `Course.Level`,
#     D_auto_pathway = `Automotive`,
#     D_animal_science_pathway = `Animal.Science`,
#     D_applied_agricultural_engineering_pathway = `Applied.Agriculture.Engineering`,
#     D_plant_science_pathway = `Plant.Science`,
#     D_carpentry_pathway = `Carpentry`,
#     D_digital_communication_pathway = `Digital.Communications`,
#     D_graphic_arts_pathway = `Graphic.Arts`,
#     D_animation_pathway = `Animation`,
#     D_commercial_photography_pathway = `Commercial.Photohraphy`,
#     D_accounting_and_financial_services_pathway = `Accounting.and.Financial.Services`,
#     D_early_learning_pathway = `Early.Learning`,
#     D_healthcare_theraputic_pathway = `Healthcare.Therapeutic`,
#     D_culinary_arts_pathway = `Culinary.Arts`,
#     D_web_development_pathway = `Web.Development`,
#     D_law_enforcement_pathway = `Law.Enforcement`,
#     D_cosmetology_pathway = `Cosmetology`,
#     D_advadnced_manufacturing_and_machinery_pathway = `Advadnced.Manufacturing.and.Machinery.(Robotics)`,
#     D_biomedical_science_pathway = `Biomedical.Science`,
#     D_engineering_pathway = `Engineering`,
#     D_cybersecurity_pathway = `CyberSecurity`,
#     .keep = "none" #"Keep-None" here will ONLY keep the columns listed above, so if you do not list a column, it will be deleted. Make sure to keep any district specific columns or other columns you need for your analysis.
# 
#   )
# print(pathways_companion)
# 
# #and check that the number of rows has not changed
# nrow(pathways_companion)
    
```

#### b. Evaluate pathways courses and N/A's and blank rows

-   **What:** Check the number of courses per pathway, and check for duplicates and N/A values in course Ids

-   **Check:** Review rows/data for any missing data.

-   **Project Action:** Gut check/track

```{r}
# #First lets count the number of courses per pathway by looking at the number of rows in each pathway with a value of "Y"
# pathways_companion_counts <- pathways_companion %>%
#   summarise(
#     D_auto_pathway = sum(D_auto_pathway == "Y"),
#     D_animal_science_pathway = sum(D_animal_science_pathway == "Y"),
#     D_applied_agricultural_engineering_pathway = sum(D_applied_agricultural_engineering_pathway == "Y"),
#     D_plant_science_pathway = sum(D_plant_science_pathway == "Y"),
#     D_carpentry_pathway = sum(D_carpentry_pathway == "Y"),
#     D_digital_communication_pathway = sum(D_digital_communication_pathway == "Y"),
#     D_graphic_arts_pathway = sum(D_graphic_arts_pathway == "Y"),
#     D_animation_pathway = sum(D_animation_pathway == "Y"),
#     D_commercial_photography_pathway = sum(D_commercial_photography_pathway == "Y"),
#     D_accounting_and_financial_services_pathway = sum(D_accounting_and_financial_services_pathway == "Y"),
#     D_early_learning_pathway = sum(D_early_learning_pathway == "Y"),
#     D_healthcare_theraputic_pathway = sum(D_healthcare_theraputic_pathway == "Y"),
#     D_culinary_arts_pathway = sum(D_culinary_arts_pathway == "Y"),
#     D_web_development_pathway = sum(D_web_development_pathway == "Y"),
#     D_law_enforcement_pathway = sum(D_law_enforcement_pathway == "Y"),
#     D_cosmetology_pathway = sum(D_cosmetology_pathway == "Y"),
#     D_advadnced_manufacturing_and_machinery_pathway = sum(D_advadnced_manufacturing_and_machinery_pathway == "Y"),
#     D_biomedical_science_pathway = sum(D_biomedical_science_pathway == "Y"),
#     D_engineering_pathway = sum(D_engineering_pathway == "Y"),
#     D_cybersecurity_pathway = sum(D_cybersecurity_pathway == "Y")
#   )
# print(pathways_companion_counts)
# 
# 
# #Now lets check N/A rows- this would result from any courses without a course ID in your pathways companion
# pathways_companion %>%
#   filter(is.na(D_course_id))
# 
# #and duplicate courses for both course ID and Course Name (should be zero)
#  pathways_companion %>%
#   group_by(D_course_id) %>%
#   filter(n() > 1) %>%
#   ungroup()

```

# Part 3: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <- "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"

ers_write_sharepoint(data = course_catalog_all_years,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "MNPS_Course_Catalog_Validated.csv",
                     drive_name = "client_work_drive")
```

## Part 4: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval=FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
