---
title: "1_raw_data_transcript_validation_Mnps"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: July 23 2024
review:
last update: April 10 2025
output: html_document 
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Start fresh
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
# Connect to SharePoint where the data files live
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
orgfiles_site     <- get_sharepoint_site(site_url = orgfiles_site_url)
internal_drive    <- orgfiles_site$get_drive("Internal")

# Read the district config — tells us where to find files, how to rename columns, etc.
config <- yaml::read_yaml(here::here("config", "districts", "tr_nashville.yaml"))
raw_data_folder_path <- config$raw_data_folder_path
source_of_truth_folder_path <- config$source_of_truth_folder_path
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. If your data is spread out over multiple files - i.e., **one file per year, go to step 1.1**

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning=F, message=F}

# Load the raw transcript data
transcript_raw <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = config$raw_files$transcript_raw
)

# Clean column names to remove spaces and special characters
transcript_raw <- transcript_raw %>% clean_names()

# #Added after the fact for SE to do some validation on past teams course coding!
course_catalog <- ers_read_sharepoint(
  folder_path = source_of_truth_folder_path,
  file_name_with_extension = config$source_of_truth_files$course_catalog
)
```


## Part 2: Validate Data

### 2.1 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. See resolution of critical/important fields that are missing in step 2.2b. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present

-   **Project Team Action:** Definitely change (see guidance in code block)

```{r}
# Rename transcript columns to standard names from tr_[district].yaml
transcript_all_yrs <- transcript_raw |>
  rename(any_of(unlist(config$column_map$transcript)))

#and lets see what we can do about the term columns.First, lets see if the course_name column contains "S1", "S2", And for quarters
transcript_all_yrs %>%
  mutate(last2 = str_sub(D_course_name, -2)) %>%   
  count(last2, sort = TRUE)  

#okay so its mostly S1 and S2


#I want a function that looks at course name, and if the courses name contains "S1" or S2", bring it into the new column
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(
    D_term = case_when(
      str_detect(D_course_name, regex("S1", ignore_case = TRUE)) ~ "Semester 1",
      str_detect(D_course_name, regex("S2", ignore_case = TRUE)) ~ "Semester 2",
      TRUE ~ NA_character_  # Default case if no match is found
    )
  )
```

#### b. Review columns

-   **What:** Manually check that you have the columns you need/would expect after renaming, and have not lost any critical data. Review the columns you are missing (as necessary).

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them for later.

### 2.1 Manual Checks

-   **What:** a handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input them into a checks table in OneNote, Word, or another tool.

    -   Do we have right year(s) of data?

        -   Check: Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

    -   Does term signify years, semesters, or quarters? What is the relationship between credits and term?

        -   Check: look at the term column to determine how term is used, and how many credits are typically earned by course. For example, if there is one row per course per semester, you would expect to see that courses are worth 0.5 credits

<!-- -->

-   Are there types of courses that tend to be worth more credits?

-   **Check:** Continue to build understanding of how credits are organized. LOok for differences by types of courses. For example, CTE or practicum courses tend to be worth more credits. That is okay, right now you are just trying to understand the structure of the file

-   **Project Team Action:** Gut check/track

```{r}
#group by credits earned and ensure that values match course notation (ie. semester courses carry .5, etc.)
transcript_all_yrs %>%
  group_by(D_creds_earn, D_term) %>%
  summarize(n = n_distinct((D_course_name))) %>% 
  pivot_wider(D_creds_earn, names_from = D_term, values_from = n) %>%
  ungroup()

```

### 2.2 Understand File Structure

#### a. Number of Total Rows

-   **What:** Understand the structure of your file by loading the number of rows

-   **Check:** The number of rows depends on how many courses you expect the student to have per year, which is explained by how many terms/periods/rotations are covered in the data. For example, if classes are organized by semester across 8 periods and repeat daily, you expect to see the number of rows be equal to 2 \* 8 \* 1 \* count of high school students. If classes are organized by quarter across 7 periods and repeat daily, you expect to see the number of rows be equal to 4 \* 7 \* 1 \* count of high school students

-   **Project Team Action:** update the Data Validation Checks table with the result of your check. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY total high school students per year and classes happen on a X term / X rotation / X period format"

```{r}
# Count number of rows
transcript_all_yrs %>% nrow() #View output of check to know what to write for your status column

#quick check before anything else- how many unique course IDs?
transcript_all_yrs %>%
  summarise(unique_course_ids = n_distinct(D_course_id)) %>% 
  print()

#how many unique course names?
transcript_all_yrs %>%
  summarise(unique_course_names = n_distinct(D_course_name)) %>% 
  print()

#can I get a list of course IDs with more than 1 unique course name?
transcript_all_yrs %>%
  group_by(D_course_id) %>%
  summarise(unique_course_names = n_distinct(D_course_name)) %>% 
  filter(unique_course_names > 1) %>% 
  print()

#filter to check this course AAD3001S1
transcript_all_yrs %>%
  filter(D_course_id == "AAD3001S1") %>% 
  select(D_course_name, D_course_id) %>% 
  distinct() %>% 
  print()

#Are any course IDs in Catalog NA?
course_catalog %>%
  filter(is.na(C_course_id_clean)) %>% 
  select(C_course_name, C_course_id_clean) %>% 
  distinct() %>% 
  print()

#Are any course IDs in the transcript NA?
transcript_all_yrs %>%
  filter(is.na(D_course_id)) %>% 
  select(D_course_name, D_course_id) %>% 
  distinct() %>% 
  print()

```

#### b. Remove Rows

-   **What:** If any of the rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one, your data has any totally or almost totally blank rows, etc.

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before.

-   **Project Team Action:** add code as required

```{r}
#Add code/comments here if resolution is required

#Remove any rows in transcript where the course ID is Na
transcript_all_yrs <- transcript_all_yrs %>%
  filter(!is.na(D_course_id))

#remove rows
# Remove rows where a specific condition is met
# For example, remove rows where age is less than 18

# transcript_all_yrs <- transcript_all_yrs %>%
#   filter(age >= 18)
# 
# # Print the result
# print(head(transcript_all_yrs))

```

#### c. Review current columns

-   **What:** View the data as a table to understand the values in each column

-   **Check:** A new tab is created with a table view of your data

-   **Project Team Action:** Do not change (run as is)

```{r}
#move data viewer to another tab to view both at once. Click on the "show in new window" button in the top left of the data viewer tab (just to the right of the left and right arrows)
transcript_all_yrs %>% View
```

### 2.5 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting

```{r}
#View duplicates
transcript_all_yrs %>%
  group_by_all() %>%
 mutate(n = n()) %>%
    filter(n>1) %>% 
    ungroup()
  
```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates.

-   **Check:** That the number of duplicates is now 0.

-   **Project Team Action:** None

```{r}
#removes duplicates
transcript_all_yrs <- transcript_all_yrs[!duplicated(transcript_all_yrs), ]

```

### 2.6 Missing & Unique values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action:** None (run as is)

```{r}
# Update blanks or NULL to NA
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** (1) Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates

```{r}
#create new data frame with column information
col_info <-  data.frame(
  Column_Name = names(transcript_all_yrs),
  Data_Type = sapply(transcript_all_yrs, class),
  Count_Rows = sapply(transcript_all_yrs, function(x) length(na.omit(x))),
  NA_Rows = sapply(transcript_all_yrs, function(x) sum(is.na(x))),
  Unique_Rows = sapply(transcript_all_yrs, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)

#calc % missing
col_info <- col_info %>%
            mutate(Percent_Missing = NA_Rows / Count_Rows * 100)

#print new data frame
col_info

## Same thing as above, but instead of NA values, look at 0 values
#create new data frame with column information
col_info_zero <-  data.frame(
  Column_Name = names(transcript_all_yrs),
  Data_Type = sapply(transcript_all_yrs, class),
  Count_Zero = sapply(transcript_all_yrs, function(x) sum(x == 0)), # Count zero values
  Total_Rows = nrow(transcript_all_yrs), # Total number of rows in transcript_combined_file
  stringsAsFactors = FALSE,
  row.names= NULL
)

col_info_zero <- col_info_zero %>%
  mutate(Percent_Zero = Count_Zero / Total_Rows * 100)

#calc % missing
col_info <- col_info %>%
            mutate(Percent_Missing = NA_Rows / Count_Rows * 100)

#print new data frame
print(col_info)

# In the chunk for 2.6b, after col_info is created and printed:
ggplot(col_info, aes(x = reorder(Column_Name, Percent_Missing), y = Percent_Missing)) +
  geom_bar(stat = "identity") +
  coord_flip() + # Makes it easier to read column names
  labs(title = "Percentage of Missing Values per Column",
       x = "Column Name",
       y = "Percent Missing") +
  theme_minimal()

# Similarly for col_info_zero:
ggplot(col_info_zero, aes(x = reorder(Column_Name, Percent_Zero), y = Percent_Zero)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Percentage of Zero Values per Column",
       x = "Column Name",
       y = "Percent Zero") +
  theme_minimal()


#lets look at some of these NA student grade level rows?
transcript_all_yrs %>%
  filter(!is.na(D_stu_grade_level)) %>%
  distinct(D_stu_grade_level) %>%
  arrange(D_stu_grade_level)

```

#### c. Review missing data for high schools / schools of interest- Can't do this for MNPS because we don't have location ID

-   **What:** Understand if we have any blanks or NAs in key fields for schools of interest (it is possible that the district provided data for other locations that aren't actually schools and do not have section IDs). If you don't have or need section IDs, you can skip this step.

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like student ID and D_location match expectations?

-   **Project Team Action:** (1) Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates

```{r}
# #investigating school codes with the most section values
# #group by D_location_id and count unique D_course_section values within each group
# transcript_all_yrs %>%
#   group_by(D_location_id) %>%
#   summarise(section_count = n_distinct(D_course_section))

```

#### d. List the counts of students per course and counts of students missing sections for each course- also skipping this for MNPS, not sure what this is doing?

-   **What:** Investigating students who are missing section IDs in schools of interest further. With the printed result. Filtered / sorted to look at the courses with the highest % of students missing sections. See if the students who weren't assigned sections were in really small sections where there would only be one section.If you don't have or need section IDs, you can skip this step.

-   **Check:** Are there any irregularities or disparities in enrollment that would cause concern for the integrity of the data?

-   **Project Team Action:** Follow up as necessary

```{r}
#Group the data by D_location_id and D_course_name, then summarize to get counts

# # If course ID is zero, set it to NA
# transcript_all_yrs <- transcript_all_yrs %>%
#   mutate(D_course_id = ifelse(D_course_id == 0, NA, D_course_id))
# 
# result_missing_sections <- transcript_all_yrs %>%
#   group_by(D_end_year, D_course_name) %>%
#   summarize(
#     count_students_per_course = n(),
#     count_students_missing_sections = sum(is.na(D_course_section))
#   ) %>%
#   ungroup() %>%
#   mutate(
#     count_students_per_course = ifelse(is.na(count_students_per_course), 0, count_students_per_course),
#     count_students_missing_sections = ifelse(is.na(count_students_missing_sections), 0,     
#     count_students_missing_sections),
#     percent_missing_sections = (count_students_missing_sections / count_students_per_course) * 100
#   )
# 
# #Print the result
# print(result_missing_sections)
```

#### 2.8. Counts by Terms and Course

-   **What:** Review the number of rows by term to ensure that we have data for the whole entire year.

-   **Check:** Are there any irregularities or disparities in term or course counts that may indicate missing data?

-   **Project Team Action:** Follow up as necessary

```{r}
#group by term and count the number of unique values in that column
transcript_all_yrs %>%
  group_by(D_term) %>%
  summarize(n = n()) %>% 
  ungroup()

#group by course name and list the unique names of each course. Are they separated out by semesters?
transcript_all_yrs %>%
  group_by(D_course_name) %>%
  summarize(n = n()) %>% 
  ungroup()

#group by course name and list the unique names of each course. Are they separated out by semesters?
transcript_all_yrs %>%
  group_by(D_course_name,D_term) %>%
  summarize(n = n()) %>% 
  pivot_wider(D_course_name,names_from = D_term,values_from = n) %>%
  ungroup()

```

### 2.7 Student counts by school by year

-   **What:** Count the number of students at each school per year to build confidence in data

-   **Check:** Are there any irregularities or disparities in enrollment or schools that would cause concern for the integrity of the data?

-   **Project Team Action:** Follow up as necessary

```{r}
#count the number of students (D_stu_id) at each school (D_location_id) per year (D_year)
students_count_by_school_year <- transcript_all_yrs %>%
  group_by(D_end_year) %>%
  summarise(
    students_count = n_distinct(D_stu_id)
  ) %>% 
  ungroup()

#using pivot wider, print the results to view number of students by year in a simple table
students_count_by_school_year <- students_count_by_school_year %>%
  pivot_wider(names_from = D_end_year, values_from = students_count) %>% 
  print()

```

### 2.8 Distinct count of students by location id

-   **What:** Count of students at each school. This check might helpful confirm any exclude/include schools for your teams analysis (like alternative schools).

-   **Check:** Check across other files while you validate them, and if necessary exclude schools with low student counts.

-   **Project Action:** Gut check/track

```{r}
#calculate the total count of distinct D_stu_id across all locations
transcript_all_yrs %>%
  summarise(
   total_students_count = n_distinct(D_stu_id)
  ) 

# #count the number of students (D_stu_id) at each location D_location_id)
# transcript_all_yrs %>%
#   group_by(D_location_id, D_location_name) %>%
#   summarise(
#     students_count = n_distinct(D_stu_id)
#   ) %>% 
#   arrange(desc(students_count)) %>% 
#   ungroup()

```

### 2.9 Distribution of students by Term

-   **What:** Count of records by term at each school. Given your project needs, you may want to just filter your original data-frame (transcript_combined_file) to only include high schools of interest for this step. If you are looking at all locations, you can skip this step.

-   **Check:** Review the tables to understand if there is an even distribution of semester 1 and semester 2 classes

-   **Project Action:** Update the numbers in the first line of code to match the location codes in your district's data

```{r}
# #Filtering only to high schools of interest for this step, but you might choose to look at distribution of students by term for all locations based on project needs
# 
# filtered_high_schools <- transcript_all_yrs[transcript_all_yrs$D_location_id %in% 
#                                               c(1, 2, 3),]
# 
# #View count of records by D_term/school
# filtered_high_schools %>%
#   filter(!is.na(D_term)) %>%
#   group_by(D_location_id, D_term) %>% 
#   count(sort = TRUE) %>%
#   arrange(D_location_id) %>%
#   group_by(D_location_id) %>% 
#   mutate(perc_D_term_at_school = round(100 * n / sum(n), 1)) %>%
#   select(D_location_id, D_term, perc_D_term_at_school) %>%
#   pivot_wider(names_from = D_term, 
#               values_from = perc_D_term_at_school, 
#               names_prefix = "D_term_") %>%
#   rename_all(~ ifelse(. == "", "D_term_unknown", .)) %>% 
#   ungroup()
# 
# filtered_high_schools %>% 
#   group_by(D_location_id) %>%
#   mutate(student_count_school = n_distinct(D_stu_id)) %>%
#   group_by(D_term, D_location_id, .add = TRUE) %>%
#   summarize(
#     student_count = n_distinct(D_stu_id),
#     student_count_school = first(student_count_school),
#     percent_students = round(100 * n_distinct(D_stu_id) / first(student_count_school), 0)
#   ) %>%
#   ungroup() %>%
#   arrange(D_location_id, D_term)

```

### 2.10 Distinct count of course IDs by student id

-   **What:** Count of course IDs by student id. If you did not filter any high schools in step 2.9, you may need to change the name of your data-frame.

-   **Check:** Review the counts courses by student id. You would expect to a mix since you have freshmen, sophomore, junior, and senior data across five years. You would also expect to see a maximum of \~# of classes students can take per year \*4

-   **Project Action:** No updates

```{r}
course_bucket_summary <- transcript_all_yrs %>% 
  group_by(D_stu_id) %>%
  summarize(Course_Count = n_distinct(D_course_name), .groups = "drop") %>% 
  mutate(Course_Count_Bucket = case_when(
    Course_Count >= 0  & Course_Count <= 6  ~ "00-06",
    Course_Count > 6  & Course_Count <= 12 ~ "07-12",
    Course_Count > 12 & Course_Count <= 18 ~ "13-18",
    Course_Count > 18 & Course_Count <= 24 ~ "19-24",
    Course_Count > 24 & Course_Count <= 30 ~ "25-30",
    Course_Count > 30 & Course_Count <= 36 ~ "31-36",
    Course_Count > 36 & Course_Count <= 42 ~ "37-42",
    Course_Count > 42 & Course_Count <= 48 ~ "43-48",
    Course_Count > 48                      ~ "49+"
  ))

# Count number of students in each bucket
bucket_counts <- course_bucket_summary %>%
  count(Course_Count_Bucket, name = "n_students") %>%
  mutate(total_students = sum(n_students),
         pct_of_total = round(100 * n_students / total_students, 2))

print(bucket_counts)

#bucket_counts is viewed:
ggplot(bucket_counts, aes(x = Course_Count_Bucket, y = n_students)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Students by Number of Courses Taken",
       x = "Number of Courses Bucket",
       y = "Number of Students") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels if crowded

#whats the average number of courses a student takes per year
median_courses_per_year <- transcript_all_yrs %>%
  group_by(D_stu_id, D_end_year) %>%
  summarize(Courses_Per_Year = n_distinct(D_course_name), .groups = "drop") %>%
  group_by(D_end_year) %>%
  summarize(Median_Courses = median(Courses_Per_Year, na.rm = TRUE), .groups = "drop")

print(median_courses_per_year)
```

### 2.101- MNPS Simon Added- count of students by credit earning
```{r}

course_bucket_summary <- transcript_all_yrs %>% 
  group_by(D_stu_id) %>%
  summarize(Total_Credits = sum(D_creds_earn, na.rm = TRUE), .groups = "drop") %>% 
  mutate(Credit_Bucket = case_when(
    Total_Credits >= 0  & Total_Credits <= 6  ~ "00-06",
    Total_Credits > 6  & Total_Credits <= 12 ~ "07-12",
    Total_Credits > 12 & Total_Credits <= 18 ~ "13-18",
    Total_Credits > 18 & Total_Credits <= 24 ~ "19-24",
    Total_Credits > 24 & Total_Credits <= 30 ~ "25-30",
    Total_Credits > 30 & Total_Credits <= 36 ~ "31-36",
    Total_Credits > 36 & Total_Credits <= 42 ~ "37-42",
    Total_Credits > 42 & Total_Credits <= 48 ~ "43-48",
    Total_Credits > 48                      ~ "49+"
  ))

# Count number of students in each bucket
bucket_counts <- course_bucket_summary %>%
  count(Credit_Bucket, name = "n_students") %>%
  mutate(total_students = sum(n_students),
         pct_of_total = round(100 * n_students / total_students, 2))

print(bucket_counts)

# Plot
ggplot(bucket_counts, aes(x = Credit_Bucket, y = n_students)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Students by Total Credits Earned",
       x = "Total Credits Bucket",
       y = "Number of Students") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



### 2.11 Distinct count of students per Section

-   **What:** Returns the number of students per section per school.If you do not have section IDs in your data or need them for analysis, you can skip this step.

-   **Check:** Visually investigate the resulting enrollment_count_by_location_id table. You are looking for wonky things - really big core sections, really small PE sections. We want to ensure we have complete data. Sort count of students per section from highest to lowest. Which courses have the largest sections? Which courses have the smallest sections?

-   **Project Action:** None

```{r}
#calculate the number of students enrolled in each course and section
transcript_all_yrs %>%
  group_by(D_course_name, D_term, D_end_year) %>%
  summarise(
    students_enrolled = n_distinct(D_stu_id)
  ) %>%
  ungroup() 

## Count the number of rows where course_section is equal to zero. 
transcript_all_yrs %>% 
  filter(D_term == 0) %>% 
  count()

# calculate the percent of this out of total rows
transcript_all_yrs %>% 
  filter(D_term == 0) %>% 
  count() / nrow(transcript_all_yrs)

# write a table with the count of values of D_course_section that equal zero by school_year
transcript_all_yrs %>% 
  filter(D_term == 0) %>% 
  group_by(D_end_year) %>% 
  count()

#I feel like this isn't fully showing me what I want it to. Lets look at the courses with the most unique students enrolled, arranged desc order
enrollment_count_by_location_id <- transcript_all_yrs %>%
  group_by(D_course_name, D_term) %>%
  summarise(
    enrollment_count = n_distinct(D_stu_id),
    .groups = "drop"
  ) %>%
  arrange(desc(enrollment_count)) %>% 
  print()

#What is the number of unique course names
enrollment_count_by_location_id %>%
  summarise(unique_course_names = n_distinct(D_course_name)) %>% 
  print()

#what is the number of unique course IDs
enrollment_count_by_location_id %>%
  summarise(unique_course_ids = n_distinct(D_course_name)) %>% 
  print()

```

### 2.12 MNPS - Testing Course Catalog
```{r}
# Step 1. Clean transcript course IDs (strip "S1"/"S2" if present)
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(course_id_cleaned = str_replace(D_course_id, "([[:punct:]]?\\s*)?(S1|S2)\\s*$", ""))

# Step 2. Get distinct IDs from both sources
course_catalog_ids <- unique(course_catalog$C_course_id_clean)
transcript_ids <- unique(transcript_all_yrs$course_id_cleaned)

# Step 3. Compare
ids_in_both       <- intersect(transcript_ids, course_catalog_ids)
ids_only_catalog  <- setdiff(course_catalog_ids, transcript_ids)
ids_only_transcript <- setdiff(transcript_ids, course_catalog_ids)

# Step 4. Counts
cat("Total unique IDs in catalog:", length(course_catalog_ids), "\n")
cat("Total unique IDs in transcript:", length(transcript_ids), "\n")
cat("IDs present in BOTH:", length(ids_in_both), "\n")
cat("IDs only in catalog:", length(ids_only_catalog), "\n")
cat("IDs only in transcript:", length(ids_only_transcript), "\n")

# Step 5. Optional: Get list of transcript courses missing from catalog
transcript_courses_not_in_catalog <- transcript_all_yrs %>%
  filter(course_id_cleaned %in% ids_only_transcript) %>%
  select(course_id_cleaned, D_course_name) %>%
  distinct()

#Create a data-frame of the unique course IDs that are only in the transcript data, not in the course catalog
transcript_courses_not_in_catalog <- transcript_all_yrs %>%
  filter(course_id_cleaned %in% ids_only_transcript) %>%
  select(course_id_cleaned, D_course_name) %>%
  distinct() %>%
  arrange(course_id_cleaned) %>%
  print()

```

###2.13- MNPS - Testing Course Names
```{r}
library(fuzzyjoin)

# Step 1. Clean transcript course names (remove trailing " S1"/" S2" if present)
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(course_name_cleaned = str_remove(D_course_name, " S?([12])\\s*\\*?$"))

#Okay great. And just as a double-check, lets look in the course_name_cleaned column and see if there are any instances of courses that still contain S1 or S2
transcript_all_yrs %>%
  filter(str_detect(course_name_cleaned, " S1| S2")) %>%
  select(D_course_name, course_name_cleaned) %>%
  distinct() %>%
  print()

# Step 2. Get distinct names from both sources
catalog_names <- unique(course_catalog$C_course_name)
transcript_names <- unique(transcript_all_yrs$course_name_cleaned)

# Step 3. Compare
names_in_both        <- intersect(transcript_names, catalog_names)
names_only_catalog   <- setdiff(catalog_names, transcript_names)
names_only_transcript <- setdiff(transcript_names, catalog_names)

# Step 4. Counts
cat("Total unique names in catalog:", length(catalog_names), "\n")
cat("Total unique names in transcript:", length(transcript_names), "\n")
cat("Names present in BOTH:", length(names_in_both), "\n")
cat("Names only in catalog:", length(names_only_catalog), "\n")
cat("Names only in transcript:", length(names_only_transcript), "\n")

# Step 5. Optional: Get list of transcript courses missing from catalog
transcript_course_names_not_in_catalog <- transcript_all_yrs %>%
  filter(course_name_cleaned %in% names_only_transcript) %>%
  select(course_name_cleaned, D_course_id) %>%
  distinct()

  
# Trying strict cleaning then comparing
clean_strings <- function(x) {
  x %>%
    str_replace_all("[[:space:]]+", " ") %>%   # collapse ALL whitespace (even tabs, non-breaking)
    str_trim() %>%                            # remove leading/trailing
    str_to_lower() %>%                        # lowercase
    stringi::stri_trans_general("latin-ascii") # remove accents/diacritics
}

catalog_names    <- clean_strings(unique(course_catalog$C_course_name))
transcript_names <- clean_strings(unique(transcript_all_yrs$course_name_cleaned))

names_only_transcript <- setdiff(transcript_names, catalog_names)
length(names_only_transcript)

# Compare again
names_in_both        <- intersect(transcript_names, catalog_names)
names_only_catalog   <- setdiff(catalog_names, transcript_names)
names_only_transcript <- setdiff(transcript_names, catalog_names)

cat("Catalog:", length(catalog_names), "\n")
cat("Transcript:", length(transcript_names), "\n")
cat("Both:", length(names_in_both), "\n")
cat("Only catalog:", length(names_only_catalog), "\n")
cat("Only transcript:", length(names_only_transcript), "\n")

```

### 2.14- MNPS Students with multiple grades in the same year
```{r}
#testing the credit assumptions hypothesis for students with multiple grades in the same year! First, create a data-frame of students who had more than one grade level in the same year
students_multiple_grades <- transcript_all_yrs %>%
  group_by(D_stu_id, D_end_year) %>%
  filter(n_distinct(D_stu_grade_level) > 1) %>%
  ungroup() %>% 
  distinct()

#and how many students is it? (knowing the number will be different than student!
students_multiple_grades %>%
  summarise(num_students = n_distinct(D_stu_id)) %>%
  print()

#create a new column called "credit failed?" with a "Y" if D_creds_earn is less than D_creds_possible, and "N" otherwise
students_multiple_grades <- students_multiple_grades %>%
  mutate(credit_failed = ifelse(D_creds_earn < D_creds_possible, "Y", "N"))

#and then count the number of students by that flag to see how many have failed enrollments
students_multiple_grades %>%
  group_by(credit_failed) %>%
  summarise(num_students = n_distinct(D_stu_id)) %>%
  ungroup() %>%
  print()

#hmmm, so 731/2, 468 which is only 30%, would expect it to be a bit higher. Lets try another way to look at this.

#In transcript all years lets create a new column that groups by student ID and sums total credits
transcript_all_yrs <- transcript_all_yrs %>%
  group_by(D_stu_id) %>%
  mutate(total_credits_earned = sum(D_creds_earn, na.rm = TRUE),
         total_credits_possible = sum(D_creds_possible, na.rm = TRUE)) %>%
  ungroup()

#and then one more new column that groups by student ID and year for total_credits_earned_year
transcript_all_yrs <- transcript_all_yrs %>%
  group_by(D_stu_id, D_end_year) %>%
  mutate(total_credits_earned_year = sum(D_creds_earn, na.rm = TRUE)) %>%
  ungroup()

#and in transcript all years create a Y/N flag for students with multiple grade levels in the same year
transcript_all_yrs <- transcript_all_yrs %>%
  group_by(D_stu_id, D_end_year) %>%
  mutate(multiple_grades_same_year = ifelse(n_distinct(D_stu_grade_level) > 1, "Y", "N")) %>%
  ungroup()

#now lets look at the average (median) number of credits earned per year for students by using the mutltiple_grades_same_year flag
median_credits_by_multiple_grades_flag <- transcript_all_yrs %>%
  group_by(multiple_grades_same_year, D_end_year) %>%
  summarize(median_credits_earned = median(total_credits_earned_year, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(multiple_grades_same_year, D_end_year) %>%
  print()

#so interesting, they are actually accumulating MORE credits. Lets do another check

#create a new column called "credit failed?" with a "Y" if D_creds_earn is less than D_creds_possible, and "N" otherwise
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(credit_failed = ifelse(D_creds_earn < D_creds_possible, "Y", "N"))


#and then count the number of students using the multiple grades flag that have any failures compared to the number of students with only one grade/year with any failures
failure_rate_test <- transcript_all_yrs %>%
  group_by(multiple_grades_same_year, credit_failed) %>%
  summarise(num_students = n_distinct(D_stu_id)) %>%
  arrange(multiple_grades_same_year, credit_failed) %>%
  print()

#so intersting, its 30% for both! Lets look at some specific student examples of students with multiple grades
stu_check_1 <- students_multiple_grades %>%
  filter(D_stu_id == "190000208") %>%
  print()

stu_check_2 <- students_multiple_grades %>%
  filter(D_stu_id == "190003328") %>%
  print()

stu_check_3 <- students_multiple_grades %>%
  filter(D_stu_id == "190044042") %>%
  print()

#what is the distribution of students by grade level in students with multiple grades?
students_multiple_grades %>%
  group_by(D_stu_grade_level) %>%
  summarise(num_students = n_distinct(D_stu_id)) %>%
  arrange(D_stu_grade_level) %>%
  print()

# Step 1: Keep all rows for students with multiple grades
students_multigrade_all <- transcript_all_yrs %>%
  filter(D_stu_id %in% students_multiple_grades$D_stu_id)

# Step 2: Identify all failed courses (any year)
failed_courses <- students_multigrade_all %>%
  filter(credit_failed == "Y") %>%
  select(D_stu_id,
         D_course_name,
         original_grade = D_stu_grade,
         original_year = D_end_year,
         original_grade_level = D_stu_grade_level)

# Step 3: Keep all courses to check for retakes (including failed ones)
all_courses <- students_multigrade_all %>%
  select(D_stu_id,
         D_course_name,
         retake_grade = D_stu_grade,
         retake_year = D_end_year,
         retake_grade_level = D_stu_grade_level)

# Step 4: Join failed courses with all courses to find retakes
retake_table <- failed_courses %>%
  inner_join(all_courses, by = c("D_stu_id", "D_course_name")) %>%
  # Remove rows where the failed course is the same attempt (same year)
  filter(retake_year != original_year) %>%
  arrange(D_stu_id, D_course_name, original_year, retake_year)

# Step 5: View the results
retake_table

#and how many unique stduents are there in the retake table?
unique_students_retake <- retake_table %>%
  summarise(unique_students = n_distinct(D_stu_id)) %>%
  print()

#only 281, out of how many total students with multiple grades?
transcript_all_yrs %>%
  filter(multiple_grades_same_year == "Y") %>%
  summarise(total_students = n_distinct(D_stu_id)) %>%
  print()

#1740, why is it a different number of students now? Oh yes because its transcript data. lets export the retake table with ERS write SharePoint

#Okay lets export this to see what we can match/add to the catalog
# processed_data_folder_path <- "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"
# 
# ers_write_sharepoint(data = retake_table,
#                      folder_path = processed_data_folder_path,
#                      file_name_with_extension = "student_retake_table.csv",
#                     drive_name = "client_work_drive")



# And because Rebel wanted to check, how many students have multiple grades in the same year in semester 2 specifically?
students_multiple_grades_semester_2 <- transcript_all_yrs %>%
  filter(D_term == "Semester 2") %>%
  group_by(D_stu_id, D_end_year) %>%
  filter(n_distinct(D_stu_grade_level) > 1) %>%
  ungroup() %>%
  summarise(total_students = n_distinct(D_stu_id)) %>% 
  print()

#1071

```

### 2.15 MNPS testing some missing course IDs
```{r}
#creating some quick data-frame checks to filter to students with enrollment in specific course IDs that were in the transcript data but missing in the course catalog.

#Create a new data-frame of any student who took course ID ELD5099
students_eld5099 <- transcript_all_yrs %>%
  filter(course_id_cleaned == "ELD5099") %>%
  select(D_stu_id, D_course_name, D_end_year, D_stu_grade_level) %>%
  distinct() %>%
  arrange(D_stu_id, D_end_year) %>% 
print()

#Create a new-data frame of any student who took course ID ENG1402S2Q4
students_eng1402s2q4 <- transcript_all_yrs %>%
  filter(course_id_cleaned== "ENG1402S2Q4") %>%
  select(D_stu_id, D_course_name, D_end_year, D_stu_grade_level) %>%
  distinct() %>%
  arrange(D_stu_id, D_end_year) %>%
print()


#Create a new-data frame of any student who took course ID ENG1923
students_ENG1923 <- transcript_all_yrs %>%
  filter(course_id_cleaned == "ENG1923") %>%
  select(D_stu_id, D_course_name, D_end_year, D_stu_grade_level) %>%
  distinct() %>% 
  arrange(D_stu_id, D_end_year) %>% 
  print()

#create a new-data frame of any student who took course ID MTH4111
students_MTH4111 <- transcript_all_yrs %>%
  filter(course_id_cleaned == "MTH4111") %>%
  select(D_stu_id, D_course_name, D_end_year, D_stu_grade_level) %>%
  distinct() %>%
  arrange(D_stu_id, D_end_year) %>% 
  print()

#create a new-data frame of any student who took course ID MZ59525AL
students_MZ59525AL <- transcript_all_yrs %>%
  filter(course_id_cleaned == "MZ59525AL") %>%
  select(D_stu_id, D_course_name, D_end_year, D_stu_grade_level) %>%
  distinct() %>%
  arrange(D_stu_id, D_end_year) %>% 
  print()

```

### 2.16 MNPS export course catalog for coding
```{r}
# Create df of unique D_course_id and D_course_name pairs
unique_course_catalog <- transcript_all_yrs %>%
  filter(!is.na(D_course_name) & !is.na(D_stu_grade)) %>%
  select(D_course_id, course_name_cleaned) %>%
  distinct() %>%
  group_by(D_course_id) %>%
  mutate(course_name_count = n()) %>%
  ungroup() |> 
  arrange(desc(course_name_count), D_course_id)
```

# Part 3: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
# Export validated transcript_all_yrs to SharePoint
ers_write_sharepoint(data = transcript_all_yrs,
                     folder_path = raw_data_folder_path,
                     file_name_with_extension = "/2. Processed Data/01_tr_transcript_validated.csv",
                     drive_name = "client_work_drive")

# Export unique course catalog to SharePoint
# ers_write_sharepoint(data = unique_course_catalog,
#                      folder_path = raw_data_folder_path,
#                      file_name_with_extension = "/1. Coding/course_catalog_2122_2425.csv",
#                      drive_name = "client_work_drive")
```

## Part 4: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval=FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
