---
title: "Transcript Pipeline"
subtitle: "Metro Nashville Public Schools"
date: today
params:
  district: "nashville"
  refresh_raw: false
  smoke_test: false
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE)
source(here::here("templates", "transcript", "00_tr_setup.R"))
```

# What this file is

One document, run top to bottom, replacing the eight scripts that used to run
in sequence. District-specific cleaning decisions live here or in
`config/districts/tr_nashville.yaml`. Reusable machinery lives in
`templates/transcript/00_tr_*.R`.

**Reading the output.** Every check prints as it runs:

- `[PASS]` — a structural check held.
- `[FAIL]` — a structural check broke. The render stops. Offending rows print,
  and stay available via `chk_offenders("<check name>")`.
- `[info]` — a count or table to eyeball. Never halts.

| # | Stage | Replaces | Status |
|---|-------|----------|--------|
| 0 | Setup and config | Part 0 of all eight | built |
| 1 | Load sources | Part 1 of all eight | built |
| 2 | Student cleaning | `01_raw_data_student_validation` | built |
| 3 | Transcript cleaning | `01_raw_data_transcript_validation` | built |
| 4 | Cross-file reconciliation | `02_cross_file_validation` §2.1 | next |
| 5 | Home school assignment | `02_cross_file_validation` §2.2 | |
| 6 | Cohort selection | `03_student_cohort_selection` | |
| 7 | Course coding join | new — the rollup that never existed | |
| 8 | Output and check report | Part 3/4 of all eight | partial |

---

# Stage 0 — Setup and configuration

```{r config}
chk_reset()
chk_stage("Stage 0 — Setup")

config <- load_config(district = params$district)

# Failed checks write their offending rows here before the run halts, because
# knitr discards a chunk's output when the chunk errors.
chk_set_output_dir(file.path(
  config$paths$output_dir %||% here::here("output", tolower(params$district)),
  "failed_checks"))

config_summary(config)
```

## 0.1 Run manifest

```{r run-manifest}
run_manifest <- tibble::tibble(
  run_time    = Sys.time(),
  district    = config$district,
  config_file = basename(config$.meta$path),
  config_md5  = substr(config$.meta$md5, 1, 8),
  refresh_raw = params$refresh_raw,
  r_version   = paste0(R.version$major, ".", R.version$minor),
  user        = Sys.info()[["user"]]
)
chk_info("Run stamped", run_manifest$config_md5)
run_manifest
```

## 0.2 The decisions this run is making

```{r config-decisions}
chk_info("Cohort entry", paste0("grade ", config$cohort$entry_grade,
                                " in ", config$cohort$entry_year))
chk_info("Years required", config$cohort$years_required)
chk_info("Duration rule", paste0(config$cohort$duration_rule, ", >= ",
                                 config$cohort$duration_min_days, " days/yr"))
chk_info("Progression rule", config$cohort$progression_rule)
chk_info("Days enrolled rule", config$cohort$days_enrolled_rule)
chk_info("Negative days treatment", config$derived$negative_days_treatment)
chk_info("Schools in scope", length(config$cohort$schools_to_include))
```

## 0.3 Check-framework self-test

```{r smoke-test}
#| eval: !expr params$smoke_test
toy <- tibble::tibble(D_stu_id = c(1, 1, 2, 3),
                      D_year   = c(2022, 2023, 2022, 2022),
                      score    = c(10, 20, 30, NA))
chk_unique(toy, c("D_stu_id", "D_year"), "Self-test: unique on id + year")
demo <- try(chk_no_na(toy, "score", "Self-test: deliberate failure"), silent = TRUE)
print(chk_offenders("Self-test: deliberate failure"))
chk_reset(); chk_stage("Stage 0 — Setup")
chk_info("Self-test complete", "framework verified, log reset")
```

---

# Stage 1 — Load sources

Each file is read from SharePoint once and cached as `.rds`. **The cache does
not know when the district re-sends a file.** Set `refresh_raw: true` when new
data lands.

```{r load-sources}
chk_stage("Stage 1 — Load sources")

transcript_raw   <- read_source("transcript_raw",   config, refresh = params$refresh_raw)
demographics_raw <- read_source("stu_demographics", config, refresh = params$refresh_raw)

course_coding_raw <- tryCatch(
  read_source("course_coding", config, refresh = params$refresh_raw,
              sheet = config$sheets$course_coding),
  error = function(e) {
    chk_info("Course coding file NOT loaded", "see detail", detail = conditionMessage(e))
    NULL
  }
)
```

## 1.1 Expected shape

```{r source-shape}
chk_stop("Transcript has expected columns",
         all(unlist(config$column_map$transcript) %in% names(transcript_raw)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$column_map$transcript), names(transcript_raw)),
           collapse = ", ")))

chk_stop("Demographics has expected columns",
         all(unlist(config$column_map$demographics) %in% names(demographics_raw)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$column_map$demographics), names(demographics_raw)),
           collapse = ", ")))

chk_info("Transcript rows",   nrow(transcript_raw))
chk_info("Demographics rows", nrow(demographics_raw))
```

---

# Stage 2 — Student cleaning

Ends at one row per student per enrolment spell. Collapsing to one row per
student-year happens at Stage 5, after home school is assigned.

## 2.1 Rename to ERS standard names

```{r stu-rename}
chk_stage("Stage 2 — Student cleaning")

stu <- demographics_raw |>
  dplyr::rename(dplyr::any_of(unlist(config$column_map$demographics))) |>
  dplyr::rename(!!config$grade_level_names$demographics := D_stu_grade_level)

chk_stop("Rename produced all standard columns",
         all(setdiff(names(config$column_map$demographics), "D_stu_grade_level")
             %in% names(stu)),
         value = paste0(ncol(stu), " columns"))

chk_no_na(stu, c("D_stu_id", "D_year"), "Student id and year always present")
chk_info("Distinct students in raw demographics", dplyr::n_distinct(stu$D_stu_id))
```

## 2.2 Dates

Where the old pipeline broke. `wdt` arrives as `m/d/yyyy`; the original asked
`as.Date()` for `%Y-%m-%d`, which returns `NA` silently, so every exit date and
therefore every duration was `NA`. Formats now come from config, unparseable
values are a structural failure, and the fill uses `coalesce()` rather than
`ifelse()` so the Date class survives.

```{r stu-dates}
stu <- stu |>
  dplyr::mutate(
    D_stu_enter_date = parse_date_strict(D_stu_enter_date,
                                         config$derived$date_formats$D_stu_enter_date,
                                         "enter date"),
    D_stu_exit_date  = parse_date_strict(D_stu_exit_date,
                                         config$derived$date_formats$D_stu_exit_date,
                                         "exit date"),
    C_exit_date_imputed = is.na(D_stu_exit_date)
  )

if (isTRUE(config$derived$force_missing_exit_date)) {
  stu <- stu |>
    dplyr::mutate(D_stu_exit_date = fill_missing_date(
      D_stu_exit_date,
      as.Date(paste0(D_year, "-", config$derived$school_year_end_mmdd)),
      label = "exit date -> year end"))
}

chk_no_na(stu, c("D_stu_enter_date", "D_stu_exit_date"), "Both dates present after fill")

chk_stop("Exit date is not before enter date",
         all(stu$D_stu_exit_date >= stu$D_stu_enter_date),
         offenders = stu |>
           dplyr::filter(D_stu_exit_date < D_stu_enter_date) |>
           dplyr::select(D_year, D_location_id, D_stu_enter_date, D_stu_exit_date))
```

`wdt` behaves like a withdrawal date, not a year-end date: spells that carry a
real one have a median of about five weeks. That means the duration threshold
is doing less work than it looks — it is close to separating "withdrew" from
"did not."

## 2.3 Days enrolled

Spells are clamped to the school year so an entry date carried over from a
prior year cannot produce a 1,200-day "year".

```{r stu-days}
stu <- stu |>
  dplyr::mutate(
    H_year_start = as.Date(paste0(D_year - 1, "-", config$derived$school_year_start_mmdd)),
    H_year_end   = as.Date(paste0(D_year,     "-", config$derived$school_year_end_mmdd)),
    C_spell_before_window = D_stu_exit_date  < H_year_start,
    C_spell_after_window  = D_stu_enter_date > H_year_end,
    H_enter_clamped = if (isTRUE(config$derived$clamp_days_to_school_year))
      pmax(D_stu_enter_date, H_year_start) else D_stu_enter_date,
    H_exit_clamped  = if (isTRUE(config$derived$clamp_days_to_school_year))
      pmin(D_stu_exit_date,  H_year_end)   else D_stu_exit_date,
    M_days_enrolled = as.numeric(H_exit_clamped - H_enter_clamped)
  )

chk_info("Spells clamped at the start of the window",
         sum(stu$D_stu_enter_date < stu$H_year_start))
chk_info("Spells clamped at the end of the window",
         sum(stu$D_stu_exit_date > stu$H_year_end))
```

### Spells that fall outside the school year

These are what produced negative durations on the first run. Entry dates in
early July, recorded against the year that had just ended.

```{r stu-outside-window}
outside <- stu |>
  dplyr::filter(C_spell_before_window | C_spell_after_window) |>
  dplyr::select(D_year, D_location_id, D_stu_enter_date, D_stu_exit_date,
                C_exit_date_imputed, C_spell_before_window, C_spell_after_window,
                M_days_enrolled)

chk_info("Spells entirely outside the school year", nrow(outside),
         data = outside |>
           dplyr::count(D_year, C_spell_before_window, C_spell_after_window,
                        name = "n_rows"))

outside |> dplyr::arrange(M_days_enrolled) |> utils::head(20)
```

```{r stu-negative-treatment}
treatment <- config$derived$negative_days_treatment

if (treatment == "drop") {
  before <- nrow(stu)
  stu <- stu |> dplyr::filter(M_days_enrolled >= 0)
  chk_info("Negative-duration spells dropped", before - nrow(stu))
} else if (treatment == "floor_zero") {
  n_floored <- sum(stu$M_days_enrolled < 0)
  stu <- stu |> dplyr::mutate(M_days_enrolled = pmax(M_days_enrolled, 0))
  chk_info("Negative-duration spells floored to zero", n_floored)
} else {
  chk_info("Negative-duration spells left as-is", sum(stu$M_days_enrolled < 0))
}

guard <- range_two_tier(stu$M_days_enrolled, config$ranges$M_days_enrolled,
                        "M_days_enrolled")
stu$M_days_enrolled <- guard$values

chk_stop("Days enrolled is never NA", !any(is.na(stu$M_days_enrolled)))
```

```{r stu-days-by-imputation}
stu |>
  dplyr::group_by(C_exit_date_imputed) |>
  dplyr::summarise(
    n_rows       = dplyr::n(),
    median_days  = stats::median(M_days_enrolled),
    pct_over_thr = round(100 * mean(M_days_enrolled >= config$cohort$duration_min_days), 1),
    .groups = "drop")
```

The threshold counts **calendar** days, not instructional ones. A typical year
has about 180 instructional days, so "200 instructional days" could never be
met by anyone — 200 can only mean calendar days.

What it actually means here: for the 75.5% of rows whose exit date is imputed
to 30 June, `M_days_enrolled` is simply `30 June - entry date`, and 30 June
minus 200 days is exactly **12 December**. So for three quarters of the file
the rule reads "enrolled on or before 12 December". For the rest, which carry a
real withdrawal date at a median of about five weeks, almost none clear it.

Worth writing up that way rather than as "enrolled 200 days", which implies a
measure of attendance this data cannot support. A genuine instructional-day
threshold would need a district school-calendar table.

## 2.4 Locations

Names carry a year stamp, are truncated at ~24 characters, and change spelling
between years. Codes are stable, so names resolve to codes once and everything
downstream matches on the code.

```{r stu-locations}
stu <- stu |>
  dplyr::mutate(
    C_location_name = clean_location_name(
      D_location_name,
      strip_prefix = config$derived$location_name_strip_prefix,
      strip_suffix = config$derived$location_name_strip_suffix,
      aliases      = config$cohort$school_name_aliases),
    C_credit_recovery_site = grepl(config$cohort$credit_recovery_name_pattern,
                                   D_location_name)
  )

chk_info("Credit recovery site rows", sum(stu$C_credit_recovery_site))

stu <- apply_location_merges(stu,
  id_col = "D_location_id", name_col = "C_location_name",
  year_col = "D_year", merges = config$cohort$location_merges)

location_xwalk <- build_location_crosswalk(
  stu, "C_location_id", "C_location_name_merged",
  include = config$cohort$schools_to_include)

included_location_ids <- attr(location_xwalk, "included_ids")
```

Hillwood and James Lawson are one institution across time, not two schools.
MNPS relocated the whole campus — all non-graduating students and the teaching
staff moved from the Hillwood building to the new Bellevue facility for August
2023 — so a student appearing at both is continuous enrolment, not a transfer.
`location_merges` remaps Hillwood's code onto James Lawson's, leaving the raw
`D_location_id` untouched and putting the merged code in `C_location_id`.

The merge is verified rather than trusted: each name must resolve to exactly
one code, and the two must not operate in the same year. Concurrent operation
would mean two schools rather than one relocation, and the run would stop.

Credit-recovery sites already carry the **parent school's** code — `S2022 CR
Cane Ridge HS` is code 182, the same as Cane Ridge High School — so matching on
the code rolls them up with no extra work. Only the display name needed fixing,
which is what canonical naming does. The flag is kept so they stay findable.

```{r stu-canonical-names}
stu <- resolve_canonical_names(stu, "C_location_id", "C_location_name_merged")

stu <- stu |>
  dplyr::mutate(C_school_in_scope = C_location_id %in% included_location_ids)

chk_stop("Credit recovery rows resolve to a parent school",
         all(stu$C_location_name_canonical[stu$C_credit_recovery_site] !=
             stu$C_location_name[stu$C_credit_recovery_site]),
         value = paste0(sum(stu$C_credit_recovery_site), " row(s)"),
         offenders = stu |>
           dplyr::filter(C_credit_recovery_site) |>
           dplyr::distinct(C_location_id, C_location_name, C_location_name_canonical))

chk_info("Rows at in-scope schools",
         paste0(format(sum(stu$C_school_in_scope), big.mark = ","), " of ",
                format(nrow(stu), big.mark = ","), " (",
                round(100 * mean(stu$C_school_in_scope), 1), "%)"))

stu |>
  dplyr::filter(C_credit_recovery_site) |>
  dplyr::count(C_location_id, C_location_name, C_location_name_canonical,
               C_school_in_scope, name = "n_rows")
```

```{r stu-inscope-table}
stu |>
  dplyr::filter(C_school_in_scope) |>
  dplyr::count(C_location_id, C_location_name_canonical, name = "n_rows") |>
  dplyr::arrange(C_location_name_canonical)
```

### Which schools exist in which years

Schools open and close. An in-scope school missing from a year is either a data
problem or a fact about the district, and nothing in the data distinguishes
them — so the exceptions are declared in config, with a reason that prints
here. An undeclared gap stops the run.

Because Hillwood merges into James Lawson, the merged location is expected in
every year and needs no exception. The override mechanism stays available for
districts where a school genuinely opens or closes without a successor.

```{r stu-school-years}
school_presence <- check_school_year_coverage(
  stu, id_col = "C_location_id", name_col = "C_location_name_canonical",
  year_col = "D_year", include_ids = included_location_ids,
  expectations = config$cohort$school_year_expectations)

school_presence
```

James Lawson should now show enrolment in all four years — 2022 and 2023 from
the Hillwood building, 2024 and 2025 from Bellevue — so no override is needed.

## 2.5 Demographics recode

Flags stay as 0/1 integers. Race goes through a map that stops on any code the
config does not cover.

```{r stu-recode}
stu <- stu |>
  dplyr::mutate(C_stu_race = apply_value_map(D_stu_race, config$value_maps$D_stu_race,
                                             "D_stu_race")) |>
  dplyr::rename(C_stu_ell_flag     = D_stu_ell,
                C_stu_swd_flag     = D_stu_swd,
                C_stu_poverty_flag = D_stu_frl)

for (f in c("C_stu_ell_flag", "C_stu_swd_flag", "C_stu_poverty_flag")) {
  chk_values_in(stu, f, c(0, 1), paste0(f, " is 0/1"))
}

stu |>
  dplyr::summarise(rows = dplyr::n(),
                   pct_ell     = round(100 * mean(C_stu_ell_flag), 1),
                   pct_swd     = round(100 * mean(C_stu_swd_flag), 1),
                   pct_poverty = round(100 * mean(C_stu_poverty_flag), 1))
```

## 2.6 Scope

The original config filtered on `D_stu_grade >= 9`, but `D_stu_grade` maps to
`percent` — the course grade. That kept anyone scoring above 9%.

```{r stu-scope}
stu <- apply_scope(stu, config$scope$demographics, id_col = "D_stu_id")
```

## 2.7 Grain

```{r stu-grain}
chk_unique(stu, c("D_stu_id", "D_year", "D_location_id", "D_stu_enter_date"),
           "One row per student / year / school / entry date")

overlaps <- stu |>
  dplyr::arrange(D_stu_id, D_year, D_location_id, H_enter_clamped) |>
  dplyr::group_by(D_stu_id, D_year, D_location_id) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::mutate(prev_exit = dplyr::lag(H_exit_clamped),
                overlaps  = !is.na(prev_exit) & H_enter_clamped < prev_exit) |>
  dplyr::ungroup()

chk_info("Repeat spells at the same school", dplyr::n_distinct(overlaps$D_stu_id))
chk_stop("No same-school spells overlap in time",
         sum(overlaps$overlaps, na.rm = TRUE) == 0,
         detail = "sum_of_spells would double-count if any overlapped",
         offenders = overlaps |> dplyr::filter(overlaps))

stu_cleaned <- stu
chk_info("Stage 2 output rows", nrow(stu_cleaned))
```

## 2.8 Trace students

```{r stu-trace}
set.seed(config$tracing$seed)
trace_ids <- unique(c(config$tracing$students,
                      sample(unique(stu_cleaned$D_stu_id), config$tracing$random_n)))

stu_cleaned |>
  dplyr::filter(D_stu_id %in% trace_ids) |>
  dplyr::select(D_stu_id, D_year, C_location_name_canonical,
                !!rlang::sym(config$grade_level_names$demographics),
                D_stu_enter_date, D_stu_exit_date, C_exit_date_imputed,
                M_days_enrolled, C_school_in_scope) |>
  dplyr::arrange(D_stu_id, D_year)
```

---

# Stage 3 — Transcript cleaning

Replaces `01_raw_data_transcript_validation.Rmd`. The dedup sections of that
script are gone: the file is already perfectly unique on student × year ×
course, with zero fully duplicated rows.

## 3.1 Rename and derive the year

The transcript has no year column. `D_year` comes from `end_year`, which
carries the same four values as the demographics `year` — that is what makes
the Stage 4 join possible. The old `02` script joined on a `D_year` that was
never created.

```{r tr-rename}
chk_stage("Stage 3 — Transcript cleaning")

tr <- transcript_raw |>
  dplyr::rename(dplyr::any_of(unlist(config$column_map$transcript))) |>
  dplyr::rename(!!config$grade_level_names$transcript := D_stu_grade_level) |>
  dplyr::mutate(D_year = .data[[config$derived$transcript_year_from]])

chk_stop("Rename produced all standard columns",
         all(setdiff(names(config$column_map$transcript), "D_stu_grade_level")
             %in% names(tr)),
         value = paste0(ncol(tr), " columns"))

# Only meaningful where the district supplies both a start and an end year.
if (all(c("D_start_year", "D_end_year") %in% names(tr))) {
  chk_stop("Every school year spans exactly one year",
           all(tr$D_end_year - tr$D_start_year == 1),
           offenders = tr |>
             dplyr::filter(D_end_year - D_start_year != 1) |>
             dplyr::count(D_start_year, D_end_year, name = "n_rows"))
} else {
  chk_info("Start/end year check skipped", "district supplies only one year column")
}

chk_stop("Transcript years match demographics years",
         setequal(unique(tr$D_year), unique(stu_cleaned$D_year)),
         value  = paste(sort(unique(tr$D_year)), collapse = ", "),
         detail = paste("demographics:", paste(sort(unique(stu_cleaned$D_year)), collapse = ", ")))

# Assert only what is required unconditionally. D_course_id nullity is handled
# by scope below, which is configured to drop those rows — asserting it here
# halted the run on a row the pipeline was about to discard anyway.
chk_no_na(tr, c("D_stu_id", "D_year"), "Student id and year always present")

chk_info("Rows with no course ID", sum(is.na(tr$D_course_id)),
         detail = "dropped by scope in 3.5",
         data = tr |> dplyr::filter(is.na(D_course_id)) |>
           dplyr::select(D_stu_id, D_year, D_course_name,
                         D_creds_earn, D_creds_possible))
```

## 3.2 Term

Derived, since there is no term column. The S1/S2 marker sits on the course
number for most rows and on the course name as a fallback.

```{r tr-term}
tr <- tr |>
  dplyr::mutate(C_term = match_first_pattern(
    sources = list(D_course_id   = D_course_id,
                   D_course_name = D_course_name),
    rules   = config$derived$term_detection$rules,
    label   = "C_term"))

chk_info("Rows with no term detected", sum(is.na(tr$C_term)),
         data = tr |> dplyr::filter(is.na(C_term)) |>
           dplyr::count(D_course_id, D_course_name,
                        D_creds_possible, D_creds_earn, name = "n_rows") |>
           dplyr::arrange(dplyr::desc(n_rows)))
```

```{r tr-noterm-credits}
tr |>
  dplyr::filter(is.na(C_term)) |>
  dplyr::count(D_creds_possible, D_creds_earn, name = "n_rows") |>
  dplyr::mutate(pct = round(100 * n_rows / sum(n_rows), 1)) |>
  dplyr::arrange(dplyr::desc(n_rows))
```

### 3.2a Are DA/DB terms, or sections?

Every no-term row is a dual-enrolment course, and their IDs come in pairs —
`SST5601DA` and `SST5601DB`, `HLT3801DA` and `HLT3801DB`. That could be a
semester marker in a different dialect, or it could be section A and section B.

The two readings make opposite predictions. If DA/DB are **terms**, a large
share of students take both halves of the same base course in one year. If they
are **sections**, essentially nobody does, because you cannot enrol in two
sections of the same course at once.

```{r tr-de-suffix-test}
de <- tr |>
  dplyr::filter(grepl("D[AB]\\*?\\s*$", D_course_id)) |>
  dplyr::mutate(de_suffix = stringr::str_extract(D_course_id, "D[AB]"),
                de_base   = sub("D[AB]\\*?\\s*$", "", D_course_id))

chk_info("Rows with a DA/DB course ID", nrow(de),
         detail = paste0(dplyr::n_distinct(de$de_base), " distinct base course(s)"))

de_both <- de |>
  dplyr::group_by(D_stu_id, D_year, de_base) |>
  dplyr::summarise(suffixes = paste(sort(unique(de_suffix)), collapse = "+"),
                   .groups = "drop") |>
  dplyr::count(suffixes, name = "n_student_course_years") |>
  dplyr::mutate(pct = round(100 * n_student_course_years /
                              sum(n_student_course_years), 1))

chk_info("Students taking both DA and DB of the same base course in one year",
         "see table",
         detail = "a large DA+DB share means term; near zero means section",
         data = de_both)

de_both
```

```{r tr-de-credits}
de |>
  dplyr::count(de_suffix, D_creds_possible, D_creds_earn, name = "n_rows") |>
  dplyr::arrange(de_suffix, dplyr::desc(n_rows))
```

Credits by term. A semester course should be worth half a credit; anything else
is worth understanding before credits get summed.

```{r tr-term-credits}
tr |>
  dplyr::count(C_term, D_creds_possible, name = "n_rows") |>
  dplyr::arrange(C_term, dplyr::desc(n_rows))
```

## 3.3 Course identifiers

The base course ID, with the semester suffix stripped, is kept alongside the
original. The Stage 7 join keys on the original ID **plus** the name, so the
base is for reporting and for a fallback if the primary join underperforms.

```{r tr-course-ids}
tr <- tr |>
  dplyr::mutate(
    C_course_id_base = stringr::str_replace(D_course_id,
                                            config$derived$course_id_strip_suffix, ""),
    C_course_name_trimmed = trimws(D_course_name),
    C_course_name_starred = grepl("\\*", D_course_name)
  )

chk_info("Distinct course IDs (as given)",  dplyr::n_distinct(tr$D_course_id))
chk_info("Distinct course IDs (base)",      dplyr::n_distinct(tr$C_course_id_base))
chk_info("Distinct course names",           dplyr::n_distinct(tr$D_course_name))
chk_info("Course names containing an asterisk", sum(tr$C_course_name_starred))

n_pairs <- tr |> dplyr::distinct(D_course_id, D_course_name) |> nrow()
chk_info("Distinct course ID + name pairs", n_pairs,
         detail = "this is the Stage 7 join key")
```

How much name variation sits behind a single course ID — the reason the join
keys on both.

```{r tr-id-name-variation}
tr |>
  dplyr::distinct(D_course_id, D_course_name) |>
  dplyr::count(D_course_id, name = "n_names") |>
  dplyr::count(n_names, name = "n_course_ids") |>
  dplyr::arrange(n_names)
```

## 3.4 Grades and credits

Two tiers of bounds, because "outside what I expected" and "cannot be a real
measurement" are different questions.

`percent` is a computed weighted average, not a rounded mark — 9,994 distinct
values, and a near-continuous run of 100.01, 100.02, 100.03 at roughly eighty
rows apiece. Grades just above 100 are **extra credit, genuinely earned**. A
single bound at 100 would either halt on those or let 981 through; separating
the tiers lets the plausible range describe the data while the hard range
guards against values that are not grades at all.

The plausible tier only reports. The hard tier gates only when
`hard_tolerance_pct` is set in config, which is deliberately left unset for
grades on this pass — read the distribution below, then choose.

```{r tr-ranges}
g_grade <- range_two_tier(tr$D_stu_grade,      config$ranges$D_stu_grade,      "D_stu_grade")
g_earn  <- range_two_tier(tr$D_creds_earn,     config$ranges$D_creds_earn,     "D_creds_earn")
g_poss  <- range_two_tier(tr$D_creds_possible, config$ranges$D_creds_possible, "D_creds_possible")

tr <- tr |>
  dplyr::mutate(
    D_stu_grade            = g_grade$values,
    D_creds_earn           = g_earn$values,
    D_creds_possible       = g_poss$values,
    C_grade_above_100      = g_grade$flag_plausible,
    C_grade_not_a_grade    = g_grade$flag_hard,
    C_creds_out_of_range   = g_earn$flag_plausible | g_poss$flag_plausible
  )
```

### Shape of the grade tail

How far above 100 does it actually go, and how thin does it get. This is the
table to read before setting a bound.

```{r tr-grade-tail}
grade_tail <- describe_tail(tr$D_stu_grade, config$grade_tail_buckets, "D_stu_grade")
grade_tail
```

```{r tr-grade-summary}
tr |>
  dplyr::filter(!is.na(D_stu_grade)) |>
  dplyr::summarise(
    rows          = dplyr::n(),
    max           = max(D_stu_grade),
    p99           = round(stats::quantile(D_stu_grade, .99), 2),
    p999          = round(stats::quantile(D_stu_grade, .999), 2),
    n_above_100   = sum(D_stu_grade > 100),
    pct_above_100 = round(100 * mean(D_stu_grade > 100), 3),
    n_above_110   = sum(D_stu_grade > 110),
    n_above_150   = sum(D_stu_grade > 150))
```

If grades above 100 cluster in particular courses, the cause is weighted or
extra-credit scoring. If they are spread evenly, it is more likely a gradebook
setting applied district-wide.

```{r tr-grade-by-course}
tr |>
  dplyr::filter(C_grade_above_100) |>
  dplyr::count(D_course_id, D_course_name, name = "n_rows") |>
  dplyr::mutate(pct_of_flagged = round(100 * n_rows / sum(n_rows), 1)) |>
  dplyr::arrange(dplyr::desc(n_rows)) |>
  utils::head(20)
```

```{r tr-grade-by-year}
tr |>
  dplyr::group_by(D_year) |>
  dplyr::summarise(rows = dplyr::n(),
                   pct_above_100 = round(100 * mean(D_stu_grade > 100, na.rm = TRUE), 2),
                   max_grade = max(D_stu_grade, na.rm = TRUE),
                   .groups = "drop")
```

The values that cannot be grades at all — and which courses they sit in. A
value like 981 is most likely a misplaced decimal on 98.1, or a points total
recorded in a percent field.

```{r tr-grade-impossible}
tr |>
  dplyr::filter(C_grade_not_a_grade) |>
  dplyr::count(D_stu_grade, D_course_id, D_course_name,
               D_creds_earn, D_creds_possible, name = "n_rows") |>
  dplyr::arrange(dplyr::desc(D_stu_grade)) |>
  utils::head(30)
```

### Credits

```{r tr-credit-outliers}
tr |>
  dplyr::filter(C_creds_out_of_range) |>
  dplyr::count(D_creds_possible, D_creds_earn, D_course_name, name = "n_rows") |>
  dplyr::arrange(dplyr::desc(D_creds_earn)) |>
  utils::head(20)
```

Credits earned can never exceed credits attempted.

```{r tr-credit-logic}
tr <- tr |>
  dplyr::mutate(C_creds_earn_exceeds_attempted =
                  !is.na(D_creds_earn) & !is.na(D_creds_possible) &
                  D_creds_earn > D_creds_possible)

chk_info("Rows where credits earned exceed credits attempted",
         sum(tr$C_creds_earn_exceeds_attempted),
         data = tr |>
           dplyr::filter(C_creds_earn_exceeds_attempted) |>
           dplyr::count(D_creds_possible, D_creds_earn, name = "n_rows") |>
           dplyr::arrange(dplyr::desc(n_rows)))

chk_info("Rows with zero credits attempted", sum(tr$D_creds_possible == 0, na.rm = TRUE))
chk_info("Rows with zero credits earned",    sum(tr$D_creds_earn == 0, na.rm = TRUE))
```

Tettra has no rule for this, so it is a decision to log rather than a rule to
follow. Before deciding, it is worth knowing whether the dominant 0.5 to 1.0
pattern is a **course-level convention** — some courses always award a full
credit regardless of what the attempted field says — or something scattered.

If a course ID shows 1.0 earned on every row, the attempted field is simply
wrong for that course and the earned figure can be trusted. If the same course
sometimes shows 0.5, the pattern is student-level and needs a different answer.

```{r tr-credit-excess-courses}
excess_courses <- tr |>
  dplyr::filter(D_creds_possible == 0.5) |>
  dplyr::group_by(D_course_id, D_course_name) |>
  dplyr::summarise(rows = dplyr::n(),
                   n_earn_1 = sum(D_creds_earn == 1.0, na.rm = TRUE),
                   .groups = "drop") |>
  dplyr::filter(n_earn_1 > 0) |>
  dplyr::mutate(pct_of_course_at_1 = round(100 * n_earn_1 / rows, 1)) |>
  dplyr::arrange(dplyr::desc(n_earn_1))

chk_info("Courses showing the 0.5-attempted / 1.0-earned pattern",
         nrow(excess_courses),
         detail = "pct_of_course_at_1 near 100 means a course-level convention")

utils::head(excess_courses, 25)
```

```{r tr-credit-excess-concentration}
excess_courses |>
  dplyr::mutate(bucket = cut(pct_of_course_at_1,
                             breaks = c(0, 5, 25, 50, 75, 95, 100),
                             include.lowest = TRUE)) |>
  dplyr::group_by(bucket) |>
  dplyr::summarise(n_courses = dplyr::n(),
                   n_rows_at_1 = sum(n_earn_1), .groups = "drop")
```

```{r tr-credit-treatment}
n_impossible <- sum(tr$D_creds_earn > config$credits$impossible_above, na.rm = TRUE)

if (config$credits$impossible_treatment == "na") {
  tr <- tr |>
    dplyr::mutate(C_creds_impossible = !is.na(D_creds_earn) &
                    D_creds_earn > config$credits$impossible_above,
                  D_creds_earn = dplyr::if_else(C_creds_impossible,
                                                NA_real_, D_creds_earn))
  chk_info("Impossible credit values set to NA", n_impossible,
           detail = paste0("earned above ", config$credits$impossible_above,
                           "; the enrolment row is kept, only the number is removed"))
} else if (config$credits$impossible_treatment == "drop") {
  before <- nrow(tr)
  tr <- tr |> dplyr::filter(is.na(D_creds_earn) |
                              D_creds_earn <= config$credits$impossible_above)
  chk_info("Rows dropped for impossible credit values", before - nrow(tr))
} else {
  chk_info("Impossible credit values left in place", n_impossible)
}

if (config$credits$earned_exceeds_attempted$treatment == "na_earned") {
  tr <- tr |> dplyr::mutate(D_creds_earn =
    dplyr::if_else(C_creds_earn_exceeds_attempted, NA_real_, D_creds_earn))
  chk_info("Earned credits blanked where they exceeded attempted",
           sum(tr$C_creds_earn_exceeds_attempted))
}
```

## 3.5 Scope and grain

Dropping a row removes whatever it was carrying, so scope reports what each
rule costs in credits as well as in rows. At one row it is immaterial; the
report exists so it cannot become material without being seen.

```{r tr-scope}
tr <- apply_scope(tr, config$scope$transcript, id_col = "D_stu_id",
                  sum_cols = c("D_creds_earn", "D_creds_possible"))

chk_no_na(tr, c("D_stu_id", "D_year", "D_course_id"),
          "Core transcript keys present after scope")

chk_unique(tr, c("D_stu_id", "D_year", "D_course_id"),
           "One row per student / year / course")

chk_stop("No fully duplicated rows", sum(duplicated(tr)) == 0)
```

### Course load and credit accumulation

A student taking 16 semester records should attempt 8.0 credits. Whether they
*earn* 8.0 is a different question, and the gap between the two is the thing
worth watching: 2024 lands on 8.0 while 2022 comes in a full credit lower.

Two explanations compete. Either students failed more in the earlier years, or
those years carry more zero-credit courses that pad the record count without
contributing credit. Putting attempted next to earned separates them — if
attempted holds at 8.0 across all years and only earned dips, it is failure.

```{r tr-load-credits}
stu_year <- tr |>
  dplyr::group_by(D_stu_id, D_year) |>
  dplyr::summarise(
    courses        = dplyr::n(),
    courses_credit_bearing = sum(D_creds_possible > 0, na.rm = TRUE),
    creds_attempted = sum(D_creds_possible, na.rm = TRUE),
    creds_earned    = sum(D_creds_earn, na.rm = TRUE),
    n_failed        = sum(D_creds_possible > 0 & D_creds_earn == 0, na.rm = TRUE),
    .groups = "drop")

stu_year |>
  dplyr::group_by(D_year) |>
  dplyr::summarise(
    students             = dplyr::n(),
    median_courses       = stats::median(courses),
    median_credit_courses = stats::median(courses_credit_bearing),
    median_attempted     = stats::median(creds_attempted),
    median_earned        = stats::median(creds_earned),
    median_failed        = stats::median(n_failed),
    mean_failed          = round(mean(n_failed), 2),
    pct_earning_all      = round(100 * mean(creds_earned >= creds_attempted), 1),
    .groups = "drop")
```

Failure rates by year, at the row level rather than the student level. If the
earlier years fail more, the credit gap explains itself.

```{r tr-failure-by-year}
tr |>
  dplyr::filter(D_creds_possible > 0) |>
  dplyr::group_by(D_year) |>
  dplyr::summarise(
    credit_bearing_rows = dplyr::n(),
    n_zero_earned       = sum(D_creds_earn == 0, na.rm = TRUE),
    pct_zero_earned     = round(100 * mean(D_creds_earn == 0, na.rm = TRUE), 1),
    .groups = "drop")
```

And the share of records carrying no credit at all, which is the competing
explanation.

```{r tr-zero-credit-by-year}
tr |>
  dplyr::group_by(D_year) |>
  dplyr::summarise(
    rows                 = dplyr::n(),
    n_zero_attempted     = sum(D_creds_possible == 0, na.rm = TRUE),
    pct_zero_attempted   = round(100 * mean(D_creds_possible == 0, na.rm = TRUE), 2),
    .groups = "drop")
```

```{r tr-credit-distribution}
stu_year |>
  dplyr::group_by(D_year) |>
  dplyr::summarise(
    p10 = stats::quantile(creds_earned, .10),
    p25 = stats::quantile(creds_earned, .25),
    p50 = stats::quantile(creds_earned, .50),
    p75 = stats::quantile(creds_earned, .75),
    p90 = stats::quantile(creds_earned, .90),
    .groups = "drop")
```

```{r tr-output}
transcript_cleaned <- tr
chk_info("Stage 3 output rows", nrow(transcript_cleaned))
chk_info("Distinct students in transcript", dplyr::n_distinct(transcript_cleaned$D_stu_id))
```

## 3.6 Coding file reconnaissance

Not the join — that is Stage 7. This just reports what the coded file contains
so the join can be written against its real column names.

```{r coding-recon}
if (!is.null(course_coding_raw)) {
  chk_info("Coding file shape",
           paste0(format(nrow(course_coding_raw), big.mark = ","), " rows x ",
                  ncol(course_coding_raw), " cols"))
  print(tibble::tibble(column = names(course_coding_raw),
                       type   = purrr::map_chr(course_coding_raw,
                                               ~ paste(class(.x), collapse = "/")),
                       n_distinct = purrr::map_int(course_coding_raw,
                                                   ~ dplyr::n_distinct(.x)),
                       pct_missing = purrr::map_dbl(course_coding_raw,
                                                    ~ round(100 * mean(is.na(.x)), 1))) |>
          as.data.frame())
} else {
  chk_info("Coding file not available", "Stage 7 will stop")
}
```

```{r coding-head}
if (!is.null(course_coding_raw)) utils::head(course_coding_raw, 10)
```

### Stage 7 join preview

Cheap to run now, and it de-risks the biggest remaining unknown. The join keys
on course ID plus name. 46,780 transcript course names carry an asterisk, so if
the coding file stripped those during cleaning, the exact match will miss on
every one of them. Better to find that out here than at Stage 7.

```{r coding-join-preview}
if (!is.null(course_coding_raw)) {
  tr_pairs <- tr |>
    dplyr::distinct(D_course_id, D_course_name)

  cd_pairs <- course_coding_raw |>
    dplyr::distinct(d_course_id, d_course_name) |>
    dplyr::rename(D_course_id = d_course_id, D_course_name = d_course_name)

  exact <- dplyr::inner_join(tr_pairs, cd_pairs, by = c("D_course_id", "D_course_name"))

  chk_info("Coding join preview — pairs in transcript", nrow(tr_pairs))
  chk_info("Coding join preview — pairs in coding file", nrow(cd_pairs))
  chk_info("Coding join preview — exact matches on ID + name",
           paste0(format(nrow(exact), big.mark = ","), " (",
                  round(100 * nrow(exact) / nrow(tr_pairs), 1), "% of transcript pairs)"))

  # Weighted by rows, not pairs: an unmatched pair on one row matters less than
  # an unmatched pair on ten thousand.
  matched_rows <- tr |>
    dplyr::semi_join(cd_pairs, by = c("D_course_id", "D_course_name")) |>
    nrow()
  chk_info("Coding join preview — transcript ROWS matched",
           paste0(format(matched_rows, big.mark = ","), " (",
                  round(100 * matched_rows / nrow(tr), 2), "%)"))

  unmatched <- tr |>
    dplyr::anti_join(cd_pairs, by = c("D_course_id", "D_course_name")) |>
    dplyr::count(D_course_id, D_course_name, name = "n_rows") |>
    dplyr::arrange(dplyr::desc(n_rows))

  chk_info("Coding join preview — unmatched pairs", nrow(unmatched),
           data = utils::head(unmatched, 25))
}
```

If the exact match falls short, the next question is whether normalising the
name closes the gap — asterisks stripped, whitespace collapsed, case folded.
That tells us whether the fix belongs in the pipeline or in the coding file.

```{r coding-join-fallback}
if (!is.null(course_coding_raw)) {
  norm <- function(x) tolower(trimws(gsub("\\s+", " ", gsub("\\*", "", x))))

  tr_norm <- tr |>
    dplyr::mutate(k_id = trimws(D_course_id), k_nm = norm(D_course_name))

  cd_norm <- course_coding_raw |>
    dplyr::distinct(d_course_id, d_course_name) |>
    dplyr::mutate(k_id = trimws(d_course_id), k_nm = norm(d_course_name)) |>
    dplyr::distinct(k_id, k_nm)

  matched_norm <- tr_norm |> dplyr::semi_join(cd_norm, by = c("k_id", "k_nm")) |> nrow()

  chk_info("Coding join preview — rows matched after name normalisation",
           paste0(format(matched_norm, big.mark = ","), " (",
                  round(100 * matched_norm / nrow(tr), 2), "%)"),
           detail = "asterisks removed, whitespace collapsed, case folded")

  # ID alone, as an upper bound on what any name-level fix could achieve.
  matched_id <- tr |>
    dplyr::semi_join(course_coding_raw |>
                       dplyr::distinct(D_course_id = d_course_id),
                     by = "D_course_id") |> nrow()
  chk_info("Coding join preview — rows matched on course ID alone",
           paste0(format(matched_id, big.mark = ","), " (",
                  round(100 * matched_id / nrow(tr), 2), "%)"),
           detail = "upper bound for any name-matching fix")
}
```

## 3.7 Trace students

```{r tr-trace}
transcript_cleaned |>
  dplyr::filter(D_stu_id %in% trace_ids) |>
  dplyr::select(D_stu_id, D_year, D_course_id, C_course_id_base, D_course_name,
                C_term, D_stu_grade, D_creds_earn, D_creds_possible) |>
  dplyr::arrange(D_stu_id, D_year, D_course_id) |>
  utils::head(40)
```

---

# Stage 4 — Cross-file reconciliation

> Next step. All 47,058 transcript students appear in demographics, so the
> exclude-flag routine in the old `02` collapses to a single check.

---

# Stage 5 — Home school assignment

---

# Stage 6 — Cohort selection

---

# Stage 7 — Course coding join

---

# Stage 8 — Output and check report

```{r final-report}
check_log <- chk_report(print_table = FALSE)
```

```{r final-table}
check_log |>
  dplyr::select(seq, stage, check, severity, result, value)
```

```{r write-log}
#| eval: false
write_check_log(config)
```
