---
title: "Transcript Pipeline"
subtitle: "Metro Nashville Public Schools"
date: today
params:
  district: "nashville"
  refresh_raw: false      # TRUE re-reads every source from SharePoint
  smoke_test: true        # TRUE runs the check-framework self-test in Stage 0
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE)
source(here::here("R", "setup.R"))
```

# What this file is

One document, run top to bottom, replacing the eight scripts that used to run
in sequence. Everything district-specific lives either here or in
`config/districts/tr_nashville.yaml`. Everything reusable lives in `R/`.

**Reading the output.** Every check prints as it runs:

- `[PASS]` — a structural check held.
- `[FAIL]` — a structural check broke. The render stops here. Offending rows
  are printed, and available afterwards via `chk_offenders("<check name>")`.
- `[info]` — a count or table to eyeball. Never halts anything.

The last section prints a summary table of every check in the run.

**Stages.**

| # | Stage | Replaces |
|---|-------|----------|
| 0 | Setup and config | Part 0 of all eight scripts |
| 1 | Load sources | Part 1 of all eight scripts |
| 2 | Student cleaning | `01_raw_data_student_validation` |
| 3 | Transcript cleaning | `01_raw_data_transcript_validation` |
| 4 | Cross-file reconciliation | `02_cross_file_validation` §2.1 |
| 5 | Home school assignment | `02_cross_file_validation` §2.2 |
| 6 | Cohort selection | `03_student_cohort_selection` |
| 7 | Course coding join | new — the rollup that never existed |
| 8 | Output and check report | Part 3/4 of all eight scripts |

---

# Stage 0 — Setup and configuration

## 0.1 Load and validate the config

`load_config()` fails on the spot, listing every problem at once, if the YAML
is missing anything the pipeline depends on. Better than discovering a missing
column map at Stage 4.

```{r config}
chk_reset()
chk_stage("Stage 0 — Setup")

config <- load_config(district = params$district)
config_summary(config)
```

## 0.2 Record the run

Stamped so any number produced here can be traced back to the exact config and
code that made it.

```{r run-manifest}
run_manifest <- tibble::tibble(
  run_time    = Sys.time(),
  district    = config$district,
  config_file = basename(config$.meta$path),
  config_md5  = substr(config$.meta$md5, 1, 8),
  refresh_raw = params$refresh_raw,
  r_version   = paste0(R.version$major, ".", R.version$minor),
  user        = Sys.info()[["user"]]
)

chk_info("Run stamped", run_manifest$config_md5,
         detail = paste("config", run_manifest$config_file))

run_manifest
```

## 0.3 Config decisions worth seeing in the report

These were hardcoded in the old scripts, in more than one place, and in one
case inconsistently. Printing them here means a reviewer can see the cohort
definition without reading any code.

```{r config-decisions}
chk_info("Cohort entry", paste0("grade ", config$cohort$entry_grade,
                                " in ", config$cohort$entry_year))
chk_info("Years required", config$cohort$years_required)
chk_info("Duration rule", paste0(config$cohort$duration_rule, ", >= ",
                                 config$cohort$duration_min_days, " days/yr"))
chk_info("Progression rule", config$cohort$progression_rule)
chk_info("Days enrolled rule", config$cohort$days_enrolled_rule)
chk_info("Schools in scope", length(config$cohort$schools_to_include),
         data = tibble::tibble(school = config$cohort$schools_to_include))

tibble::tibble(
  alias     = names(config$cohort$school_name_aliases),
  resolves_to = unlist(config$cohort$school_name_aliases)
)
```

## 0.4 Check-framework self-test

Runs against fabricated data, touches nothing real. Confirms `[PASS]`,
`[info]`, and `[FAIL]` all behave before any district data is involved. Set
`smoke_test: false` in the header once you trust it.

```{r smoke-test}
#| eval: !expr params$smoke_test

toy <- tibble::tibble(
  D_stu_id = c(1, 1, 2, 3),
  D_year   = c(2022, 2023, 2022, 2022),
  score    = c(10, 20, 30, NA)
)

# Should PASS
chk_unique(toy, c("D_stu_id", "D_year"), "Self-test: toy is unique on id + year")
chk_rows_stable(toy, toy, "Self-test: row count stable")
chk_n_distinct(toy, "D_stu_id", "Self-test: distinct students")

# Should print an info table
chk_info("Self-test: rows by year",
         dplyr::n_distinct(toy$D_year),
         data = dplyr::count(toy, D_year))

# Should FAIL — wrapped so the render survives the demonstration
demo <- try(chk_no_na(toy, "score", "Self-test: deliberate failure"), silent = TRUE)
cat("\nFailure was caught. Offending rows retained:\n")
print(chk_offenders("Self-test: deliberate failure"))
```

## 0.5 Reset after the self-test

Clears the toy checks so the report only contains real ones.

```{r post-smoke-reset}
#| eval: !expr params$smoke_test
chk_reset()
chk_stage("Stage 0 — Setup")
chk_info("Self-test complete", "framework verified, log reset")
```

---

# Stage 1 — Load sources

> Not built yet. Next step.

Reads the transcript, demographics, and coded course file once each, caching
them locally as `.rds`. Re-runs read from cache in under a second. Set
`refresh_raw: true` in the header when the district sends new data.

```{r load-sources}
#| eval: false
chk_stage("Stage 1 — Load sources")

transcript_raw <- read_source("transcript_raw",   config, refresh = params$refresh_raw)
demographics_raw <- read_source("stu_demographics", config, refresh = params$refresh_raw)
course_coding_raw <- read_source("course_coding",  config, refresh = params$refresh_raw,
                                 sheet = config$sheets$course_coding)
```

---

# Stage 2 — Student cleaning

> Not built yet.

---

# Stage 3 — Transcript cleaning

> Not built yet.

---

# Stage 4 — Cross-file reconciliation

> Not built yet.

---

# Stage 5 — Home school assignment

> Not built yet.

---

# Stage 6 — Cohort selection

> Not built yet.

---

# Stage 7 — Course coding join

> Not built yet.

---

# Stage 8 — Output and check report

```{r final-report}
chk_report()
```

```{r write-log}
#| eval: false
write_check_log(config)
```
