---
title: "Transcript Pipeline"
subtitle: "Metro Nashville Public Schools"
date: today
params:
  district: "nashville"
  refresh_raw: false # Change this to "true" if the district has sent new raw files and you want to re-read them
  smoke_test: false
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE)
# Bootstrap: read helpers_path directly from YAML, source setup, then load full config
.cfg_raw     <- yaml::read_yaml(here::here("config", "districts", paste0("tr_", params$district, ".yaml")))
.helpers_dir <- here::here(.cfg_raw$paths$helpers_path)
source(file.path(.helpers_dir, "00_tr_setup.R"))
config <- load_config(params$district)
```

# What this file is

One document, run top to bottom, replacing the eight scripts that used to run
in sequence. District-specific cleaning decisions live here or in
`config/districts/tr_nashville.yaml`. Reusable machinery lives in
`templates/transcript/00_tr_*.R`.

**Reading the output.** Every check prints as it runs:

- `[PASS]` — a structural check held.
- `[FAIL]` — a structural check broke. The render stops. Offending rows print,
  and stay available via `chk_offenders("<check name>")`.
- `[info]` — a count or table to eyeball. Never halts.

**Diagnostics run before gates.** When a structural check fails mid-render,
knitr discards the whole chunk's output — so a gate placed above the table that
explains it destroys exactly the evidence you need. Every stage therefore
measures and describes first, and gates last. Failed checks also write their
offending rows to `output/<district>/failed_checks/`.

| # | Stage | Replaces | Status |
|---|-------|----------|--------|
| 0 | Setup and config | Part 0 of all eight | built |
| 1 | Load sources | Part 1 of all eight | built |
| 2 | Student cleaning | `01_raw_data_student_validation` | built |
| 3 | Transcript cleaning | `01_raw_data_transcript_validation` | built |
| 4 | Cross-file reconciliation | `02_cross_file_validation` §2.1 | built |
| 5 | Home school assignment | `02_cross_file_validation` §2.2 | built |
| 6 | Cohort selection | `03_student_cohort_selection` | built |
| 7 | Course coding join | new — the rollup that never existed | built |
| 8 | Output and check report | `03_transcript_rollup` + Part 4 of all eight | built |

---

# Stage 0 — Setup and configuration

```{r config}
chk_reset()
chk_stage("Stage 0 — Setup")

config <- load_config(district = params$district)

# Failed checks write their offending rows here before the run halts, because
# knitr discards a chunk's output when the chunk errors.
chk_set_output_dir(file.path(cache_dir(config), "failed_checks"))

config_summary(config)
```

## 0.1 Run manifest

```{r run-manifest}
run_manifest <- tibble(
  run_time    = Sys.time(),
  district    = config$district,
  config_file = basename(config$.meta$path),
  config_md5  = substr(config$.meta$md5, 1, 8),
  refresh_raw = params$refresh_raw,
  r_version   = paste0(R.version$major, ".", R.version$minor),
  user        = Sys.info()[["user"]]
)
chk_info("Run stamped", run_manifest$config_md5)
run_manifest
```

## 0.2 The decisions this run is making

```{r config-decisions}
chk_info("Cohort entry", paste0("grade ", config$cohort$entry_grade,
                                " in ", config$cohort$entry_year))
chk_info("Years required", config$cohort$years_required)
chk_info("Duration rule", paste0(config$cohort$duration_rule, ", >= ",
                                 config$cohort$duration_min_days, " days/yr"))
chk_info("Progression rule", config$cohort$progression_rule)
chk_info("Days enrolled rule", config$cohort$days_enrolled_rule)
chk_info("Negative days treatment", config$derived$negative_days_treatment)
chk_info("Schools in scope", length(config$cohort$schools_to_include))
```

## 0.3 Check-framework self-test

```{r smoke-test}
#| eval: !expr params$smoke_test
toy <- tibble(D_stu_id = c(1, 1, 2, 3),
                      D_year   = c(2022, 2023, 2022, 2022),
                      score    = c(10, 20, 30, NA))
chk_unique(toy, c("D_stu_id", "D_year"), "Self-test: unique on id + year")
demo <- try(chk_no_na(toy, "score", "Self-test: deliberate failure"), silent = TRUE)
print(chk_offenders("Self-test: deliberate failure"))
chk_reset(); chk_stage("Stage 0 — Setup")
chk_info("Self-test complete", "framework verified, log reset")
```

---

# Stage 1 — Load sources

Each file is read from SharePoint once and cached as `.rds`. **The cache does
not know when the district re-sends a file.** Set `refresh_raw: true` when new
data lands.

```{r load-sources}
chk_stage("Stage 1 — Load sources")

transcript_raw   <- read_source("transcript_raw",   config, refresh = params$refresh_raw)
demographics_raw <- read_source("stu_demographics", config, refresh = params$refresh_raw)

course_coding_raw <- tryCatch(
  read_source("course_coding", config, refresh = params$refresh_raw,
              sheet = config$sheets$course_coding),
  error = function(e) {
    chk_info("Course coding file NOT loaded", "see detail", detail = conditionMessage(e))
    NULL
  }
)
```

## 1.1 Expected shape

```{r source-shape}
chk_stop("Transcript has expected columns",
         all(unlist(config$column_map$transcript) %in% names(transcript_raw)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$column_map$transcript), names(transcript_raw)),
           collapse = ", ")))

chk_stop("Demographics has expected columns",
         all(unlist(config$column_map$demographics) %in% names(demographics_raw)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$column_map$demographics), names(demographics_raw)),
           collapse = ", ")))

chk_info("Transcript rows",   nrow(transcript_raw))
chk_info("Demographics rows", nrow(demographics_raw))
```

---

# Stage 2 — Student cleaning

Ends at one row per student per enrolment spell. Collapsing to one row per
student-year happens at Stage 5, after home school is assigned.

## 2.1 Rename to ERS standard names

```{r stu-rename}
chk_stage("Stage 2 — Student cleaning")

stu <- demographics_raw |>
  rename(any_of(unlist(config$column_map$demographics))) |>
  rename(!!config$grade_level_names$demographics := D_stu_grade_level)

chk_stop("Rename produced all standard columns",
         all(setdiff(names(config$column_map$demographics), "D_stu_grade_level")
             %in% names(stu)),
         value = paste0(ncol(stu), " columns"))

chk_no_na(stu, c("D_stu_id", "D_year"), "Student id and year always present")
chk_info("Distinct students in raw demographics", n_distinct(stu$D_stu_id))
```

## 2.2 Dates

Where the old pipeline broke. `wdt` arrives as `m/d/yyyy`; the original asked
`as.Date()` for `%Y-%m-%d`, which returns `NA` silently, so every exit date and
therefore every duration was `NA`. Formats now come from config, unparseable
values are a structural failure, and the fill uses `coalesce()` rather than
`ifelse()` so the Date class survives.

```{r stu-dates}
stu <- stu |>
  mutate(
    D_stu_enter_date = parse_date_strict(D_stu_enter_date,
                                         config$derived$date_formats$D_stu_enter_date,
                                         "enter date"),
    D_stu_exit_date  = parse_date_strict(D_stu_exit_date,
                                         config$derived$date_formats$D_stu_exit_date,
                                         "exit date"),
    C_exit_date_imputed = is.na(D_stu_exit_date)
  )

if (isTRUE(config$derived$force_missing_exit_date)) {
  stu <- stu |>
    mutate(D_stu_exit_date = fill_missing_date(
      D_stu_exit_date,
      as.Date(paste0(D_year, "-", config$derived$school_year_end_mmdd)),
      label = "exit date -> year end"))
}

chk_no_na(stu, c("D_stu_enter_date", "D_stu_exit_date"), "Both dates present after fill")

chk_stop("Exit date is not before enter date",
         all(stu$D_stu_exit_date >= stu$D_stu_enter_date),
         offenders = stu |>
           filter(D_stu_exit_date < D_stu_enter_date) |>
           select(D_year, D_location_id, D_stu_enter_date, D_stu_exit_date))
```

`wdt` behaves like a withdrawal date, not a year-end date: spells that carry a
real one have a median of about five weeks. That means the duration threshold
is doing less work than it looks — it is close to separating "withdrew" from
"did not."

## 2.3 Days enrolled

Spells are clamped to the school year so an entry date carried over from a
prior year cannot produce a 1,200-day "year".

```{r stu-days}
stu <- stu |>
  mutate(
    H_year_start = as.Date(paste0(D_year - 1, "-", config$derived$school_year_start_mmdd)),
    H_year_end   = as.Date(paste0(D_year,     "-", config$derived$school_year_end_mmdd)),
    C_spell_before_window = D_stu_exit_date  < H_year_start,
    C_spell_after_window  = D_stu_enter_date > H_year_end,
    H_enter_clamped = if (isTRUE(config$derived$clamp_days_to_school_year))
      pmax(D_stu_enter_date, H_year_start) else D_stu_enter_date,
    H_exit_clamped  = if (isTRUE(config$derived$clamp_days_to_school_year))
      pmin(D_stu_exit_date,  H_year_end)   else D_stu_exit_date,
    M_days_enrolled = as.numeric(H_exit_clamped - H_enter_clamped)
  )

chk_info("Spells clamped at the start of the window",
         sum(stu$D_stu_enter_date < stu$H_year_start))
chk_info("Spells clamped at the end of the window",
         sum(stu$D_stu_exit_date > stu$H_year_end))
```

### Spells that fall outside the school year

These are what produced negative durations on the first run. Entry dates in
early July, recorded against the year that had just ended.

```{r stu-outside-window}
outside <- stu |>
  filter(C_spell_before_window | C_spell_after_window) |>
  select(D_year, D_location_id, D_stu_enter_date, D_stu_exit_date,
                C_exit_date_imputed, C_spell_before_window, C_spell_after_window,
                M_days_enrolled)

chk_info("Spells entirely outside the school year", nrow(outside),
         data = outside |>
           count(D_year, C_spell_before_window, C_spell_after_window,
                        name = "n_rows"))

outside |> arrange(M_days_enrolled) |> head(20) |> print(width = Inf)
```

```{r stu-negative-treatment}
treatment <- config$derived$negative_days_treatment

if (treatment == "drop") {
  before <- nrow(stu)
  stu <- stu |> filter(M_days_enrolled >= 0)
  chk_info("Negative-duration spells dropped", before - nrow(stu))
} else if (treatment == "floor_zero") {
  n_floored <- sum(stu$M_days_enrolled < 0)
  stu <- stu |> mutate(M_days_enrolled = pmax(M_days_enrolled, 0))
  chk_info("Negative-duration spells floored to zero", n_floored)
} else {
  chk_info("Negative-duration spells left as-is", sum(stu$M_days_enrolled < 0))
}

guard <- range_two_tier(stu$M_days_enrolled, config$ranges$M_days_enrolled,
                        "M_days_enrolled")
stu$M_days_enrolled <- guard$values

chk_stop("Days enrolled is never NA", !any(is.na(stu$M_days_enrolled)))
```

```{r stu-days-by-imputation}
stu |>
  group_by(C_exit_date_imputed) |>
  summarise(
    n_rows       = n(),
    median_days  = median(M_days_enrolled),
    pct_over_thr = round(100 * mean(M_days_enrolled >= config$cohort$duration_min_days), 1),
    .groups = "drop")
```

The threshold counts **calendar** days, not instructional ones. A typical year
has about 180 instructional days, so "200 instructional days" could never be
met by anyone — 200 can only mean calendar days.

What it actually means here: for the 75.5% of rows whose exit date is imputed
to 30 June, `M_days_enrolled` is simply `30 June - entry date`, and 30 June
minus 200 days is exactly **12 December**. So for three quarters of the file
the rule reads "enrolled on or before 12 December". For the rest, which carry a
real withdrawal date at a median of about five weeks, almost none clear it.

Worth writing up that way rather than as "enrolled 200 days", which implies a
measure of attendance this data cannot support. A genuine instructional-day
threshold would need a district school-calendar table.

## 2.4 Locations

Names carry a year stamp, are truncated at ~24 characters, and change spelling
between years. Codes are stable, so names resolve to codes once and everything
downstream matches on the code.

```{r stu-locations}
stu <- stu |>
  mutate(
    C_location_name = clean_location_name(
      D_location_name,
      strip_prefix = config$derived$location_name_strip_prefix,
      strip_suffix = config$derived$location_name_strip_suffix,
      aliases      = config$cohort$school_name_aliases),
    C_credit_recovery_site = grepl(config$cohort$credit_recovery_name_pattern,
                                   D_location_name)
  )

chk_info("Credit recovery site rows", sum(stu$C_credit_recovery_site))

stu <- apply_location_merges(stu,
  id_col = "D_location_id", name_col = "C_location_name",
  year_col = "D_year", merges = config$cohort$location_merges)

location_xwalk <- build_location_crosswalk(
  stu, "C_location_id", "C_location_name_merged",
  include = config$cohort$schools_to_include)

included_location_ids <- attr(location_xwalk, "included_ids")
```

Hillwood and James Lawson are one institution across time, not two schools.
MNPS relocated the whole campus — all non-graduating students and the teaching
staff moved from the Hillwood building to the new Bellevue facility for August
2023 — so a student appearing at both is continuous enrolment, not a transfer.
`location_merges` remaps Hillwood's code onto James Lawson's, leaving the raw
`D_location_id` untouched and putting the merged code in `C_location_id`.

The merge is verified rather than trusted: each name must resolve to exactly
one code, and the two must not operate in the same year. Concurrent operation
would mean two schools rather than one relocation, and the run would stop.

Credit-recovery sites already carry the **parent school's** code — `S2022 CR
Cane Ridge HS` is code 182, the same as Cane Ridge High School — so matching on
the code rolls them up with no extra work. Only the display name needed fixing,
which is what canonical naming does. The flag is kept so they stay findable.

```{r stu-canonical-names}
stu <- resolve_canonical_names(stu, "C_location_id", "C_location_name_merged")

stu <- stu |>
  mutate(C_school_in_scope = C_location_id %in% included_location_ids)

chk_stop("Credit recovery rows resolve to a parent school",
         all(stu$C_location_name_canonical[stu$C_credit_recovery_site] !=
             stu$C_location_name[stu$C_credit_recovery_site]),
         value = paste0(sum(stu$C_credit_recovery_site), " row(s)"),
         offenders = stu |>
           filter(C_credit_recovery_site) |>
           distinct(C_location_id, C_location_name, C_location_name_canonical))

chk_info("Rows at in-scope schools",
         paste0(format(sum(stu$C_school_in_scope), big.mark = ","), " of ",
                format(nrow(stu), big.mark = ","), " (",
                round(100 * mean(stu$C_school_in_scope), 1), "%)"))

stu |>
  filter(C_credit_recovery_site) |>
  count(C_location_id, C_location_name, C_location_name_canonical,
               C_school_in_scope, name = "n_rows") |> 
  print(width = Inf)
```

```{r stu-inscope-table}
stu |>
  filter(C_school_in_scope) |>
  count(C_location_id, C_location_name_canonical, name = "n_rows") |>
  arrange(C_location_name_canonical) |> 
  print(width = Inf)
```

### Which schools exist in which years

Schools open and close. An in-scope school missing from a year is either a data
problem or a fact about the district, and nothing in the data distinguishes
them — so the exceptions are declared in config, with a reason that prints
here. An undeclared gap stops the run.

Because Hillwood merges into James Lawson, the merged location is expected in
every year and needs no exception. The override mechanism stays available for
districts where a school genuinely opens or closes without a successor.

```{r stu-school-years}
school_presence <- check_school_year_coverage(
  stu, id_col = "C_location_id", name_col = "C_location_name_canonical",
  year_col = "D_year", include_ids = included_location_ids,
  expectations = config$cohort$school_year_expectations)

explore_gt(school_presence)
```

James Lawson should now show enrolment in all four years — 2022 and 2023 from
the Hillwood building, 2024 and 2025 from Bellevue — so no override is needed.

## 2.5 Demographics recode

Flags stay as 0/1 integers. Race goes through a map that stops on any code the
config does not cover.

```{r stu-recode}
stu <- stu |>
  mutate(C_stu_race = apply_value_map(D_stu_race, config$value_maps$D_stu_race,
                                             "D_stu_race")) |>
  rename(C_stu_ell_flag     = D_stu_ell,
                C_stu_swd_flag     = D_stu_swd,
                C_stu_poverty_flag = D_stu_frl)

for (f in c("C_stu_ell_flag", "C_stu_swd_flag", "C_stu_poverty_flag")) {
  chk_values_in(stu, f, c(0, 1), paste0(f, " is 0/1"))
}

stu |>
  summarise(rows = n(),
                   pct_ell     = round(100 * mean(C_stu_ell_flag), 1),
                   pct_swd     = round(100 * mean(C_stu_swd_flag), 1),
                   pct_poverty = round(100 * mean(C_stu_poverty_flag), 1))
```

## 2.6 Scope

The original config filtered on `D_stu_grade >= 9`, but `D_stu_grade` maps to
`percent` — the course grade. That kept anyone scoring above 9%.

```{r stu-scope}
stu <- apply_scope(stu, config$scope$demographics, id_col = "D_stu_id")
```

## 2.7 Grain

```{r stu-grain}
chk_unique(stu, c("D_stu_id", "D_year", "D_location_id", "D_stu_enter_date"),
           "One row per student / year / school / entry date")

overlaps <- stu |>
  arrange(D_stu_id, D_year, D_location_id, H_enter_clamped) |>
  group_by(D_stu_id, D_year, D_location_id) |>
  filter(n() > 1) |>
  mutate(prev_exit = lag(H_exit_clamped),
                overlaps  = !is.na(prev_exit) & H_enter_clamped < prev_exit) |>
  ungroup()

chk_info("Repeat spells at the same school", n_distinct(overlaps$D_stu_id))
chk_stop("No same-school spells overlap in time",
         sum(overlaps$overlaps, na.rm = TRUE) == 0,
         detail = "sum_of_spells would double-count if any overlapped",
         offenders = overlaps |> filter(overlaps))

stu_cleaned <- stu
chk_info("Stage 2 output rows", nrow(stu_cleaned))
```

## 2.8 Trace students

```{r stu-trace}
set.seed(config$tracing$seed)
trace_ids <- unique(c(config$tracing$students,
                      sample(unique(stu_cleaned$D_stu_id), config$tracing$random_n)))

stu_cleaned |>
  filter(D_stu_id %in% trace_ids) |>
  select(D_stu_id, D_year, C_location_name_canonical,
                !!sym(config$grade_level_names$demographics),
                D_stu_enter_date, D_stu_exit_date, C_exit_date_imputed,
                M_days_enrolled, C_school_in_scope) |>
  arrange(D_stu_id, D_year) |> 
  print(width = Inf)
```

---

# Stage 3 — Transcript cleaning

Replaces `01_raw_data_transcript_validation.Rmd`. The dedup sections of that
script are gone: the file is already perfectly unique on student × year ×
course, with zero fully duplicated rows.

## 3.1 Rename and derive the year

The transcript has no year column. `D_year` comes from `end_year`, which
carries the same four values as the demographics `year` — that is what makes
the Stage 4 join possible. The old `02` script joined on a `D_year` that was
never created.

```{r tr-rename}
chk_stage("Stage 3 — Transcript cleaning")

tr <- transcript_raw |>
  rename(any_of(unlist(config$column_map$transcript))) |>
  rename(!!config$grade_level_names$transcript := D_stu_grade_level) |>
  mutate(D_year = .data[[config$derived$transcript_year_from]])

chk_stop("Rename produced all standard columns",
         all(setdiff(names(config$column_map$transcript), "D_stu_grade_level")
             %in% names(tr)),
         value = paste0(ncol(tr), " columns"))

# Only meaningful where the district supplies both a start and an end year.
if (all(c("D_start_year", "D_end_year") %in% names(tr))) {
  chk_stop("Every school year spans exactly one year",
           all(tr$D_end_year - tr$D_start_year == 1),
           offenders = tr |>
             filter(D_end_year - D_start_year != 1) |>
             count(D_start_year, D_end_year, name = "n_rows"))
} else {
  chk_info("Start/end year check skipped", "district supplies only one year column")
}

chk_stop("Transcript years match demographics years",
         setequal(unique(tr$D_year), unique(stu_cleaned$D_year)),
         value  = paste(sort(unique(tr$D_year)), collapse = ", "),
         detail = paste("demographics:", paste(sort(unique(stu_cleaned$D_year)), collapse = ", ")))

# Assert only what is required unconditionally. D_course_id nullity is handled
# by scope below, which is configured to drop those rows — asserting it here
# halted the run on a row the pipeline was about to discard anyway.
chk_no_na(tr, c("D_stu_id", "D_year"), "Student id and year always present")

chk_info("Rows with no course ID", sum(is.na(tr$D_course_id)),
         detail = "dropped by scope in 3.5",
         data = tr |> filter(is.na(D_course_id)) |>
           select(D_stu_id, D_year, D_course_name,
                         D_creds_earn, D_creds_possible))
```

## 3.2 Term

Derived, since there is no term column. The S1/S2 marker sits on the course
number for most rows and on the course name as a fallback.

```{r tr-term}
tr <- tr |>
  mutate(C_term = match_first_pattern(
    sources = list(D_course_id   = D_course_id,
                   D_course_name = D_course_name),
    rules   = config$derived$term_detection$rules,
    label   = "C_term"))

chk_info("Rows with no term detected", sum(is.na(tr$C_term)),
         data = tr |> filter(is.na(C_term)) |>
           count(D_course_id, D_course_name,
                        D_creds_possible, D_creds_earn, name = "n_rows") |>
           arrange(desc(n_rows)))
```

```{r tr-noterm-credits}
tr |>
  filter(is.na(C_term)) |>
  count(D_creds_possible, D_creds_earn, name = "n_rows") |>
  mutate(pct = round(100 * n_rows / sum(n_rows), 1)) |>
  arrange(desc(n_rows))
```

### 3.2a Are DA/DB terms, or sections?

Every no-term row is a dual-enrolment course, and their IDs come in pairs —
`SST5601DA` and `SST5601DB`, `HLT3801DA` and `HLT3801DB`. That could be a
semester marker in a different dialect, or it could be section A and section B.

The two readings make opposite predictions. If DA/DB are **terms**, a large
share of students take both halves of the same base course in one year. If they
are **sections**, essentially nobody does, because you cannot enrol in two
sections of the same course at once.

```{r tr-de-suffix-test}
de <- tr |>
  filter(grepl("D[AB]\\*?\\s*$", D_course_id)) |>
  mutate(de_suffix = str_extract(D_course_id, "D[AB]"),
                de_base   = sub("D[AB]\\*?\\s*$", "", D_course_id))

chk_info("Rows with a DA/DB course ID", nrow(de),
         detail = paste0(n_distinct(de$de_base), " distinct base course(s)"))

de_both <- de |>
  group_by(D_stu_id, D_year, de_base) |>
  summarise(suffixes = paste(sort(unique(de_suffix)), collapse = "+"),
                   .groups = "drop") |>
  count(suffixes, name = "n_student_course_years") |>
  mutate(pct = round(100 * n_student_course_years /
                              sum(n_student_course_years), 1))

chk_info("Students taking both DA and DB of the same base course in one year",
         "see table",
         detail = "a large DA+DB share means term; near zero means section",
         data = de_both)

de_both
```

```{r tr-de-credits}
de |>
  count(de_suffix, D_creds_possible, D_creds_earn, name = "n_rows") |>
  arrange(de_suffix, desc(n_rows))
```

The DA+DB overlap test above is confounded, and the first run showed why. A
two-semester sequence uses **different base IDs** — `ENG4039` is English Comp I
and `ENG4040` is English Comp II — so a student working through a sequence never
appears as DA+DB on one base. Low overlap was never evidence for sections.

The direct check was available all along. Most DA/DB rows already picked up a
term from their course name, and those names are explicit: `ENG4039DA` is
"DE English Comp I **S1**" and `ENG4040DB` is "DE English Comp II **S2**". If
DA maps to Semester 1 and DB to Semester 2 almost every time, the suffix is a
term marker and the 749 unresolved rows are simply names truncated before the
S1/S2 survived.

```{r tr-de-term-crosstab}
de_term <- de |>
  count(de_suffix, C_term, name = "n_rows") |>
  group_by(de_suffix) |>
  mutate(pct_of_suffix = round(100 * n_rows / sum(n_rows), 1)) |>
  ungroup()

chk_info("DA/DB suffix against the term read from the course name", "see table",
         detail = "DA -> Semester 1 and DB -> Semester 2 means the suffix is a term marker",
         data = de_term)

de_term
```

```{r tr-de-name-check}
de |>
  filter(!is.na(C_term)) |>
  mutate(name_says = case_when(
    grepl("S1\\s*\\*?\\s*$", D_course_name) ~ "S1",
    grepl("S2\\s*\\*?\\s*$", D_course_name) ~ "S2",
    TRUE ~ "other")) |>
  count(de_suffix, name_says, name = "n_rows")
```

```{r tr-de-truncation}
tr |>
  filter(is.na(C_term)) |>
  mutate(name_len = nchar(D_course_name)) |>
  count(name_len, name = "n_rows") |>
  arrange(desc(n_rows))
```

Credits by term. A semester course should be worth half a credit; anything else
is worth understanding before credits get summed.

```{r tr-term-credits}
tr |>
  count(C_term, D_creds_possible, name = "n_rows") |>
  arrange(C_term, desc(n_rows))
```

## 3.3 Course identifiers

The base course ID, with the semester suffix stripped, is kept alongside the
original. The Stage 7 join keys on the original ID **plus** the name, so the
base is for reporting and for a fallback if the primary join underperforms.

```{r tr-course-ids}
tr <- tr |>
  mutate(
    C_course_id_base = str_replace(D_course_id,
                                            config$derived$course_id_strip_suffix, ""),
    C_course_name_trimmed = trimws(D_course_name),
    C_course_name_starred = grepl("\\*", D_course_name)
  )

chk_info("Distinct course IDs (as given)",  n_distinct(tr$D_course_id))
chk_info("Distinct course IDs (base)",      n_distinct(tr$C_course_id_base))
chk_info("Distinct course names",           n_distinct(tr$D_course_name))
chk_info("Course names containing an asterisk", sum(tr$C_course_name_starred))

n_pairs <- tr |> distinct(D_course_id, D_course_name) |> nrow()
chk_info("Distinct course ID + name pairs", n_pairs,
         detail = "this is the Stage 7 join key")
```

How much name variation sits behind a single course ID — the reason the join
keys on both.

```{r tr-id-name-variation}
tr |>
  distinct(D_course_id, D_course_name) |>
  count(D_course_id, name = "n_names") |>
  count(n_names, name = "n_course_ids") |>
  arrange(n_names)
```

## 3.4 Grades and credits

Two tiers of bounds, because "outside what I expected" and "cannot be a real
measurement" are different questions.

`percent` is a computed weighted average, not a rounded mark — 9,994 distinct
values, and a near-continuous run of 100.01, 100.02, 100.03 at roughly eighty
rows apiece. Grades just above 100 are **extra credit, genuinely earned**. A
single bound at 100 would either halt on those or let 981 through; separating
the tiers lets the plausible range describe the data while the hard range
guards against values that are not grades at all.

The plausible tier only reports. The hard tier gates only when
`hard_tolerance_pct` is set in config, which is deliberately left unset for
grades on this pass — read the distribution below, then choose.

```{r tr-ranges}
# Keep the raw value before anything is blanked, so a decision to cap, drop or
# restore later does not require re-reading the source.
tr <- tr |> mutate(D_stu_grade_raw = D_stu_grade)

g_grade <- range_two_tier(tr$D_stu_grade,      config$ranges$D_stu_grade,      "D_stu_grade")
g_earn  <- range_two_tier(tr$D_creds_earn,     config$ranges$D_creds_earn,     "D_creds_earn")
g_poss  <- range_two_tier(tr$D_creds_possible, config$ranges$D_creds_possible, "D_creds_possible")

tr <- tr |>
  mutate(
    D_stu_grade            = g_grade$values,
    D_creds_earn           = g_earn$values,
    D_creds_possible       = g_poss$values,
    C_grade_above_100      = g_grade$flag_plausible,
    C_grade_not_a_grade    = g_grade$flag_hard,
    C_creds_out_of_range   = g_earn$flag_plausible | g_poss$flag_plausible
  )
```

### Shape of the grade tail

How far above 100 does it actually go, and how thin does it get. This is the
table to read before setting a bound.

```{r tr-grade-tail}
grade_tail <- describe_tail(tr$D_stu_grade, config$grade_tail_buckets, "D_stu_grade")
grade_tail
```

```{r tr-grade-summary}
tr |>
  filter(!is.na(D_stu_grade)) |>
  summarise(
    rows          = n(),
    max           = max(D_stu_grade),
    p99           = round(quantile(D_stu_grade, .99), 2),
    p999          = round(quantile(D_stu_grade, .999), 2),
    n_above_100   = sum(D_stu_grade > 100),
    pct_above_100 = round(100 * mean(D_stu_grade > 100), 3),
    n_above_110   = sum(D_stu_grade > 110),
    n_above_150   = sum(D_stu_grade > 150))
```

If grades above 100 cluster in particular courses, the cause is weighted or
extra-credit scoring. If they are spread evenly, it is more likely a gradebook
setting applied district-wide.

```{r tr-grade-by-course}
tr |>
  filter(C_grade_above_100) |>
  count(D_course_id, D_course_name, name = "n_rows") |>
  mutate(pct_of_flagged = round(100 * n_rows / sum(n_rows), 1)) |>
  arrange(desc(n_rows)) |>
  head(20)
```

```{r tr-grade-by-year}
tr |>
  group_by(D_year) |>
  summarise(rows = n(),
                   pct_above_100 = round(100 * mean(D_stu_grade > 100, na.rm = TRUE), 2),
                   max_grade = max(D_stu_grade, na.rm = TRUE),
                   .groups = "drop")
```

The values that cannot be grades at all — and which courses they sit in. A
value like 981 is most likely a misplaced decimal on 98.1, or a points total
recorded in a percent field.

```{r tr-grade-impossible}
# D_stu_grade has already been blanked by the hard-range rule, so show the
# preserved raw value instead of a column of NAs.
tr |>
  filter(C_grade_not_a_grade) |>
  count(D_stu_grade_raw, D_course_id, D_course_name,
               D_creds_earn, D_creds_possible, name = "n_rows") |>
  arrange(desc(D_stu_grade_raw)) |>
  head(30)
```

### Credits

```{r tr-credit-outliers}
tr |>
  filter(C_creds_out_of_range) |>
  count(D_creds_possible, D_creds_earn, D_course_name, name = "n_rows") |>
  arrange(desc(D_creds_earn)) |>
  head(20)
```

Credits earned can never exceed credits attempted.

```{r tr-credit-logic}
tr <- tr |>
  mutate(C_creds_earn_exceeds_attempted =
                  !is.na(D_creds_earn) & !is.na(D_creds_possible) &
                  D_creds_earn > D_creds_possible)

chk_info("Rows where credits earned exceed credits attempted",
         sum(tr$C_creds_earn_exceeds_attempted),
         data = tr |>
           filter(C_creds_earn_exceeds_attempted) |>
           count(D_creds_possible, D_creds_earn, name = "n_rows") |>
           arrange(desc(n_rows)))

chk_info("Rows with zero credits attempted", sum(tr$D_creds_possible == 0, na.rm = TRUE))
chk_info("Rows with zero credits earned",    sum(tr$D_creds_earn == 0, na.rm = TRUE))
```

Tettra has no rule for this, so it is a decision to log rather than a rule to
follow. Before deciding, it is worth knowing whether the dominant 0.5 to 1.0
pattern is a **course-level convention** — some courses always award a full
credit regardless of what the attempted field says — or something scattered.

If a course ID shows 1.0 earned on every row, the attempted field is simply
wrong for that course and the earned figure can be trusted. If the same course
sometimes shows 0.5, the pattern is student-level and needs a different answer.

```{r tr-credit-excess-courses}
excess_courses <- tr |>
  filter(D_creds_possible == 0.5) |>
  group_by(D_course_id, D_course_name) |>
  summarise(rows = n(),
                   n_earn_1 = sum(D_creds_earn == 1.0, na.rm = TRUE),
                   .groups = "drop") |>
  filter(n_earn_1 > 0) |>
  mutate(pct_of_course_at_1 = round(100 * n_earn_1 / rows, 1)) |>
  arrange(desc(n_earn_1))

chk_info("Courses showing the 0.5-attempted / 1.0-earned pattern",
         nrow(excess_courses),
         detail = "pct_of_course_at_1 near 100 means a course-level convention")

head(excess_courses, 25)
```

```{r tr-credit-excess-concentration}
excess_courses |>
  mutate(bucket = cut(pct_of_course_at_1,
                             breaks = c(0, 5, 25, 50, 75, 95, 100),
                             include.lowest = TRUE)) |>
  group_by(bucket) |>
  summarise(n_courses = n(),
                   n_rows_at_1 = sum(n_earn_1), .groups = "drop")
```

The mass sits in the middle buckets, so this is student-level rather than a
course convention. But the courses carrying it are uniformly **S2 and
asterisked** — `Integrated Math II S2 *`, `Biology I S2 *`, `World History &
Geog S2 *`. That points at credit recovery: a student fails the first semester,
retakes it alongside the second, and the S2 record is credited with both while
attempted stays at 0.5.

```{r tr-credit-excess-shape}
tr |>
  filter(C_creds_earn_exceeds_attempted) |>
  count(C_term, C_course_name_starred, name = "n_rows") |>
  mutate(pct = round(100 * n_rows / sum(n_rows), 1)) |>
  arrange(desc(n_rows))
```

The decisive test. If the S2 row is carrying a recovered credit, the same
student's Semester 1 row for the same base course should show **zero** earned.

```{r tr-credit-recovery-test}
s1_companion <- tr |>
  filter(C_term == "Semester 1") |>
  distinct(D_stu_id, D_year, C_course_id_base, .keep_all = TRUE) |>
  select(D_stu_id, D_year, C_course_id_base, s1_creds_earn = D_creds_earn)

recovery <- tr |>
  filter(C_creds_earn_exceeds_attempted, C_term == "Semester 2") |>
  left_join(s1_companion, by = c("D_stu_id", "D_year", "C_course_id_base")) |>
  mutate(s1_status = case_when(
    is.na(s1_creds_earn) ~ "no S1 record",
    s1_creds_earn == 0   ~ "S1 earned zero",
    TRUE                 ~ "S1 earned credit")) |>
  count(s1_status, name = "n_rows") |>
  mutate(pct = round(100 * n_rows / sum(n_rows), 1))

chk_info("S2 rows with excess credit — what happened in their S1", "see table",
         detail = "mostly 'S1 earned zero' means the S2 row carries a recovered credit",
         data = recovery)

recovery
```

### What does the asterisk mean

Credit recovery is one case of it, but not the whole story: 46,780 rows carry
an asterisk against only 6,889 showing excess credit. If asterisked rows fail
more often, or cluster in Semester 2, the marker is about repeating a course
rather than about recovery specifically.

```{r tr-asterisk-profile}
tr |>
  group_by(C_course_name_starred) |>
  summarise(
    rows            = n(),
    pct_semester_2  = round(100 * mean(C_term == "Semester 2", na.rm = TRUE), 1),
    pct_zero_earned = round(100 * mean(D_creds_earn == 0, na.rm = TRUE), 1),
    median_grade    = median(D_stu_grade, na.rm = TRUE),
    pct_excess_cred = round(100 * mean(C_creds_earn_exceeds_attempted), 1),
    .groups = "drop")
```

```{r tr-asterisk-by-year}
tr |>
  group_by(D_year) |>
  summarise(rows = n(),
                   pct_starred = round(100 * mean(C_course_name_starred), 1),
                   .groups = "drop")
```

Do asterisked courses come in pairs with an unasterisked version of the same
course? If a course exists both ways, the asterisk is a variant of the same
course rather than a course in its own right.

```{r tr-asterisk-pairs}
starred_base <- tr |>
  mutate(name_plain = trimws(gsub("[*]", "", D_course_name))) |>
  distinct(D_course_id, name_plain, C_course_name_starred) |>
  count(D_course_id, name_plain, name = "n_variants") |>
  count(n_variants, name = "n_courses")

chk_info("Course ID + plain name combinations by number of asterisk variants",
         "see table",
         detail = "2 means the course exists both with and without the marker",
         data = starred_base)

starred_base
```

```{r tr-credit-treatment}
n_impossible <- sum(tr$D_creds_earn > config$credits$impossible_above, na.rm = TRUE)

if (config$credits$impossible_treatment == "na") {
  tr <- tr |>
    mutate(C_creds_impossible = !is.na(D_creds_earn) &
                    D_creds_earn > config$credits$impossible_above,
                  D_creds_earn = if_else(C_creds_impossible,
                                                NA_real_, D_creds_earn))
  chk_info("Impossible credit values set to NA", n_impossible,
           detail = paste0("earned above ", config$credits$impossible_above,
                           "; the enrolment row is kept, only the number is removed"))
} else if (config$credits$impossible_treatment == "drop") {
  before <- nrow(tr)
  tr <- tr |> filter(is.na(D_creds_earn) |
                              D_creds_earn <= config$credits$impossible_above)
  chk_info("Rows dropped for impossible credit values", before - nrow(tr))
} else {
  chk_info("Impossible credit values left in place", n_impossible)
}

if (config$credits$earned_exceeds_attempted$treatment == "na_earned") {
  tr <- tr |> mutate(D_creds_earn =
    if_else(C_creds_earn_exceeds_attempted, NA_real_, D_creds_earn))
  chk_info("Earned credits blanked where they exceeded attempted",
           sum(tr$C_creds_earn_exceeds_attempted))
}
```

## 3.5 Scope and grain

Dropping a row removes whatever it was carrying, so scope reports what each
rule costs in credits as well as in rows. At one row it is immaterial; the
report exists so it cannot become material without being seen.

```{r tr-scope}
tr <- apply_scope(tr, config$scope$transcript, id_col = "D_stu_id",
                  sum_cols = c("D_creds_earn", "D_creds_possible"))

chk_no_na(tr, c("D_stu_id", "D_year", "D_course_id"),
          "Core transcript keys present after scope")

chk_unique(tr, c("D_stu_id", "D_year", "D_course_id"),
           "One row per student / year / course")

chk_stop("No fully duplicated rows", sum(duplicated(tr)) == 0)
```

### Course load and credit accumulation

A student taking 16 semester records should attempt 8.0 credits. Whether they
*earn* 8.0 is a different question, and the gap between the two is the thing
worth watching: 2024 lands on 8.0 while 2022 comes in a full credit lower.

Two explanations compete. Either students failed more in the earlier years, or
those years carry more zero-credit courses that pad the record count without
contributing credit. Putting attempted next to earned separates them — if
attempted holds at 8.0 across all years and only earned dips, it is failure.

```{r tr-load-credits}
stu_year <- tr |>
  group_by(D_stu_id, D_year) |>
  summarise(
    courses        = n(),
    courses_credit_bearing = sum(D_creds_possible > 0, na.rm = TRUE),
    creds_attempted = sum(D_creds_possible, na.rm = TRUE),
    creds_earned    = sum(D_creds_earn, na.rm = TRUE),
    n_failed        = sum(D_creds_possible > 0 & D_creds_earn == 0, na.rm = TRUE),
    .groups = "drop")

stu_year |>
  group_by(D_year) |>
  summarise(
    students             = n(),
    median_courses       = median(courses),
    median_credit_courses = median(courses_credit_bearing),
    median_attempted     = median(creds_attempted),
    median_earned        = median(creds_earned),
    median_failed        = median(n_failed),
    mean_failed          = round(mean(n_failed), 2),
    pct_earning_all      = round(100 * mean(creds_earned >= creds_attempted), 1),
    .groups = "drop")
```

Failure rates by year, at the row level rather than the student level. If the
earlier years fail more, the credit gap explains itself.

```{r tr-failure-by-year}
tr |>
  filter(D_creds_possible > 0) |>
  group_by(D_year) |>
  summarise(
    credit_bearing_rows = n(),
    n_zero_earned       = sum(D_creds_earn == 0, na.rm = TRUE),
    pct_zero_earned     = round(100 * mean(D_creds_earn == 0, na.rm = TRUE), 1),
    .groups = "drop")
```

And the share of records carrying no credit at all, which is the competing
explanation.

```{r tr-zero-credit-by-year}
tr |>
  group_by(D_year) |>
  summarise(
    rows                 = n(),
    n_zero_attempted     = sum(D_creds_possible == 0, na.rm = TRUE),
    pct_zero_attempted   = round(100 * mean(D_creds_possible == 0, na.rm = TRUE), 2),
    .groups = "drop")
```

```{r tr-credit-distribution}
stu_year |>
  group_by(D_year) |>
  summarise(
    p10 = quantile(creds_earned, .10),
    p25 = quantile(creds_earned, .25),
    p50 = quantile(creds_earned, .50),
    p75 = quantile(creds_earned, .75),
    p90 = quantile(creds_earned, .90),
    .groups = "drop")
```

```{r tr-output}
transcript_cleaned <- tr
chk_info("Stage 3 output rows", nrow(transcript_cleaned))
chk_info("Distinct students in transcript", n_distinct(transcript_cleaned$D_stu_id))
```

## 3.6 Coding file reconnaissance

Not the join — that is Stage 7. This just reports what the coded file contains
so the join can be written against its real column names.

```{r coding-recon}
if (!is.null(course_coding_raw)) {
  chk_info("Coding file shape",
           paste0(format(nrow(course_coding_raw), big.mark = ","), " rows x ",
                  ncol(course_coding_raw), " cols"))
  print(tibble(column = names(course_coding_raw),
                       type   = map_chr(course_coding_raw,
                                               ~ paste(class(.x), collapse = "/")),
                       n_distinct = map_int(course_coding_raw,
                                                   ~ n_distinct(.x)),
                       pct_missing = map_dbl(course_coding_raw,
                                                    ~ round(100 * mean(is.na(.x)), 1))) |>
          as.data.frame())
} else {
  chk_info("Coding file not available", "Stage 7 will stop")
}
```

```{r coding-head}
if (!is.null(course_coding_raw)) head(course_coding_raw, 10)
```

### Stage 7 join preview

Cheap to run now, and it de-risks the biggest remaining unknown. The join keys
on course ID plus name. 46,780 transcript course names carry an asterisk, so if
the coding file stripped those during cleaning, the exact match will miss on
every one of them. Better to find that out here than at Stage 7.

```{r coding-join-preview}
if (!is.null(course_coding_raw)) {
  tr_pairs <- tr |>
    distinct(D_course_id, D_course_name)

  cd_pairs <- course_coding_raw |>
    distinct(d_course_id, d_course_name) |>
    rename(D_course_id = d_course_id, D_course_name = d_course_name)

  exact <- inner_join(tr_pairs, cd_pairs, by = c("D_course_id", "D_course_name"))

  chk_info("Coding join preview — pairs in transcript", nrow(tr_pairs))
  chk_info("Coding join preview — pairs in coding file", nrow(cd_pairs))
  chk_info("Coding join preview — exact matches on ID + name",
           paste0(format(nrow(exact), big.mark = ","), " (",
                  round(100 * nrow(exact) / nrow(tr_pairs), 1), "% of transcript pairs)"))

  # Weighted by rows, not pairs: an unmatched pair on one row matters less than
  # an unmatched pair on ten thousand.
  matched_rows <- tr |>
    semi_join(cd_pairs, by = c("D_course_id", "D_course_name")) |>
    nrow()
  chk_info("Coding join preview — transcript ROWS matched",
           paste0(format(matched_rows, big.mark = ","), " (",
                  round(100 * matched_rows / nrow(tr), 2), "%)"))

  unmatched <- tr |>
    anti_join(cd_pairs, by = c("D_course_id", "D_course_name")) |>
    count(D_course_id, D_course_name, name = "n_rows") |>
    arrange(desc(n_rows))

  chk_info("Coding join preview — unmatched pairs", nrow(unmatched),
           data = head(unmatched, 25))
}
```

If the exact match falls short, the next question is whether normalising the
name closes the gap — asterisks stripped, whitespace collapsed, case folded.
That tells us whether the fix belongs in the pipeline or in the coding file.

```{r coding-join-fallback}
if (!is.null(course_coding_raw)) {
  norm <- function(x) tolower(trimws(gsub("\\s+", " ", gsub("\\*", "", x))))

  tr_norm <- tr |>
    mutate(k_id = trimws(D_course_id), k_nm = norm(D_course_name))

  cd_norm <- course_coding_raw |>
    distinct(d_course_id, d_course_name) |>
    mutate(k_id = trimws(d_course_id), k_nm = norm(d_course_name)) |>
    distinct(k_id, k_nm)

  matched_norm <- tr_norm |> semi_join(cd_norm, by = c("k_id", "k_nm")) |> nrow()

  chk_info("Coding join preview — rows matched after name normalisation",
           paste0(format(matched_norm, big.mark = ","), " (",
                  round(100 * matched_norm / nrow(tr), 2), "%)"),
           detail = "asterisks removed, whitespace collapsed, case folded")

  # ID alone, as an upper bound on what any name-level fix could achieve.
  matched_id <- tr |>
    semi_join(course_coding_raw |>
                       distinct(D_course_id = d_course_id),
                     by = "D_course_id") |> nrow()
  chk_info("Coding join preview — rows matched on course ID alone",
           paste0(format(matched_id, big.mark = ","), " (",
                  round(100 * matched_id / nrow(tr), 2), "%)"),
           detail = "upper bound for any name-matching fix")
}
```

## 3.7 Trace students

```{r tr-trace}
transcript_cleaned |>
  filter(D_stu_id %in% trace_ids) |>
  select(D_stu_id, D_year, D_course_id, C_course_id_base, D_course_name,
                C_term, D_stu_grade, D_creds_earn, D_creds_possible) |>
  arrange(D_stu_id, D_year, D_course_id) |>
  head(40)
```

---

# Stage 4 — Cross-file reconciliation

Replaces `02_cross_file_validation.Rmd` §2.1. That section built exclusion
flags on both sides and removed mismatched students. Most of it is unnecessary
here: every transcript student appears in demographics, so nothing is lost from
the transcript side.

What still needs doing is quieter. The join key is **student and year**, not
student, and a student present in both files can still be missing the specific
year a transcript row belongs to. That is what this stage measures.

## 4.1 Student-level overlap

```{r x-student-overlap}
chk_stage("Stage 4 — Cross-file reconciliation")

tr_ids  <- unique(transcript_cleaned$D_stu_id)
stu_ids <- unique(stu_cleaned$D_stu_id)

pct_tr_found <- 100 * mean(tr_ids %in% stu_ids)

chk_info("Students in transcript",     length(tr_ids))
chk_info("Students in demographics",   length(stu_ids))
chk_info("Students in demographics only", length(setdiff(stu_ids, tr_ids)))

chk_info("Transcript students found in demographics",
         paste0(round(pct_tr_found, 2), "%"),
         detail = "gated in 4.5")
```

## 4.2 Student-year overlap

The real join key. A student can appear in both files and still have a
transcript row for a year with no matching enrolment record.

```{r x-student-year-measure}
tr_sy  <- transcript_cleaned |> distinct(D_stu_id, D_year)
stu_sy <- stu_cleaned         |> distinct(D_stu_id, D_year)

orphan_sy <- anti_join(tr_sy, stu_sy, by = c("D_stu_id", "D_year"))
pct_sy    <- 100 * (1 - nrow(orphan_sy) / nrow(tr_sy))

chk_info("Transcript student-years",   nrow(tr_sy))
chk_info("Demographics student-years", nrow(stu_sy))
chk_info("Transcript student-years with no enrolment record",
         paste0(format(nrow(orphan_sy), big.mark = ","), " (",
                round(100 - pct_sy, 2), "%)"))
```

```{r x-orphan-by-year}
orphan_sy |>
  count(D_year, name = "orphan_student_years") |>
  left_join(count(tr_sy, D_year, name = "transcript_student_years"),
                   by = "D_year") |>
  mutate(pct = round(100 * orphan_student_years / transcript_student_years, 2))
```

### Where do the orphan years sit in each student's history

Every one of these students **is** in demographics — just not for that year. So
the question is which year. If the orphan years fall *before* a student's first
enrolled year, the likely cause is transferred-in credit: a student who moved
into the district carries a transcript from a prior school, and MNPS has no
enrolment record for those earlier years. A gap in the middle of an enrolled
span would mean something quite different.

```{r x-orphan-position}
stu_span <- stu_sy |>
  group_by(D_stu_id) |>
  summarise(first_enrolled = min(D_year),
                   last_enrolled  = max(D_year),
                   n_enrolled_yrs = n_distinct(D_year),
                   .groups = "drop")

orphan_pos <- orphan_sy |>
  left_join(stu_span, by = "D_stu_id") |>
  mutate(position = case_when(
    is.na(first_enrolled)    ~ "student absent from demographics entirely",
    D_year <  first_enrolled ~ "before first enrolled year",
    D_year >  last_enrolled  ~ "after last enrolled year",
    TRUE                     ~ "gap inside enrolled span"))

chk_info("Orphan student-years by position in the student's history", "see table",
         detail = "mostly 'before first enrolled year' points to transferred-in credit")

orphan_pos |>
  count(position, name = "n_student_years") |>
  mutate(pct = round(100 * n_student_years / sum(n_student_years), 1)) |>
  arrange(desc(n_student_years))
```

```{r x-orphan-position-by-year}
orphan_pos |>
  count(D_year, position, name = "n_student_years") |>
  arrange(D_year, desc(n_student_years))
```

### How much of the transcript is affected

Student-years overstate this if the orphan years carry few courses. Rows and
credits are the figures that matter for any analysis built on them.

```{r x-orphan-weight}
orphan_rows <- transcript_cleaned |>
  semi_join(orphan_sy, by = c("D_stu_id", "D_year"))

chk_info("Transcript rows in orphan student-years",
         paste0(format(nrow(orphan_rows), big.mark = ","), " of ",
                format(nrow(transcript_cleaned), big.mark = ","), " (",
                round(100 * nrow(orphan_rows) / nrow(transcript_cleaned), 2), "%)"))

orphan_rows |>
  group_by(D_year) |>
  summarise(rows = n(),
                   students = n_distinct(D_stu_id),
                   median_courses_per_student = median(table(D_stu_id)),
                   creds_earned = round(sum(D_creds_earn, na.rm = TRUE), 1),
                   .groups = "drop")
```

If these are transferred credits, the course names should look unlike the MNPS
catalogue — or should be the same courses a student would already have taken
elsewhere.

```{r x-orphan-courses}
orphan_rows |>
  count(D_course_id, D_course_name, name = "n_rows") |>
  arrange(desc(n_rows)) |>
  head(20)
```

A student's full picture, for the ones with the most orphan years. Reading a
few of these end to end is usually faster than any summary.

```{r x-orphan-examples}
worst <- orphan_pos |>
  count(D_stu_id, name = "orphan_years") |>
  arrange(desc(orphan_years)) |>
  head(5)

bind_rows(
  tr_sy  |> semi_join(worst, by = "D_stu_id") |>
    mutate(source = "transcript"),
  stu_sy |> semi_join(worst, by = "D_stu_id") |>
    mutate(source = "enrolment")
) |>
  count(D_stu_id, D_year, source, name = "n") |>
  pivot_wider(names_from = source, values_from = n, values_fill = 0) |>
  arrange(D_stu_id, D_year)
```

### The other direction: enrolment with no coursework

Enrolment records with no transcript rows are not a problem to fix — they are a
population to understand. A student enrolled all year at an in-scope school
with no course records means the transcript extract is incomplete, which is a
district question. A short stay at an out-of-scope school means a student
passing through, which is not.

```{r x-orphan-enrolment}
orphan_enrol <- anti_join(stu_sy, tr_sy, by = c("D_stu_id", "D_year"))

chk_info("Enrolment student-years with no transcript rows", nrow(orphan_enrol),
         data = orphan_enrol |> count(D_year, name = "n_student_years"))

stu_cleaned |>
  semi_join(orphan_enrol, by = c("D_stu_id", "D_year")) |>
  group_by(D_year) |>
  summarise(
    spells        = n(),
    median_days   = median(M_days_enrolled),
    pct_in_scope  = round(100 * mean(C_school_in_scope), 1),
    pct_full_year = round(100 * mean(M_days_enrolled >=
                                       config$cohort$duration_min_days), 1),
    .groups = "drop")
```

```{r x-orphan-enrolment-schools}
stu_cleaned |>
  semi_join(orphan_enrol, by = c("D_stu_id", "D_year")) |>
  count(C_location_name_canonical, C_school_in_scope, name = "n_spells") |>
  arrange(desc(n_spells)) |>
  head(20)
```

## 4.3 Grade level agreement

Both files carry a grade level, which is why the old `02` ended up with
`.x`/`.y` columns. They are kept apart by name here, so they can be compared
rather than silently overwritten.

First, is the transcript grade level even stable within a student-year?

```{r x-grade-stability}
# The MODAL grade, not the first. 1,768 student-years carry more than one
# transcript grade level - a senior recovering an 11th-grade course keeps the
# 11th-grade level on that row - so first() returns whichever row the
# file happened to order first and manufactures disagreement that isn't there.
tr_grade <- transcript_cleaned |>
  count(D_stu_id, D_year, D_stu_grade_level_transcript, name = "n_rows") |>
  group_by(D_stu_id, D_year) |>
  mutate(n_grades = n()) |>
  slice_max(n_rows, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(grade_tr = D_stu_grade_level_transcript)

chk_info("Student-years with more than one transcript grade level",
         sum(tr_grade$n_grades > 1),
         data = transcript_cleaned |>
           semi_join(tr_grade |> filter(n_grades > 1),
                            by = c("D_stu_id", "D_year")) |>
           count(D_stu_id, D_year, D_stu_grade_level_transcript,
                        name = "n_rows") |>
           head(20))
```

Demographics can legitimately carry more than one grade level per student-year
when a student changes school mid-year, so the comparison uses the grade from
the spell with the most days enrolled.

```{r x-grade-compare}
stu_grade <- stu_cleaned |>
  group_by(D_stu_id, D_year) |>
  slice_max(M_days_enrolled, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(D_stu_id, D_year,
                grade_demo = !!sym(config$grade_level_names$demographics))

grade_cmp <- tr_grade |>
  inner_join(stu_grade, by = c("D_stu_id", "D_year")) |>
  mutate(diff = grade_tr - grade_demo)

chk_info("Student-years where grade level agrees",
         paste0(round(100 * mean(grade_cmp$diff == 0), 2), "% (",
                format(sum(grade_cmp$diff == 0), big.mark = ","), " of ",
                format(nrow(grade_cmp), big.mark = ","), ")"))

grade_cmp |>
  count(diff, name = "n_student_years") |>
  mutate(pct = round(100 * n_student_years / sum(n_student_years), 2)) |>
  arrange(diff)
```

```{r x-grade-disagree-detail}
grade_cmp |>
  filter(diff != 0) |>
  count(grade_demo, grade_tr, name = "n_student_years") |>
  arrange(desc(n_student_years)) |>
  head(20)
```

## 4.4 Reconciliation flags

Nothing is dropped here. Flags are attached so any downstream stage can decide,
and so the decision is visible rather than buried in a filter.

```{r x-flags}
transcript_reconciled <- transcript_cleaned |>
  mutate(
    C_has_enrolment_record = paste(D_stu_id, D_year) %in%
      paste(stu_sy$D_stu_id, stu_sy$D_year)
  )

chk_info("Transcript rows with an enrolment record",
         paste0(format(sum(transcript_reconciled$C_has_enrolment_record),
                       big.mark = ","), " of ",
                format(nrow(transcript_reconciled), big.mark = ","), " (",
                round(100 * mean(transcript_reconciled$C_has_enrolment_record), 2), "%)"))

chk_rows_stable(transcript_cleaned, transcript_reconciled,
                "Stage 4 changed no rows")
```

## 4.5 Gates

Every structural check for this stage runs here, after the tables above have
printed. If one fails, the evidence is already on the page.

```{r x-gates}
pct_students <- 100 * mean(unique(transcript_cleaned$D_stu_id) %in%
                             unique(stu_cleaned$D_stu_id))

chk_stop("Transcript students found in demographics",
         pct_students >= config$thresholds$min_student_match_pct,
         value  = paste0(round(pct_students, 2), "%"),
         detail = paste0("threshold is ", config$thresholds$min_student_match_pct, "%"),
         offenders = tibble(
           D_stu_id = setdiff(unique(transcript_cleaned$D_stu_id),
                              unique(stu_cleaned$D_stu_id))))

chk_stop("Transcript student-years found in demographics",
         pct_sy >= config$thresholds$min_student_year_match_pct,
         value  = paste0(round(pct_sy, 2), "%"),
         detail = paste0("threshold is ", config$thresholds$min_student_year_match_pct,
                         "%; see 4.2 for what the unmatched years are"),
         offenders = orphan_pos)

chk_stop("Grade level agreement is not systematically broken",
         mean(grade_cmp$diff == 0) >= 0.90,
         value  = paste0(round(100 * mean(grade_cmp$diff == 0), 2), "%"),
         detail = "below 90% would mean the two files disagree about who is in which grade",
         offenders = grade_cmp |> filter(diff != 0))
```

---

# Stage 5 — Home school assignment

Replaces `02_cross_file_validation.Rmd` §2.2. Collapses enrolment spells to
**one row per student per year**, assigning each student-year a single home
school, then joins those attributes onto the transcript.

About 10,700 student-years involve more than one school, so this is a real
decision rather than a formality: the home school determines which school a
student's outcomes are attributed to.

## 5.1 Days by school

```{r hs-days-by-school}
chk_stage("Stage 5 — Home school assignment")

spell_school <- stu_cleaned |>
  group_by(D_stu_id, D_year, C_location_id) |>
  summarise(
    days_at_school = sum(M_days_enrolled),
    n_spells       = n(),
    first_enter    = min(D_stu_enter_date),
    last_exit      = max(D_stu_exit_date),
    in_scope       = any(C_school_in_scope),
    .groups = "drop")

chk_unique(spell_school, c("D_stu_id", "D_year", "C_location_id"),
           "One row per student / year / school after summing spells")

spell_school |>
  count(D_stu_id, D_year, name = "schools") |>
  count(schools, name = "n_student_years") |>
  mutate(pct = round(100 * n_student_years / sum(n_student_years), 2))
```

## 5.2 Ties

A student splitting a year evenly between two schools has no natural home
school. The rule has to be declared rather than left to row order, or the same
data produces different answers on different runs.

```{r hs-ties}
ties <- spell_school |>
  group_by(D_stu_id, D_year) |>
  filter(n() > 1, days_at_school == max(days_at_school)) |>
  filter(n() > 1) |>
  ungroup()

chk_info("Student-years where two or more schools tie on days",
         n_distinct(paste(ties$D_stu_id, ties$D_year)),
         detail = paste0("resolved by: ", config$student_year_collapse$tiebreak))

ties |>
  group_by(D_stu_id, D_year) |>
  summarise(schools_tied = n(),
                   days_each    = first(days_at_school),
                   any_in_scope = any(in_scope),
                   .groups = "drop") |>
  count(schools_tied, any_in_scope, name = "n_student_years")
```

## 5.3 Assign the home school

```{r hs-assign}
# The tiebreak is materialised as a column rather than an inline `if` inside
# arrange(), which would resolve to a length-1 FALSE under one config setting
# and a full column under the other.
home_school <- spell_school |>
  mutate(tie_pref = case_when(
    config$student_year_collapse$tiebreak == "prefer_included" ~ as.integer(in_scope),
    TRUE ~ 0L)) |>
  group_by(D_stu_id, D_year) |>
  arrange(desc(days_at_school),
                 desc(tie_pref),
                 C_location_id,
                 .by_group = TRUE) |>
  slice(1) |>
  ungroup() |>
  select(D_stu_id, D_year,
                C_home_school_id      = C_location_id,
                C_home_school_days    = days_at_school,
                C_home_school_in_scope = in_scope)

chk_unique(home_school, c("D_stu_id", "D_year"),
           "One home school per student per year")

chk_ids_stable(stu_cleaned, home_school, "D_stu_id",
               "No students lost assigning home school")
```

```{r hs-share-of-year}
home_school |>
  left_join(
    spell_school |>
      group_by(D_stu_id, D_year) |>
      summarise(days_all_schools = sum(days_at_school), .groups = "drop"),
    by = c("D_stu_id", "D_year")) |>
  mutate(share = if_else(days_all_schools > 0,
                                       C_home_school_days / days_all_schools, NA_real_)) |>
  summarise(
    student_years        = n(),
    median_share         = round(median(share, na.rm = TRUE), 3),
    pct_home_school_100  = round(100 * mean(share >= 0.999, na.rm = TRUE), 1),
    pct_home_school_lt75 = round(100 * mean(share < 0.75, na.rm = TRUE), 1))
```

A home school holding well under three quarters of a student's year is a weak
attribution. The share is carried forward so any analysis can exclude or
sensitivity-test those students rather than treating every attribution as
equally solid.

## 5.4 Collapse to one row per student-year

```{r hs-collapse}
flag_fn <- if (config$student_year_collapse$flag_rule == "any") max else first

student_year <- stu_cleaned |>
  group_by(D_stu_id, D_year) |>
  summarise(
    M_days_enrolled_year = sum(M_days_enrolled),
    n_spells             = n(),
    n_schools            = n_distinct(C_location_id),
    C_stu_race           = first(C_stu_race),
    n_races              = n_distinct(C_stu_race),
    C_stu_ell_flag       = flag_fn(C_stu_ell_flag),
    C_stu_swd_flag       = flag_fn(C_stu_swd_flag),
    C_stu_poverty_flag   = flag_fn(C_stu_poverty_flag),
    C_exit_date_imputed  = all(C_exit_date_imputed),
    .groups = "drop") |>
  join_checked(home_school, by = c("D_stu_id", "D_year"),
               check = "Home school onto student-year")

chk_unique(student_year, c("D_stu_id", "D_year"),
           "One row per student per year")

chk_info("Student-years with more than one recorded race",
         sum(student_year$n_races > 1),
         detail = "race should be constant within a student-year")
```

The grade level for the year comes from the home-school spell, so it describes
where the student actually spent the year.

```{r hs-grade-level}
grade_source <- stu_cleaned |>
  inner_join(home_school |> select(D_stu_id, D_year, C_home_school_id),
                    by = c("D_stu_id", "D_year")) |>
  filter(C_location_id == C_home_school_id) |>
  group_by(D_stu_id, D_year) |>
  slice_max(M_days_enrolled, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(D_stu_id, D_year,
                C_stu_grade_level = !!sym(config$grade_level_names$demographics),
                C_home_school_name = C_location_name_canonical)

student_year <- student_year |>
  join_checked(grade_source, by = c("D_stu_id", "D_year"),
               check = "Grade level and school name onto student-year")

chk_no_na(student_year, c("C_home_school_id", "C_stu_grade_level"),
          "Home school and grade level present on every student-year")
```

```{r hs-student-year-summary}
student_year |>
  group_by(D_year) |>
  summarise(
    student_years   = n(),
    pct_multischool = round(100 * mean(n_schools > 1), 1),
    median_days     = median(M_days_enrolled_year),
    pct_in_scope    = round(100 * mean(C_home_school_in_scope), 1),
    pct_ell         = round(100 * mean(C_stu_ell_flag), 1),
    pct_swd         = round(100 * mean(C_stu_swd_flag), 1),
    pct_poverty     = round(100 * mean(C_stu_poverty_flag), 1),
    .groups = "drop")
```

## 5.5 Join onto the transcript

The rollup the old pipeline never actually built. Transcript rows in orphan
student-years — coursework from before a transfer student enrolled — will not
match, which is correct: MNPS has no enrolment record to attach.

```{r hs-join-transcript}
transcript_student <- transcript_reconciled |>
  join_checked(student_year, by = c("D_stu_id", "D_year"),
               check = "Student-year attributes onto transcript")

chk_rows_stable(transcript_reconciled, transcript_student,
                "Join did not change the transcript row count")

chk_info("Transcript rows with a home school",
         paste0(format(sum(!is.na(transcript_student$C_home_school_id)),
                       big.mark = ","), " (",
                round(100 * mean(!is.na(transcript_student$C_home_school_id)), 2), "%)"))
```

```{r hs-join-by-year}
transcript_student |>
  group_by(D_year) |>
  summarise(
    rows            = n(),
    pct_with_school = round(100 * mean(!is.na(C_home_school_id)), 2),
    pct_in_scope    = round(100 * mean(C_home_school_in_scope, na.rm = TRUE), 1),
    .groups = "drop")
```

## 5.6 Trace students

```{r hs-trace}
student_year |>
  filter(D_stu_id %in% trace_ids) |>
  select(D_stu_id, D_year, C_home_school_name, C_stu_grade_level,
                n_spells, n_schools, C_home_school_days, M_days_enrolled_year,
                C_home_school_in_scope) |>
  arrange(D_stu_id, D_year)
```

## 5.7 Gates

```{r hs-gates}
chk_stop("Every student-year has exactly one home school",
         nrow(student_year) == n_distinct(paste(student_year$D_stu_id,
                                                       student_year$D_year)))

chk_stop("Student-year count matches the distinct student-years in enrolment",
         nrow(student_year) == nrow(distinct(stu_cleaned, D_stu_id, D_year)),
         value = paste0(format(nrow(student_year), big.mark = ","), " rows"))

chk_stop("Days at home school never exceed total days enrolled",
         all(student_year$C_home_school_days <= student_year$M_days_enrolled_year),
         offenders = student_year |>
           filter(C_home_school_days > M_days_enrolled_year) |>
           select(D_stu_id, D_year, C_home_school_days, M_days_enrolled_year))
```

---

# Stage 6 — Cohort selection

Replaces `03_student_cohort_selection.Rmd`. Selects students who entered grade
9 in 2022 and were continuously enrolled through 2025, applying the criteria
one at a time so each rule's cost is visible.

Every rule here comes from config, and the funnel below reports what each one
removes. A criterion that drops far more students than expected is the easiest
way to publish a cohort that quietly describes a narrower population than it
claims to.

## 6.1 Starting universe

```{r co-universe}
chk_stage("Stage 6 — Cohort selection")

cohort_years <- sort(as.integer(names(config$cohort$progression)))
expected_grade <- tibble(
  D_year      = cohort_years,
  grade_expected = as.integer(unlist(config$cohort$progression)))

chk_info("Cohort years", paste(cohort_years, collapse = ", "))
expected_grade
```

```{r co-entrants}
entrants_all <- student_year |>
  filter(D_year == config$cohort$entry_year,
                C_stu_grade_level == config$cohort$entry_grade)

chk_info("Students in the entry grade in the entry year, district-wide",
         nrow(entrants_all),
         detail = paste0("grade ", config$cohort$entry_grade, " in ",
                         config$cohort$entry_year))

chk_info("Of those, at an in-scope home school",
         paste0(format(sum(entrants_all$C_home_school_in_scope), big.mark = ","),
                " (", round(100 * mean(entrants_all$C_home_school_in_scope), 1), "%)"))
```

Starting the funnel district-wide would make the school criterion do two jobs
at once — removing students who were never at a cohort school alongside those
who left one. Those are different phenomena and deserve separate rows. The
universe is therefore set by config; the entrants who fall outside it are
reported here rather than absorbed into a later step.

```{r co-out-of-scope-entrants}
entrants_all |>
  filter(!C_home_school_in_scope) |>
  count(C_home_school_name, name = "entrants") |>
  arrange(desc(entrants))
```

```{r co-universe-set}
universe_ids <- switch(config$cohort$universe_rule,
  entry_grade_any_school = entrants_all$D_stu_id,
  entry_grade_in_scope   = entrants_all$D_stu_id[entrants_all$C_home_school_in_scope],
  stop("Unknown universe_rule: ", config$cohort$universe_rule, call. = FALSE))

entrants <- tibble(D_stu_id = unique(universe_ids))

chk_info("Universe rule", config$cohort$universe_rule)

funnel <- funnel_new(entrants$D_stu_id,
  paste0("Grade ", config$cohort$entry_grade, " in ", config$cohort$entry_year,
         if (config$cohort$universe_rule == "entry_grade_in_scope")
           ", in-scope school" else ", district-wide"))
```

## 6.2 Enrolled in every required year

```{r co-presence}
cohort_years_data <- student_year |>
  semi_join(entrants, by = "D_stu_id") |>
  filter(D_year %in% cohort_years)

present <- cohort_years_data |>
  group_by(D_stu_id) |>
  summarise(years_present = n_distinct(D_year), .groups = "drop")

chk_info("Years of enrolment among entrants", "see table",
         data = count(present, years_present, name = "n_students") |>
           arrange(desc(years_present)))

funnel <- funnel_add(funnel,
  paste0("Enrolled in all ", config$cohort$years_required, " years"),
  present$D_stu_id[present$years_present == config$cohort$years_required])
```

## 6.3 Duration

```{r co-duration}
window_days <- as.numeric(
  as.Date(paste0(config$cohort$entry_year, "-", config$derived$school_year_end_mmdd)) -
  as.Date(paste0(config$cohort$entry_year - 1, "-", config$derived$school_year_start_mmdd)))

duration_by_year <- cohort_years_data |>
  mutate(
    share_of_year = M_days_enrolled_year / window_days,
    ok = if (config$cohort$duration_basis == "share")
      share_of_year >= config$cohort$duration_min_share
    else M_days_enrolled_year >= config$cohort$duration_min_days)

chk_info("Duration basis",
         paste0(config$cohort$duration_basis, " — ",
                if (config$cohort$duration_basis == "share")
                  paste0(">= ", 100 * config$cohort$duration_min_share, "% of a ",
                         window_days, "-day window")
                else paste0(">= ", config$cohort$duration_min_days, " days")))

duration_by_year |>
  group_by(D_year) |>
  summarise(students = n(),
                   median_days = median(M_days_enrolled_year),
                   pct_meeting = round(100 * mean(ok), 1),
                   .groups = "drop")
```

The Tettra rule is 75% of the school year. On this window that is about
`r round(0.75 * 250)` days, against the inherited 200. The comparison below
shows what the choice costs before it is made.

```{r co-duration-sensitivity}
map_dfr(c(150, 180, 200, 225, 250, 275), function(thr) {
  ok_ids <- cohort_years_data |>
    mutate(ok = M_days_enrolled_year >= thr) |>
    summarise_year_rule("D_stu_id", "D_year", config$cohort$duration_rule, cohort_years) |>
    filter(passes) |>
    pull(D_stu_id)
  tibble(threshold_days = thr,
                 pct_of_window  = round(100 * thr / window_days, 1),
                 students_passing = length(intersect(funnel$ids, ok_ids)))
})
```

```{r co-duration-apply}
duration_pass <- duration_by_year |>
  summarise_year_rule("D_stu_id", "D_year", config$cohort$duration_rule, cohort_years)

funnel <- funnel_add(funnel,
  paste0("Duration (", config$cohort$duration_rule, ")"),
  duration_pass$D_stu_id[duration_pass$passes])
```

## 6.4 Grade progression

A student must be in the expected grade each year. Retention and
reclassification both break this, which is the point — but the count should be
plausible rather than enormous.

```{r co-progression}
progression_by_year <- cohort_years_data |>
  left_join(expected_grade, by = "D_year") |>
  mutate(ok = C_stu_grade_level == grade_expected)

progression_by_year |>
  group_by(D_year, grade_expected) |>
  summarise(students = n(),
                   pct_on_grade = round(100 * mean(ok), 1),
                   .groups = "drop")
```

```{r co-progression-offgrade}
progression_by_year |>
  filter(!ok) |>
  count(D_year, grade_expected, C_stu_grade_level, name = "n_students") |>
  arrange(D_year, desc(n_students))
```

```{r co-progression-apply}
progression_pass <- progression_by_year |>
  summarise_year_rule("D_stu_id", "D_year", config$cohort$progression_rule, cohort_years)

funnel <- funnel_add(funnel,
  paste0("Grade progression (", config$cohort$progression_rule, ")"),
  progression_pass$D_stu_id[progression_pass$passes])
```

## 6.5 School in scope

```{r co-school}
school_rule <- config$cohort$school_rule

school_by_year <- cohort_years_data |>
  mutate(ok = C_home_school_in_scope)

school_pass <- switch(school_rule,
  in_scope_all_years  = summarise_year_rule(school_by_year, "D_stu_id", "D_year",
                                            "all_years", cohort_years),
  in_scope_entry_year = summarise_year_rule(school_by_year, "D_stu_id", "D_year",
                                            "bookends", config$cohort$entry_year),
  in_scope_any_year   = school_by_year |>
    group_by(D_stu_id) |>
    summarise(passes = any(ok), .groups = "drop"),
  stop("Unknown school_rule: ", school_rule, call. = FALSE))

chk_info("School rule", school_rule)

step_label <- if (config$cohort$universe_rule == "entry_grade_in_scope")
  "Stayed in scope every year" else paste0("Home school in scope (", school_rule, ")")

funnel <- funnel_add(funnel, step_label, school_pass$D_stu_id[school_pass$passes])
```

With an in-scope universe this step now measures only students who **started**
at a cohort school and later moved out — which is what "left the cohort
schools" should mean. Previously it also swept up everyone who was never at
one.

```{r co-school-note}
invisible(NULL)
```

## 6.6 The funnel

```{r co-funnel}
funnel_table <- funnel_report(funnel, config$cohort$benchmarks$step_retention_pct)

chk_info("Cohort selected", length(funnel$ids),
         detail = paste0(round(100 * length(funnel$ids) / funnel$start, 1),
                         "% of the starting universe"))

funnel_table
```

`vs_benchmark` compares each step's retention against the Lubbock figure of
`r config$cohort$benchmarks$step_retention_pct`% from the Tettra guidance. A
step far below it is not necessarily wrong, but it is worth being able to
explain.

## 6.7 Who gets excluded

The attrition matters as much as the cohort. If a rule removes mobile students,
poor students or English learners disproportionately, the cohort describes a
narrower experience than its size suggests — and every equity finding built on
it inherits that narrowing.

```{r co-attrition-groups}
entry_profile <- entrants_all |>
  semi_join(entrants, by = "D_stu_id") |>
  mutate(in_cohort = D_stu_id %in% funnel$ids)

chk_info("Entrants profiled for attrition", nrow(entry_profile))

bind_rows(
  entry_profile |> group_by(group = "All entrants") |>
    summarise(n = n(), pct_retained = round(100 * mean(in_cohort), 1),
                     .groups = "drop"),
  entry_profile |> group_by(group = paste0("Race: ", C_stu_race)) |>
    summarise(n = n(), pct_retained = round(100 * mean(in_cohort), 1),
                     .groups = "drop"),
  entry_profile |> group_by(group = paste0("ELL: ", C_stu_ell_flag)) |>
    summarise(n = n(), pct_retained = round(100 * mean(in_cohort), 1),
                     .groups = "drop"),
  entry_profile |> group_by(group = paste0("SWD: ", C_stu_swd_flag)) |>
    summarise(n = n(), pct_retained = round(100 * mean(in_cohort), 1),
                     .groups = "drop"),
  entry_profile |> group_by(group = paste0("Poverty: ", C_stu_poverty_flag)) |>
    summarise(n = n(), pct_retained = round(100 * mean(in_cohort), 1),
                     .groups = "drop")
) |>
  arrange(group)
```

```{r co-attrition-schools}
entry_profile |>
  group_by(C_home_school_name) |>
  summarise(entrants = n(),
                   cohort   = sum(in_cohort),
                   pct_retained = round(100 * mean(in_cohort), 1),
                   .groups = "drop") |>
  arrange(pct_retained)
```

## 6.8 Assign the cohort school

A student can change home school across the four years. The attribution rule is
declared rather than assumed, and the multi-school flag is kept so any analysis
can separate stayers from movers.

```{r co-cohort-school}
cohort_ids <- tibble(D_stu_id = funnel$ids)

cohort_school <- student_year |>
  semi_join(cohort_ids, by = "D_stu_id") |>
  filter(D_year %in% cohort_years) |>
  group_by(D_stu_id) |>
  mutate(n_home_schools = n_distinct(C_home_school_id)) |>
  arrange(desc(C_home_school_days), D_year, .by_group = TRUE) |>
  slice(switch(config$cohort$four_year_school_rule,
                      most_days = 1L,
                      entry_year = 1L,
                      final_year = 1L,
                      modal = 1L)) |>
  ungroup() |>
  select(D_stu_id,
                C_cohort_school_id   = C_home_school_id,
                C_cohort_school_name = C_home_school_name,
                C_multiple_home_schools = n_home_schools) |>
  mutate(C_multiple_home_schools_flag = C_multiple_home_schools > 1)

chk_unique(cohort_school, "D_stu_id", "One cohort school per student")

chk_info("Cohort students attending more than one home school",
         paste0(sum(cohort_school$C_multiple_home_schools_flag), " (",
                round(100 * mean(cohort_school$C_multiple_home_schools_flag), 1), "%)"))

cohort_school |>
  count(C_cohort_school_name, name = "cohort_students") |>
  arrange(desc(cohort_students))
```

## 6.9 The cohort table

```{r co-table}
cohort <- cohort_school |>
  join_checked(entry_profile |>
                 select(D_stu_id, C_stu_race, C_stu_ell_flag,
                               C_stu_swd_flag, C_stu_poverty_flag),
               by = "D_stu_id", check = "Entry demographics onto cohort") |>
  mutate(C_cohort_entry_year = config$cohort$entry_year,
                C_cohort_include    = TRUE)

chk_unique(cohort, "D_stu_id", "One row per cohort student")
chk_info("Cohort size", nrow(cohort))

cohort |>
  summarise(students    = n(),
                   pct_ell     = round(100 * mean(C_stu_ell_flag), 1),
                   pct_swd     = round(100 * mean(C_stu_swd_flag), 1),
                   pct_poverty = round(100 * mean(C_stu_poverty_flag), 1),
                   pct_movers  = round(100 * mean(C_multiple_home_schools_flag), 1))
```

## 6.10 Criteria table

Tettra's method is a **flag with everyone retained**, not a filter: "9th graders
who attended a high school in the district for any amount of time — these are
all the students in the data file". Both sets are objects of interest, and the
comprehensiveness test it asks for needs the excluded students.

This assembles one row per entrant carrying every criterion outcome, so the
attrition question stays answerable from the exported files rather than only
from this report.

```{r co-criteria}
cohort_criteria <- entrants |>
  left_join(
    present |> transmute(D_stu_id,
      C_met_years = years_present == config$cohort$years_required),
    by = "D_stu_id") |>
  left_join(
    duration_pass |> rename(C_met_duration = passes), by = "D_stu_id") |>
  left_join(
    progression_pass |> rename(C_met_progression = passes), by = "D_stu_id") |>
  left_join(
    school_pass |> rename(C_met_school = passes), by = "D_stu_id") |>
  mutate(across(starts_with("C_met_"),
                              ~ coalesce(.x, FALSE))) |>
  mutate(
    C_stu_cohort = if_else(D_stu_id %in% funnel$ids, "include", "exclude"),
    C_cohort_exclusion_reason = case_when(
      C_stu_cohort == "include" ~ NA_character_,
      !C_met_years              ~ "Not enrolled in all required years",
      !C_met_duration           ~ "Did not meet duration",
      !C_met_progression        ~ "Did not progress on grade",
      !C_met_school             ~ "Left the in-scope schools",
      TRUE                      ~ "Other"))

chk_unique(cohort_criteria, "D_stu_id", "One criteria row per entrant")

chk_info("Cohort flag counts", "see table",
         data = count(cohort_criteria, C_stu_cohort, name = "students"))

cohort_criteria |>
  count(C_cohort_exclusion_reason, name = "students") |>
  mutate(pct = round(100 * students / sum(students), 1)) |>
  arrange(desc(students))
```

The reason column reports the **first** criterion a student failed, in funnel
order. A student can fail several; this names where they first dropped out,
which is what makes the counts sum to the funnel.

## 6.11 Trace students

```{r co-trace}
student_year |>
  filter(D_stu_id %in% trace_ids, D_year %in% cohort_years) |>
  select(D_stu_id, D_year, C_home_school_name, C_stu_grade_level,
                M_days_enrolled_year, C_home_school_in_scope) |>
  mutate(in_cohort = D_stu_id %in% funnel$ids) |>
  arrange(D_stu_id, D_year)
```

## 6.12 Gates

```{r co-gates}
chk_stop("Cohort is not empty", nrow(cohort) > 0)

chk_stop("Every cohort student is in the entry grade in the entry year",
         all(cohort$D_stu_id %in% entrants$D_stu_id))

chk_stop("Every cohort student has a school attribution",
         !any(is.na(cohort$C_cohort_school_id)))

chk_stop("Cohort schools are all in scope",
         all(cohort$C_cohort_school_id %in% included_location_ids),
         offenders = cohort |>
           filter(!C_cohort_school_id %in% included_location_ids) |>
           count(C_cohort_school_id, C_cohort_school_name, name = "n_students"))

chk_info("Overall cohort retention against the Tettra snapshot benchmark",
         paste0(round(100 * nrow(cohort) / funnel$start, 1), "% vs ",
                config$cohort$benchmarks$snapshot_retention_pct, "% benchmark"))
```

---

# Stage 7 — Course coding join

The rollup the old pipeline never actually built. `02` sized the mismatch
between transcript and catalogue but never joined them, so no analysis could
cut by ERS subject, rigor or credit type.

Stage 3.6 established that exact matching on course ID **plus** name reaches
99.99% of rows, so the primary join is exact and the fallback handles a
residue of about 125 rows.

## 7.1 Prepare the coded file

```{r cd-prepare}
chk_stage("Stage 7 — Course coding join")

chk_stop("Coding file is available", !is.null(course_coding_raw),
         detail = "Stage 1 could not read it; see the Stage 1 log")

coding <- course_coding_raw |>
  rename(any_of(unlist(config$column_map$course_coding))) |>
  rename_with(~ sub("^c_", "C_", .x), starts_with("c_"))

chk_info("Coded fields available", ncol(coding),
         data = tibble(column = names(coding)))

missing_fields <- setdiff(config$coding_join$required_coded_fields, names(coding))
chk_stop("Required coded fields present",
         length(missing_fields) == 0,
         detail = if (length(missing_fields) > 0)
           paste("Missing:", paste(missing_fields, collapse = ", ")))
```

The join key must be unique in the coded file or the join multiplies transcript
rows. `join_checked` would catch that, but catching it here names the problem
more clearly.

```{r cd-key-unique}
chk_unique(coding, config$coding_join$keys,
           "Coded file is unique on course ID + name")
```

## 7.2 Primary join

```{r cd-join}
transcript_coded <- transcript_student |>
  join_checked(coding, by = config$coding_join$keys,
               check = "Course coding onto transcript",
               report_unmatched = TRUE)

chk_rows_stable(transcript_student, transcript_coded,
                "Coding join did not change the transcript row count")

transcript_coded <- transcript_coded |>
  mutate(C_coding_source = if_else(
    is.na(C_course_subject_area), NA_character_, "id_and_name"))

chk_info("Rows coded on the primary key",
         paste0(format(sum(!is.na(transcript_coded$C_course_subject_area)),
                       big.mark = ","), " (",
                round(100 * mean(!is.na(transcript_coded$C_course_subject_area)), 3), "%)"))
```

## 7.3 Fallback on course ID

Only where a course ID maps to exactly one coded course. `BUS3718S1` answers to
four different names in the transcript, so an unconditional ID fallback would
hand it whichever coding happened to come first.

```{r cd-fallback}
if (config$coding_join$fallback == "id_when_unambiguous") {

  unambiguous <- coding |>
    group_by(D_course_id) |>
    filter(n_distinct(C_course_id) == 1) |>
    slice(1) |>
    ungroup() |>
    select(-D_course_name)

  chk_info("Course IDs mapping to exactly one coded course",
           paste0(format(nrow(unambiguous), big.mark = ","), " of ",
                  format(n_distinct(coding$D_course_id), big.mark = ",")))

  coded_cols <- setdiff(names(unambiguous), "D_course_id")

  filled <- transcript_coded |>
    filter(is.na(C_course_subject_area)) |>
    select(-all_of(coded_cols), -C_coding_source) |>
    inner_join(unambiguous, by = "D_course_id") |>
    mutate(C_coding_source = "id_only")

  chk_info("Rows recovered by the ID fallback", nrow(filled),
           data = filled |> count(D_course_id, D_course_name, name = "n_rows") |>
             arrange(desc(n_rows)) |> head(20))

  transcript_coded <- bind_rows(
    transcript_coded |> filter(!is.na(C_course_subject_area) |
                                        !D_course_id %in% filled$D_course_id),
    filled) |>
    distinct(D_stu_id, D_year, D_course_id, .keep_all = TRUE)

  chk_rows_stable(transcript_student, transcript_coded,
                  "Fallback did not change the transcript row count")
}
```

```{r cd-coverage}
transcript_coded |>
  count(C_coding_source, name = "rows") |>
  mutate(pct = round(100 * rows / sum(rows), 3)) |>
  arrange(desc(rows))
```

```{r cd-uncoded}
uncoded <- transcript_coded |>
  filter(is.na(C_course_subject_area))

chk_info("Rows still uncoded", nrow(uncoded),
         data = uncoded |> count(D_course_id, D_course_name, name = "n_rows") |>
           arrange(desc(n_rows)))
```

## 7.4 What the coding gives us

The first look at the transcript through ERS categories rather than district
course names.

```{r cd-subject-area}
transcript_coded |>
  filter(!is.na(C_course_subject_area)) |>
  group_by(C_course_subject_area) |>
  summarise(rows = n(),
                   students = n_distinct(D_stu_id),
                   creds_earned = round(sum(D_creds_earn, na.rm = TRUE), 0),
                   .groups = "drop") |>
  arrange(desc(rows))
```

```{r cd-rigor}
transcript_coded |>
  filter(!is.na(C_course_rigor)) |>
  count(C_course_rigor, C_course_rigor_detail, name = "rows") |>
  arrange(desc(rows))
```

The EPSO, work-based learning and industry-credential flags are the fields the
Arnold hypotheses turn on, so it is worth seeing their scale now.

```{r cd-epso}
transcript_coded |>
  filter(!is.na(C_course_subject_area)) |>
  group_by(D_year) |>
  summarise(
    rows          = n(),
    pct_epso      = round(100 * mean(C_course_epso_flag, na.rm = TRUE), 2),
    pct_wbl       = round(100 * mean(C_course_wbl_flag, na.rm = TRUE), 2),
    pct_ic        = round(100 * mean(C_course_ic_flag, na.rm = TRUE), 2),
    .groups = "drop")
```

```{r cd-epso-type}
transcript_coded |>
  filter(!is.na(C_course_epso_type)) |>
  count(C_course_epso_type, name = "rows") |>
  arrange(desc(rows))
```

## 7.5 Coverage where it matters

Overall coverage can look excellent while a specific population is poorly
covered. The cohort is the population every headline number will be built on.

```{r cd-coverage-cohort}
transcript_coded |>
  mutate(in_cohort = D_stu_id %in% cohort$D_stu_id) |>
  group_by(in_cohort) |>
  summarise(rows = n(),
                   pct_coded = round(100 * mean(!is.na(C_course_subject_area)), 3),
                   .groups = "drop")
```

```{r cd-coverage-year}
transcript_coded |>
  group_by(D_year) |>
  summarise(rows = n(),
                   pct_coded = round(100 * mean(!is.na(C_course_subject_area)), 3),
                   .groups = "drop")
```

## 7.6 Trace students

```{r cd-trace}
transcript_coded |>
  filter(D_stu_id %in% trace_ids) |>
  select(D_stu_id, D_year, D_course_id, D_course_name, C_term,
                C_course_subject_area, C_course_subject, C_course_rigor,
                D_creds_earn, C_coding_source) |>
  arrange(D_stu_id, D_year, D_course_id) |>
  head(40)
```

## 7.7 Gates

```{r cd-gates}
pct_coded <- 100 * mean(!is.na(transcript_coded$C_course_subject_area))

chk_stop("Course coding match rate",
         pct_coded >= config$thresholds$min_course_coding_match_pct,
         value  = paste0(round(pct_coded, 3), "%"),
         detail = paste0("threshold is ",
                         config$thresholds$min_course_coding_match_pct, "%"),
         offenders = uncoded |>
           count(D_course_id, D_course_name, name = "n_rows"))

chk_unique(transcript_coded, c("D_stu_id", "D_year", "D_course_id"),
           "Coded transcript is still one row per student / year / course")

chk_rows_stable(transcript_student, transcript_coded,
                "Stage 7 preserved every transcript row")
```

---

# Stage 8 — Output and check report

One table carries everything: the coded transcript at student x year x course,
with the analysis fields and both selection flags attached. The snapshot export
is a filtered copy of that same table rather than a separately built object, so
the two cannot drift apart.

## 8.1 Analysis fields

The legacy `04_transcript_analysis` recomputed these on every run. Computing
them once, here, with checks, means every analysis starts from the same
definitions.

```{r out-analysis-fields}
chk_stage("Stage 8 — Output")

bucket_grade <- function(x, buckets) {
  out <- rep(NA_character_, length(x))
  for (b in buckets) {
    hit <- is.na(out) & !is.na(x) & x >= b$min
    out[hit] <- b$label
  }
  out
}

transcript_out <- transcript_coded |>
  mutate(
    C_course_failed    = !is.na(D_creds_possible) & D_creds_possible > 0 &
                          !is.na(D_creds_earn) & D_creds_earn == 0,
    C_grade_pct_rollup = bucket_grade(D_stu_grade,
                                      config$analysis_fields$grade_pct_buckets),
    C_course_marked_repeat = C_course_name_starred
  )

chk_info("Failed course rows",
         paste0(format(sum(transcript_out$C_course_failed), big.mark = ","), " (",
                round(100 * mean(transcript_out$C_course_failed), 1), "%)"))

transcript_out |>
  count(C_grade_pct_rollup, name = "rows") |>
  arrange(desc(rows))
```

### Credit recovery

A later-term row earning more credit than it attempted, where the same
student's earlier-term row for the same base course earned zero. No hardcoded
passing mark.

```{r out-credit-recovery}
earlier_failed <- transcript_out |>
  filter(C_term == "Semester 1") |>
  distinct(D_stu_id, D_year, C_course_id_base, .keep_all = TRUE) |>
  transmute(D_stu_id, D_year, C_course_id_base,
                   s1_failed = !is.na(D_creds_earn) & D_creds_earn == 0)

transcript_out <- transcript_out |>
  left_join(earlier_failed,
                   by = c("D_stu_id", "D_year", "C_course_id_base")) |>
  mutate(
    C_credit_recovery_flag = coalesce(
      C_creds_earn_exceeds_attempted & C_term == "Semester 2" & s1_failed, FALSE)) |>
  select(-s1_failed)

chk_info("Credit recovery rows",
         paste0(format(sum(transcript_out$C_credit_recovery_flag), big.mark = ","),
                " across ",
                format(n_distinct(
                  transcript_out$D_stu_id[transcript_out$C_credit_recovery_flag]),
                  big.mark = ","), " students"))

transcript_out |>
  filter(C_credit_recovery_flag) |>
  count(D_year, name = "rows")
```

### Mobility

`04` built `C_pure_cohort` from two transfer flags. Both are derivable from
what Stage 5 already computed.

```{r out-mobility}
mobility <- student_year |>
  arrange(D_stu_id, D_year) |>
  group_by(D_stu_id) |>
  mutate(prev_school = lag(C_home_school_id)) |>
  summarise(
    C_transferred_within_year  = any(n_schools > 1),
    C_transferred_across_years = any(!is.na(prev_school) &
                                       prev_school != C_home_school_id),
    .groups = "drop") |>
  mutate(C_pure_cohort = !C_transferred_within_year &
                                !C_transferred_across_years)

chk_info("Students who never changed school", 
         paste0(format(sum(mobility$C_pure_cohort), big.mark = ","), " (",
                round(100 * mean(mobility$C_pure_cohort), 1), "%)"))

transcript_out <- transcript_out |>
  join_checked(mobility, by = "D_stu_id", check = "Mobility flags onto transcript")
```

## 8.2 Selection flags

```{r out-selection-flags}
# C_stu_cohort and C_stu_snapshot use Tettra's field names and its
# include/exclude values, so anyone working from the guidance finds what they
# expect. The logical twins exist because "include" does not sum.
transcript_out <- transcript_out |>
  mutate(
    C_cohort_include = D_stu_id %in% cohort$D_stu_id,
    C_stu_cohort     = if_else(C_cohort_include, "include", "exclude"),
    C_eoy_grade      = C_stu_grade_level)

chk_info("Transcript rows belonging to cohort students",
         paste0(format(sum(transcript_out$C_cohort_include), big.mark = ","), " (",
                round(100 * mean(transcript_out$C_cohort_include), 1), "%)"))
```

## 8.3 Snapshot selection

Independent of the cohort: a single-year population, selected on its own
criteria. Set `snapshot$enabled: false` for a cohort-only project.

```{r out-snapshot}
if (isTRUE(config$snapshot$enabled)) {

  snap_year <- config$snapshot$year
  snap_base <- student_year |> filter(D_year == snap_year)

  chk_info(paste0("Snapshot ", snap_year, " — students enrolled"), nrow(snap_base))

  snap_funnel <- funnel_new(snap_base$D_stu_id,
                            paste0("Enrolled in ", snap_year))

  snap_funnel <- funnel_add(snap_funnel, "Grade level in range",
    snap_base$D_stu_id[snap_base$C_stu_grade_level %in% config$snapshot$grade_levels])

  if (isTRUE(config$snapshot$require_school_in_scope)) {
    snap_funnel <- funnel_add(snap_funnel, "Home school in scope",
      snap_base$D_stu_id[snap_base$C_home_school_in_scope])
  }

  # The legacy script summed days without group_by, so every student received
  # the same total and the flag was uniform. Grouping is what makes this a
  # per-student test.
  if (config$snapshot$duration_basis == "credits_attempted") {
    creds <- transcript_out |>
      filter(D_year == snap_year) |>
      group_by(D_stu_id) |>
      summarise(creds_attempted = sum(D_creds_possible, na.rm = TRUE),
                       .groups = "drop")
    snap_funnel <- funnel_add(snap_funnel,
      paste0("Attempted >= ", config$snapshot$credits_attempted_min, " credits"),
      creds$D_stu_id[creds$creds_attempted >= config$snapshot$credits_attempted_min])
  } else {
    thr <- if (config$snapshot$duration_basis == "share")
      config$snapshot$duration_min_share * window_days else config$snapshot$duration_min_days
    snap_funnel <- funnel_add(snap_funnel,
      paste0("Enrolled >= ", round(thr), " days"),
      snap_base$D_stu_id[snap_base$M_days_enrolled_year >= thr])
  }

  snapshot_ids <- snap_funnel$ids
  chk_info("Snapshot selected", length(snapshot_ids),
           detail = paste0(round(100 * length(snapshot_ids) / snap_funnel$start, 1),
                           "% of students enrolled in ", snap_year))
} else {
  snapshot_ids <- character(0)
  chk_info("Snapshot", "disabled in config")
}
```

```{r out-snapshot-funnel}
if (isTRUE(config$snapshot$enabled)) {
  funnel_report(snap_funnel, config$cohort$benchmarks$snapshot_retention_pct)
}
```

```{r out-snapshot-flag}
transcript_out <- transcript_out |>
  mutate(
    C_stu_snapshot_include = D_stu_id %in% snapshot_ids &
                             D_year == config$snapshot$year,
    C_stu_snapshot = if_else(C_stu_snapshot_include, "include", "exclude"))

chk_info("Transcript rows in the snapshot",
         format(sum(transcript_out$C_stu_snapshot_include), big.mark = ","))
```

## 8.4 Final table

Columns ordered by prefix so the table reads consistently across districts:
`D_` district-reported, `C_` ERS-created, `M_` measures.

```{r out-order}
transcript_out <- transcript_out |>
  select(starts_with("D_"),
                starts_with("C_"),
                starts_with("M_"),
                everything())

chk_unique(transcript_out, c("D_stu_id", "D_year", "D_course_id"),
           "Output is one row per student / year / course")

chk_rows_stable(transcript_coded, transcript_out,
                "Stage 8 added columns without changing rows")

chk_info("Output shape",
         paste0(format(nrow(transcript_out), big.mark = ","), " rows x ",
                ncol(transcript_out), " cols"))

tibble(column = names(transcript_out),
               type   = map_chr(transcript_out, ~ class(.x)[1]),
               pct_missing = map_dbl(transcript_out,
                                            ~ round(100 * mean(is.na(.x)), 1)))
```

## 8.5 Scope the export

Everything above ran on the full transcript, because the diagnostics are only
meaningful against the whole file. What leaves the pipeline is a different
question, and config decides it.

Filtering to the cohort does more than shrink the file. A cohort student is
enrolled in every required year by definition, so the orphan rows — coursework
transferred in from before a student enrolled — disappear along with their
missing home school and missing demographics. Every row in the export is fully
populated.

```{r out-scope}
rows_rule <- config$outputs$coded_transcript$rows

keep <- switch(rows_rule,
  all                = rep(TRUE, nrow(transcript_out)),
  cohort_only        = transcript_out$C_cohort_include,
  cohort_or_snapshot = transcript_out$C_cohort_include |
                       transcript_out$C_stu_snapshot_include,
  stop("Unknown outputs$coded_transcript$rows: ", rows_rule, call. = FALSE))

if (isTRUE(config$outputs$coded_transcript$restrict_to_cohort_years)) {
  keep <- keep & transcript_out$D_year %in% cohort_years
}

transcript_export <- transcript_out |> filter(keep)

chk_info("Export row rule", rows_rule)
chk_info("Rows exported",
         paste0(format(nrow(transcript_export), big.mark = ","), " of ",
                format(nrow(transcript_out), big.mark = ","), " (",
                round(100 * nrow(transcript_export) / nrow(transcript_out), 1), "%)"))
chk_info("Students exported", n_distinct(transcript_export$D_stu_id))
```

The export should be complete on every field that was patchy in the full file.
If it is not, a student got into the cohort who should not have.

```{r out-export-completeness}
tibble(
  measure = c("Rows", "Students", "Years",
              "% with a home school", "% coded", "% with an enrolment record"),
  value = c(
    format(nrow(transcript_export), big.mark = ","),
    format(n_distinct(transcript_export$D_stu_id), big.mark = ","),
    paste(sort(unique(transcript_export$D_year)), collapse = ", "),
    paste0(round(100 * mean(!is.na(transcript_export$C_home_school_id)), 2), "%"),
    paste0(round(100 * mean(!is.na(transcript_export$C_course_subject_area)), 2), "%"),
    paste0(round(100 * mean(transcript_export$C_has_enrolment_record), 2), "%")))
```

```{r out-export-by-year}
transcript_export |>
  group_by(D_year) |>
  summarise(rows = n(),
                   students = n_distinct(D_stu_id),
                   median_courses = round(n() / n_distinct(D_stu_id), 1),
                   creds_earned = round(sum(D_creds_earn, na.rm = TRUE), 0),
                   .groups = "drop")
```

```{r out-export-gates}
chk_stop("Export is not empty", nrow(transcript_export) > 0)

if (rows_rule == "cohort_only") {
  chk_stop("Every exported student is in the cohort",
           all(transcript_export$D_stu_id %in% cohort$D_stu_id))

  # Completeness only follows when the export is confined to the cohort years.
  # A district supplying more years than the cohort spans can legitimately give
  # a cohort student pre-enrolment coursework, and that is not a failure.
  n_no_school <- sum(is.na(transcript_export$C_home_school_id))
  if (isTRUE(config$outputs$coded_transcript$restrict_to_cohort_years) ||
      all(unique(transcript_export$D_year) %in% cohort_years)) {
    chk_stop("Every exported row has a home school", n_no_school == 0,
             detail = "within the cohort years a cohort student is enrolled every year",
             offenders = transcript_export |>
               filter(is.na(C_home_school_id)) |>
               count(D_stu_id, D_year, name = "n_rows"))
  } else {
    chk_info("Exported rows outside the cohort years with no home school",
             n_no_school,
             detail = "pre-enrolment coursework for cohort students; expected")
  }
}

chk_unique(transcript_export, c("D_stu_id", "D_year", "D_course_id"),
           "Export is one row per student / year / course")
```

## 8.6 The student-year export

Tettra keeps every student and distinguishes with a flag, and asks that the
cohort be tested for comprehensiveness against the full 9th-grade population.
A cohort-only transcript cannot answer that, so the denominator lives here
instead — every student-year, with every criterion outcome and the reason each
excluded student dropped out.

Narrow and short enough that keeping everyone costs nothing.

```{r out-student-year}
student_year_export <- student_year |>
  left_join(cohort_criteria, by = "D_stu_id") |>
  left_join(mobility, by = "D_stu_id") |>
  left_join(cohort |> select(D_stu_id, C_cohort_school_id,
                                           C_cohort_school_name),
                   by = "D_stu_id") |>
  mutate(
    C_stu_cohort   = coalesce(C_stu_cohort, "exclude"),
    C_stu_snapshot = if_else(D_stu_id %in% snapshot_ids &
                                      D_year == config$snapshot$year,
                                    "include", "exclude"),
    C_entry_cohort_candidate = D_stu_id %in% entrants$D_stu_id)

chk_unique(student_year_export, c("D_stu_id", "D_year"),
           "Student-year export is one row per student per year")

chk_info("Student-year export",
         paste0(format(nrow(student_year_export), big.mark = ","), " rows x ",
                ncol(student_year_export), " cols"))

student_year_export |>
  filter(C_entry_cohort_candidate, D_year == config$cohort$entry_year) |>
  count(C_stu_cohort, C_cohort_exclusion_reason, name = "students") |>
  arrange(desc(students))
```

## 8.7 Write

The snapshot is filtered from the same full table rather than from the export,
so a snapshot student who is not in the cohort is still written out.
## 8.7 Write to SharePoint

Nothing is written locally. `put()` reports each file into the check log, so
the run report records exactly what left the pipeline and how big it was.

```{r out-write}
out_folder <- config$raw_data_folder_path
out_prefix <- config$outputs$subfolder %||% "/"

put <- function(data, name) {
  if (is.null(data) || nrow(data) == 0) {
    chk_info(paste0("Output SKIPPED (empty): ", name), 0)
    return(invisible(NULL))
  }
  path <- paste0(out_prefix, name)
  ers_write_sharepoint(data                     = data,
                       folder_path              = out_folder,
                       file_name_with_extension = path)
  chk_info(paste0("Output written: ", name),
           paste0(format(nrow(data), big.mark = ","), " rows x ",
                  ncol(data), " cols"),
           detail = paste0(out_folder, path))
  invisible(data)
}

if (isTRUE(config$outputs$coded_transcript$write)) {
  put(transcript_export, config$outputs$coded_transcript$file_name)
}

if (isTRUE(config$outputs$student_year$write)) {
  put(student_year_export, config$outputs$student_year$file_name)
}

if (isTRUE(config$snapshot$enabled) && isTRUE(config$outputs$snapshot$write)) {
  put(transcript_out |> dplyr::filter(C_stu_snapshot_include),
      config$outputs$snapshot$file_name)
}
```

## 8.8 Check report

```{r final-report}
check_log <- chk_report(print_table = FALSE)
```

```{r final-table}
check_log |>
  select(seq, stage, check, severity, result, value)
```