---
title: "3_student_cohort_selection"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: August 12 2024
review:
last update: April 14 2025
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Start fresh
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
gc()
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in the student and transcript data that you will need to select your snapshot of students.

-   **Project Team Action:** Replace folder path with the folder path where your files are and replace specific files names as necessary.

```{r, warning=F, message=F}

raw_data_folder_path <- "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"


stu_transcript_rollup <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "stu_transcript_rollup_cohort_only_updated.csv",
  drive_name = "client_work_drive")


```

# Part II: Choosing Cohort of Students- MNPS Updated

-   **What:** The following code helps you establish variables you will need to identify a "cohort" of students, or 4 year group of students started as freshman in one school year and graduating 4 years later. Your team may have specific constraints within this methodology (ie. students need to progress or "persist" traditionally, matriculating from 9th to 10th, etc.), but the following code is meant to apply to all teams. You will have to change the years based on which specific cohort or cohorts of students you want to select. NOTE- 2025 MNPS team took a slighly different methodoloical approach, so these checks will be heavily editted to what you might experience on a traditional transcript analysis project

-   **Project Team Action:** Update years as necessary for your specific analysis cohort

###2.1 Create Student Cohort Data Frame

####a. Student counts before filtering

-   **What:** If you have not already, you should read through the Tettra pages outlining student cohort slection here: <https://app.tettra.co/teams/ersknowledge/pages/wip-deep-dive-cohort-analysis>. The following code contains validation checks before filtering, including number of students, number of high schoolers, and number of students by location for included schools

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Updated based on year/specific snapshot group of students you want to select. Gut check/track validation results with team.

```{r}
# #Starting with your 9th grade students (this example uses 2021), check how many total students will "start" the cohort
# stu_transcript_rollup %>% 
#   filter(D_year == 2022 & C_eoy_grade == "9") %>% 
#   distinct() 
# 
# #count the number of students by specific locations
# validated_student_data %>% 
#   group_by(D_location_name) %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   arrange(desc(count_students)) %>% 
#   ungroup()
# 
# #count the number of 12th graders in 2024 (who will be your cohort graduates)
# validated_student_data %>% 
#   filter(D_year == 2024 & D_stu_grade == "12") %>% 
#   distinct() 

```

#### b. Check student enter and exit dates

-   **What:** Convert enter and exit dates to be as dates (if they aren't already), and check for any students with more than one row

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Gut check/track

```{r}
# # convert enter and exit date to be as dates
# validated_student_data <- validated_student_data %>%
#   mutate(D_stu_enter_date = as.Date(D_stu_enter_date),
#          D_stu_exit_date = as.Date(D_stu_exit_date))
# 
# #check student enter dates
# validated_student_data %>% 
#   group_by(D_stu_enter_date) %>% 
#   count() %>% 
#   arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
#   ungroup()
# 
# # Check student exit dates
# validated_student_data %>% 
#   group_by(D_stu_exit_date) %>% 
#   count() %>% 
#   arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
#   ungroup()

```

#### c. Creating Exclusion Critera, Location, Duration, Persistence

-   **What:** There are three major exclusion criteria and filters for transcript analysis: Location (the school the student attended), Duration (how long they were in the school per year), and Persistence (If they matriculated traditionally from one grade to the next.) These steps may vary based on specific team methodology and needs. For some districts, ERS has used a 75% criteria, where students need to be present 75% of the year (in each school year). In other districts, that has been lowered to 10%. Other districts have ignored entry and exit dates all together and used credits attained as the determination for selecting students. You should outline with your team the options for selecting a cohort, and move forward with the best option based on your specific analytic questions. The code below shows examples of selecting students based on some of these methodologies.

####i. Location (Home School)

-   **What:** The first exclusion criteria for cohort selection is home school- asigning a student one location ID based on the location ID they spent the most time at. This is done by creating a new column that calculates the duration of stay at each school, and then filtering to the maximum duration of stay for each student. Then, you can easily filter to schools of interest for your analysis. Similar to snapshot, the cohort home school assignment is based on whatever school a student spent the majority of their time enrolled in the district at (so number of years when looking specifically at cohort students.)

-   **Check:** Make sure no student has more than 1 home school, and that you have cleared and checked this methodology with your team.

```{r}
# # Step-by-step version with safe_left_join at the end
# student_home_school <- validated_student_data %>%
#   group_by(D_stu_id) %>% 
#   # Step 1: Calculate number of days per row (entry-exit pair)
#   mutate(H_stay_duration = as.numeric(D_stu_exit_date - D_stu_enter_date)) %>%
# 
#   # Step 2: Summarize total stay duration per student-location combo
#   group_by(D_stu_id, D_location_id) %>%
#   summarize(H_total_days_at_location = sum(H_stay_duration, na.rm = TRUE), .groups = "drop") %>%
# 
#   # Step 3: For each student, find the location with the most days (break ties randomly)
#   group_by(D_stu_id) %>%
#   slice_max(order_by = H_total_days_at_location, n = 1, with_ties = TRUE) %>%  # keeps all ties
#   slice_sample(n = 1) %>%  # randomly pick one if tied
#   ungroup() %>%
# 
#   # Step 4: Rename the chosen location as the "home school"
#   rename(H_home_school = D_location_id) %>%
#   select(D_stu_id, H_home_school)
# 
# # Step 5: Use safe_left_join to merge back onto the original dataset
# validated_student_data <- safe_left_join(validated_student_data, student_home_school, by = "D_stu_id")
# 
# # Step 6: Add back in home school name
# validated_student_data <- validated_student_data %>%
#   mutate(C_home_school_name = case_when(
#     H_home_school == "1" ~ "Waikea High School",
#     H_home_school == "2" ~ "Hilo High School",
#     H_home_school == "3" ~ "Konaeawena HS",
#     H_home_school == "4" ~ "Kamehameha High",
#     H_home_school == "5" ~ "Hawaii Preparatory Academy",
#     TRUE ~ "Other"
#   ))
# 
# # count total students in the cohort
# validated_student_data %>% 
#   summarize(count_students = n_distinct(D_stu_id))
# 
# #Get student counts by grade
# validated_student_data %>% 
#   group_by(C_eoy_grade) %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   ungroup()
# 
# #count total students by home school
# validated_student_data %>% 
#   group_by(C_home_school_name) %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   ungroup()
# 
# #If necessary, filter to only home schools of interest/filter out excluded locations (skip this if you already have a filtered data set)
# # validated_student_data <- validated_student_data %>% 
# #   filter(C_home_school_code == "1"| C_home_school_code == "29"| C_home_school_code == "39")


```

#### ii. Duration

-   **What:** Filtering to students who spent a certain amount of time at their home school in each year. This can be done in a number of ways, check with your team on the appropriate methodology. The example below shows a 75% threshold, but this can be adjusted based on your needs.

-   **NOTE:** If your team is using a credit attempt cutoff (see Tettra for examples and guidance), skip this step and see option step in part 3.1 C.

```{r}
#First, lets double-check that no student has more than one value of M_days_enrolled
check_days_enrolled <- stu_transcript_rollup %>%
  group_by(D_stu_id, D_year) %>%
  summarise(n_unique_days = n_distinct(M_days_enrolled), .groups = "drop") %>%
  filter(n_unique_days > 1) %>% 
  print()

#Great, as we would expect

#and double check the number of distinct students in the cohort
stu_transcript_rollup %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()


schools_to_include <- c(
  "Hunters Lane High",
  "McGavock High",
  "John Overton High",
  "Stratford",
  "Cane Ridge High School",
  "Whites Creek High School",
  "Glencliff High School",
  "Antioch High School",
  "Hillsboro High",
  "PearlCohn High",
  "Maplewood High",
  "James Lawson High",
  "James Lawson High School",
  "Hillwood High School"
)


# Step 1: Calculate total stay duration per student per year
student_duration_flags <- stu_transcript_rollup %>%
  group_by(D_stu_id, D_year) %>%
  summarize(
    H_sum_stay_duration_year = max(M_days_enrolled, na.rm = TRUE),  # instead of sum()
    .groups = "drop"
  ) %>%
  mutate(
    duration_flag = if_else(H_sum_stay_duration_year >= 200, "include", "exclude")
  ) %>%
  select(-H_sum_stay_duration_year)

# Step 2: Pivot to create one column per year
student_flags_wide <- student_duration_flags %>%
  pivot_wider(
    names_from = D_year,
    values_from = duration_flag,
    names_prefix = "duration_"
  )

# Step 3: Safe join back to original data
stu_transcript_rollup <- safe_left_join(stu_transcript_rollup, student_flags_wide, by = "D_stu_id")


#count of students by include/exclude flag
student_flags_wide %>%
  pivot_longer(
    cols = starts_with("duration_"),
    names_to = "year",
    names_prefix = "duration_",
    values_to = "flag"
  ) %>%
  count(year, flag) %>%
  arrange(year, flag)

#make sure no student has both include and exclude flag
stu_transcript_rollup %>% 
  group_by(D_stu_id) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  filter(count_students > 1) 

#create a summary column that looks at the duration flags for each year, and if a student is present for each of the years you care about for your analysis (ie. 2021-2024 for this example), marks them as "include" for a new column called "H_cohort_duration_include"
stu_transcript_rollup <- stu_transcript_rollup %>%
  mutate(
    H_cohort_duration_include = case_when(
duration_2022 == "include" & duration_2023 == "include" & duration_2024 == "include" & duration_2025 == "include" ~ "include",
      TRUE ~ "exclude"
    )
  )

#count the number of distinct students who meet this criteria
stu_transcript_rollup %>% 
  filter(H_cohort_duration_include == "include") %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()


#grouping by D_creds_earned, what is the Median number of credits earned by students with the include flag as opposed to students with the exclude flag?
stu_transcript_rollup %>%
  group_by(D_stu_id, H_cohort_duration_include) %>%
  summarize(total_credits = sum(D_creds_earn, na.rm = TRUE), .groups = "drop") %>%
  group_by(H_cohort_duration_include) %>%
  summarize(
    median_credits_earned = median(total_credits, na.rm = TRUE),
    mean_credits_earned   = mean(total_credits, na.rm = TRUE),
    count_students        = n_distinct(D_stu_id),
    .groups = "drop"
  ) %>%
  print()

#can we look at an example of some students who were present for 10 days or fewer in the same year
stu_transcript_rollup %>%
  filter(H_cohort_duration_include == "exclude") %>%
  select(D_stu_id, D_year, M_days_enrolled, H_cohort_duration_include) %>%
  arrange(D_stu_id, D_year) %>%
  distinct() %>% 
  print()


# filter to D_stu_id 190173946
student_check_1 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190173946) %>%
  print()

#2
student_check_2 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190162750) %>%
  print()

#3
student_check_3 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190181214) %>%
  print()

#Rebels Student (2), my student 4
student_check_4 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190137047) %>%
  print()


#Can we put student checks 1, 2, and 3, and now 4 together in one data-frame?
student_checks <- bind_rows(student_check_1, student_check_2, student_check_3, student_check_4) %>%
  arrange(D_stu_id, D_year) %>%
  print()

# #export for team to look over
# ers_write_sharepoint(data = student_checks,
#                      folder_path = raw_data_folder_path,
#                      file_name_with_extension = "student_checks_duration_exclude.csv")


#Are all students with exclude flags for duration also students who had any C_year_home_school enrollments at an exclude high school?
# Tag each school as included/excluded
stu_transcript_rollup <- stu_transcript_rollup %>%
  mutate(
    school_included = if_else(C_year_home_school %in% schools_to_include, "included_school", "excluded_school")
  )

# For students excluded by duration flag, see if they ever attended an excluded school
exclude_school_check <- stu_transcript_rollup %>%
  filter(H_cohort_duration_include == "exclude") %>%
  group_by(D_stu_id) %>%
  summarize(
    ever_excluded_school = any(school_included == "excluded_school"),
    ever_included_school = any(school_included == "included_school"),
    .groups = "drop"
  )

# Count how many excluded students fall into each category
exclude_school_check %>%
  count(ever_excluded_school)

#quick check of semester, because I hadn't thought to check that- How many of these excluded students have only "semester 1" enrollments?
# Step 1: year-level check for excluded students only
semester_check_excluded_year <- stu_transcript_rollup %>%
  filter(H_cohort_duration_include == "exclude") %>%
  group_by(D_stu_id, D_year) %>%
  summarize(
    only_semester_1 = all(D_term == "Semester 1"),
    only_semester_2 = all(D_term == "Semester 2"),
    both_semesters  = n_distinct(D_term) == 2,
    .groups = "drop"
  )

# Step 2: collapse to student-level with MECE logic
semester_check_excluded_student <- semester_check_excluded_year %>%
  group_by(D_stu_id) %>%
  summarize(
    ever_only_semester = any(only_semester_1 | only_semester_2),
    ever_both          = any(both_semesters),
    .groups = "drop"
  ) %>%
  mutate(
    category = case_when(
      ever_only_semester ~ "only_one_semester",
      ever_both ~ "both_semesters",
      TRUE ~ "neither"
    )
  )

# Step 3: count excluded students by category
semester_summary_excluded <- semester_check_excluded_student %>%
  count(category)

print(semester_summary_excluded)

#double-check for all students
semester_check_by_year <- stu_transcript_rollup %>%
  group_by(D_stu_id, D_year) %>%
  summarize(
    only_semester_1 = all(D_term == "Semester 1"),
    only_semester_2 = all(D_term == "Semester 2"),
    both_semesters  = n_distinct(D_term) == 2,
    .groups = "drop"
  )

# Collapse to student-level with MECE logic
semester_check_by_student <- semester_check_by_year %>%
  group_by(D_stu_id) %>%
  summarize(
    ever_only_semester = any(only_semester_1 | only_semester_2),
    ever_both          = any(both_semesters),
    .groups = "drop"
  ) %>%
  mutate(
    category = case_when(
      ever_only_semester ~ "only_one_semester",
      ever_both ~ "both_semesters",
      TRUE ~ "neither"   # fallback, in case of strange data
    )
  )

# Count students by final category
semester_summary_all_years <- semester_check_by_student %>%
  count(category)

print(semester_summary_all_years)


#Great, and lastly does any student who would otherwise be include for the 200 days threhsold only have one semester of enrollment?
# Student-level categorization
semester_check_students <- stu_transcript_rollup %>%
  filter(H_cohort_duration_include == "include") %>%
  group_by(D_stu_id) %>%
  summarize(
    only_semester_1 = all(D_term == "Semester 1"),
    only_semester_2 = all(D_term == "Semester 2"),
    both_semesters  = n_distinct(D_term) == 2,
    .groups = "drop"
  ) %>%
  mutate(
    category = case_when(
      only_semester_1 ~ "only_semester_1",
      only_semester_2 ~ "only_semester_2",
      both_semesters  ~ "both_semesters",
      TRUE            ~ "neither"
    )
  )

# Summary counts by category
semester_check_summary <- semester_check_students %>%
  group_by(category) %>%
  summarise(
    distinct_students = n_distinct(D_stu_id),
    .groups = "drop"
  )

# If you want to inspect the student IDs in "neither"
semester_check_students %>%
  filter(category == "neither") %>%
  select(D_stu_id)


#student_check_1
student_check_1 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190180653) %>%
  print()

student_check_2 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190180786) %>%
  print()

student_check_3 <- stu_transcript_rollup %>%
  filter(D_stu_id == 190183015) %>%
  print()

#Got it, it looks like in each case there are rows of enrollment in Spanish that for some reason aren't carrying a grade, and also aren't displaying the semester enrollment right for those 7 students, but otherwise they look good. I can follow up on this after full cohort selection!

```

####iii. MNPS check- Duration buckets- Don't need this anymore
```{r}
# #Lets create a bucket system to look at M_Days_enrolled to determine the number of distinct students by each bucket
# 
# #NOTE THAT THIS DOUBLE-COUNTS, I"M JUST TRYING TO GET AN IDEA OF THE DISTRIBUTION
# # Create enrollment day buckets and count unique students
# Days_enrolled_buckets <- stu_transcript_rollup %>%
#   mutate(H_stay_duration_bucket = case_when(
#     M_days_enrolled >= 0 & M_days_enrolled <= 50   ~ "0-50",
#     M_days_enrolled > 50 & M_days_enrolled <= 100  ~ "51-100",
#     M_days_enrolled > 100 & M_days_enrolled <= 150 ~ "101-150",
#     M_days_enrolled > 150 & M_days_enrolled <= 200 ~ "151-200",
#     M_days_enrolled > 200 & M_days_enrolled <= 250 ~ "201-250",
#     M_days_enrolled > 250 & M_days_enrolled <= 300 ~ "251-300",
#     M_days_enrolled > 300                            ~ "300+",
#     TRUE ~ "Unknown"
#   )) %>%
#   group_by(H_stay_duration_bucket) %>%
#   summarize(n_students = n_distinct(D_stu_id), .groups = "drop") %>%
#   arrange(H_stay_duration_bucket)
# 
# # View the table
# print(Days_enrolled_buckets)
# 
# 
# #okay, so most are 300+ which is great and want we want. Second highest is less than 50 though, so what if we included just more than 50 days? how many would we lose?
# 


```


### d. Indivudal Student Checks

```{r}
#check a few student ids
# ---- Function: Check Random Students ----
check_random_students <- function(data, student_id_col = "D_stu_id", n = 5, seed = NULL) {
  # Optional seed for reproducibility
  if (!is.null(seed)) set.seed(seed)
  
  # Sample n unique student IDs
  random_ids <- sample(unique(data[[student_id_col]]), n)

  # Filter and return rows for those students, grouped for easy viewing
  data %>%
    filter(.data[[student_id_col]] %in% random_ids) %>%
    arrange(.data[[student_id_col]]) %>%
    group_by(.data[[student_id_col]])
}


# Check 5 random students
check_random_students(stu_transcript_rollup, n = 5, seed = 123)

#These look normal to me!

```

### e. Persistence

-   **What:** The persistence criteria is to check that students are progressing traditionally from freshman to sophomore to junior to senior year across 4 years, because in the cohort analysis we are hoping to understand the patterns of "traditional" students. Some teams have modified this based on their needs, such as making the criteria just include starting 9th grade and finishing 12th grade within 4 years (meaning repeating a grade as long as you caught up was not a problem). The methodology below shows the traditional progression criteria, but this can be modified based on your needs.

- **MNPS Update:** MNPS is going to test both traditional 9, 10, 11, 12, and also just 9th in 2021/2022 and 12th in 2024/2025 to see implications

```{r}
#First confirm number of students in the cohort
stu_transcript_rollup %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

#2756

# Step 1: Create year-level flags at the student level
stu_transcript_rollup <- stu_transcript_rollup %>%
  group_by(D_stu_id) %>%
  mutate(
    H_stu_progression_traditional_2022 = if_else(any(C_eoy_grade == "9" & D_year == 2022), "include", "exclude"),
    H_stu_progression_traditional_2023 = if_else(any(C_eoy_grade == "10" & D_year == 2023), "include", "exclude"),
    H_stu_progression_traditional_2024 = if_else(any(C_eoy_grade == "11" & D_year == 2024), "include", "exclude"),
    H_stu_progression_traditional_2025 = if_else(any(C_eoy_grade == "12" & D_year == 2025), "include", "exclude")
  ) %>%
  ungroup()

# Step 2: Create an overall progression flag at the student level
stu_transcript_rollup <- stu_transcript_rollup %>%
  group_by(D_stu_id) %>%
  mutate(
    H_cohort_progression_include = case_when(
      all(H_stu_progression_traditional_2022 == "include",
          H_stu_progression_traditional_2023 == "include",
          H_stu_progression_traditional_2024 == "include",
          H_stu_progression_traditional_2025 == "include") ~ "include",
      TRUE ~ "exclude"
    )
  ) %>%
  ungroup()

# Step 3: Count included students
stu_transcript_rollup %>%
  filter(H_cohort_progression_include == "include") %>%
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

#2727

  # Summary table: counts of included/excluded students by year
# 2x4 table: rows = include/exclude, cols = 2022–2025, values = distinct students
progression_summary <- stu_transcript_rollup %>%
  select(
    D_stu_id,
    H_stu_progression_traditional_2022,
    H_stu_progression_traditional_2023,
    H_stu_progression_traditional_2024,
    H_stu_progression_traditional_2025
  ) %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%             # one row per student
  tidyr::pivot_longer(
    cols = starts_with("H_stu_progression_traditional_"),
    names_to = "year",
    names_prefix = "H_stu_progression_traditional_",
    values_to = "flag"
  ) %>%
  dplyr::group_by(year, flag) %>%
  dplyr::summarise(student_count = dplyr::n_distinct(D_stu_id), .groups = "drop") %>%
  dplyr::mutate(year = as.integer(year)) %>%
  tidyr::pivot_wider(
    names_from = year,
    values_from = student_count,
    values_fill = 0
  ) %>%
  dplyr::select(flag, `2022`, `2023`, `2024`, `2025`)

progression_summary

#wow, super small compared to other districts. I wonder if RT and Team will want to just keep it as this.

#Methodology 2- 9th in 2022 and 12th in 2025
# Bookends methodology: 9th in 2022 AND 12th in 2025
bookends_summary <- stu_transcript_rollup %>%
  select(D_stu_id,
         H_stu_progression_traditional_2022,
         H_stu_progression_traditional_2025) %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  mutate(
    H_cohort_progression_bookends = case_when(
      H_stu_progression_traditional_2022 == "include" &
      H_stu_progression_traditional_2025 == "include" ~ "include",
      TRUE ~ "exclude"
    )
  ) %>%
  count(H_cohort_progression_bookends,
        name = "student_count")

bookends_summary

#okay, would still exclude 26 students, which makes sense given what we saw with traditional being that the most common case was repeating 11th grade, or not having 12th grade enorllment in 2025. Great! Will present to Rebel


```

### f. Filter to cohort by critera

-   **What:** The final step is to filter the data to just the students you want to include in your analysis. This is done by filtering to students who meet the location, duration, and presistence critera you have established. FOR MNPS- so that Rebel can look at additional flags, I am not going to filter just yet, but rather leave the unfiltered students in with the criteria columns for her to explore

```{r}
#If necessary, filter to just specific home-school locations
# student_data <- student_data %>%
#   filter(C_home_school_code == "1"| C_home_school_code == "29"| C_home_school_code == "39")

#How many students would meet the criteria for both duration (200 days) and persistance?
stu_transcript_rollup %>%
  filter(H_cohort_duration_include == "include" & H_cohort_progression_include == "include") %>%
  summarize(count_students = n_distinct(D_stu_id)) %>%
  print()

#2306

#Filter to students who met the duration criteria, or were present in the district for 75% of the year for each year of your analysis
# validated_student_data <- validated_student_data %>%
#   filter(H_cohort_duration_include == "include")

#Filter to students who met the persistence criteria, or progressed traditionally from 9th to 12th grade in each of the years of your analysis
# validated_student_data <- validated_student_data %>%
#   filter(H_cohort_progression_include == "include")

# #Count the new unique number of students, which should now be your cohort number
# validated_student_data %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   print()
# 
# #Check the count by year, which should be the same number in each year
# validated_student_data %>% 
#   group_by(D_year) %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   ungroup() %>% 
#   print()
# 
# #Drop the helper columns for location, duration, and persistence 
# validated_student_data <- validated_student_data %>%
#   select(-starts_with("H_stay_duration"),
#          -starts_with("H_stu_progression"),
#          -starts_with("duration_"))

```

# Part III: Create transcript data frame

## 3.1 Merge Transcript and Student Cohort Data Frames

-   **What:** Return & remove student ids that exist in cohort file but not transcript file, and vice versa; this will make merging easier later on

-   **Check:** Number of rows at end is as expected

-   **Project Team Action:** Gut check/track

```{r}
# #get the IDs that are in the student cohort file but not in the transcript file, should be a minimal number
# stu_ids_missing <- setdiff(validated_student_data$D_stu_id, transcript_roll_up$D_stu_id) %>% print()
# 
# #are these students we would have otherwise included in cohort
# included_students <- validated_student_data %>%
#   filter(D_stu_id %in% stu_ids_missing & C_stu_cohort == "include") %>%
#   pull(D_stu_id) %>% 
#   print()
# 
# #remove any student IDs from the transcript file that we do not have student data for. NOTE- this will also remove any data for students not included in the cohort from the transcript file, so if you want to do any analysis on those students or students overall, you will have to use just the transcript data frame!
# transcript_roll_up <- transcript_roll_up %>%
#   filter(D_stu_id %in% validated_student_data$D_stu_id)
# 
# #count distinct students to make sure number validates
# transcript_roll_up %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   print()
# 
# #Using safe-left join, join the student data into the transcript data by student ID and year
# transcript_roll_up <- transcript_roll_up %>%
#   safe_left_join(validated_student_data, by = c("D_stu_id", "D_year"))
# 
# #Count the number of students, should be your "cohort" number
# transcript_roll_up %>% 
#   summarize(count_students = n_distinct(D_stu_id)) %>% 
#   print()

```

### 3.2 File Cleaning

-   **What:** The following code cleans the transcript data frame to remove any columns that are not necessary for analysis. This is a good time to check in with your team on what columns are necessary for your analysis, and remove any unnecessary columns.

```{r}
# #Rename the data-frame
# merged_transcript_cohort <- transcript_roll_up
# 
# #Remove any columns that are not necessary for analysis (update as necessary for your team). These might be columns you didn't remove before, any district specific columns or checks you added or no longer need, or you might slip this step all together
# merged_transcript_cohort <- merged_transcript_cohort %>%
#   select(-starts_with("D_stu_enter_date"),
#          -starts_with("D_stu_exit_date"),
#          -starts_with("H_stay_duration"),
#          -starts_with("H_stu_progression")
#     )

```

# Part IV: Export Tables / Files

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <-  "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"


ers_write_sharepoint(data = stu_transcript_rollup,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "MNPS_Draft_Student_Cohort.csv",
                    drive_name = "client_work_drive")

```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval=FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
