---
title: "3_student_snapshot_selection"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: August 13 2024
review:
last update: April 14 2025
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Start fresh
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
gc()
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in the student and transcript data that you will need to select your snapshot of students.

-   **Project Team Action:** Replace folder path with the folder path where your files are and replace specific files names as necessary.

```{r, warning=F, message=F}

raw_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

#bring in file
validated_student_data <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "HPS_stu_crossvalidated.csv",
  drive_name = "internal_drive")


#bring in file
transcript_roll_up <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "HPS_transcript_crossvalidated.csv",
  drive_name = "internal_drive")


```

# Part II: Choosing Snapshot of Students

## 2.1 Create student snapshot for current year (e.g. SY 2024)

### a. Student counts before filtering

-   **What:** If you have not already, you should read through the Tettra pages outlining student snapshot selection here: <https://app.tettra.co/teams/ersknowledge/pages/wip-deep-dive-snapshot-analysis>. The following code contains validation checks before filtering, including number of students, number of students in high schools, and number of students by location for included schools

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Update based on year/specific snapshot group of students you want to select. Gut check/track validation results with team.

```{r}
#First, filter your data to just the snapshot year you want to work with. The example here shows 2024, but if you are using this script in later years, you will need to update this to the year you are working with
student_snapshot <- validated_student_data %>% 
  filter(D_year == 2024)

#count distinct students before filtering
student_snapshot %>% 
  summarize(count_students = n_distinct(D_stu_id)) 

#count of total HS students (in the case that your data frame contains non-high school students)
student_snapshot %>% 
  filter(D_stu_grade_level %in% c(9, 10, 11, 12)) %>%
  summarize(count_students = n_distinct(D_stu_id)) 


#Count of students by location IDs before filtering- note that this count is NOT totally unique, it could be including students who have a row for each location
student_snapshot %>% 
  filter(D_stu_grade_level %in% c(9,10,11,12)) %>%
  group_by(D_location_name) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()


```

### b. Check student enter and exit dates

-   **What:** Convert enter and exit dates to be as dates (if they aren't already), and check for any students with more than one row

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Gut check/track

```{r}

#correct data format
#update your date format below, this one work for "05/28/2024" 
#if your date format is "2023-05-28" then you would want to adjust the below to: "%y_%m_%d,
correct_date_format <-"%m/%d/%Y"

# convert enter and exit date to be as dates (if not already)
student_snapshot <- student_snapshot %>%
  mutate(D_stu_enter_date = as.Date(D_stu_enter_date, format = correct_date_format),
         D_stu_exit_date = as.Date(D_stu_exit_date,format =correct_date_format))

#check student enter dates
student_snapshot %>% 
  group_by(D_stu_enter_date) %>% 
  count() %>% 
  arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
  ungroup()

# Check student exit dates
student_snapshot %>% 
  group_by(D_stu_exit_date) %>% 
  count() %>% 
  arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
  ungroup()

```

### c. Creating Exclusion Criteria: Location, Duration, Persistence

-   **What:** There are three major exclusion criteria and filters for transcript analysis:
    1.  Location (the school the student attended),
    2.  Duration (how long they were in the school per year), and;
    3.  Persistence (If they matriculated traditionally from one grade to the next.)
-   These steps may vary based on specific team methodology and needs. For some districts, ERS has used a 75% criteria, where students need to be present 75% of the year. In other districts, that has been lowered to 10%. Other districts have ignored entry and exit dates all together and used credits attained as the determination for selecting students. You should outline with your team the options for selecting a snapshot, and move forward with the best option based on your specific analytic questions. The code below shows examples of selecting students based on some of these methodologies.

#### i. Location (Home School)

-   **What:** The first exclusion criteria for snapshot selection is home school- assigning a student one location ID based on the location ID they spent the most time at. This is done by creating a new column that calculates the duration of stay at each school, and then filtering to the maximum duration of stay for each student. Then, you can easily filter to schools of interest for your analysis.

-   **Check:** Make sure no student has more than 1 home school, and that you have cleared and checked this methodology with your team.

```{r}
# Step-by-step version with safe_left_join at the end
student_home_school <- student_snapshot %>%
  # Step 1: Calculate number of days per row (entry-exit pair)
  mutate(H_stay_duration = as.numeric(D_stu_exit_date - D_stu_enter_date)) %>%

  # Step 2: Summarize total stay duration per student-location combo
  group_by(D_stu_id, D_location_id) %>%
  summarize(H_total_days_at_location = sum(H_stay_duration, na.rm = TRUE), .groups = "drop") %>%

  # Step 3: For each student, find the location with the most days (break ties randomly)
  group_by(D_stu_id) %>%
  slice_max(order_by = H_total_days_at_location, n = 1, with_ties = TRUE) %>%  # keeps all ties
  slice_sample(n = 1) %>%  # randomly pick one if tied
  ungroup() %>%

  # Step 4: Rename the chosen location as the "home school"
  rename(H_home_school = D_location_id) %>%
  select(D_stu_id, H_home_school)

# Step 5: Use safe_left_join to merge back onto the original dataset
student_snapshot <- safe_left_join(student_snapshot, student_home_school, by = "D_stu_id")

# Step 6: Add back in home school name
student_snapshot <- student_snapshot %>%
  mutate(C_home_school_name = case_when(
    H_home_school == "1" ~ "Waikea High School",
    H_home_school == "2" ~ "Hilo High School",
    H_home_school == "3" ~ "Konaeawena HS",
    H_home_school == "4" ~ "Kamehameha High",
    H_home_school == "5" ~ "Hawaii Preparatory Academy",
    TRUE ~ "Other"
  ))

# count total students in the snapshot
student_snapshot %>% 
  summarize(count_students = n_distinct(D_stu_id))

#Get student counts by grade
student_snapshot %>% 
  group_by(C_eoy_grade) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#count total students by home school
student_snapshot %>% 
  group_by(C_home_school_name) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#If necessary, filter to only home schools of interest/filter out excluded locations (skip this if you already have a filtered data set)
# student_snapshot <- student_snapshot %>% 
#   filter(C_home_school_code == "1"| C_home_school_code == "29"| C_home_school_code == "39")


```

#### ii. Duration

-   **What:** Filtering to students who spent a certain amount of time at their home school. This can be done in a number of ways, check with your team on the appropriate methodology. The example below shows a 75% threshold, but this can be adjusted based on your needs.

-   **NOTE:** If your team is using a credit attempt cutoff (see Tettra for examples and guidance), skip this step and see option step in part 3.1 C.

```{r}
#Using Sum stay duration (the total number of days a student is present that you created in the last step), create a new include/exclude flag based on the number of days a student was present at their home school. This is done by filtering to students who spent 75% of the year at their home school. This can be adjusted based on your needs.
student_snapshot <- student_snapshot %>%
  mutate(H_sum_stay_duration_year = sum(M_days_enrolled, na.rm = TRUE)) %>%
  mutate(C_stu_snapshot = if_else(H_sum_stay_duration_year >= 0.75 * 180, "include", "exclude")) %>%
  select(-H_sum_stay_duration_year)


#count of students by include/exclude flag
student_snapshot %>% 
  group_by(C_stu_snapshot) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#make sure no student has both include and exclude flag
student_snapshot %>% 
  group_by(D_stu_id) %>% 
  summarize(count_students = n_distinct(C_stu_snapshot)) %>% 
  filter(count_students > 1) 

```

### d. Individual Student Checks

```{r}
#check a few student ids
# ---- Function: Check Random Students ----
check_random_students <- function(data, student_id_col = "D_stu_id", n = 5, seed = NULL) {
  # Optional seed for reproducibility
  if (!is.null(seed)) set.seed(seed)
  
  # Sample n unique student IDs
  random_ids <- sample(unique(data[[student_id_col]]), n)

  # Filter and return rows for those students, grouped for easy viewing
  data %>%
    filter(.data[[student_id_col]] %in% random_ids) %>%
    arrange(.data[[student_id_col]]) %>%
    group_by(.data[[student_id_col]])
}


# Check 5 random students
check_random_students(student_snapshot, n = 5, seed = 123)

#counts of unique students by include flag by home school, checking that no home school has an outsized proportion of excluded students
student_snapshot %>% 
  group_by(C_home_school_name,C_school_exclude_flag) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  pivot_wider(names_from = C_school_exclude_flag,values_from = count_students)%>%
  ungroup() 


```

### e. Persistence

-   **What:** The persistence criteria only applies in cohort analysis, and is generally not used in the snapshot analysis. If you are looking at more than one specific group of students, check the Tettra page for cohort analysis for more information on how to do this. For snapshot, there may be specific cases in which you only want "true" freshman (no repeaters) etc.

### f. Filter to snapshot by criteria

-   **What:** The final step is to filter the data to just the students you want to include in your analysis. This is done by filtering to students who meet the location, duration, and persistence criteria you have established.

```{r}
#Filter the snapshot data-frame just to students who have home_schools you want to include, and who meet the duration criteria your team established
student_snapshot <- student_snapshot %>% 
  filter(C_stu_snapshot == "include") %>% 
  select(-C_stu_snapshot) %>% 
  distinct()

#count of students in the snapshot
student_snapshot %>% 
  summarize(count_students = n_distinct(D_stu_id))

```

# Part III: Merge in Transcript Data

## 3.1 Creating snapshot transcript data frames

### a. Create Student Table

-   **What:** Using your cleaned and validated student data-frame, join in transcript data to create your data-frame for analysis. This is done by checking the number of students that match across the files, and then merging the transcript data with the student data. It is fairly common to have a slight mismatch in the number of students across data sources (about 10%), but if the number if higher than that you may want to check with your team for how to proceed.

-   **Check:** Number of rows at end is as expected

```{r}
#before transformation, count the number of unique snapshot students
student_snapshot %>% 
  summarize(count_students = n_distinct(D_stu_id))

#Now that you have the criteria you need, you can filter the student snapshot data to just the rows of data you need to add into the transcript information. We want this to be just one row per student.  Edit the columns as necessary for your specific analysis (if you have any district bespoke columns)
student_snapshot_unique <- student_snapshot %>% 
  select(D_stu_id, 
         C_home_school_name, 
         C_eoy_grade, 
         D_stu_race, 
         D_stu_frl, 
         D_stu_swd, 
         D_stu_ell) %>% 
  distinct()

#and then filter to just unique rows, such that there is one row per student
student_snapshot_unique <- student_snapshot_unique %>% 
  group_by(D_stu_id) %>% 
  slice(1) %>% 
  ungroup()

#and that unique number should be the same as before
student_snapshot_unique %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>%
  print()


```

### b. Check mismatching ids b/w Student Table and Transcript

```{r}

#Identify the student IDs that are present in the student snapshot  unique data frame but not in the transcript data frame
stu_ids_missing_from_transcript <- setdiff(student_snapshot_unique$D_stu_id, transcript_roll_up$D_stu_id)


#If you have a large number of missing IDs, you can run the following code for some checks-

#Are the IDs that are not matching all at one specific location?
student_snapshot_unique %>% 
  filter(D_stu_id %in% stu_ids_missing_from_transcript) %>% 
  group_by(C_home_school_name) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#Are the IDs that are not matching all at one specific grade?
student_snapshot_unique %>% 
  filter(D_stu_id %in% stu_ids_missing_from_transcript) %>% 
  group_by(C_eoy_grade) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()


#Once you feel good about moving forward or excluding any of these students, you can move on
```

### c. Filter the transcript to just snapshot year data (e.g., 2024)

-   **What:** Before merging transcript and student data, we want to make sure that we are using the same year (only year) of data for the snapshot.

-   **Check:** Number of rows for in snapshot year should match number of rows after filtering the file

-   **Project Team Action:** Update the snapshot year for your year of analysis

```{r}

#snapshot year
snapshot_year = 2024

#check how many transcript rows are from snapshot year
transcript_roll_up %>% 
  filter(D_year == snapshot_year) %>% 
  nrow() %>%
  print()

#filter the transcript data to just the rows you need for the analysis
transcript_2024 <- transcript_roll_up %>% 
  filter(D_year == snapshot_year)

```

### d. Create transcript data frame with only students you have transcript data for

-   **What:** Merge transcript snapshot and student snapshot data to bring in demographic info, etc.

-   **Check:** Number of rows in merged file should be equal to number of rows of students you have transcript data for. If you have a large number of students missing, check the code above to see if there are any patterns in the missing students.

-   **Project Team Action:** Update Snapshot Analysis Set-up table

```{r}
# Step 1: Filter the student snapshot to only students who appear in the transcript data
student_snapshot_unique <- student_snapshot_unique %>% 
  filter(D_stu_id %in% transcript_roll_up$D_stu_id)


# Step 2: Perform a safe left join of transcript data with the filtered snapshot
merged_transcript_snapshot <- safe_left_join(transcript_2024, student_snapshot_unique, by = "D_stu_id")

# Step 3: Sanity check – count number of unique students after merge
merged_transcript_snapshot %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()


```

### e. Preliminary checks on student credit accumulation and enrollment

-   **What:** Add some preliminary checks on student credit accumulation and course enrollment here before exporting the data for analysis, ensuring that the join worked well.

```{r}
# First lets look at overall student credit accumulation by creating buckets, and counting the distinct number of
 student_credit_accumulation <- merged_transcript_snapshot %>%
  group_by(`D_stu_id`) %>%
  # Summarize the number of credits earned per student per year
  summarize(credits_earned = sum(`D_creds_earn`, na.rm = TRUE)) %>%
  ungroup() %>%
  # Create a new column that buckets the number of credits earned per year
  mutate(credit_bucket = case_when(
    credits_earned == 0 ~ "0",
    credits_earned == 1 ~ "1",
    credits_earned == 2 ~ "2",
    credits_earned == 3 ~ "3",
    credits_earned == 4 ~ "4",
    credits_earned == 5 ~ "5",
    credits_earned == 6 ~ "6",
    credits_earned == 7 ~ "7",
    credits_earned == 8 ~ "8",
    credits_earned == 9 ~ "9",
    TRUE ~ "10+"
  )) %>%
  # Summarize the number of students in each credit bucket per year
  group_by(credit_bucket) %>%
  summarize(students = n_distinct(D_stu_id))

# View the result
print(student_credit_accumulation)


#Now lets check overall course taking to see that it matches similarly what we would expect
student_course_taking <- merged_transcript_snapshot %>%
  group_by(`D_stu_id`) %>%
  # Summarize the number of courses a student took by counting the number of unique course IDs
  summarize(courses_taken = n_distinct(`D_course_id`)) %>%

  # Create a new column that buckets the number of courses that students are taking
  mutate(courses_taken = case_when(
    courses_taken == 0 ~ "0",
    courses_taken <= 5 ~ "0-5",
    courses_taken <= 10 ~ "6-10",
    courses_taken <= 15 ~ "11-15",
    courses_taken <= 20 ~ "16-20",
    TRUE ~ "20+"
  )) %>%
  # Summarize the number of students in each credit bucket per year
  group_by(courses_taken) %>%
  summarize(students = n_distinct(D_stu_id))

# View the result
print(student_course_taking)

#And finally lets look at the percentage of students who took a pathway course, by first grouping by distinct student Ids, then counting the number of rows in which a student took a pathway course
pathway_summary <- merged_transcript_snapshot %>%
  # Ensure the student-course pairs are unique
  group_by(D_stu_id) %>%
  summarize(took_pathway = any(D_pathways_flag == "Y"), na_count = sum(is.na(D_pathways_flag))) %>%
  ungroup() %>%
  # Categorize the students based on the summary
  mutate(category = case_when(
    took_pathway ~ "Took at least one pathway course",
    na_count > 0 & !took_pathway ~ "N/A",
    TRUE ~ "Did not take any pathway course"
  )) %>%
  # Summarize the counts for each category
  count(category) %>%
  # Calculate percentages
  mutate(percentage = n / sum(n) * 100)

# Print the summary table
print(pathway_summary)

```

### c. Duration Exclusion- Credit Option (OPTIONAL)

-   **What:** If your team is using a credits attempted methodology approach to the duration exclusion criteria, the following code will allow you to set a limit for the number of credits attempted in a given year to be included in the analysis. This is an optional step, and should be discussed with your team before proceeding.

```{r}
#create a column that counts the number of credits attempted per student per year
merged_transcript_snapshot <- merged_transcript_snapshot %>%
  group_by(D_stu_id) %>%
  mutate(credits_attempted = sum(D_creds_possible, na.rm = TRUE)) %>%
  ungroup()

#create a new include/exclude credits columns for students who attempted XX credits per year (replace based on team needs!)
merged_transcript_snapshot <- merged_transcript_snapshot %>%
  mutate(C_credits_exclusion = if_else(credits_attempted >= 4, "include", "exclude")) %>%
  select(-credits_attempted)

#count of students by include/exclude flag
merged_transcript_snapshot %>% 
  group_by(C_credits_exclusion) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#make sure no student has both include and exclude flag
merged_transcript_snapshot %>% 
  group_by(D_stu_id) %>% 
  summarize(count_students = n_distinct(C_credits_exclusion)) %>% 
  filter(count_students > 1)

#filter out students who do not meet the criteria
merged_transcript_snapshot %>% 
  filter(C_credits_exclusion == "include") %>% 
  select(-C_credits_exclusion) 

```

# Part IV: Export Tables / Files

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

ers_write_sharepoint(data = merged_transcript_snapshot,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "HPS_merged_transcript_snapshot.csv",
                    drive_name = "internal_drive")
```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval=FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
