---
title: "2_cross_file_validation_MNPS"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: "July 30 2024"
review:
last update: "April 10 2025"
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. Note that for this script, you will be validating across your already validated data sources and files, so you should read in multiple files below. At this point, column names should be cleaned, match ERS common naming conventions, and be in the correct format.

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning = F, message = F}

raw_data_folder_path <- "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"

#transcript data
transcript_data_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "MNPS_Transcript_Validated.csv",
  drive_name = "client_work_drive")

#student data
student_data_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "MNPS_Stu_Rollup.csv",
  drive_name = "client_work_drive")

#course catalog data
past_course_catalog <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "03_course_coding_24-25 - CODED 10.24.xlsx",
                   drive_name = "client_work_drive")

```

### 1.4 View your data

**ACTION: Replace "raw_data" with the same name of your data frame you used above, and run the code to inspect your data.**

```{r, eval = interactive()}
View(transcript_data_validated)
view(student_data_validated)
```

## Part 2: Validate Data

### 2.1 Run ers_column_info on your data

**ACTION: Replace "raw_data" with the name of your data frame**

```{r}
ers_column_info(student_data_validated)
ers_column_info(transcript_data_validated)
ers_column_info(past_course_catalog)

```

### 2.2 Student counts- I actually don't think this is needed here because you do it in the next step?

#### a. Return distinct student counts in student, transcript, and school files

-   **What:** Comparing student counts across files to check that they are comparable. You might see different counts across data-sources due to data integrity differences or reporting differences. For example, enrollment files often contain more records than stored grades or transcript files.

-   **Check:** Are the student counts similar? How many overlapp between files?

-   **Project Team Action:** Gut check/track

```{r}
# # Function: takes a dataset + label and returns distinct counts by year + overall
# calculate_dynamic_counts <- function(data, label) {
#   # Get all unique years
#   years <- data %>% 
#     distinct(D_year) %>% 
#     pull(D_year)
#   
#   # Calculate per-year counts
#   per_year_counts <- map_dfr(years, function(y) {
#     data %>%
#       filter(D_year == y) %>%
#       summarise(
#         Source = label,
#         Year = as.character(y),
#         Distinct_Student_Count = n_distinct(D_stu_id)
#       )
#   })
# 
#   # Add all-years summary
#   all_years_count <- tibble(
#     Source = label,
#     Year = "All",
#     Distinct_Student_Count = n_distinct(data$D_stu_id)
#   )
# 
#   bind_rows(per_year_counts, all_years_count)
# }
# 
# 
# # Named list of datasets and their labels
# datasets <- list(
#   "Student Data" = student_data_validated,
#   "Transcript Data" = transcript_data_validated
# )
# 
# # Apply function to all datasets
# compare_student_counts <- imap_dfr(datasets, ~ calculate_dynamic_counts(.x, .y)) %>%
#    pivot_wider(
#     names_from = Year,
#     values_from = Distinct_Student_Count
#   ) %>% 
#   print()
 
```


 

#### c. Return student ids that don't match between student and transcript files

-   **What:** This is an important step to return student ids that don't match between files

-   **Check:** Count of unique student ids should match your counts from step 2.1a (see your global environment under values for these new counts and compare). After returning the differing sets of student ids, check your work by choosing a few student ids to search for in each file. For example, if XXX is returned as a student id existing in transcript but not student, search for that in in both to confirm

-   **Project Team Action:** Gut check/track. Discuss whether it makes sense to exclude these students with your team

```{r}
#get the unique D_stu_id values in student file
stu_ids_stu_data <- unique(student_data_validated$D_stu_id) 

#get the unique D_stu_id values in transcript file
stu_ids_transcript_data <- unique(transcript_data_validated$D_stu_id)

#get the ids that are different
#this line of code returns student ids that exist in student data but not transcript data
stu_ids_in_stu_not_transcript <- setdiff(stu_ids_stu_data, stu_ids_transcript_data) 

#this line of code returns student ids that exist in transcript data but not student data
stu_ids_in_transcript_not_stu <- setdiff(stu_ids_transcript_data, stu_ids_stu_data)

# Get the IDs that are the same in both files
stu_ids_in_both <- intersect(stu_ids_stu_data, stu_ids_transcript_data)

# Count the number of distinct IDs that are the same between the two files
count_same_ids <- length(stu_ids_in_both)

# Count the number of distinct IDs that are different between the two files
count_different_ids <- length(unique(c(stu_ids_in_stu_not_transcript, stu_ids_in_transcript_not_stu)))

# Print the counts
cat("Number of distinct IDs that are the same between the two files:", count_same_ids, "\n")
cat("Number of distinct IDs that are different between the two files:", count_different_ids, "\n")


```

#### d. Filter to rows of student ids that exist in transcript file but not student file, and vice versa

-   **What:** Filters to rows of student ids that exist in transcript file but not student file, and vice versa

-   **Check:** Check that distinct counts in new filtered sets of student ids match distinct counts from step 2.1c

-   **Project Team Action:** Gut check/track

```{r}
#filter to student ids that exist in student file but not transcript file
filtered_stu_ids_stu_data <- student_data_validated[student_data_validated$D_stu_id %in% stu_ids_in_stu_not_transcript, ]

#filter to student ids that exist in transcript file but not student file
filtered_stu_ids_trancript_data <- transcript_data_validated[transcript_data_validated$D_stu_id %in% stu_ids_in_transcript_not_stu, ]

#check that distinct count in filtered_diff_stu_ids data frames matches distinct count in diff_stu_ids
count_filtered_stu_ids_stu_data <- filtered_stu_ids_stu_data %>%
  distinct(D_stu_id)

# create a table with the counts of year in filtered_stu_ids_transcript_data
 filtered_stu_ids_trancript_data %>%
  group_by(D_end_year) %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  ungroup()

```

#### e. Explore exit codes and enter/exit dates of mismatched student ids

-   **What:** To determine if these mismatched student ids should be excluded and removed, explore the students' enter and exit dates and exit code descriptions (if you have them) in your filtered student data. In Lubbock, for example, these appeared to be students who left the district before they could get any grades. We removed these students (see next steps)

-   **Check:** You are checking to see if these mismatched student ids are students who left the district to go to another educational institution before they could receive any grades

-   **Project Team Action:** Discuss whether it makes sense to exclude these students with your team

#### f. Add exclude flag to the student ids that don't match & remove students

-   **What:** Add exclude flag to students who don't match between files and remove them

-   **Check:** After removing the students, check that the number of rows is equal to the original number of rows minus the number of rows in your filtered data (when you filtered to the mismatched student ids). For example, if you had 2,000 rows of students who needed to be excluded from student file, and 60,000 rows, you should have 58,000 rows after removing those students from the file

-   **Project Team Action:** None (run as is)

#####MNPS CODE PATCH-
```{r}
#09/15 Going back to add a flag for students who ever transferred WITHIN the same school year by looking at the student demo data
student_data_validated <- student_data_validated %>% 
  group_by(D_stu_id, D_year) %>% 
  mutate(C_transfered_within_year_flag = ifelse(n_distinct(D_location_name) > 1, "Y", "N")) %>% 
  ungroup()
    
#Count the flag
student_data_validated %>%
  filter(C_transfered_within_year_flag == "Y") %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#out of how many students total
student_data_validated %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#okay so 8.800 out of 51,000, about 17% of students in the demo data transferred at least once within a school year. When we check this flag later in the cohort, we should expect to see a similar number.


```

##### i. Student File

```{r}
#add column with exclude flag
 student_data_validated <- student_data_validated %>%
   mutate(C_stu_exclude_flag = ifelse(D_stu_id %in% filtered_stu_ids_stu_data$D_stu_id, "exclude", "include"))
 
# #check count with exclude flag to make sure it matches # rows in filtered student data set
 exclude_count_stu_data <- sum(student_data_validated$C_stu_exclude_flag == "exclude")

#check distinct count with exclude flag to make sure it matches
student_data_validated %>%
   filter(C_stu_exclude_flag == "exclude") %>%
   summarise(distinct_count = n_distinct(D_stu_id))


#check a student id that was a mismatch and should have exclude flag
 stu_id_deep_dive_1 <- student_data_validated %>%
   filter(D_stu_id == "XXX") %>% #update with student ID from your file
   print()

#removing excluded students from data
student_data_validated <- student_data_validated %>% 
   filter(C_stu_exclude_flag == "include")
 
# #remove C_stu_exclude_flag column since you don't need it anymore
student_data_validated <- select(student_data_validated, -C_stu_exclude_flag)

#so from this point onwards, student data validated should ONLY have data that includes students that we have transcript data for. Does not imply a 1:1 year match, but does imply that we have transcript data for them.

#Now double-check that the exclusion work by seeing if any students not in the transcript file are still in the student file
 check_stu_ids <- student_data_validated %>%
   filter(D_stu_id %in% filtered_stu_ids_stu_data$D_stu_id) %>% 
   print()
 
 #Great!

```

##### ii. Transcript File

```{r}
#add column with exclude flag
 transcript_data_validated <- transcript_data_validated %>%
   mutate(C_stu_exclude_flag = ifelse(D_stu_id %in% filtered_stu_ids_trancript_data$D_stu_id, "exclude", "include"))
 
#check count with exclude flag to make sure it matches
 exclude_count_transcript_data <- sum(transcript_data_validated$C_stu_exclude_flag == "exclude")

#check distinct count with exclude flag to make sure it matches
 distinct_exclude_count_transcript_data <- transcript_data_validated %>%
   filter(C_stu_exclude_flag == "exclude") %>%
   summarise(distinct_count = n_distinct(D_stu_id))
 
# #check a student id that was a mismatch and should have exclude flag
 stu_id_deep_dive_2 <- transcript_data_validated %>%
   filter(D_stu_id == "XXX") #update with student ID from your file
 
#removing excluded students from data
 transcript_data_validated <- transcript_data_validated %>% 
   filter(C_stu_exclude_flag == "include")
 
#remove C_stu_exclude_flag column since you don't need it anymore
 transcript_data_validated <- select(transcript_data_validated, -C_stu_exclude_flag)


```

#### g. Check to make sure all student ids match now

-   **What:** Check that count of distinct student ids in student and transcript file match

-   **Check:** Check counts on unique student ids in each file type by looking in your environment under "values". There should not be any ids that are different between files now

-   **Project Team Action:** None (run as is)

```{r}
#get the unique D_stu_id values in student file
 stu_ids_stu_data2 <- unique(student_data_validated$D_stu_id) 
 
#get the unique D_stu_id values in transcript file
 stu_ids_transcript_data2 <- unique(transcript_data_validated$D_stu_id)

#get the ids that are different
 check_diff_stu_ids <- setdiff(stu_ids_stu_data2, stu_ids_transcript_data2 )
 
 check_diff_transcript_ids <- setdiff(stu_ids_transcript_data2, stu_ids_stu_data2)
 
# Count the number of distinct IDs that are different between the two files
count_different_ids <- length(unique(c(check_diff_stu_ids, check_diff_transcript_ids)))

# Print the counts
cat("Number of distinct IDs that are different between the two files:", count_different_ids, "\n")
 
```

####h. Checking student year AND ID between both files
```{r}
#realizing what is throwing off my code later on is that the unique combination of student ID and YEAR needs to match across both files, not JUST student ID. 

#in transcript, rename D_end_year to D_year so that we have a common column name to match on, and remove D_entry_year
transcript_data_validated <- transcript_data_validated %>%
  rename(D_year = D_end_year) %>%
  select(-D_start_year)

#make sure the D_stu_id columns are the same data type (character)
student_data_validated <- student_data_validated %>%
  mutate(D_stu_id = as.character(D_stu_id))

transcript_data_validated <- transcript_data_validated %>%
  mutate(D_stu_id = as.character(D_stu_id))

# Get unique student-year pairs in each file
stu_year_pairs_student <- student_data_validated %>%
  distinct(D_stu_id, D_year)

stu_year_pairs_transcript <- transcript_data_validated %>%
  distinct(D_stu_id, D_year)

# Pairs that are only in student data
diff_student_only <- anti_join(stu_year_pairs_student, stu_year_pairs_transcript,
                               by = c("D_stu_id", "D_year"))

# Pairs that are only in transcript data
diff_transcript_only <- anti_join(stu_year_pairs_transcript, stu_year_pairs_student,
                                  by = c("D_stu_id", "D_year"))

# Pairs that are in both files
matching_pairs <- inner_join(stu_year_pairs_student, stu_year_pairs_transcript,
                             by = c("D_stu_id", "D_year"))

# Students that appear in both datasets
common_students <- intersect(student_data_validated$D_stu_id, transcript_data_validated$D_stu_id)

# Student-year pairs that are in both datasets (already computed as matching_pairs)
# Now identify mismatched pairs for those students
diff_year_pairs <- bind_rows(
  anti_join(stu_year_pairs_student, stu_year_pairs_transcript, by = c("D_stu_id", "D_year")),
  anti_join(stu_year_pairs_transcript, stu_year_pairs_student, by = c("D_stu_id", "D_year"))
) %>%
  filter(D_stu_id %in% common_students)

# Print counts
cat("Number of student-year pairs only in student data:", nrow(diff_student_only), "\n")
cat("Number of student-year pairs only in transcript data:", nrow(diff_transcript_only), "\n")
cat("Number of student-year pairs in both files:", nrow(matching_pairs), "\n")
cat("Number of student IDs with mismatched years:", n_distinct(diff_year_pairs$D_stu_id), "\n")
cat("Number of mismatched student-year pairs:", nrow(diff_year_pairs), "\n")
# Validation before removal-

# Unique students in original datasets
orig_students_student <- n_distinct(student_data_validated$D_stu_id)
orig_students_transcript <- n_distinct(transcript_data_validated$D_stu_id)

# Unique students in filtered datasets
filtered_students_student <- n_distinct(student_data_validated$D_stu_id)
filtered_students_transcript <- n_distinct(transcript_data_validated$D_stu_id)

# Students removed
removed_students_student <- orig_students_student - filtered_students_student
removed_students_transcript <- orig_students_transcript - filtered_students_transcript

cat("Number of unique students removed from student data:", removed_students_student, "\n")
cat("Number of unique students removed from transcript data:", removed_students_transcript, "\n")


#and filter them out
# Step 1: Get the student-year pairs present in both files
valid_pairs <- inner_join(
  student_data_validated %>% distinct(D_stu_id, D_year),
  transcript_data_validated %>% distinct(D_stu_id, D_year),
  by = c("D_stu_id", "D_year")
)

# Step 2: Filter both datasets to only include valid student-year pairs
student_data_validated <- student_data_validated %>%
  semi_join(valid_pairs, by = c("D_stu_id", "D_year"))

transcript_data_validated <- transcript_data_validated %>%
  semi_join(valid_pairs, by = c("D_stu_id", "D_year"))

# Step 3: Validation checks

# 3a. Count of student-year pairs in student data
cat("Student data rows after filtering:", nrow(student_data_validated), "\n")

# 3b. Count of student-year pairs in transcript data
cat("Transcript data rows after filtering:", nrow(transcript_data_validated), "\n")

# 3c. Ensure no missing pairs remain
check_missing_student <- anti_join(student_data_validated, transcript_data_validated, 
                                   by = c("D_stu_id", "D_year"))
check_missing_transcript <- anti_join(transcript_data_validated, student_data_validated, 
                                      by = c("D_stu_id", "D_year"))

cat("Student data rows with missing transcript match:", nrow(check_missing_student), "\n")
cat("Transcript data rows with missing student match:", nrow(check_missing_transcript), "\n")



```


### 2.2 MNPS Updated- Append files and check Locations

-   **What:** Because a. The transcript file did not have location codes attached and b. we found a larger than expected number of missing courses when validating the transcript course IDs and names to the previously coded course-catalog, we are going to leapfrog a bit here compared to normal methodology.

A. First, we need to make sure that the student demographic file is one row per student per year, so we have to assign students a "year_home_school", which is whatever school they spent the majority of the year at (using M_Days_enrolled) 
B. Then, we are going to append the student data to the transcript data
C. Next, we are going to assign 4 home-school using m-days enrolled/year_home_school to see where students spent the majority of their 4 years in MNPS
D. Next, we will look at the list of home schools and create a new data-frame excluding any home-schools that are not AON, but not any student enrollment at schools that are not AON (team still has to make a choice on this methodology)
E. Finally, we will re-test the overlap of course IDs in transcript and course catalog to re-size the issue and see how many courses may need to be re-coded

#### A. Assign home school year
```{r}
#Check student 4 before transformations
student_check_4 <- student_data_validated %>%
  filter(D_stu_id == 190137047) %>%
  print()


#Before transformations, check the number of unique students in the file
n_distinct(student_data_validated$D_stu_id) %>% 
  print()


#Do we have any students with NA values for D_location_name?
student_data_validated %>%
  filter(is.na(D_location_name)) %>%
  print()


#Using M_days_enrolled, create a new column that takes the D_location_name_clean value for the row that each individual student spent the majority of the year at (highest value for M_days_enrolled)
student_data_validated <- student_data_validated %>%
  group_by(D_stu_id, D_year) %>%
  mutate(
    C_year_home_school = D_location_name_clean[which.max(M_days_enrolled)]
  ) %>%
  ungroup()

#and double check that the number of unique students in the file hasn't changed
n_distinct(student_data_validated$D_stu_id) %>% 
  print()

student_check_4 <- student_data_validated %>%
  filter(D_stu_id == 190137047) %>%
  print()


#and that there are no NA values of D_location_name_clean
student_data_validated %>%
  filter(is.na(D_location_name_clean)) %>%
  print()
```

####B. SE MNPS CODE FIX
```{r}
#By jumping ahead in the cohort selection earlier than we normally would, I made an error that lead to M_days_enrolled not working quite like it should. Since we already have home_school set by the right methodology now, we need to do a couple of things. First we need to ensure that M_days_enroled is consistent per student per year. So we need to group by D_stu_id and D_year, Sum any unique values of M_days_enrolled, and use THAT as M_days_enrolled so that every student has ONE accuate number per year
student_data_validated <- student_data_validated %>%
  group_by(D_stu_id, D_year) %>%
  mutate(M_days_enrolled = sum(unique(M_days_enrolled))) %>%
  ungroup()

#double-check student 4
student_check_4 <- student_data_validated %>%
  filter(D_stu_id == 190137047) %>%
  print()

#amazing, okay that should fix it!

#great its the same, and then lets create a new student data frame that drops D_location_name, D_stu_entry_date, D_stu_exit_date, and D_location_name_clean, so that we can get to a unique row per student per year
student_data_unique <- student_data_validated %>%
  select(-D_location_name, -D_stu_enter_date, -D_stu_exit_date, -D_location_name_clean)

#Great, and now using student_data_unique, keep only unique row combinations of D_stu_id, D_year, and C_year_home_school (forcing 1 row per year per student)
student_data_unique <- student_data_unique %>%
  distinct(D_stu_id, D_year, C_year_home_school, .keep_all = TRUE)

#double-check student 4
student_check_4 <- student_data_unique %>%
  filter(D_stu_id == 190137047) %>%
  print()

#Yup- All set now! Okay, going to re-run this whole script, then back to cohort selection to see what impact this update has. 

#and double-check that student count is the same
n_distinct(student_data_unique$D_stu_id) %>% 
  print()

#and that student_data_unique has no NA values for C_year_home_school
student_data_unique %>%
  filter(is.na(C_year_home_school)) %>%
  print()

#Step 1 done!

```


#### B. Append student data to the transcript data
```{r}
#before we merge, double-check the number of unique students in each data-frame
demo_student_count <- n_distinct(student_data_validated$D_stu_id) %>% 
  print()

transcript_student_count <- n_distinct(transcript_data_validated$D_stu_id) %>% 
  print()

#great


#and before join just double-check that matching columns are standard
transcript_data_validated <- transcript_data_validated %>%
  mutate(
    D_stu_id = str_trim(as.character(D_stu_id)),
    D_year   = as.integer(D_year)
  )

student_data_unique <- student_data_unique %>%
  mutate(
    D_stu_id = str_trim(as.character(D_stu_id)),
    D_year   = as.integer(D_year)
  )

#Matching by student ID and year, we want to leftjoin the student data to the transcript data-frame. It will be a 1
student_transcript_rollup <- transcript_data_validated %>%
  safe_left_join(student_data_unique, by = c("D_stu_id", "D_year"))

#check that the join worked by counting the number of unique students in the new data-frame
n_distinct(student_transcript_rollup$D_stu_id) %>% 
  print()

#check that there are no NA values for C_year_home_school
student_transcript_rollup %>%
  filter(is.na(C_year_home_school)) %>%
  print()


#Rename D_stu_grade_level.x to D_stu_grade_level_transcript, and D_stu_grade_level.y to D_stu_grade_level_demo. This is part of the ongoing multiple grades per year thing that we will need to solve later!
student_transcript_rollup <- student_transcript_rollup %>%
  rename(D_stu_grade_level_transcript = D_stu_grade_level.x,
         D_stu_grade_level_demo = D_stu_grade_level.y)


student_check_4 <- student_transcript_rollup %>%
  filter(D_stu_id == 190137047) %>%
  print()


```


#### C. Assign 4 year home_school

```{r}
#Alright, first grouping by student ID and year, we want to create a new column that counts the number of distinct year
student_transcript_rollup <- student_transcript_rollup %>%
  group_by(D_stu_id) %>%
  mutate(M_years_enrolled = n_distinct(D_year)) %>%
  ungroup()

#Now create a new-data-frame that filters to just students who have been enrolled for 4 years AND students who were 9th graders in 2021-2022
eligible_students <- student_transcript_rollup %>%
  filter(M_years_enrolled == 4 & C_9th_in_22 == 1) %>%
  distinct(D_stu_id)

#Keep ALL rows for those students
cohort_selection <- student_transcript_rollup %>%
  filter(D_stu_id %in% eligible_students$D_stu_id)

#and now how many students in cohort
n_distinct(cohort_selection$D_stu_id) %>% 
  print()


#Do we have any students with only NA values for C_year_home_school?
student_transcript_rollup %>%
  filter(is.na(C_year_home_school)) %>%
  select(D_stu_id, D_year) %>%
  print()

#First lets look at the list of all of the unique location names
cohort_selection %>%
  group_by(C_year_home_school) %>%
  distinct(C_year_home_school) %>%
  print()

schools_to_include <- c(
  "Hunters Lane High",
  "McGavock High",
  "John Overton High",
  "Stratford",
  "Cane Ridge High School",
  "Whites Creek High School",
  "Glencliff High School",
  "Antioch High School",
  "Hillsboro High",
  "PearlCohn High",
  "Maplewood High",
  "James Lawson High",
  "James Lawson High School",
  "Hillwood High"
)

#09/12 GOING BACK TO CREATE A FLAG-
#Very important, we want to create a new flag that is "Y" if a student has more than 1 unique Year Home School in their 4 years in MNPS
cohort_selection <- cohort_selection %>%
  group_by(D_stu_id) %>%
  mutate(M_unique_year_home_schools = n_distinct(C_year_home_school)) %>%
  ungroup() %>%
  mutate(C_multiple_year_home_schools_flag = ifelse(M_unique_year_home_schools > 1, "Y", "N")) %>%
  select(-M_unique_year_home_schools)

#How many students with that flag?
cohort_selection %>%
  filter(C_multiple_year_home_schools_flag == "Y") %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()


#Okay, so now we want to assign students a 4 year home school, or grouping by D_stu_id and C_year_home_school, create a new column called C_four_year_home_school that takes whichever C_year_home_school value the student spent the majority of their 4 years at. So if a student has 3 rows with C_year_home_school = "School A" and 1 row with C_year_home_school = "School B", they get assigned "School A" as their C_four_year_home_school. If a student has 2 and 2, they get assigned randomly.
cohort_selection <- cohort_selection %>%
  group_by(D_stu_id, C_year_home_school) %>%
  mutate(M_years_at_school = n()) %>%
  ungroup() %>%
  group_by(D_stu_id) %>%
  slice_max(M_years_at_school, n = 1, with_ties = TRUE) %>%
  # prioritize schools in the "include" list when there's a tie
  mutate(priority = if_else(C_year_home_school %in% schools_to_include, 1, 0)) %>%
  slice_max(priority, n = 1, with_ties = TRUE) %>% 
  # if still tied among include schools (or all exclude), break ties randomly
  slice_sample(n = 1) %>%
  mutate(C_four_year_home_school = C_year_home_school) %>%
  select(D_stu_id, C_four_year_home_school) %>%
  right_join(cohort_selection, by = "D_stu_id")


#Double check that no student has more than 1 four year home school
cohort_selection %>%
  group_by(D_stu_id) %>%
  summarise(count = n_distinct(C_four_year_home_school)) %>%
  filter(count > 1) %>%
  print()


 
```

####D. Look at Unique Home School names and filter to just AON schools
```{r}
#First lets look at all of the unique values for home_school
cohort_selection %>%
  group_by(C_four_year_home_school) %>%
  distinct(C_four_year_home_school) %>%
  print()

#Okay great. Now lets filter to just the AON schools
# Step 1: Create a vector of schools to include
schools_to_include <- c(
  "Hunters Lane High",
  "McGavock High",
  "John Overton High",
  "Stratford",
  "Cane Ridge High School",
  "Whites Creek High School",
  "Glencliff High School",
  "Antioch High School",
  "Hillsboro High",
  "PearlCohn High",
  "Maplewood High",
  "James Lawson High",
  "James Lawson High School",
  "Hillwood High"
)

#ungroup
cohort_selection <- cohort_selection %>%
  ungroup()

#How many students in those schools (AON home schools?)
cohort_selection %>%
  filter(C_four_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#3,014, thats a big cut

#How many students with a location other than an AON home school?
cohort_selection %>%
  filter(!C_four_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#1954

#what are the most common non-AON home schools?
cohort_selection %>%
  filter(!C_four_year_home_school %in% schools_to_include) %>%
  group_by(C_four_year_home_school) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  arrange(desc(distinct_students)) %>%
  print(n = Inf)

#and how many students in the cohort had ANY enrollment at non-include schools?
cohort_selection %>%
  filter(!C_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#and how many WERE included
cohort_selection %>%
  filter(C_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

# Step 2: Filter the cohort_selection dataframe
cohort_selection_filtered <- cohort_selection %>%
  filter(C_four_year_home_school %in% schools_to_include)

# Optional: check distinct schools to make sure only the included ones remain
distinct(cohort_selection_filtered, C_four_year_home_school)

#Awesome

#quickly rename "James Lawson High" to "James Lawson High School" because they are the same
cohort_selection <- cohort_selection %>%
  mutate(C_four_year_home_school = ifelse(C_four_year_home_school == "James Lawson High", "James Lawson High School", C_four_year_home_school))


#and how many students in the cohort had ANY enrollment at non-include schools?
cohort_selection_filtered %>%
  filter(!C_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#and how many WERE included
cohort_selection_filtered %>%
  filter(C_year_home_school %in% schools_to_include) %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()


```

####D2. MNPS Check- non AON home school students
```{r}
#Doing a check for MNPS to check of students at an AON home high school with enrollment at an alt school location, how many were AON for 10, 11, 12, and just non AON for 9th?

# MNPS check:
# Of students at an AON home high school with enrollment at an alt school location,
# how many were AON for 10, 11, 12, and non-AON for 9th only?

students_aon_10_11_12 <- cohort_selection %>%
  group_by(D_stu_id) %>%
  summarise(
    had_non_aon_9th  = any(!C_year_home_school %in% schools_to_include & D_year == 2022),
    had_non_aon_10th = any(!C_year_home_school %in% schools_to_include & D_year == 2023),
    had_non_aon_11th = any(!C_year_home_school %in% schools_to_include & D_year == 2024),
    had_non_aon_12th = any(!C_year_home_school %in% schools_to_include & D_year == 2025),
    was_aon_9th      = any(C_year_home_school %in% schools_to_include & D_year == 2022),
    was_aon_10th     = any(C_year_home_school %in% schools_to_include & D_year == 2023),
    was_aon_11th     = any(C_year_home_school %in% schools_to_include & D_year == 2024),
    was_aon_12th     = any(C_year_home_school %in% schools_to_include & D_year == 2025)
  ) %>%
  filter(
    had_non_aon_9th,                  # must have non-AON in 9th
    !had_non_aon_10th,                # no non-AON in 10th
    !had_non_aon_11th,                # no non-AON in 11th
    !had_non_aon_12th,                # no non-AON in 12th
    was_aon_10th & was_aon_11th & was_aon_12th,  # AON in 10th–12th
    !was_aon_9th                      # and NOT AON in 9th
  )

# summarize how many students fit that criteria
students_aon_10_11_12 %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#None was not correct, its 91 students!

#Can we give those 91 students a flag just to help track?
cohort_selection <- cohort_selection %>%
  mutate(C_aon_10_11_12_flag = ifelse(D_stu_id %in% students_aon_10_11_12$D_stu_id, "YES", "NO"))

# #great, lets check a couple of those students
# student_check_5 <- cohort_selection %>%
#   filter(D_stu_id == 190164794) %>%
#   print()
# 
# 
# student_check_6 <- cohort_selection %>%
#   filter(D_stu_id == 190172101) %>%
#   print()
# 
# 
# student_check_7 <- cohort_selection %>%
#   filter(D_stu_id == 190184979) %>%
#   print()


#check Rebel's student example-
#190173225

student_check_7 <- cohort_selection %>%
  filter(D_stu_id == 190173225) %>%
  print()

#All fixed now!

#Pert team methodology, we want to exclude all students with any enrollment at a non AON school except for those who are "YES" for C_aon_10_11_12_flag

# --- Create the flag directly on cohort_selection ---
cohort_selection <- cohort_selection %>%
  mutate(C_aon_10_11_12_flag = ifelse(D_stu_id %in% students_aon_10_11_12$D_stu_id, "YES", "NO"))

# --- If cohort_selection_filtered is supposed to be your working dataset, 
#     then re-derive it from cohort_selection AFTER you add the flag ---
cohort_selection_filtered <- cohort_selection

# Step 1: identify students with any non-AON enrollment
students_with_nonAON <- cohort_selection_filtered %>%
  filter(!C_year_home_school %in% schools_to_include) %>%
  distinct(D_stu_id) 

# Step 2: exclude all rows for those students 
# EXCEPT if they have the special flag
cohort_selection_filtered <- cohort_selection_filtered %>%
  filter(!(D_stu_id %in% students_with_nonAON$D_stu_id & 
           C_aon_10_11_12_flag != "YES"))

# Step 3: confirm the distinct count
cohort_selection_filtered %>% 
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#doublecheck Rebels student
student_check_7 <- cohort_selection_filtered %>%
  filter(D_stu_id == 190173225) %>%
  print()

#2665, great so that validates all set now!

#and very quickly- how many students will "Hillwood High" as a home school
cohort_selection_filtered %>% 
  filter(C_four_year_home_school == "Hillwood High") %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#and how many at James Lawson
cohort_selection_filtered %>% 
  filter(C_four_year_home_school == "James Lawson High School" | C_four_year_home_school == "James Lawson High") %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#okay 9

#194, okay we want to re-asign all of those students to "James Lawson High School" since they are the same school
cohort_selection_filtered <- cohort_selection_filtered %>%
  mutate(C_four_year_home_school = ifelse(C_four_year_home_school == "Hillwood High", "James Lawson High School", C_four_year_home_school))

#and now check how many at James Lawson (should be 205)
cohort_selection_filtered %>% 
  filter(C_four_year_home_school == "James Lawson High School" | C_four_year_home_school == "James Lawson High") %>%
  summarise(distinct_students = n_distinct(D_stu_id)) %>%
  print()

#and re-assign "James Lawson High" to "James Lawson High School", again theyre the same thing
cohort_selection_filtered <- cohort_selection_filtered %>%
  mutate(C_four_year_home_school = ifelse(C_four_year_home_school == "James Lawson High", "James Lawson High School", C_four_year_home_school))

#Double check unique values of C_four_year_home_school
distinct(cohort_selection_filtered, C_four_year_home_school) %>%
  print()

```



#### E. Re-test the overlap of course IDs in transcript and course catalog to re-size the issue and see how many courses may need to be re-coded
```{r}
# Step 1. Clean transcript course IDs (strip "S1"/"S2" if present)
cohort_selection <- cohort_selection %>%
  mutate(course_id_cleaned = str_replace(D_course_id, "([[:punct:]]?\\s*)?(S1|S2)\\s*$", ""))

# Step 2. Get distinct IDs from both sources
course_catalog_ids <- unique(past_course_catalog$C_course_id_clean)
cohort_ids <- unique(cohort_selection$course_id_cleaned)

# Step 3. Compare
ids_in_both       <- intersect(cohort_ids, course_catalog_ids)
ids_only_catalog  <- setdiff(course_catalog_ids, cohort_ids)
ids_only_transcript <- setdiff(cohort_ids, course_catalog_ids)

# Step 4. Counts
cat("Total unique IDs in catalog:", length(course_catalog_ids), "\n")
cat("Total unique IDs in transcript:", length(cohort_ids), "\n")
cat("IDs present in BOTH:", length(ids_in_both), "\n")
cat("IDs only in catalog:", length(ids_only_catalog), "\n")
cat("IDs only in transcript:", length(ids_only_transcript), "\n")


#Create a data-frame of the unique course IDs that are only in the transcript data, not in the course catalog
transcript_courses_not_in_catalog <- cohort_selection %>%
  ungroup() %>%  # <- important
  filter(course_id_cleaned %in% ids_only_transcript) %>%
  distinct(course_id_cleaned, D_course_name) %>%
  arrange(course_id_cleaned)

print(transcript_courses_not_in_catalog)

#great, lets export that

# ers_write_sharepoint(transcript_courses_not_in_catalog,
#                       folder_path = raw_data_folder_path,
#                       file_name_with_extension = "MNPS_Transcript_Courses_Not_in_Catalog.csv",
#                       drive_name = "client_work_drive")


```

### 2.3- Methodology 2- Sizing the catalog/transcript issue if we exclude ALL rows of alternative enrollment

-   **What:** This would just be if we wanted to look at students who ONLY had AON enrollment and no rows of wnrollment at an alertantive school, which in the end I don't think we will want to do so we're all set. Leaving code in as reference!

```{r}
# #This is just a quick test to see how different the issue might be if we were to exclude all rows at exclude schools for a students enrollment
# 
# #Step 1: Create a new data-frame called "include_only_cohort" that first just filters to students present for 4 years
# Include_only_cohort_selection <- student_transcript_rollup %>% 
#   filter(M_years_enrolled == 4) 
# 
# #then filter out any C_year_home_school that is not in the schools_to_include list
# Include_only_cohort_selection <- Include_only_cohort_selection %>%
#   filter(C_year_home_school %in% schools_to_include)
# 
# #now size the issue
# 
# # Step 1. Clean transcript course IDs (strip "S1"/"S2" if present)
# Include_only_cohort_selection <- Include_only_cohort_selection %>%
#   mutate(course_id_cleaned = str_replace(D_course_id, "([[:punct:]]?\\s*)?(S1|S2)\\s*$", ""))
# 
# # Step 2. Get distinct IDs from both sources
# course_catalog_ids <- unique(past_course_catalog$C_course_id_clean)
# Include_only_cohort_ids <- unique(Include_only_cohort_selection$course_id_cleaned)
# 
# # Step 3. Compare
# ids_in_both       <- intersect(Include_only_cohort_ids, course_catalog_ids)
# ids_only_catalog  <- setdiff(course_catalog_ids, Include_only_cohort_ids)
# ids_only_transcript <- setdiff(Include_only_cohort_ids, course_catalog_ids)
# 
# # Step 4. Counts
# cat("Total unique IDs in catalog:", length(course_catalog_ids), "\n")
# cat("Total unique IDs in transcript:", length(Include_only_cohort_ids), "\n")
# cat("IDs present in BOTH:", length(ids_in_both), "\n")
# cat("IDs only in catalog:", length(ids_only_catalog), "\n")
# cat("IDs only in transcript:", length(ids_only_transcript), "\n")


```



### 2.4 Student Grade Level

#### a. More than one grade level

-   **What:** Addressing the student grade level mismatch by taking the grade-level from the student demographic file as the source of truth

-   **Check:** Check so that we can work to develop a source of truth for student grade level

-   **Project Team Action:** None (run as is)

```{r}
#filter for rows where there are multiple distinct D_stu_grade values within the same D_year within the cohort
multiple_grades <- cohort_selection_filtered %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade_level_demo) > 1) %>%
  distinct() %>% 
  print() %>% 
  ungroup()

#none! now check transcript grade level
#filter for rows where there are multiple distinct D_stu_grade values within the same D_year within the cohort
multiple_grades <- cohort_selection_filtered %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade_level_transcript) > 1) %>%
  distinct() %>% 
  print() %>% 
  ungroup()

#how many distinct students?
multiple_grades_transcript <- cohort_selection_filtered %>% 
  group_by(D_stu_id, D_year) %>%
 filter(n_distinct(D_stu_grade_level_transcript) > 1)


print((n_distinct(multiple_grades_transcript$D_stu_id)))

#okay quite a few, which would mean that grade level between the transcript and demo file aren't matching

#get unique D_stu_id values with more than one D_stu_grade value within the same D_year
unique_multiple_grades <- unique(multiple_grades$D_stu_id)

# create a new dataframe with just these students
multiple_grades_df <- cohort_selection_filtered %>%
  filter(D_stu_id %in% unique_multiple_grades) %>% 
  print()

#count distinct students in the multiple grades dataframe
count_distinct_multiple_grades <- multiple_grades_df %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()

```

#### b. MNPS- Difference between transcript and demo grade levels

-   **What:** Investigate cases where students show a different grade level in the transcript file and student demo file for the same year

-   **Project Team Action:** Decide on source of truth for student grade level(D_stu_id)

```{r}
# Step 1: Identify students with different values of D_stu_grade_level_transcript and D_stu_grade_level_demo within the same year
cohort_selection_filtered <- cohort_selection_filtered %>%
  mutate(grade_mismatch = ifelse(D_stu_grade_level_transcript != D_stu_grade_level_demo, TRUE, FALSE))

#how many students with TRUE for grade_mismatch
mismatch_count <- cohort_selection_filtered %>%
  filter(grade_mismatch == TRUE) %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()

#75

#quick check- how many unique students in cohort?
cohort_selection_filtered %>% 
  ungroup()

print((n_distinct(cohort_selection_filtered$D_stu_id)))

#2,812


```

#### c. Pull grade level that students finished the year on

-   **What:** To create one source of truth for student grade level, so we will pull the grade level from the demo file and make a new variable for C_eoy_grade

-   **Check:** There are checks built in after each step including, checking that students don't have multiple rows with the same exit date by year

-   **Project Team Action:** Adjust as necessary for your district context and analysis.

```{r}
# Create a new variable called C_eoy_grade that groups by student ID and year and selects the grade level that the student spent the most time in (aka D_stu_grade for the student with the highest value of M_days)
cohort_selection_filtered <- cohort_selection_filtered %>%
  group_by(`D_stu_id`, `D_year`) %>%
  mutate(C_eoy_grade = D_stu_grade_level_demo) %>% 
  ungroup()


# Flag rows where the two grade level columns differ
grade_mismatches <- cohort_selection_filtered %>%
  filter(C_eoy_grade != D_stu_grade_level_demo)

# Print count and mismatches
cat("Number of student-year rows with grade mismatches:", nrow(grade_mismatches), "\n")

grade_mismatches %>% print()

#none, great


# check for how many IDs still have multiple rows for the same student ID and year with different values of grade after removing summer school

# should be zero
multiple_grades <- cohort_selection_filtered %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(C_eoy_grade) > 1) %>%
  distinct() %>%
  ungroup()

# Print result
print(multiple_grades)

# Optional: check how many problematic student-year combos exist
cat("Number of student-year combos with >1 grade level:", nrow(multiple_grades), "\n")

#great



```

#### d. Check grade level distribution of C_eoy_grade

```{r}
# Create a count of unique student IDs by year by C_eoy_grade
# Create a count of unique student IDs by year by C_eoy_grade
grade_counts_eoy <- cohort_selection_filtered %>% 
  filter(C_eoy_grade %in% c(9, 10, 11, 12)) %>% 
  group_by(D_year, C_eoy_grade) %>% 
  summarize(n = n_distinct(D_stu_id), .groups = "drop") %>% 
  arrange(D_year, C_eoy_grade) %>% 
  print()

#this is kind of an early peak at the "persistence" exclusion category for cohort. So of

grade_counts_eoy <- cohort_selection_filtered %>% 
  filter(C_eoy_grade %in% c(9, 10, 11, 12)) %>% 
  group_by(D_year, C_eoy_grade) %>% 
  summarize(n = n()) %>% 
  arrange(D_year, C_eoy_grade) %>%
  ungroup() %>% 
  print()



```


###2.5 MNPS quick check on M_days_enrolled
```{r}
# #Grouping by year, how many students would be excluded for having fewer than 274 days enrolled in a given school year (testing 75% threshold?)
# cohort_selection_filtered %>%
#   group_by(D_year) %>%
#   summarise(students_below_274_days = n_distinct(D_stu_id[M_days_enrolled < 274]),
#             total_students = n_distinct(D_stu_id),
#             pct_below_274_days = students_below_274_days / total_students * 100) %>%
#   print()
# 
# 
# #75%
# cohort_selection_filtered %>%
#   group_by(D_stu_id) %>%
#   summarise(any_below_274 = any(M_days_enrolled < 274), .groups = "drop") %>%
#   summarise(
#     students_below_274_days = n_distinct(D_stu_id[any_below_274]),
#     total_students = n_distinct(D_stu_id),
#     pct_below_274_days = students_below_274_days / total_students * 100
#   ) %>%
#   print()
# 
# #1275, 44%....
# 
# cohort_selection_filtered %>%
#   group_by(D_stu_id) %>%
#   summarise(any_below_183 = any(M_days_enrolled < 183), .groups = "drop") %>%
#   summarise(
#     students_below_183_days = n_distinct(D_stu_id[any_below_183]),
#     total_students = n_distinct(D_stu_id),
#     pct_below_183_days = students_below_183_days / total_students * 100
#   ) %>%
#   print()
# 
# 
# #1171, 40.8%
# 
# 
# #can we look at some of these student examples of students who are below?
# duration_exclusion_75 <- cohort_selection_filtered %>%
#   group_by(D_stu_id) %>%
#   filter(any(M_days_enrolled < 274)) %>%
#   ungroup() %>%
#   print()
# 
# 
# #Student Example One- 190008899
# student_check_1 <- duration_exclusion_75 %>%
#   filter(D_stu_id == "190008899") %>%
#   arrange(D_year) %>%
#   print()
# 
# #student example two
# student_check_2 <- duration_exclusion_75 %>%
#   filter(D_stu_id == "190109601") %>%
#   arrange(D_year) %>%
#   print()
# 
# 
# #student example three
# student_check_3 <- duration_exclusion_75 %>%
#   filter(D_stu_id == "190125355") %>%
#   arrange(D_year) %>%
#   print()
# 
# #This is making me realize a few things... First is that the summary of 4 years might not be totally right 
# 

#double check number of distinct students in the cohort before exporting!
n_distinct(cohort_selection_filtered$D_stu_id) %>% 
  print()

#new new NEW cohort- 2756

#great


#check infamous student 4
student_check_4 <- cohort_selection_filtered %>%
  filter(D_stu_id == 190137047) %>%
  print()

#okay so its somewhere in this file 

#Really quickly can we rename the C_multiple_years_home_school_flag to "C_transfered_across_years_flag"?
cohort_selection_filtered <- cohort_selection_filtered %>%
  rename(C_transferred_across_years_flag = C_multiple_year_home_schools_flag)

```

# Part 3: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <-  "District Partners/Nashville/Nashville - HS Pathways - 2025/1. Data & Analysis - Secure/Transcript Analysis"


ers_write_sharepoint(data = cohort_selection_filtered,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "stu_transcript_rollup_cohort_only_updated.csv",
                    drive_name = "client_work_drive")

```

## Part 4: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))

```
