---
title: "1_raw_data_transcript_validation"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: July 23 2024
review:
last update: April 10 2025
output: html_document 
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. If your data is spread out over multiple files - i.e., **one file per year, go to step 1.1**

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning=F, message=F}

raw_data_folder_path <- "District Partners/Detroit Public Schools/26-27 HS Redesign Implementation/1. Data & Analysis - Secure/Fall 2026 Transcript Investigation"

# Bring in file
# transcript_raw <- ers_read_sharepoint(folder_path = raw_data_folder_path,
#                    file_name_with_extension = "/0. Raw Data/Transcript YE23-YE26/Stored Grades YE26 - Corrected.csv")

# Clean column names to remove spaces and special characters
# transcript_raw  <- transcript_raw  %>% clean_names()

# Bring in location coding
location_coded <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                    file_name_with_extension = "/1. Coding/location_coded.xlsx")

```

## 1.1 Bring in data for Multiple Files (skip if you are working with one year of data/one file)

-   **What:** The code below grabs all csv files in a folder and stacks the data into one data frame. Use this if your data is spread out across multiple files (e.g., student_dem_2019, student_dem_2020, student_dem_2021). Otherwise, Jump to Part II

-   **Project Team Action:**

    -   Make sure that all relevant files are in one folder in SharePoint.

    -   Replace the file path below with your file path (in the code chunk below the first one)

-   **NOTES:** *This long function will most likely be replaced by an ers_tools package by end of March 2025. Check back for details!*

```{r}
#1. Run the function
ers_read_sharepoint_folder <- function(folder_path,
                                file_name_with_extension = NULL,
                                sheet_name = NULL,
                                drive_name = "client_work_drive",
                                skip_rows = 0,
                                na_values = c(""))
{
  # Clean the folder path and extract the drive name from it if not explicitly provided
  cleaned <- erstools::ers_sharepoint_path_clean(folder_path)
  folder_path <- cleaned$path_clean  # Cleaned folder path for SharePoint
  if (missing(drive_name)) {
    drive_name <- cleaned$drive_name
  }
  # Assign Drive
  orgfiles_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles") %>% suppressMessages
  external_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/external") %>% suppressMessages
  data_hub_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/ers-data") %>% suppressMessages
  client_work_drive <- orgfiles_site$get_drive("Client Work")
  internal_drive <- orgfiles_site$get_drive("Internal")
  external_drive <- external_site$get_drive("Documents")
  data_hub_drive <- data_hub_site$get_drive("Documents")
  drive <- switch(drive_name,
                  client_work_drive = client_work_drive,
                  internal_drive = internal_drive,
                  external_drive = external_drive,
                  data_hub_drive = data_hub_drive,
                  stop("Invalid drive name. Use 'client_work_drive', 'internal_drive', 'external_drive', or 'data_hub_drive'."))
  # Error message if folder path doesn't exist
  folder_item <- try(drive$get_item(folder_path), silent = TRUE)
  if (inherits(folder_item, "try-error")) {
    stop("Folder path does not exist: '", folder_path, "'")
  }
  # STEPS FOR READING IN WHOLE FOLDER OF FILES
  if (is.null(file_name_with_extension)) {
    files <- drive$list_items(folder_path)
    valid_files <- files[grepl("\\.(csv|xls|xlsx)$", files$name, ignore.case = TRUE), ]
    # Error message if no files in folder
    if (nrow(valid_files) == 0) {
      stop("No valid files (.csv, .xls, .xlsx) found in the folder.")
    }
    # Create a data list and use fread or read_excel to read in the files
    data_list <- list()
    for (file in valid_files$name) {
      file_path <- file.path(folder_path, file)
      temp_file <- tempfile(fileext = paste0(".", tools::file_ext(file)))
      drive$download_file(file_path, dest = temp_file)
      file_data <- if (tools::file_ext(file) == "csv") {
        fread(temp_file, skip = skip_rows, na.strings = na_values)
      } else {
        read_excel(temp_file, na = na_values, guess_max = 100000)
      }
      data_list[[tools::file_path_sans_ext(file)]] <- file_data
    }
    #assign dataframes to environment
    return(invisible(data_list))
  }
  #PROCESS SINGLE FILE ONLY
  file_name_with_extension <- gsub("%20", " ", file_name_with_extension)
  data_path <- file.path(folder_path, file_name_with_extension)
  #Error message if file doesn't exist
  file_item <- try(drive$get_item(data_path), silent = TRUE)
  if (inherits(file_item, "try-error")) {
    stop("File '", file_name_with_extension, "' does not exist in folder: '", folder_path, "'")
  }
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  temp_file <- tempfile(fileext = file_extension)
  drive$download_file(data_path, dest = temp_file)
  raw_data <-
    if (tools::file_ext(file_name_with_extension) == "csv") {
      fread(temp_file, skip = skip_rows, na.strings = na_values)
    } else if (tools::file_ext(file_name_with_extension) %in% c("xls", "xlsx")) {
      read_excel(temp_file, sheet = sheet_name, skip = skip_rows, na = na_values, guess_max = 100000)
    } else if (tools::file_ext(file_name_with_extension) == "shp") {
      # Create a temporary directory
      temp_dir <- tempdir()
      # Define possible extensions
      shp_components <- c("shp", "shx", "dbf", "prj", "cpg")
      # Download each component
      for (ext in shp_components) {
        file_path <- file.path(temp_dir,
                               paste0(tools::file_path_sans_ext(file_name_with_extension), ".", ext))
        drive$download_file(file.path(folder_path,
                                      paste0(tools::file_path_sans_ext(file_name_with_extension),
                                             ".", ext)), dest = file_path)
      }
    } else {
      stop("Unsupported file format. Supported formats are xls, xlsx, csv, shp: '", file_name_with_extension, "'")
    }
  return(raw_data)
}
```

```{r}
#get the list of files
data_list <- ers_read_sharepoint_folder("District Partners/Detroit Public Schools/26-27 HS Redesign Implementation/1. Data & Analysis - Secure/Fall 2026 Transcript Investigation/0. Raw Data")

# Use list2env to assign to your environment
list2env(data_list, envir = .GlobalEnv)

# Bind all tables into a single dataframe
transcript_all_yrs <- bind_rows(data_list, .id = "source_file")  # Adds a column indicating source file (optional)

# Clean column names to remove spaces and special characters
transcript_all_yrs<- transcript_all_yrs %>% clean_names()

# View combined data
ers_column_info(transcript_all_yrs)
```

## Part 2: Validate Data

### 2.1 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. See resolution of critical/important fields that are missing in step 2.2b. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present

-   **Project Team Action:** Definitely change (see guidance in code block)

```{r}
#if the ERS Standard column is present in your data then type your district's column on the right hand side (it is okay if the names are identical)
#if the ERS Standard column is not present in your data then type NA
#you can also exclude the columns you do not need for your analysis by simply not renaming those columns in this step (the mutate function with the .keep = none argument only retains columns that follow the ERS column tagging convention)


transcript_all_yrs <- transcript_all_yrs %>%
  mutate(
    # critical
    D_stu_id = dpscd_masked_student_id,
    D_course_id = course_number,
    D_course_name = course_name,
    D_location_id = eem_code,
    D_year = year_end,
    D_term = store_code,  # term code from PowerSchool
    D_teacher_id = NA,
    D_stu_grade = grade,
    D_creds_earn = credits_earned,
    D_creds_possible = potential_credits,
    #nice to have
    D_stu_grade_level = grade_level,
    D_course_section = NA,
    # D_location_name = NA, # Detroit: pulled in through location_coded
    D_course_subject = NA,
    D_course_subject_desc = NA,
    D_course_credit_recovery = NA,
    D_course_pathway = NA,
    D_cte_flag = NA,
    D_course_rigor = NA,
    .keep = "none"  # This drops any column not explicitly created above
  )

```

#### b. Review columns

-   **What:** Manually check that you have the columns you need/would expect after renaming, and have not lost any critical data. Review the columns you are missing (as necessary).

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them for later.

### 2.1 Manual Checks

-   **What:** a handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input them into a checks table in OneNote, Word, or another tool.

    -   Do we have right year(s) of data?

        -   Check: Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

    -   Does term signify years, semesters, or quarters? What is the relationship between credits and term?

        -   Check: look at the term column to determine how term is used, and how many credits are typically earned by course. For example, if there is one row per course per semester, you would expect to see that courses are worth 0.5 credits

<!-- -->

-   Are there types of courses that tend to be worth more credits?

-   **Check:** Continue to build understanding of how credits are organized. LOok for differences by types of courses. For example, CTE or practicum courses tend to be worth more credits. That is okay, right now you are just trying to understand the structure of the file

-   **Project Team Action:** Gut check/track

```{r}
#group by credits earned and ensure that values match course notation (ie. semester courses carry .5, etc.)
transcript_all_yrs %>%
  group_by(D_creds_earn, D_term) %>%
  summarize(n = n_distinct((D_course_name))) %>% 
  pivot_wider(D_creds_earn, names_from = D_term, values_from = n) %>%
  ungroup()

```

### 2.2 Understand File Structure

#### a. Number of Total Rows

-   **What:** Understand the structure of your file by loading the number of rows

-   **Check:** The number of rows depends on how many courses you expect the student to have per year, which is explained by how many terms/periods/rotations are covered in the data. For example, if classes are organized by semester across 8 periods and repeat daily, you expect to see the number of rows be equal to 2 \* 8 \* 1 \* count of high school students. If classes are organized by quarter across 7 periods and repeat daily, you expect to see the number of rows be equal to 4 \* 7 \* 1 \* count of high school students

-   **Project Team Action:** update the Data Validation Checks table with the result of your check. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY total high school students per year and classes happen on a X term / X rotation / X period format"

```{r}
# Count number of rows
transcript_all_yrs %>% nrow() #View output of check to know what to write for your status column
```

#### b. Remove Rows

-   **What:** If any of the rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one, your data has any totally or almost totally blank rows, etc.

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before.

-   **Project Team Action:** add code as required

```{r}
#Add code/comments here if resolution is required

#remove rows
# Remove rows where a specific condition is met
# For example, remove rows where age is less than 18

# transcript_all_yrs <- transcript_all_yrs %>%
#   filter(age >= 18)
# 
# # Print the result
# print(head(transcript_all_yrs))

```

```{r}
# Detroit: Merge in location names
location_coded <- location_coded |>
  mutate(D_location_id = str_pad(D_location_id, width = 5, side = "left", pad = "0"))

transcript_all_yrs <- transcript_all_yrs %>%
  left_join(
    location_coded %>%
      select(D_location_id, D_location_name),
    by = "D_location_id")

# Check
transcript_all_yrs |> 
  group_by(D_location_name, D_location_id) |>
  summarise(n = n())
```

```{r}
# Detroit: Filter records with missing location name
transcript_all_yrs <- transcript_all_yrs |>
  filter(!is.na(D_location_name))
```


#### c. Review current columns

-   **What:** View the data as a table to understand the values in each column

-   **Check:** A new tab is created with a table view of your data

-   **Project Team Action:** Do not change (run as is)

```{r}
#move data viewer to another tab to view both at once. Click on the "show in new window" button in the top left of the data viewer tab (just to the right of the left and right arrows)
transcript_all_yrs %>% View
```

### 2.5 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting

```{r}
#View duplicates
transcript_all_yrs %>%
  group_by_all() %>%
 mutate(n = n()) %>%
    filter(n>1) %>% 
    ungroup()
  
```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates.

-   **Check:** That the number of duplicates is now 0.

-   **Project Team Action:** None

```{r}
#removes duplicates
transcript_all_yrs <- transcript_all_yrs[!duplicated(transcript_all_yrs), ]

```

### 2.6 Missing & Unique values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action:** None (run as is)

```{r}
# Update blanks or NULL to NA
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** (1) Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates

```{r}
# Create new data frame with column information
col_info <-  data.frame(
  Column_Name = names(transcript_all_yrs),
  Data_Type = sapply(transcript_all_yrs, class),
  Count_Rows = sapply(transcript_all_yrs, function(x) length(na.omit(x))),
  NA_Rows = sapply(transcript_all_yrs, function(x) sum(is.na(x))),
  Unique_Rows = sapply(transcript_all_yrs, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)

# Calc % missing
col_info <- col_info %>%
            mutate(Percent_Missing = NA_Rows / Count_Rows * 100)

# Print new data frame
col_info

# Same thing as above, but instead of NA values, look at 0 values
# Create new data frame with column information
col_info_zero <-  data.frame(
  Column_Name = names(transcript_all_yrs),
  Data_Type = sapply(transcript_all_yrs, class),
  Count_Zero = sapply(transcript_all_yrs, function(x) sum(x == 0)), # Count zero values
  Total_Rows = nrow(transcript_all_yrs), # Total number of rows in transcript_combined_file
  stringsAsFactors = FALSE,
  row.names= NULL
)

col_info_zero <- col_info_zero %>%
  mutate(Percent_Zero = Count_Zero / Total_Rows * 100)

# Count of zero values of section ID by location name 
transcript_all_yrs %>%
  group_by(D_location_name) %>%
  summarize(
    zero_count = sum(D_course_section == 0)
  ) %>% 
   ungroup()


# Calc % missing
col_info <- col_info %>%
            mutate(Percent_Missing = NA_Rows / Count_Rows * 100)

# Print new data frame
print(col_info)


# In the chunk for 2.6b, after col_info is created and printed:
ggplot(col_info, aes(x = reorder(Column_Name, Percent_Missing), y = Percent_Missing)) +
  geom_bar(stat = "identity") +
  coord_flip() + # Makes it easier to read column names
  labs(title = "Percentage of Missing Values per Column",
       x = "Column Name",
       y = "Percent Missing") +
  theme_minimal()

# Similarly for col_info_zero:
ggplot(col_info_zero, aes(x = reorder(Column_Name, Percent_Zero), y = Percent_Zero)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Percentage of Zero Values per Column",
       x = "Column Name",
       y = "Percent Zero") +
  theme_minimal()
```

#### c. Review missing data for high schools / schools of interest

-   **What:** Understand if we have any blanks or NAs in key fields for schools of interest (it is possible that the district provided data for other locations that aren't actually schools and do not have section IDs). If you don't have or need section IDs, you can skip this step.

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like student ID and D_location match expectations?

-   **Project Team Action:** (1) Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates

```{r}
# Investigating school codes with the most section values
# Group by D_location_id and count unique D_course_section values within each group
transcript_all_yrs %>%
  group_by(D_location_id) %>%
  summarise(section_count = n_distinct(D_course_section))

```

#### d. List the counts of students per course and counts of students missing sections for each course

-   **What:** Investigating students who are missing section IDs in schools of interest further. With the printed result. Filtered / sorted to look at the courses with the highest % of students missing sections. See if the students who weren't assigned sections were in really small sections where there would only be one section.If you don't have or need section IDs, you can skip this step.

-   **Check:** Are there any irregularities or disparities in enrollment that would cause concern for the integrity of the data?

-   **Project Team Action:** Follow up as necessary

```{r}
# Group the data by D_location_id and D_course_name, then summarize to get counts

# If course ID is zero, set it to NA
transcript_all_yrs <- transcript_all_yrs %>%
  mutate(D_course_id = ifelse(D_course_id == 0, NA, D_course_id))

result_missing_sections <- transcript_all_yrs %>%
  group_by(D_year, D_location_id, D_course_name) %>%
  summarize(
    count_students_per_course = n(),
    count_students_missing_sections = sum(is.na(D_course_section))
  ) %>%
  ungroup() %>%
  mutate(
    count_students_per_course = ifelse(is.na(count_students_per_course), 0, count_students_per_course),
    count_students_missing_sections = ifelse(is.na(count_students_missing_sections), 0,     
    count_students_missing_sections),
    percent_missing_sections = (count_students_missing_sections / count_students_per_course) * 100
  )

# Print the result
print(result_missing_sections)
```

#### 2.8. Counts by Terms and Course

-   **What:** Review the number of rows by term to ensure that we have data for the whole entire year.

-   **Check:** Are there any irregularities or disparities in term or course counts that may indicate missing data?

-   **Project Team Action:** Follow up as necessary

```{r}
# Group by term and count the number of unique values in that column
transcript_all_yrs %>%
  group_by(D_term) %>%
  summarize(n = n()) %>% 
  ungroup()

# Group by course name and list the unique names of each course. Are they seperated out by semesters?
transcript_all_yrs %>%
  group_by(D_course_name) %>%
  summarize(n = n()) %>% 
  ungroup()

# Group by course name and list the unique names of each course. Are they seperated out by semesters?
transcript_all_yrs %>%
  group_by(D_course_name, D_term) %>%
  summarize(n = n()) %>% 
  pivot_wider(D_course_name,names_from = D_term,values_from = n) %>%
  ungroup()

```

### 2.7 Student counts by school by year

-   **What:** Count the number of students at each school per year to build confidence in data

-   **Check:** Are there any irregularities or disparities in enrollment or schools that would cause concern for the integrity of the data?

-   **Project Team Action:** Follow up as necessary

```{r}
# Count the number of students (D_stu_id) at each school (D_location_id) per year (D_year)
students_count_by_school_year <- transcript_all_yrs %>%
  group_by(D_location_id, D_year) %>%
  summarise(
    students_count = n_distinct(D_stu_id)
  ) %>% 
  ungroup()

# Using pivot wider, print the results to view number of students by year in a simple table
students_count_by_school_year <- students_count_by_school_year %>%
  pivot_wider(names_from = D_year, values_from = students_count) %>% 
  print()

```

### 2.8 Distinct count of students by location id

-   **What:** Count of students at each school. This check might helpful confirm any exclude/include schools for your teams analysis (like alternative schools).

-   **Check:** Check across other files while you validate them, and if necessary exclude schools with low student counts.

-   **Project Action:** Gut check/track

```{r}
# Calculate the total count of distinct D_stu_id across all locations
transcript_all_yrs %>%
  summarise(
   total_students_count = n_distinct(D_stu_id)
  ) 

# Count the number of students (D_stu_id) at each location D_location_id)
transcript_all_yrs %>%
  group_by(D_location_id, D_location_name) %>%
  summarise(
    students_count = n_distinct(D_stu_id)
  ) %>% 
  arrange(desc(students_count)) %>% 
  ungroup()

```

### 2.9 Distribution of students by Term

-   **What:** Count of records by term at each school. Given your project needs, you may want to just filter your original data-frame (transcript_combined_file) to only include high schools of interest for this step. If you are looking at all locations, you can skip this step.

-   **Check:** Review the tables to understand if there is an even distribution of semester 1 and semester 2 classes

-   **Project Action:** Update the numbers in the first line of code to match the location codes in your district's data

```{r}
# Filtering only to high schools of interest for this step, but you might choose to look at distribution of students by term for all locations based on project needs

# Detroit: already filtered for neighborhood high schools, so look at all locations
codes <- unique(transcript_all_yrs$D_location_id)
filtered_high_schools <- transcript_all_yrs[transcript_all_yrs$D_location_id %in% codes]

# View count of records by D_term/school
filtered_high_schools %>%
  filter(!is.na(D_term)) %>%
  group_by(D_location_id, D_term) %>% 
  count(sort = TRUE) %>%
  arrange(D_location_id) %>%
  group_by(D_location_id) %>% 
  mutate(perc_D_term_at_school = round(100 * n / sum(n), 1)) %>%
  select(D_location_id, D_term, perc_D_term_at_school) %>%
  pivot_wider(names_from = D_term, 
              values_from = perc_D_term_at_school, 
              names_prefix = "D_term_") %>%
  rename_all(~ ifelse(. == "", "D_term_unknown", .)) %>% 
  ungroup()

filtered_high_schools %>% 
  group_by(D_location_id) %>%
  mutate(student_count_school = n_distinct(D_stu_id)) %>%
  group_by(D_term, D_location_id, .add = TRUE) %>%
  summarize(
    student_count = n_distinct(D_stu_id),
    student_count_school = first(student_count_school),
    percent_students = round(100 * n_distinct(D_stu_id) / first(student_count_school), 0)
  ) %>%
  ungroup() %>%
  arrange(D_location_id, D_term)

```

### 2.10 Distinct count of course IDs by student id

-   **What:** Count of course IDs by student id. If you did not filter any high schools in step 2.9, you may need to change the name of your data-frame.

-   **Check:** Review the counts courses by student id. You would expect to a mix since you have freshmen, sophomore, junior, and senior data across five years. You would also expect to see a maximum of \~# of classes students can take per year \*4

-   **Project Action:** No updates

```{r}
course_bucket_summary <- transcript_all_yrs %>% 
  group_by(D_stu_id) %>%
  summarize(Course_Count = n_distinct(D_course_name), .groups = "drop") %>% 
  mutate(Course_Count_Bucket = case_when(
    Course_Count == 0 ~ "00",
    Course_Count <= 2 ~ "01-02",
    Course_Count <= 4 ~ "03-04",
    Course_Count <= 6 ~ "05-06",
    Course_Count <= 8 ~ "07-08",
    Course_Count <= 10 ~ "09-10",
    Course_Count <= 12 ~ "11-12",
    Course_Count <= 16 ~ "13-16",
    Course_Count <= 20 ~ "17-20",
    Course_Count <= 24 ~ "21-24",
    Course_Count <= 28 ~ "25-28",
    Course_Count <= 32 ~ "29-32",
    Course_Count > 32 ~ "33+"
  ))

# Count number of students in each bucket
bucket_counts <- course_bucket_summary %>%
  count(Course_Count_Bucket, name = "n_students") %>%
  mutate(total_students = sum(n_students),
         pct_of_total = round(100 * n_students / total_students, 2))

print(bucket_counts)

# Bucket_counts is viewed:
ggplot(bucket_counts, aes(x = Course_Count_Bucket, y = n_students)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Students by Number of Courses Taken",
       x = "Number of Courses Bucket",
       y = "Number of Students") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels if crowded

```

### 2.11 Distinct count of students per Section

-   **What:** Returns the number of students per section per school.If you do not have section IDs in your data or need them for analysis, you can skip this step.

-   **Check:** Visually investigate the resulting enrollment_count_by_location_id table. You are looking for wonky things - really big core sections, really small PE sections. We want to ensure we have complete data. Sort count of students per section from highest to lowest. Which courses have the largest sections? Which courses have the smallest sections?

-   **Project Action:** None

```{r}
# Calculate the total number of students for each location
filtered_high_schools %>%
  group_by(D_location_id) %>%
  summarise(total_students = n_distinct(D_stu_id)) %>% 
  ungroup()


# Calculate the number of students enrolled in each course and section at each location
filtered_high_schools %>%
  group_by(D_location_id, D_course_name, D_course_section, D_year) %>%
  summarise(
    students_enrolled = n_distinct(D_stu_id)
  ) %>%
  ungroup() 


## Count the number of rows where course_section is equal to zero. 
filtered_high_schools %>% 
  filter(D_course_section == 0) %>% 
  count()

# Calculate the percent of this out of total rows
filtered_high_schools %>% 
  filter(D_course_section == 0) %>% 
  count() / nrow(filtered_high_schools)

# Write a table with the count of values of D_course_section that equal zero by school_year
filtered_high_schools %>% 
  filter(D_course_section == 0) %>% 
  group_by(D_year) %>% 
  count()

```

# Part 3: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
# processed_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

ers_write_sharepoint(data = transcript_all_yrs,
                     folder_path = raw_data_folder_path,
                     file_name_with_extension = "/2. Processed Data/01_transcript_validated_23_26.csv")
#                    drive_name = "internal_drive")
```

## Part 4: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval=FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
