---
title: "3_student_cohort_selection"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: August 12 2024
review:
last update: April 14 2025
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
gc()
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in the student and transcript data that you will need to select your snapshot of students.

-   **Project Team Action:** Replace folder path with the folder path where your files are and replace specific files names as necessary.

```{r, warning = F, message = F}

raw_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

#bring in file
validated_student_data <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "HPS_stu_crossvalidated.csv",
  drive_name = "internal_drive")


#bring in file
transcript_roll_up <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "HPS_transcript_crossvalidated.csv",
  drive_name = "internal_drive")


```

# Part II: Choosing Cohort of Students

-   **What:** The following code helps you establish variables you will need to identify a "cohort" of students, or 4 year group of students started as freshman in one school year and graduating 4 years later. Your team may have specific constraints within this methodology (ie. students need to progress or "persist" traditionally, matriculating from 9th to 10th, etc.), but the following code is meant to apply to all teams. You will have to change the years based on which specific cohort or cohorts of students you want to select. For example, a team may want to select students who started as 9th graders in SY 2020 and graduated in SY 2023, or a team may want to select students who started as 9th graders in SY 2018 and graduated in SY 2021.The following code is set to students starting as 9th grades in 2021.

-   **Project Team Action:** Update years as necessary for your specific analysis cohort

###2.1 Create Student Cohort Data Frame

####a. Student counts before filtering

-   **What:** If you have not already, you should read through the Tettra pages outlining student cohort slection here: <https://app.tettra.co/teams/ersknowledge/pages/wip-deep-dive-cohort-analysis>. The following code contains validation checks before filtering, including number of students, number of high schoolers, and number of students by location for included schools

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Updated based on year/specific snapshot group of students you want to select. Gut check/track validation results with team.

```{r}
#Starting with your 9th grade students (this example uses 2021), check how many total students will "start" the cohort
validated_student_data %>% 
  filter(D_year == 2021 & D_stu_grade == "9") %>% 
  distinct() 

#count the number of students by specific locations
validated_student_data %>% 
  group_by(D_location_name) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  arrange(desc(count_students)) %>% 
  ungroup()

#count the number of 12th graders in 2024 (who will be your cohort graduates)
validated_student_data %>% 
  filter(D_year == 2024 & D_stu_grade == "12") %>% 
  distinct() 

```

#### b. Check student enter and exit dates

-   **What:** Convert enter and exit dates to be as dates (if they aren't already), and check for any students with more than one row

-   **Check:** Check counts before and after, make sure code block prints without error

-   **Project Team Action:** Gut check/track

```{r}
# convert enter and exit date to be as dates
validated_student_data <- validated_student_data %>%
  mutate(D_stu_enter_date = as.Date(D_stu_enter_date),
         D_stu_exit_date = as.Date(D_stu_exit_date))

#check student enter dates
validated_student_data %>% 
  group_by(D_stu_enter_date) %>% 
  count() %>% 
  arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
  ungroup()

# Check student exit dates
validated_student_data %>% 
  group_by(D_stu_exit_date) %>% 
  count() %>% 
  arrange(desc(n)) %>%  # Arrange by count 'n' in descending order
  ungroup()

```

#### c. Creating Exclusion Critera, Location, Duration, Persistence

-   **What:** There are three major exclusion criteria and filters for transcript analysis: Location (the school the student attended), Duration (how long they were in the school per year), and Persistence (If they matriculated traditionally from one grade to the next.) These steps may vary based on specific team methodology and needs. For some districts, ERS has used a 75% criteria, where students need to be present 75% of the year (in each school year). In other districts, that has been lowered to 10%. Other districts have ignored entry and exit dates all together and used credits attained as the determination for selecting students. You should outline with your team the options for selecting a cohort, and move forward with the best option based on your specific analytic questions. The code below shows examples of selecting students based on some of these methodologies.

####i. Location (Home School)

-   **What:** The first exclusion criteria for cohort selection is home school- asigning a student one location ID based on the location ID they spent the most time at. This is done by creating a new column that calculates the duration of stay at each school, and then filtering to the maximum duration of stay for each student. Then, you can easily filter to schools of interest for your analysis. Similar to snapshot, the cohort home school assignment is based on whatever school a student spent the majority of their time enrolled in the district at (so number of years when looking specifically at cohort students.)

-   **Check:** Make sure no student has more than 1 home school, and that you have cleared and checked this methodology with your team.

```{r}
# Step-by-step version with safe_left_join at the end
student_home_school <- validated_student_data %>%
  group_by(D_stu_id) %>% 
  # Step 1: Calculate number of days per row (entry-exit pair)
  mutate(H_stay_duration = as.numeric(D_stu_exit_date - D_stu_enter_date)) %>%

  # Step 2: Summarize total stay duration per student-location combo
  group_by(D_stu_id, D_location_id) %>%
  summarize(H_total_days_at_location = sum(H_stay_duration, na.rm = TRUE), .groups = "drop") %>%

  # Step 3: For each student, find the location with the most days (break ties randomly)
  group_by(D_stu_id) %>%
  slice_max(order_by = H_total_days_at_location, n = 1, with_ties = TRUE) %>%  # keeps all ties
  slice_sample(n = 1) %>%  # randomly pick one if tied
  ungroup() %>%

  # Step 4: Rename the chosen location as the "home school"
  rename(H_home_school = D_location_id) %>%
  select(D_stu_id, H_home_school)

# Step 5: Use safe_left_join to merge back onto the original dataset
validated_student_data <- safe_left_join(validated_student_data, student_home_school, by = "D_stu_id")

# Step 6: Add back in home school name
validated_student_data <- validated_student_data %>%
  mutate(C_home_school_name = case_when(
    H_home_school == "1" ~ "Waikea High School",
    H_home_school == "2" ~ "Hilo High School",
    H_home_school == "3" ~ "Konaeawena HS",
    H_home_school == "4" ~ "Kamehameha High",
    H_home_school == "5" ~ "Hawaii Preparatory Academy",
    TRUE ~ "Other"
  ))

# count total students in the cohort
validated_student_data %>% 
  summarize(count_students = n_distinct(D_stu_id))

#Get student counts by grade
validated_student_data %>% 
  group_by(C_eoy_grade) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#count total students by home school
validated_student_data %>% 
  group_by(C_home_school_name) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup()

#If necessary, filter to only home schools of interest/filter out excluded locations (skip this if you already have a filtered data set)
# validated_student_data <- validated_student_data %>% 
#   filter(C_home_school_code == "1"| C_home_school_code == "29"| C_home_school_code == "39")


```

#### ii. Duration

-   **What:** Filtering to students who spent a certain amount of time at their home school in each year. This can be done in a number of ways, check with your team on the appropriate methodology. The example below shows a 75% threshold, but this can be adjusted based on your needs.

-   **NOTE:** If your team is using a credit attempt cutoff (see Tettra for examples and guidance), skip this step and see option step in part 3.1 C.

```{r}

# Step 1: Calculate total stay duration per student per year
student_duration_flags <- validated_student_data %>%
  group_by(D_stu_id, D_year) %>%
  summarize(H_sum_stay_duration_year = sum(M_days_enrolled, na.rm = TRUE), .groups = "drop") %>%
  mutate(duration_flag = if_else(H_sum_stay_duration_year >= 0.75 * 180, "include", "exclude")) %>%
  select(-H_sum_stay_duration_year)

# Step 2: Pivot to create one column per year
student_flags_wide <- student_duration_flags %>%
  pivot_wider(
    names_from = D_year,
    values_from = duration_flag,
    names_prefix = "duration_"
  )

# Step 3: Safe join back to original data
validated_student_data <- safe_left_join(validated_student_data, student_flags_wide, by = "D_stu_id")


#count of students by include/exclude flag
student_flags_wide %>%
  pivot_longer(
    cols = starts_with("duration_"),
    names_to = "year",
    names_prefix = "duration_",
    values_to = "flag"
  ) %>%
  count(year, flag) %>%
  arrange(year, flag)

#make sure no student has both include and exclude flag
validated_student_data %>% 
  group_by(D_stu_id) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  filter(count_students > 1) 

#create a summary column that looks at the duration flags for each year, and if a student is present for each of the years you care about for your analysis (ie. 2021-2024 for this example), marks them as "include" for a new column called "H_cohort_duration_include"
validated_student_data <- validated_student_data %>%
  mutate(
    H_cohort_duration_include = case_when(
      duration_2021 == "include" & duration_2022 == "include" & duration_2023 == "include" & duration_2024 == "include" ~ "include",
      TRUE ~ "exclude"
    )
  )

#count the number of distinct students who meet this criteria
validated_student_data %>% 
  filter(H_cohort_duration_include == "include") %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()


```

### d. Indivudal Student Checks

```{r}
#check a few student ids
# ---- Function: Check Random Students ----
check_random_students <- function(data, student_id_col = "D_stu_id", n = 5, seed = NULL) {
  # Optional seed for reproducibility
  if (!is.null(seed)) set.seed(seed)
  
  # Sample n unique student IDs
  random_ids <- sample(unique(data[[student_id_col]]), n)

  # Filter and return rows for those students, grouped for easy viewing
  data %>%
    filter(.data[[student_id_col]] %in% random_ids) %>%
    arrange(.data[[student_id_col]]) %>%
    group_by(.data[[student_id_col]])
}


# Check 5 random students
check_random_students(validated_student_data, n = 5, seed = 123)

```

### e. Persistence

-   **What:** The persistence criteria is to check that students are progressing traditionally from freshman to sophomore to junior to senior year across 4 years, because in the cohort analysis we are hoping to understand the patterns of "traditional" students. Some teams have modified this based on their needs, such as making the criteria just include starting 9th grade and finishing 12th grade within 4 years (meaning repeating a grade as long as you caught up was not a problem). The methodology below shows the traditional progression criteria, but this can be modified based on your needs.

```{r}
# Step 1: Create a new column for each year indicating if the student is in the expected grade
validated_student_data <- validated_student_data %>%
  mutate(
    H_stu_progression_2021 = if_else(D_stu_grade == "9" & D_year == 2021, "include", "exclude"),
    H_stu_progression_2022 = if_else(D_stu_grade == "10" & D_year == 2022, "include", "exclude"),
    H_stu_progression_2023 = if_else(D_stu_grade == "11" & D_year == 2023, "include", "exclude"),
    H_stu_progression_2024 = if_else(D_stu_grade == "12" & D_year == 2024, "include", "exclude")
  )


#check counts of students in these groups. At this point, these are not yet "cohort" students
validated_student_data %>%
  pivot_longer(
    cols = starts_with("H_stu_progression_"),
    names_to = "year",
    names_prefix = "H_stu_progression_",
    values_to = "flag"
  ) %>%
  distinct(D_stu_id, year, flag) %>%
  count(year, flag) %>%
  arrange(year, flag)

# Step 2: Create a new column that checks if the student is in the expected grade for all years
validated_student_data <- validated_student_data %>%
  mutate(
    H_cohort_progression_include = case_when(
      H_stu_progression_2021 == "include" & 
      H_stu_progression_2022 == "include" & 
      H_stu_progression_2023 == "include" & 
      H_stu_progression_2024 == "include" ~ "include",
      TRUE ~ "exclude"
    )
  )

#count the number of students included and excluded (NOTE- we have experienced a wide range by district on this, so be sure to check with your team on the expected number of students.)
validated_student_data %>% 
  filter(H_cohort_progression_include == "include") %>% 
  summarize(count_students = n_distinct(D_stu_id))

```

### f. Filter to cohort by critera

-   **What:** The final step is to filter the data to just the students you want to include in your analysis. This is done by filtering to students who meet the location, duration, and presistence critera you have established.

```{r}
#If necessary, filter to just specific home-school locations
# student_data <- student_data %>%
#   filter(C_home_school_code == "1"| C_home_school_code == "29"| C_home_school_code == "39")


#Filter to students who met the duration criteria, or were present in the district for 75% of the year for each year of your analysis
validated_student_data <- validated_student_data %>%
  filter(H_cohort_duration_include == "include")

#Filter to students who met the persistence criteria, or progressed traditionally from 9th to 12th grade in each of the years of your analysis
validated_student_data <- validated_student_data %>%
  filter(H_cohort_progression_include == "include")

#Count the new unique number of students, which should now be your cohort number
validated_student_data %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

#Check the count by year, which should be the same number in each year
validated_student_data %>% 
  group_by(D_year) %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup() %>% 
  print()

#Drop the helper columns for location, duration, and persistence 
validated_student_data <- validated_student_data %>%
  select(-starts_with("H_stay_duration"),
         -starts_with("H_stu_progression"),
         -starts_with("duration_"))

```

# Part III: Create transcript data frame

## 3.1 Merge Transcript and Student Cohort Data Frames

-   **What:** Return & remove student ids that exist in cohort file but not transcript file, and vice versa; this will make merging easier later on

-   **Check:** Number of rows at end is as expected

-   **Project Team Action:** Gut check/track

```{r}
#get the IDs that are in the student cohort file but not in the transcript file, should be a minimal number
stu_ids_missing <- setdiff(validated_student_data$D_stu_id, transcript_roll_up$D_stu_id) %>% print()

#are these students we would have otherwise included in cohort
included_students <- validated_student_data %>%
  filter(D_stu_id %in% stu_ids_missing & C_stu_cohort == "include") %>%
  pull(D_stu_id) %>% 
  print()

#remove any student IDs from the transcript file that we do not have student data for. NOTE- this will also remove any data for students not included in the cohort from the transcript file, so if you want to do any analysis on those students or students overall, you will have to use just the transcript data frame!
transcript_roll_up <- transcript_roll_up %>%
  filter(D_stu_id %in% validated_student_data$D_stu_id)

#count distinct students to make sure number validates
transcript_roll_up %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

#Using safe-left join, join the student data into the transcript data by student ID and year
transcript_roll_up <- transcript_roll_up %>%
  safe_left_join(validated_student_data, by = c("D_stu_id", "D_year"))

#Count the number of students, should be your "cohort" number
transcript_roll_up %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

```

### 3.2 File Cleaning

-   **What:** The following code cleans the transcript data frame to remove any columns that are not necessary for analysis. This is a good time to check in with your team on what columns are necessary for your analysis, and remove any unnecessary columns.

```{r}
#Rename the data-frame
merged_transcript_cohort <- transcript_roll_up

#Remove any columns that are not necessary for analysis (update as necessary for your team). These might be columns you didn't remove before, any district specific columns or checks you added or no longer need, or you might slip this step all together
merged_transcript_cohort <- merged_transcript_cohort %>%
  select(-starts_with("D_stu_enter_date"),
         -starts_with("D_stu_exit_date"),
         -starts_with("H_stay_duration"),
         -starts_with("H_stu_progression")
    )

```

# Part IV: Export Tables / Files

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/District%20Partners/Columbus%20City/22-23%20Financial%20Equity%20&%20Funding%20System%20Design/1.%20Data%20&%20Analysis%20-%20Secure/Processed%20Data%20for%20R"

ers_write_sharepoint(data = merged_transcript_cohort,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "merged_transcript_cohort.csv")
```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
