---
title: "Course Schedule Pipeline"
date: today
params:
  district: "nashville"
  refresh_raw: false      # TRUE re-reads every source from SharePoint; FALSE uses cached .rds files
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE)
# params exists only during a Quarto render. Running chunks interactively in
# Positron has no params object, so default it here and interactive runs
# behave the same as a render.
if (!exists("params")) {params <- list(district = "nashville", refresh_raw = FALSE)}
config <- yaml::read_yaml(here::here("config", "districts", paste0("cs_", params$district, ".yaml")))
source(here::here(config$helpers_path, "00_cs_setup.R"))
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue("\n## {config$district} · {config$year}\n\n"))
```

# What this file is

One document, run top to bottom, replacing `01_cs_explore.qmd` and
`02_cs_processing_and_rollups.qmd`. Those two were joined by a file —
`01_cs_data_post_explore.csv` was written to SharePoint at the end of the first
and read back at the start of the second. That seam is gone; the data stays in
memory.

The engine is unchanged. Every weight, explosion, flag, resolution and rollup
still comes from `templates/course_schedule/helpers/`.

**Two ways of checking, both reporting to one log.**

- `check_gt()` renders a table with a PASS / CHECK badge. Most checks are this:
  evidence to eyeball.
- `chk_stop()` halts. Reserved for structural failures — a join that changed
  the row count, an engine that returned nothing.

Both now write to a single check log, printed at the end, so a long render
finishes with a summary rather than a badge you had to have been watching for.

**No namespace prefixes.** Packages are attached in `00_cs_setup.R`, tidyverse
last so dplyr wins any collision with data.table. If you add a package, add it
there rather than reaching for `pkg::` inline.

**On `tr_safe()`.** It swallows any error and returns a fallback. That is right
for an input the district may not have supplied and wrong for the engine, where
it turned a failure into `NULL` and let every later stage degrade quietly. It is
now `cs_optional()` for the former and `cs_require()` for the latter.

| # | Stage | Replaces | Status |
|---|-------|----------|--------|
| 00 | Setup and config | Part 0 of both files | built |
| 01 | Explore and prep | `01_cs_explore.qmd` | built |
| 02 | Coding and exclusions | `02_...` §02 | built |
| 03 | Weights, minutes, DB resolution | `02_...` §03 | built |
| 04 | Rollups and export | `02_...` §04 | built |

---

# Stage 00 — Setup and configuration

```{r cfg}
chk_reset()
chk_stage("Stage 00 — Setup")

chk_info("District", config$district)
chk_info("Year", config$year)
chk_info("Time format", config$time_format)
chk_info("Enrollment type", config$intake$enrollment_type)
chk_info("Class ID recipe", paste(unlist(config$class_id$recipe), collapse = " + "))
chk_info("Cache", cache_dir(config))

cs_set_run_through(config)
```

## Config readiness

The config is filled in over several sittings, because you cannot write a
column map before you have seen the columns or choose a teacher-HR mode before
you know what the HR file contains. This says which tiers are done.

**Stage 01 runs no matter what is blank.** Later tiers gate the stages that
need them.

```{r cfg-readiness}
explore_gt(config_readiness(config), "Config readiness")
```

Config values the run depends on, printed so a reviewer can see the decisions
without reading any code.

```{r cfg-decisions}
tibble(
  setting = c("Term weights", "Rotation default", "Period default",
              "Bell minutes default", "Coding join key",
              "Uncoded threshold", "Teacher DB consolidate max",
              "Self-contained threshold"),
  value = c(
    paste(names(config$weights$term), unlist(config$weights$term),
          sep = "=", collapse = ", "),
    as.character(config$weights$rotation$default),
    as.character(config$weights$period$default),
    as.character(config$bell_minutes$default),
    paste(unlist(config$coding$join_key), collapse = " + "),
    paste0(config$coding$uncoded_flag_threshold_pct, "%"),
    as.character(config$teacher_db$consolidate_max_size),
    paste0(config$self_contained$swd_ell_threshold_pct, "%")))
```

```{r cfg-guard}
required <- c("district", "year", "sources",
              "column_map", "class_id", "time_format", "weights",
              "bell_minutes", "coding", "exclusions")
missing_keys <- required[!vapply(required, function(k) !is.null(config[[k]]), logical(1))]

chk_stop("Config has every required section",
         length(missing_keys) == 0,
         detail = if (length(missing_keys) > 0)
           paste("Missing:", paste(missing_keys, collapse = ", ")))

# Assign first, then check. An assignment inside chk_stop()'s condition would
# bind inside that function's frame, not here, and the object would vanish.
spec         <- cs_require(expr_spec(config),         "Expression spec builds")
term_expands <- cs_require(term_expands_spec(config), "Term expansion spec builds")
```

---

# Stage 01 — Explore and prep

Replaces `01_cs_explore.qmd`. Same profiling and the same prep recipe; what has
gone is the export at the end, since Stage 02 now runs in the same session.

## 1.1 Load sources

Each file is read from SharePoint once and cached as `.rds`. **The cache does
not know when the district re-sends a file** — set `refresh_raw: true` in the
header when new data lands.

The course schedule is required. Everything else is optional: a district that
has not sent demographics or HR data should still get a render, with the gap
reported rather than an error.

```{r load}
chk_stage("Stage 01 — Explore and prep")

cs_raw <- cs_require(
  read_source("course_schedule", config, refresh = params$refresh_raw),
  "Course schedule loaded")

stu_demographics <- cs_optional(
  read_source("demographics",    config, refresh = params$refresh_raw), "demographics")
teacher_hr <- cs_optional(
  read_source("teacher_hr",      config, refresh = params$refresh_raw), "teacher HR")
course_grades_raw <- cs_optional(
  read_source("course_grades",   config, refresh = params$refresh_raw), "course grades")
location_coding <- cs_optional(
  read_source("location_coding", config, refresh = params$refresh_raw), "location coding")
cs_course_coding <- cs_optional(
  read_source("course_coding",   config, refresh = params$refresh_raw), "course coding")
```

## 1.2 Raw column profile

Truly raw — original headers, nothing mapped. This is what the `column_map` is
written against.

```{r raw-profile}
explore_gt(explore_profile(cs_raw), "Course schedule · raw column profile")
```

```{r raw-profile-dem}
if (!is.null(stu_demographics))
  explore_gt(explore_profile(stu_demographics), "Student demographics · raw")
```

```{r raw-profile-hr}
if (!is.null(teacher_hr))
  explore_gt(explore_profile(teacher_hr), "Teacher HR · raw")
```

```{r raw-profile-grades}
if (!is.null(course_grades_raw))
  explore_gt(explore_profile(course_grades_raw), "Course grades · raw")
```

### Add-on files get renamed here too

The HR column map used to sit inside `add_ons`, where nothing read it — the
file went to the add-on with its original headers and the helper looked for a
`D_` column that had never been created. Both add-on files now rename on the
same path as everything else, which is what makes them visible during
exploration rather than at Stage 04.

```{r prep-addon-renames}
if (run_stage("01")) {
  if (!is.null(teacher_hr) && length(config$column_map$teacher_hr) > 0) {
    hr_map <- unlist(config$column_map$teacher_hr)
    teacher_hr <- teacher_hr |> rename(any_of(hr_map))

    # Coerce employee ID to match the type in the course schedule (numeric)
    if ("D_employee_id" %in% names(teacher_hr))
      teacher_hr <- teacher_hr |> mutate(D_employee_id = as.numeric(D_employee_id))

    landed <- names(hr_map)[names(hr_map) %in% names(teacher_hr)]
    data.frame(
      standard_name   = names(hr_map),
      district_column = unname(hr_map),
      landed          = ifelse(names(hr_map) %in% names(teacher_hr), "yes", "MISSING")
    ) |>
      check_gt(
        title  = "1.2a · Teacher HR column map",
        note   = glue("Mode is '{config$add_ons$teacher_hr$mode}'. Experience mode needs D_employee_id and D_tch_yrs_experience; licensure needs D_employee_license_type and D_employee_license_subject; evaluation needs D_employee_rating."),
        status = if (length(landed) == length(hr_map)) "PASS" else "CHECK")
  }

  if (!is.null(course_grades_raw) && length(config$column_map$course_grades) > 0) {
    cg_map <- unlist(config$column_map$course_grades)
    course_grades_raw <- course_grades_raw |> rename(any_of(cg_map))
    chk_info("Course grades renamed", paste(names(cg_map), collapse = ", "))
  }
}
```

## 1.3 Time columns

The columns that might encode when a class meets. Getting this wrong makes
every weight wrong, so it is worth looking rather than assuming.

```{r raw-time-peek}
time_candidates <- intersect(
  unlist(config$column_map$course_schedule[
    c("D_period", "D_term", "D_course_section_id")]),
  names(cs_raw))

if (length(time_candidates) > 0) {
  explore_render(setNames(
    lapply(time_candidates, function(cl) explore_value_counts(cs_raw, cl, top = 25)),
    time_candidates))
}
```

## 1.4 Duplicates in the raw file

```{r raw-dupes}
if (run_stage("01")) {
  n_raw   <- nrow(cs_raw)
  n_uniq  <- nrow(distinct(cs_raw))

  data.frame(
    measure = c("Rows as received", "Distinct rows", "Exact duplicate rows"),
    value   = c(n_raw, n_uniq, n_raw - n_uniq)
  ) |>
    check_gt(
      title  = "1.4 · Exact duplicate rows in the raw file",
      note   = "Fully identical rows are dropped during prep. A large count is worth asking the district about \u2014 it usually means an export ran twice or a join upstream fanned out.",
      status = if (n_raw == n_uniq) "PASS" else "CHECK")
}
```

## 1.5 Prep

Rename to standard names, merge demographics and location coding, apply scope,
flag the student snapshot, and apply the derivations. Each step reports what it
cost.

### Rename

```{r prep-rename}
cs_map <- unlist(config$column_map$course_schedule)

cs_prepped <- cs_raw |>
  rename(any_of(cs_map)) |>
  select(any_of(names(cs_map))) |>
  distinct()

mapped   <- names(cs_map)[names(cs_map) %in% names(cs_prepped)]
unmapped <- setdiff(names(cs_map), names(cs_prepped))

data.frame(
  standard_name   = names(cs_map),
  district_column = unname(cs_map),
  landed          = ifelse(names(cs_map) %in% names(cs_prepped), "yes", "MISSING")
) |>
  check_gt(
    title  = "1.5a · Did every mapped column land?",
    note   = "A MISSING row means column_map names a district column that is not in the file. Fix the map here rather than patching downstream.",
    status = if (length(unmapped) == 0) "PASS" else "CHECK")

chk_info("Rows after dropping exact duplicates",
         paste0(format(nrow(cs_prepped), big.mark = ","), " of ",
                format(nrow(cs_raw), big.mark = ",")))
```

### Student demographics

```{r prep-demographics}
if (!is.null(stu_demographics)) {
  dem_cfg <- config$stu_demographics_join

  demographics_cleaned <- stu_demographics |>
    rename(any_of(unlist(config$column_map$demographics))) |>
    select(all_of(c(dem_cfg$key, unlist(dem_cfg$brings_in)))) |>
    distinct()

  if (!is.null(dem_cfg$recode)) {
    for (col_name in names(dem_cfg$recode)) {
      mapping <- dem_cfg$recode[[col_name]]
      demographics_cleaned <- demographics_cleaned |>
        mutate(!!col_name := unlist(mapping)[as.character(.data[[col_name]])])
    }
  }

  # join_checked rather than left_join: a demographics file with duplicate
  # student rows would silently multiply the schedule, and that is exactly the
  # kind of thing a row-count check exists to catch.
  cs_prepped <- cs_prepped |>
    join_checked(demographics_cleaned, by = dem_cfg$key,
                 check = "Student demographics onto schedule")
} else {
  chk_info("Demographics merge skipped", "no demographics file")
}
```

### Location coding

```{r prep-location}
if (!is.null(location_coding)) {
  cs_prepped <- cs_prepped |>
    join_checked(location_coding, by = "D_location_id",
                 check = "Location coding onto schedule")
} else {
  chk_info("Location coding merge skipped", "no location file")
}
```

### Scope

```{r prep-scope}
cs_prepped <- apply_scope(cs_prepped, config$scope, id_col = "D_stu_id")
```

### Student snapshot

For longitudinal enrollment data, each row is flagged on whether the student
was enrolled on that term's snapshot date. Skipped entirely for districts whose
extract is already a snapshot.

```{r prep-snapshot}
if (!is.null(config$intake$enrollment_type) &&
    config$intake$enrollment_type != "snapshot") {

  cs_prepped <- cs_require(apply_stu_snapshot(cs_prepped, config),
                           "Student snapshot flag applied")

  cs_prepped |>
    count(D_term, C_stu_snapshot, name = "n_records") |>
    pivot_wider(names_from = C_stu_snapshot, values_from = n_records,
                values_fill = 0) |>
    mutate(snapshot_date = unlist(config$intake$snapshot_dates)[as.character(D_term)]) |>
    check_gt(
      title = "1.5b · Snapshot inclusion by term",
      note  = "Each term uses its own snapshot date. A term with almost nothing included usually means the date sits outside that term's enrolment window \u2014 check it against the enter/exit distribution before trusting it.")
} else {
  chk_info("Snapshot flagging skipped", "extract is already a snapshot")
}
```

### Derivations

Fields the pipeline computes rather than reads. These were hardcoded in the old
prep chunk, which put a district-specific assumption inside a file meant to be
reusable.

```{r prep-derivations}
null_to_zero <- intersect(unlist(config$derivations$null_to_zero %||% character()),
                          names(cs_prepped))

if (length(null_to_zero) > 0) {
  before <- vapply(null_to_zero, function(c) sum(is.na(cs_prepped[[c]])), numeric(1))
  cs_prepped <- cs_prepped |>
    mutate(across(all_of(null_to_zero), ~ coalesce(.x, 0)))
  chk_info("Null flags set to zero", "see table",
           data = tibble(column = null_to_zero, n_filled = unname(before)))
}

for (nm in names(config$derivations$constants %||% list())) {
  cs_prepped[[nm]] <- config$derivations$constants[[nm]]
  chk_info(paste0("Constant applied: ", nm),
           config$derivations$constants[[nm]],
           detail = "declared in config$derivations$constants")
}
```

## 1.6 Duplicates after prep

Confirms the merges did not fan out.

```{r prep-dupes}
if (run_stage("01")) {
  n_prep  <- nrow(cs_prepped)
  n_prepu <- nrow(distinct(cs_prepped))

  data.frame(
    measure = c("Rows after prep", "Distinct rows", "Exact duplicate rows"),
    value   = c(n_prep, n_prepu, n_prep - n_prepu)
  ) |>
    check_gt(
      title  = "1.6 · Duplicates after prep",
      note   = "Prep should not create duplicates. If this is non-zero when 1.4 was zero, a merge fanned out.",
      status = if (n_prep == n_prepu) "PASS" else "CHECK")
}
```

```{r prep-grain}
if (run_stage("01")) {
  key_cols <- intersect(c("D_stu_id", "D_course_id", "D_course_section_id",
                          "D_term", "D_period"), names(cs_prepped))

  dupe_keys <- cs_prepped |>
    count(across(all_of(key_cols)), name = "n_rows") |>
    filter(n_rows > 1)

  dupe_keys |>
    count(n_rows, name = "n_key_combinations") |>
    check_gt(
      title  = paste0("1.6b \u00B7 Rows per ", paste(key_cols, collapse = " + ")),
      note   = "More than one row per student per section per time slot means the grain is not what the pipeline assumes. Worth understanding before class sizes are computed.",
      status = if (nrow(dupe_keys) == 0) "PASS" else "CHECK")
}
```

## 1.7 Prepped profile

```{r prep-profile}
explore_gt(explore_profile(cs_prepped), "Course schedule · after prep")
```

```{r prep-time-peek}
time_cols_now <- intersect(config$time_cols[[config$time_format]], names(cs_prepped))

if (length(time_cols_now) > 0) {
  explore_render(setNames(
    lapply(time_cols_now, function(cl) explore_value_counts(cs_prepped, cl, top = 25)),
    time_cols_now))
}
```

## 1.8 Coding template

The one-time bootstrap for a district with no coding file yet. Off by default;
set `coding_template$write: true`, render once, then set it back.

The columns it groups by come from config. The previous version hardcoded
`D_course_subject_id`, a Detroit field absent from the Nashville data, so this
chunk would have errored if it had ever been run here.

```{r coding-template}
#| eval: !expr isTRUE(config$coding_template$write)

tpl_group <- intersect(unlist(config$coding_template$group_by), names(cs_prepped))

chk_stop("Coding template group_by columns exist",
         length(tpl_group) == length(unlist(config$coding_template$group_by)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$coding_template$group_by), names(cs_prepped)),
           collapse = ", ")))

coding_template <- cs_prepped |>
  distinct(across(all_of(tpl_group)))

for (f in unlist(config$coding_template$uncoded_fields)) {
  coding_template[[f]] <- "UNCODED"
}

chk_info("Coding template rows", nrow(coding_template))

ers_write_sharepoint(
  data = coding_template,
  folder_path = config$folders$raw,
  file_name_with_extension = config$coding_template$file_name)
```

## 1.9 Handoff

`cs_data` is what Stage 02 picks up. In the old pipeline this was a CSV on
SharePoint; now it is an object in memory, and the row count below is the last
number Stage 01 is responsible for.

```{r prep-handoff}
cs_data <- cs_prepped

data.frame(
  measure = c("Rows", "Columns", "Distinct students", "Distinct teachers", "Distinct schools"),
  value   = c(format(nrow(cs_data), big.mark = ","),
              ncol(cs_data),
              format(n_distinct(cs_data$D_stu_id), big.mark = ","),
              format(n_distinct(cs_data$D_employee_id), big.mark = ","),
              format(n_distinct(cs_data$D_location_id), big.mark = ","))
) |>
  check_gt(title = "1.9 · Stage 01 handoff summary")
```

---

# Stage 02 — Coding and exclusions

Merges the coded course file onto the schedule, applies the exclusion rules,
and builds the class ID. Everything downstream depends on these three, so this
is where the gates are strictest.

## 2.0 Exclusion logic

A record is excluded from a metric if **any** rule for that metric matches it.
Rules naming a column that does not exist are skipped rather than erroring, so
a district missing an optional coded field still runs.

```{r 02-exclusion-fn}
if (run_stage("02")) {
  chk_stage("Stage 02 — Coding and exclusions")

  apply_exclusions <- function(df, rules) {
    for (field in names(rules)) {
      should_exclude <- rep(FALSE, nrow(df))

      for (rule in rules[[field]]) {
        if (!rule$col %in% names(df)) next

        column_values <- df[[rule$col]]
        rule_values   <- rule$val

        matches_rule <- switch(
          rule$op,
          eq     = column_values == rule_values,
          ne     = column_values != rule_values,
          `in`   = column_values %in% rule_values,
          not_in = !(column_values %in% rule_values),
          rep(FALSE, length(column_values)))

        matches_rule[is.na(matches_rule)] <- FALSE
        should_exclude <- should_exclude | matches_rule
      }

      df[[field]] <- ifelse(should_exclude, "Exclude", "Include")
    }
    df
  }
}
```

## 2.1 Merge the coding file

The coding key has to be unique or the join multiplies the schedule, and the
row count has to hold. Both are structural, so `join_checked()` stops rather
than reporting. Unmatched courses are expected and get reported instead — those
are courses that still need coding.

```{r 02-merge-coding}
if (run_stage("02")) {
  join_key <- if (!is.null(config$coding$join_key))
    unlist(config$coding$join_key) else "D_course_name"

  chk_stop("Coding file is available", !is.null(cs_course_coding),
           detail = "Stage 01 could not read it; see the Stage 01 log")

  coded_cols <- setdiff(names(cs_course_coding), join_key)

  cs_data <- cs_data |>
    select(-any_of(coded_cols)) |>
    join_checked(cs_course_coding, by = join_key,
                 check = "Course coding onto schedule")

  unmatched_courses <- cs_data |>
    filter(is.na(.data[[coded_cols[1]]])) |>
    distinct(across(all_of(join_key))) |>
    arrange(across(all_of(join_key)))

  merge_diag <- list(
    key         = paste(join_key, collapse = " + "),
    n_unmatched = sum(is.na(cs_data[[coded_cols[1]]])),
    unmatched_courses = as.data.frame(unmatched_courses))

  chk_info("Course rows with no coding", merge_diag$n_unmatched)
}
```

```{r 02-apply-exclusions}
if (run_stage("02")) {
  cs_data <- apply_exclusions(cs_data, config$exclusions)

  for (field in names(config$exclusions)) {
    if (field %in% names(cs_data)) {
      chk_info(paste0("Exclusion rule: ", field),
               paste0(format(sum(cs_data[[field]] == "Exclude"), big.mark = ","),
                      " of ", format(nrow(cs_data), big.mark = ","), " excluded (",
                      round(100 * mean(cs_data[[field]] == "Exclude"), 1), "%)"))
    }
  }
}
```

```{r 02-class-id}
if (run_stage("02")) {
  cs_data$C_class_id <- cs_require(
    build_class_id(cs_data, config$class_id$recipe),
    "Class ID built", min_rows = 0)

  has_cols <- function(...) all(c(...) %in% names(cs_data))

  chk_info("Distinct classes", n_distinct(cs_data$C_class_id))

  # Hoisted out of check 2.2, which is display-gated: an object several checks
  # depend on cannot live inside a chunk that may not run.
  time_cols <- config$time_cols[[config$time_format]]
}
```

```{r}
#| output: asis
#| echo: false
cs_heading("02", "## 2.2 Checks")
```

```{r}
#| label: chk-2.0
if (run_stage("02")) {
  expected_fields <- names(config$column_map$course_schedule)
  data.frame(
    field   = expected_fields,
    present = ifelse(expected_fields %in% names(cs_data), "yes", "MISSING")
  ) |>
    check_gt(
      title  = "2.0 \u00B7 Did standardization from Stage 01 land?",
      note   = "Every field in column_map should be present with its D_ name. If one shows MISSING, fix the rename map in Stage 01 \u2014 don't patch it downstream.",
      status = if (length(setdiff(expected_fields, names(cs_data))) == 0) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-2.1
if (run_stage("02")) {
  cs_data |>
    count(D_term, name = "n_records") |>
    arrange(D_term) |>
    mutate(weight = sapply(as.character(D_term), \(t) {
      w <- config$weights$term[[t]]; if (is.null(w)) "MISSING from config" else as.character(w) })) |>
    check_gt(
      title  = "2.1 \u00B7 Term structure \u2014 is every term value weightable?",
      note   = "Every term in the data needs a configured weight. If any shows MISSING from config, add it to weights.term.",
      status = if (all(as.character(unique(na.omit(cs_data$D_term))) %in% names(config$weights$term))) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-2.2
if (run_stage("02")) {
  included_data <- cs_data
  if (has_cols("C_course_time_exclude", "C_class_size_and_teacher_load_exclude"))
    included_data <- included_data |>
      filter(C_course_time_exclude == "Include",
             C_class_size_and_teacher_load_exclude == "Include")

  records_per_student <- included_data |>
    count(D_location_name, D_stu_id, name = "records_per_student")

  included_data |>
    group_by(D_location_name) |>
    summarise(n_students = n_distinct(D_stu_id), n_teachers = n_distinct(D_employee_id),
              terms = paste(sort(unique(na.omit(D_term))), collapse = ", "),
              n_time_slots = n_distinct(!!!syms(time_cols)), .groups = "drop") |>
    left_join(records_per_student |> group_by(D_location_name) |>
                summarise(avg_records_per_student = round(mean(records_per_student), 1),
                          .groups = "drop"),
              by = "D_location_name") |>
    arrange(D_location_name) |>
    check_gt(
      title = "2.2 \u00B7 Records per student, by school",
      note  = "Per-school record counts and time-slot variety should look plausible for HS schedules. A school far off the others is worth a look.")
}
```

```{r}
#| label: chk-2.3
if (run_stage("02")) {
  demographic_fields <- unlist(config$stu_demographics_join$brings_in)
  data.frame(
    field      = demographic_fields,
    present    = ifelse(demographic_fields %in% names(cs_data), "yes", "no"),
    pct_filled = vapply(demographic_fields, function(f)
      if (f %in% names(cs_data)) round(100 * mean(!is.na(cs_data[[f]])), 1) else NA_real_,
      numeric(1))
  ) |>
    check_gt(
      title  = "2.3 \u00B7 Student demographics \u2014 present and populated?",
      note   = "The demographic fields from the Stage 01 merge should exist and be well-filled. Low pct_filled may be real (a district that doesn't track a flag) or a merge miss.",
      status = if (all(demographic_fields %in% names(cs_data))) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-2.4
if (run_stage("02")) {
  data.frame(
    check  = c("join key", "unmatched course rows"),
    result = c(merge_diag$key, format(merge_diag$n_unmatched, big.mark = ","))) |>
    check_gt(
      title  = "2.4 \u00B7 Did the course-coding merge land cleanly?",
      note   = "Key uniqueness and row stability are gated above \u2014 a failure there halts the run. What's left to judge is whether any courses are still uncoded.",
      status = if (merge_diag$n_unmatched == 0) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-2.4b
if (run_stage("02")) {
  merge_diag$unmatched_courses |>
    check_gt(title = "2.4 \u00B7 Courses that didn't get coded")
}
```

```{r}
#| label: chk-2.5
if (run_stage("02")) {
  pct_uncoded <- 100 * mean(cs_data$C_course_subject_area == "UNCODED", na.rm = TRUE)

  cs_data |>
    count(C_course_subject_area, name = "n_records") |>
    mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
    arrange(C_course_subject_area) |>
    check_gt(
      title  = glue("2.5 \u00B7 Coding output \u2014 nothing left uncoded? ({round(pct_uncoded, 1)}% uncoded)"),
      note   = glue("Uncoded should be \u2264 {config$coding$uncoded_flag_threshold_pct}% of records, and distributions should look plausible. High UNCODED means courses need coding."),
      status = if (pct_uncoded <= config$coding$uncoded_flag_threshold_pct) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-2.5b
if (run_stage("02")) {
  cs_data |>
    count(C_course_subject_area, C_course_subject, name = "n_records") |>
    mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
    arrange(C_course_subject_area) |>
    check_gt(title = "2.5 \u00B7 Subject \u2014 % of records")
}
```

```{r}
#| label: chk-2.6
if (run_stage("02")) {
  cs_data |>
    filter(C_course_time_exclude == "Exclude" |
           C_class_size_and_teacher_load_exclude == "Exclude") |>
    count(C_course_subject, D_course_name, C_course_time_exclude,
          C_class_size_and_teacher_load_exclude, name = "n_courses", sort = TRUE) |>
    arrange(C_course_subject) |>
    check_gt(
      title = "2.6 \u00B7 Exclusions \u2014 right thing dropping, for the right reason?",
      note  = "Every excluded course should trace to a coding value that justifies it (lunch, study hall, non-instructional). Anything here that looks like a real class means a coding fix.")
}
```

```{r}
#| label: chk-2.7
if (run_stage("02")) {
  non_standard <- cs_data |>
    count(across(all_of(c("D_location_name", time_cols))), C_course_format,
          name = "n_records") |>
    pivot_wider(names_from = C_course_format, values_from = n_records, values_fill = 0) |>
    filter(`Standard` == 0) |>
    arrange(across(all_of(c("D_location_name", time_cols))))

  non_standard |>
    check_gt(
      title = "2.7 \u00B7 Which time blocks don't carry real instruction?",
      note  = "Time-slots with no standard-format records (below), or very few distinct courses (next table), are identified so they fall out of instructional analysis.")
}
```

```{r}
#| label: chk-2.7b
if (run_stage("02")) {
  max_courses <- config$coding$max_unique_courses %||% 5

  cs_data |>
    group_by(across(all_of(c("D_location_name", time_cols)))) |>
    mutate(n_courses = n_distinct(D_course_name)) |>
    ungroup() |>
    filter(n_courses <= max_courses, !C_course_subject_area %in% c("Core", "Non-Core")) |>
    count(across(all_of(c("D_location_name", time_cols))), C_course_subject,
          D_course_name, n_courses, name = "n_records") |>
    arrange(D_location_name, n_courses, desc(n_records)) |>
    check_gt(title = glue("2.7 \u00B7 Low-diversity time-slots (\u2264 {max_courses} courses, excl. Core and Non-Core)"))
}
```

```{r}
#| label: chk-2.8
if (run_stage("02")) {
  cols <- config$class_id$recipe[config$class_id$recipe %in% names(cs_data)]
  cs_data |>
    filter(!is.na(C_class_id)) |>
    group_by(across(all_of(c("C_class_id", cols)))) |>
    summarise(course = first(D_course_name), n_students = n_distinct(D_stu_id),
              .groups = "drop") |>
    arrange(desc(n_students)) |>
    (\(d) slice_sample(d, n = min(8, nrow(d))))() |>
    check_gt(
      title = glue("2.8 \u00B7 Class ID \u2014 what defines a class here? ({paste(config$class_id$recipe, collapse = ' + ')})"),
      note  = "The recipe should mark a place-in-time without over-splitting real classes. Use SECTION, never the teacher \u2014 a teacher in the id hides teacher double-bookings.")
}
```

## 2.3 Gates

```{r 02-gates}
if (run_stage("02")) {
  chk_stop("Every mapped column survived to Stage 02",
           length(setdiff(names(config$column_map$course_schedule), names(cs_data))) == 0,
           detail = paste("Missing:", paste(
             setdiff(names(config$column_map$course_schedule), names(cs_data)),
             collapse = ", ")))

  chk_stop("Every term in the data has a configured weight",
           all(as.character(unique(na.omit(cs_data$D_term))) %in% names(config$weights$term)),
           detail = paste("Unweighted:", paste(
             setdiff(as.character(unique(na.omit(cs_data$D_term))),
                     names(config$weights$term)), collapse = ", ")))

  chk_stop("Exclusion fields were created",
           all(names(config$exclusions) %in% names(cs_data)))

  chk_no_na(cs_data, "C_class_id", "Every row has a class ID")
}
```

---

# Stage 03 — Weights, minutes and DB resolution

The engine. Every call here is `cs_require()` rather than `tr_safe()`: if the
resolution pipeline fails, the run stops. Previously a failure returned `NULL`
and every later stage degraded quietly to empty rollups.

## 3.0 Weights and minutes

```{r 03-weights}
if (run_stage("03")) {
  chk_stage("Stage 03 — Weights, minutes and DB resolution")

  cs_data <- cs_require(derive_class_weight_minutes(cs_data, config, spec),
                        "Class weights and minutes derived")

  cs_data$M_stu_class_weight <- cs_data$M_cls_class_weight
  cs_data$M_tch_class_weight <- cs_data$M_cls_class_weight

  chk_no_na(cs_data, "M_cls_class_weight", "Every row has a class weight")

  chk_stop("Class weights are non-negative",
           all(cs_data$M_cls_class_weight >= 0, na.rm = TRUE),
           offenders = cs_data |> filter(M_cls_class_weight < 0) |>
             count(D_location_name, D_term, name = "n_rows"))
}
```

## 3.1 Resolution engine

```{r 03-engine}
if (run_stage("03")) {
  cs_data$H_row_id_before_explosions <- seq_len(nrow(cs_data))

  exploded    <- cs_require(explode_for_resolution(cs_data, config, spec),
                            "Explosion for resolution")
  flagged_all <- cs_require(flag_double_bookings(exploded, config),
                            "Double-booking flags (all rows)")
  flagged     <- cs_require(flag_double_bookings(exploded, config, filter_to_db = TRUE),
                            "Double-booking flags (DB rows only)", min_rows = 0)
  res         <- cs_require(resolve_double_bookings(flagged, config),
                            "Resolution engine", min_rows = 0)
  resolved    <- res$resolved
  mapping     <- res$mapping
  loadoff     <- cs_require(teacher_loadoff(resolved, config), "Teacher load-off",
                            min_rows = 0)
  unified     <- cs_require(unify_resolution(resolved, cs_data, loadoff),
                            "Unification")
}
```

```{r 03-precompute}
if (run_stage("03")) {
  flagged_students <- unique(flagged_all$D_stu_id[flagged_all$H_db_student_row == 1])
  flagged_teachers <- unique(flagged_all$D_employee_id[flagged_all$H_db_teacher_row == 1])

  pct_of <- function(a, b) { r <- round(100 * a / b, 1); r[is.na(b) | b == 0] <- NA_real_; r }

  multi_teacher <- resolved |>
    filter(startsWith(C_class_id_post_teachers, "CONS_")) |>
    group_by(C_class_id_post_teachers) |>
    summarise(n_tch = n_distinct(D_employee_id), .groups = "drop") |>
    filter(n_tch > 1)

  consol_sum      <- consolidation_summary(mapping)
  teach_db_before <- length(flagged_teachers)
  teach_db_after  <- {
    rf <- resolved |> mutate(C_class_id = C_class_id_post_teachers) |>
      select(-any_of(c("H_db_teacher_row", "H_db_student_row")))
    rf <- flag_teacher_db(rf)
    n_distinct(rf$D_employee_id[rf$H_db_teacher_row == 1])
  }
  sec_before <- n_distinct(mapping$C_class_id_original)
  sec_after  <- n_distinct(mapping$C_class_id_resolved)

  mono <- with(resolved,
    all(M_class_weight_adj_same_course >= M_class_weight_post_class    - 1e-9, na.rm = TRUE) &&
    all(M_class_weight_post_class      >= M_class_weight_post_teachers - 1e-9, na.rm = TRUE) &&
    all(M_class_weight_post_teachers   >= M_class_weight_post_student  - 1e-9, na.rm = TRUE))

  n_matched <- cs_data |>
    semi_join(distinct(resolved, H_row_id_before_explosions),
              by = "H_row_id_before_explosions") |> nrow()
  unif_ok <- isTRUE(nrow(unified) == nrow(resolved) + (nrow(cs_data) - n_matched))

  stu_before <- exploded |> group_by(D_stu_id) |>
    summarise(b = sum(M_class_weight_exploded, na.rm = TRUE), .groups = "drop")
  stu_after <- resolved |> group_by(D_stu_id) |>
    summarise(a = sum(M_class_weight_post_student, na.rm = TRUE), .groups = "drop")
  stu_ba <- stu_before |> left_join(stu_after, by = "D_stu_id") |>
    mutate(a = coalesce(a, b))

  # Example pickers are cosmetic: a missing example weakens a trace table, it
  # does not invalidate anything. tr_safe is the right tool here.
  ex_student <- tr_safe(resolved |> filter(H_db_student_row == 1) |>
                          slice_head(n = 1) |> pull(D_stu_id), NA)
  ex_cons    <- tr_safe(consol_sum[["Example consolidations (largest first)"]]$C_class_id_consolidated[1], NA) %||% NA

  trace_cols <- function(d) d |> select(any_of(c(
    "D_stu_id", "D_course_name", "C_period_exploded", "C_rotation_exploded",
    "C_class_id_post_teachers", "M_class_weight_exploded", "M_class_weight_post_class",
    "M_class_weight_post_teachers", "M_class_weight_post_student",
    "M_student_count_adjusted")))
}
```

```{r}
#| output: asis
#| echo: false
cs_heading("03", "## 3.2 Checks")
```

```{r}
#| label: chk-3.0
if (run_stage("03")) {
  tk   <- config$time_cols[[config$time_format]]
  keep <- intersect(c("C_class_id", "D_location_name", tk, "D_term",
                      "D_course_name", "M_cls_class_weight", "M_cls_class_minutes"),
                    names(cs_data))
  base <- cs_data |> filter(!is.na(C_class_id)) |> distinct(across(all_of(keep)))
  uniq <- base |> distinct(M_cls_class_weight, .keep_all = TRUE)
  d <- if (nrow(uniq) < 10)
    bind_rows(uniq, base |> anti_join(uniq, by = "C_class_id") |>
                slice_sample(n = 10 - nrow(uniq))) else uniq
  d <- d |> arrange(desc(M_cls_class_weight))
  d$time   <- apply(d[intersect(tk, names(d))], 1, function(r) paste(r[!is.na(r)], collapse = " "))
  d$course <- d$D_course_name

  d |> select(any_of(c("D_location_name", "time", "D_term", "course",
                       "M_cls_class_weight", "M_cls_class_minutes"))) |>
    check_gt(
      title = "3.0 \u00B7 Class weight & minutes \u2014 do a few real classes look right?",
      note  = "term \u00D7 rotation \u00D7 period should reconcile by hand; minutes come from the bell schedule. They're separate on purpose. Shows every distinct weight value.")
}
```

```{r}
#| label: chk-3.1
if (run_stage("03")) {
  cs_data |>
    filter(C_course_time_exclude == "Include") |>
    distinct(D_location_name, D_stu_id, C_class_id, M_stu_class_weight) |>
    group_by(D_location_name, D_stu_id) |>
    summarise(total_weight = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |>
    group_by(D_location_name) |>
    summarise(n_students = n_distinct(D_stu_id),
              avg_weight_per_student = round(mean(total_weight), 2), .groups = "drop") |>
    check_gt(
      title = "3.1 \u00B7 Weight per student, by school",
      note  = "Sums should reflect the schedule structure (blocks vs. periods) \u2014 NOT forced to a single target. Teacher version in the next table.")
}
```

```{r}
#| label: chk-3.1b
if (run_stage("03")) {
  cs_data |>
    filter(C_class_size_and_teacher_load_exclude == "Include") |>
    distinct(D_location_name, D_employee_id, C_class_id, M_tch_class_weight) |>
    group_by(D_location_name, D_employee_id) |>
    summarise(total_weight = sum(M_tch_class_weight, na.rm = TRUE), .groups = "drop") |>
    group_by(D_location_name) |>
    summarise(n_teachers = n_distinct(D_employee_id),
              avg_weight_per_teacher = round(mean(total_weight), 2), .groups = "drop") |>
    check_gt(title = "3.1 \u00B7 Weight per teacher, by school")
}
```

```{r}
#| label: chk-3.2
if (run_stage("03")) {
  cs_data |>
    filter(C_course_time_exclude == "Include") |>
    distinct(D_location_name, D_stu_id, C_class_id, M_stu_class_weight, M_cls_class_minutes) |>
    group_by(D_location_name, D_stu_id) |>
    summarise(total_weight = sum(M_stu_class_weight, na.rm = TRUE),
              total_minutes_per_cycle = sum(M_cls_class_minutes, na.rm = TRUE),
              .groups = "drop") |>
    group_by(D_location_name) |>
    summarise(avg_weight = round(mean(total_weight), 1),
              avg_minutes_per_cycle = round(mean(total_minutes_per_cycle), 0),
              .groups = "drop") |>
    check_gt(
      title = "3.2 \u00B7 Minutes layer (separate) \u2014 instructional time, by school",
      note  = "Avg weight and avg minutes side by side. weight \u00D7 minutes should be comparable across schools even when period lengths differ.")
}
```

```{r}
#| label: chk-3.3
if (run_stage("03")) {
  n_inc_stu <- n_distinct(cs_data$D_stu_id[cs_data$C_course_time_exclude == "Include"])
  n_inc_tch <- n_distinct(cs_data$D_employee_id[cs_data$C_class_size_and_teacher_load_exclude == "Include"])

  avg_class_size <- cs_data |> filter(C_class_size_and_teacher_load_exclude == "Include") |>
    distinct(C_class_id, D_stu_id) |> count(C_class_id, name = "size") |>
    summarise(v = round(mean(size), 1)) |> pull(v)
  avg_stu_w <- cs_data |> filter(C_course_time_exclude == "Include") |>
    distinct(D_stu_id, C_class_id, M_stu_class_weight) |> group_by(D_stu_id) |>
    summarise(t = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |>
    summarise(v = round(mean(t), 1)) |> pull(v)
  avg_tch_w <- cs_data |> filter(C_class_size_and_teacher_load_exclude == "Include") |>
    distinct(D_employee_id, C_class_id, M_tch_class_weight) |> group_by(D_employee_id) |>
    summarise(t = sum(M_tch_class_weight, na.rm = TRUE), .groups = "drop") |>
    summarise(v = round(mean(t), 1)) |> pull(v)

  data.frame(
    metric = c("Students double-booked", "Teachers double-booked", "Avg class size",
               "Avg weight / student", "Avg weight / teacher"),
    before = c(paste0(length(flagged_students), " (", pct_of(length(flagged_students), n_inc_stu), "%)"),
               paste0(length(flagged_teachers), " (", pct_of(length(flagged_teachers), n_inc_tch), "%)"),
               avg_class_size, avg_stu_w, avg_tch_w),
    after  = rep("pending resolution", 5)) |>
    check_gt(
      title = "3.3 \u00B7 Before-resolution snapshot",
      note  = "The starting picture \u2014 double-bookings by type, avg class size, avg weights, all unresolved. The 'after' fills in once resolution runs (3.6).")
}
```

```{r}
#| label: chk-3.4
if (run_stage("03")) {
  consol_sum[["Component-size distribution"]] |>
    check_gt(
      title  = glue("3.4 \u00B7 consolidate_sections() \u2014 overlapping sections merge into one class (teacher DBs {teach_db_before} \u2192 {teach_db_after} \u00B7 sections {sec_before} \u2192 {sec_after})"),
      note   = "Every CONS_ class must be exactly one teacher's merged sections. If any spans >1 teacher, the CONS_ id is colliding \u2014 flag it, don't edit the helper.",
      status = if (nrow(multi_teacher) == 0) "PASS" else "FAIL")
}
```

```{r}
#| label: chk-3.5
if (run_stage("03")) {
  resolved |> filter(C_course_time_exclude == "Include") |>
    group_by(D_location_name, D_stu_id) |>
    summarise(exploded = sum(M_class_weight_exploded, na.rm = TRUE),
              post_class = sum(M_class_weight_post_class, na.rm = TRUE),
              post_teachers = sum(M_class_weight_post_teachers, na.rm = TRUE),
              post_student = sum(M_class_weight_post_student, na.rm = TRUE),
              .groups = "drop") |>
    group_by(D_location_name) |>
    summarise(n_db_students = n_distinct(D_stu_id),
              across(c(exploded, post_class, post_teachers, post_student),
                     ~ round(mean(.x), 2)), .groups = "drop") |>
    check_gt(
      title  = "3.5 \u00B7 resolve_stage() \u2014 weight splits at conflicts, only ever shrinking",
      note   = "adj \u2265 post_class \u2265 post_teachers \u2265 post_student everywhere. The columns show a double-booked student's weight moving down through each stage.",
      status = if (mono) "PASS" else "FAIL")
}
```

```{r}
#| label: chk-3.6
if (run_stage("03")) {
  data.frame(
    metric = c("Students double-booked", "Teachers double-booked", "Distinct sections",
               "Avg weight / student"),
    before = c(length(flagged_students), teach_db_before, sec_before, round(mean(stu_ba$b), 2)),
    after  = c(0, teach_db_after, sec_after, round(mean(stu_ba$a), 2))) |>
    check_gt(
      title  = "3.6 \u00B7 unify_resolution() \u2014 before \u2192 after",
      note   = "Double-bookings resolve to zero and sections consolidate to their true count. Row accounting in the next table confirms nothing was lost or duplicated.",
      status = if (unif_ok) "PASS" else "FAIL")
}
```

```{r}
#| label: chk-3.6b
if (run_stage("03")) {
  unification_summary(unified, resolved, cs_data) |>
    check_gt(title = "3.6 \u00B7 Row accounting (unified = resolved + non-DB)")
}
```

```{r}
#| label: chk-3.7
if (run_stage("03")) {
  trace_cols(trace_record(resolved, stu_id = ex_student)) |>
    check_gt(
      title = glue("3.7 \u00B7 A student double-booking, traced end to end (stu {ex_student})"),
      note  = "Follow one student's weight across the stage columns \u2014 it should sum to one class-worth, no time lost or double-counted.")
}
```

```{r}
#| label: chk-3.7b
if (run_stage("03")) {
  trace_cols(trace_record(resolved, class_id = ex_cons)) |>
    check_gt(
      title = glue("3.7 \u00B7 A teacher consolidation, traced ({ex_cons})"),
      note  = "The teacher's sections merge into one CONS_ class, but each student keeps only their own meetings \u2014 the union never inflates anyone's time.")
}
```

## 3.3 Gates

```{r 03-gates}
if (run_stage("03")) {
  chk_stop("Weight only ever shrinks through the resolution stages", mono,
           detail = "adj >= post_class >= post_teachers >= post_student must hold on every row")

  chk_stop("Every consolidated class belongs to exactly one teacher",
           nrow(multi_teacher) == 0,
           value = paste0(nrow(multi_teacher), " colliding CONS_ id(s)"),
           offenders = multi_teacher)

  chk_stop("Unification accounts for every row", unif_ok,
           value  = paste0(format(nrow(unified), big.mark = ","), " unified rows"),
           detail = paste0("expected ", format(nrow(resolved), big.mark = ","),
                           " resolved + ", format(nrow(cs_data) - n_matched, big.mark = ","),
                           " non-DB"))
}
```

---

# Stage 04 — Rollups and export

## 4.0 Rollup layer

```{r 04-rollups}
if (run_stage("04")) {
  chk_stage("Stage 04 — Rollups and export")

  unified <- cs_require(add_minutes(unified, config, spec),        "Minutes added")
  unified <- cs_require(add_class_size(unified, config),           "Class size added")
  unified <- cs_require(add_class_primaries(unified),              "Class primaries added")
  unified <- cs_require(add_class_fields(unified),                 "Class fields added")
  unified <- cs_require(add_class_weight_times_size(unified),      "Weight x size added")
  unified <- cs_require(add_teacher_metrics(unified, config),      "Teacher metrics added")
  unified <- cs_require(add_teacher_primary_fields(unified),       "Teacher primary fields added")
  unified <- cs_require(add_teacher_flags(unified),                "Teacher flags added")
}
```

## 4.1 Optional add-ons

Genuinely optional, so `cs_optional()`: a district without an HR file gets the
core rollups with those fields as NA, and the report says so.

```{r 04-addons}
if (run_stage("04")) {
  hr_on     <- isTRUE(config$add_ons$teacher_hr$enabled)
  grades_on <- isTRUE(config$add_ons$course_grades$enabled)
  grades_rollup <- NULL

  # Uses the object loaded and renamed in Stage 01 rather than re-reading the
  # file. That is what closes the gap the old pipeline had: the add-on now sees
  # D_ columns because the rename happened on the standard path.
  if (hr_on) {
    if (!is.null(teacher_hr)) {
      unified <- cs_require(apply_teacher_hr(unified, teacher_hr, config),
                            "Teacher HR applied")
    } else {
      chk_info("Add-on skipped: teacher_hr", "enabled in config but no file loaded")
      hr_on <- FALSE
    }
  } else {
    chk_info("Add-on skipped: teacher_hr", "disabled in config")
  }

  if (grades_on) {
    if (!is.null(course_grades_raw) && !is.null(cs_course_coding)) {
      gr <- cs_require(process_course_grades(course_grades_raw, cs_course_coding, config),
                       "Course grades processed", min_rows = 0)
      unified <- apply_course_grades(unified, gr$student_proficiency)
      grades_rollup <- gr$course_grades_rollup
    }
  } else {
    chk_info("Add-on skipped: course_grades", "disabled in config")
  }
}
```

## 4.2 Build the rollups

```{r 04-build}
if (run_stage("04")) {
  teacher_rollup <- cs_require(build_teacher_rollup(unified, config), "Teacher rollup built")
  class_rollup   <- cs_require(build_class_rollup(unified, teacher_rollup, config),
                               "Class rollup built")
  student_rollup <- cs_require(build_student_rollup(unified, config), "Student rollup built")
}
```

```{r 04-precompute}
if (run_stage("04")) {
  size_invariant <- all(class_rollup$M_cls_class_size <= class_rollup$M_cls_num_stu + 1e-9,
                        na.rm = TRUE)

  cls_bucket_na <- class_rollup |> filter(C_class_size_exclude == "Include") |>
    summarise(n = sum(is.na(C_cls_class_size_bucket))) |> pull(n)
  tch_bucket_na <- teacher_rollup |> summarise(n = sum(is.na(M_tch_load_bucket))) |> pull(n)
  bucket_levels <- levels(.bucketize(0, config$buckets$class_size))
  util_na       <- teacher_rollup |> summarise(n = sum(is.na(M_tch_utilization))) |> pull(n)

  want_class <- c("C_class_id","C_cls_course_subject","C_cls_grade","C_cls_course_name",
                  "M_cls_class_size","C_cls_class_size_bucket","M_cls_class_weight",
                  "M_meeting_minutes","M_tch_utilization","M_tch_load")
  want_tch   <- c("D_employee_id","C_tch_course_subject","C_tch_location_name",
                  "C_teacher_load_exclude","M_tch_num_periods","M_tch_utilization",
                  "M_tch_load","M_tch_load_bucket","M_tch_num_preps")
  want_stu   <- c("D_stu_id","C_class_id","C_course_subject","M_class_weight",
                  "M_cls_class_size","C_proficiency")

  miss_class  <- setdiff(want_class, names(class_rollup))
  miss_tch    <- setdiff(want_tch,   names(teacher_rollup))
  miss_stu    <- setdiff(want_stu,   names(student_rollup))
  contract_ok <- length(miss_class) == 0 && length(miss_tch) == 0 && length(miss_stu) == 0

  dist_by_school <- class_rollup |>
    filter(C_class_size_exclude == "Include") |>
    group_by(C_cls_location_name, C_class_id) |>
    summarise(bucket = C_cls_class_size_bucket[which.max(M_cls_class_size)], .groups = "drop") |>
    count(C_cls_location_name, bucket) |>
    group_by(C_cls_location_name) |>
    mutate(pct = round(100 * n / sum(n), 1)) |>
    ungroup() |> select(-n) |>
    pivot_wider(names_from = bucket, values_from = pct, values_fill = 0, names_expand = TRUE) |>
    select(C_cls_location_name, any_of(bucket_levels)) |>
    rename(School = C_cls_location_name)
}
```

```{r}
#| output: asis
#| echo: false
cs_heading("04", "## 4.3 Checks")
```

```{r}
#| label: chk-4.1
if (run_stage("04")) {
  data.frame(
    rollup = c("class", "teacher", "student", if (!is.null(grades_rollup)) "course_grades"),
    rows   = c(nrow(class_rollup), nrow(teacher_rollup), nrow(student_rollup),
               if (!is.null(grades_rollup)) nrow(grades_rollup)),
    cols   = c(ncol(class_rollup), ncol(teacher_rollup), ncol(student_rollup),
               if (!is.null(grades_rollup)) ncol(grades_rollup))) |>
    check_gt(
      title = "4.1 \u00B7 Rollups built \u2014 shapes look right",
      note  = "One row per class-slot, per teacher, and per student-class-slot. Class rows > distinct classes is expected (grain is class \u00D7 term \u00D7 period \u00D7 rotation).")
}
```

```{r}
#| label: chk-4.2
if (run_stage("04")) {
  dist_by_school |>
    check_gt(
      title  = "4.2 \u00B7 Class size \u2014 distribution by school (% of classes per size band)",
      note   = "Adjusted size never exceeds raw headcount (double-booked students count fractionally). Each school's row shows the share of its classes in each size band.",
      status = if (size_invariant) "PASS" else "FAIL")
}
```

```{r}
#| label: chk-4.2b
if (run_stage("04")) {
  class_rollup |>
    summarise(classes = n_distinct(C_class_id),
              avg_raw_headcount = round(mean(M_cls_num_stu, na.rm = TRUE), 1),
              avg_adjusted_size = round(mean(M_cls_class_size, na.rm = TRUE), 1)) |>
    check_gt(title = "4.2 \u00B7 Adjusted vs raw headcount")
}
```

```{r}
#| label: chk-4.3
if (run_stage("04")) {
  data.frame(
    check = c("class-size NA buckets", "teacher-load NA buckets"),
    n     = c(cls_bucket_na, tch_bucket_na)) |>
    check_gt(
      title  = "4.3 \u00B7 Class-size & teacher-load buckets populated",
      note   = "Every included class needs a size bucket; every teacher a load bucket. If n > 0, a value falls outside the config buckets \u2014 extend the top bucket in config$buckets.",
      status = if (isTRUE(cls_bucket_na == 0 && tch_bucket_na == 0)) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-4.4
if (run_stage("04")) {
  teacher_rollup |>
    group_by(C_tch_location_name) |>
    summarise(teachers = n(), avg_util = round(mean(M_tch_utilization, na.rm = TRUE), 2),
              na = sum(is.na(M_tch_utilization)), .groups = "drop") |>
    check_gt(
      title  = "4.4 \u00B7 Teacher utilization populated",
      note   = "Every teacher needs a utilization value. NA means a school has no full_utilization entry \u2014 confirm config$full_utilization$by_school keys exactly match the school names above.",
      status = if (isTRUE(util_na == 0)) "PASS" else "CHECK")
}
```

```{r}
#| label: chk-4.5
if (run_stage("04")) {
  data.frame(
    rollup  = c("class", "teacher", "student"),
    missing = c(ifelse(length(miss_class) == 0, "none", paste(miss_class, collapse = ", ")),
                ifelse(length(miss_tch)   == 0, "none", paste(miss_tch,   collapse = ", ")),
                ifelse(length(miss_stu)   == 0, "none", paste(miss_stu,   collapse = ", ")))) |>
    check_gt(
      title  = "4.5 \u00B7 VISTA column contract",
      note   = "Each rollup must have the columns VISTA expects (add-on fields present even when NA). Anything under 'missing' will break VISTA \u2014 add it to the rollup builder.",
      status = if (contract_ok) "PASS" else "FAIL")
}
```

```{r}
#| label: chk-4.6
if (run_stage("04")) {
  data.frame(
    add_on  = c("teacher_hr", "course_grades"),
    enabled = c(hr_on, grades_on),
    mode    = c(config$add_ons$teacher_hr$mode %||% NA, "\u2014"),
    result  = c(if (hr_on && "C_tch_novice_indicator" %in% names(teacher_rollup))
                  "novice indicator populated" else "skipped / NA",
                if (!is.null(grades_rollup))
                  glue("{nrow(grades_rollup)} grade rows; proficiency merged") else "skipped / NA")) |>
    check_gt(
      title = "4.6 \u00B7 Add-on status",
      note  = "The optional add-ons run (or skip) based on config$add_ons. Skipped is fine \u2014 the core rollups carry those fields as NA and VISTA drops the metrics that need them.")
}
```

## 4.4 Gates

```{r 04-gates}
if (run_stage("04")) {
  chk_stop("Adjusted class size never exceeds raw headcount", size_invariant,
           offenders = class_rollup |>
             filter(M_cls_class_size > M_cls_num_stu + 1e-9) |>
             select(any_of(c("C_class_id", "M_cls_class_size", "M_cls_num_stu"))))

  chk_stop("VISTA column contract holds", contract_ok,
           detail = paste("Missing \u2014 class:", paste(miss_class, collapse = ", "),
                          "| teacher:", paste(miss_tch, collapse = ", "),
                          "| student:", paste(miss_stu, collapse = ", ")))

  chk_stop("Rollups are not empty",
           nrow(class_rollup) > 0 && nrow(teacher_rollup) > 0 && nrow(student_rollup) > 0)
}
```

## 4.5 Write to SharePoint

Nothing is written locally. `put()` reports each file into the check log, so the
run report records exactly what left the pipeline.

```{r 04-export}
if (run_stage("04")) {
  out_folder <- config$folders[[config$outputs$folder %||% "raw"]]
  out_prefix <- config$outputs$subfolder %||% "/"

  put <- function(data, name) {
    if (is.null(data) || nrow(data) == 0) {
      chk_info(paste0("Output SKIPPED (empty): ", name), 0)
      return(invisible(NULL))
    }
    path <- paste0(out_prefix, name)
    ers_write_sharepoint(data = data, folder_path = out_folder,
                         file_name_with_extension = path)
    chk_info(paste0("Output written: ", name),
             paste0(format(nrow(data), big.mark = ","), " rows x ", ncol(data), " cols"),
             detail = paste0(out_folder, path))
    invisible(data)
  }

  if (isTRUE(config$outputs$class_rollup$write))
    put(class_rollup,   config$outputs$class_rollup$file_name)
  if (isTRUE(config$outputs$teacher_rollup$write))
    put(teacher_rollup, config$outputs$teacher_rollup$file_name)
  if (isTRUE(config$outputs$student_rollup$write))
    put(student_rollup, config$outputs$student_rollup$file_name)
  if (!is.null(grades_rollup) && isTRUE(config$outputs$grades_rollup$write))
    put(grades_rollup,  config$outputs$grades_rollup$file_name)
}
```
