---
title: "Course Schedule Processing — Step 1: Explore"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

<!---
## What this file does (high level)

This is Stage 01 — the very first step. It takes the RAW data files exactly
as received from the district and does three things:

1. **Profiles** every column so you can see what you're working with
   (types, distinct values, missing data, etc.)
2. **Preps** the raw data by renaming columns to standard names, merging
   in student demographics, and filtering to in-scope rows.
3. **Exports** the prepped data for Stage 02.

This is the most hands-on stage — you'll iterate on the prep steps until
the re-profiled data looks right. No coding or exclusion logic happens here;
that's all in Stage 02.
--->


```{r setup-libraries}
# Start fresh
rm(list = ls())
# Core data manipulation
library(tidyverse)
library(data.table)
library(lubridate)
library(glue)
library(stringr)
# Reading files
library(openxlsx)
library(readxl)
library(here)
# SharePoint access
library(Microsoft365R)
library(erstools)
# Formatting and display
library(gt)
library(htmltools)
library(scales)
```

```{r setup-helpers-and-config}
# Load helper functions for profiling/exploring data
source(here::here("templates", "course_schedule", "helpers", "01_cs_explore_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "01_cs_review_helpers.R"))

# Read the district config — tells us where to find files, how to rename columns, etc.
config <- yaml::read_yaml(here::here("config", "districts", "nashville.yaml"))
raw_data_folder_path <- config$raw_data_folder_path
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue::glue("\n## {config$district} · {config$year}\n\n"))
```

```{r load-raw-data}
# Connect to SharePoint where the data files live
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
orgfiles_site     <- get_sharepoint_site(site_url = orgfiles_site_url)
internal_drive    <- orgfiles_site$get_drive("Internal")

# Load the raw course schedule — exactly as received from the district.
# No renaming, no filtering, no transformations yet.
cs_raw <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = config$raw_files$cs_raw
)

# Load student demographic data (SWD, ELL, FRL, race, etc.)
# This will be merged onto the course schedule during prep.
# Wrapped in tr_safe() so the doc renders even if the file doesn't exist yet.
stu_demographics <- tr_safe(ers_read_sharepoint(
    folder_path = raw_data_folder_path,
    file_name_with_extension = config$raw_files$stu_demographics
  ),
  otherwise = NULL
)

location_coding <- tr_safe(ers_read_sharepoint(
    folder_path = raw_data_folder_path,
    file_name_with_extension = config$raw_files$location_coding
  ),
  otherwise = NULL
)
```

## What this pass is for

The **exploratory** half of Tier 0 — raw files, district headers, no config.
Two jobs: (1) see what you have well enough to author `column_map`, and (2)
work out the **prep recipe** (join / filter / derive) that shapes the raw file
into the fields the pipeline expects. The loop is: **profile raw → prep edits →
re-profile → iterate** until the prepped profile looks right.

## Course schedule — column profile (raw)

Each column with type, distinct count, missingness, min/max, and — for
low-distinct columns — the full value set inline.

```{r cs-profile}
explore_gt(explore_profile(cs_raw), "Raw course schedule — every column")
```

## Course schedule — peek at candidate time columns

High-cardinality columns show only a sample above. List the ones that look like
they encode **time** (term / rotation / period / expression / course number)
and inspect their top values — including whether term is hiding inside the
course number.

```{r cs-peek}
# ---- Which columns might represent time/scheduling? ----
# Edit this list to match the column names you see in the profile above.
# These are the ones we want to inspect more closely.
peek_cols <- c("Term", "Rotation", "Period", "Expression", "Period Expression", "per", "term")

peek <- peek_cols[peek_cols %in% names(cs_raw)]

if (length(peek) == 0) {
  htmltools::HTML(
    "<em>None of the named columns are present — adjust peek_cols to match the profile above.</em>"
  )
} else {
  explore_render(
    setNames(
      lapply(peek, function(col_name) explore_value_counts(cs_raw, col_name, top = 40)),
      peek
    )
  )
}
```

## Student demographics coding — column profile (raw)

The join source for student demographic flags (SWD, ELL, FRL). Find the column here
that matches the student ID column in the CS profile — that's your join key.

```{r dem-profile}
explore_gt(explore_profile(stu_demographics), "Raw student demographics — every column")
```

## Prep edits (staging)

Your scratch space to shape the raw data into what the pipeline needs. Each edit
below eventually **promotes** to a durable home — this block is where you get it
right first. Fill in the `<...>` placeholders from the profiles above, then set
this chunk's `eval` to `true`.

```{r prep}
#| eval: true

# Re-read config in case you've updated it between renders
config <- yaml::read_yaml(here::here("config", "districts", "nashville.yaml"))

# ---- Step 1: RENAME columns ----
# The config's "column_map" tells us which district column name maps to which
# standard ERS name (e.g., "Student Number" -> "D_stu_id").
# This makes the rest of the pipeline work regardless of what the district
# called their columns.
cs_prepped <- cs_raw |>
  rename(any_of(unlist(config$column_map$course_schedule))) |>
  # Drop any columns that weren't mapped to a D_ name
  select(any_of(names(unlist(config$column_map$course_schedule))))

# ---- Step 2: MERGE student demographics ----
# Join demographic data (race, grade, SWD, ELL, FRL) onto the schedule.
# This lets us analyze time-use patterns by student subgroup later.
if (exists("stu_demographics")) {
  dem_cfg <- config$stu_demographics_join

  # Rename demographic columns to standard names, then keep only what we need
  demographics_cleaned <- stu_demographics |>
    rename(any_of(unlist(config$column_map$demographics))) |>
    select(all_of(c(dem_cfg$key, dem_cfg$brings_in)))

  # Apply any value recodes specified in config
  # (e.g., "Y"/"N" -> "Yes"/"No", or numeric codes -> labels)
  if (!is.null(dem_cfg$recode)) {
    for (col_name in names(dem_cfg$recode)) {
      mapping <- dem_cfg$recode[[col_name]]
      demographics_cleaned <- demographics_cleaned |>
        mutate(!!col_name := unlist(mapping)[as.character(.data[[col_name]])])
    }
  }

  # Join demographics onto the course schedule by student ID
  cs_prepped <- cs_prepped |>
    left_join(demographics_cleaned, by = dem_cfg$key)
}
# Location coding join (if available)
if (exists("location_coding")) {
  # Join demographics onto the course schedule by student ID
  cs_prepped <- cs_prepped |>
    left_join(location_coding, by = "D_location_id")
}
# ---- Step 3: FILTER to in-scope rows ----
# Drop rows where key columns are null/NA. For example, if a row has no
# student ID or no school name, it's not usable.
null_cols <- config$scope$drop_when_null

cs_prepped <- cs_prepped |>
  filter(if_all(all_of(null_cols), ~ !is.na(.) & . != "NULL"))

# Apply any additional filter expressions from config (as raw R strings).
if (!is.null(config$scope$filters)) {
  for (expr_str in config$scope$filters) {
    cs_prepped <- cs_prepped |> filter(!!rlang::parse_expr(expr_str))
  }
}

# ---- Step 4: OTHER edits ----
# eg., updating null values to 0, etc.
cs_prepped <- cs_prepped |>
  mutate(
    D_stu_ell_flag     = ifelse(is.na(D_stu_ell_flag),     0, D_stu_ell_flag),
    D_stu_swd_flag     = ifelse(is.na(D_stu_swd_flag),     0, D_stu_swd_flag),
    D_stu_poverty_flag = ifelse(is.na(D_stu_poverty_flag), 0, D_stu_poverty_flag),
    D_rotation         = "A"
  )
```


## Reinvestigate — profile after edits

Re-profile the prepped data to confirm each edit did what you expected: school
names populated, only the right schools remain, and `D_term` holds valid term
values. Iterate the prep block until this looks right.

```{r prep-profile}
if (exists("cs_prepped")) {
  explore_gt(explore_profile(cs_prepped), "After prep edits — re-profiled")
} else {
  htmltools::HTML(
    "<em>Fill in the prep chunk above and set its eval to true, then re-render.</em>"
  )
}
```

```{r prep-peek}
# ---- Re-inspect time columns after renaming ----
# Now that columns have been renamed to D_ names, check that they look right.
peek_cols <- c(
  "D_term", "D_rotation", "D_period", "D_expression", "D_location_name", "D_stu_grade")

if (!exists("cs_prepped")) {
  htmltools::HTML("<em>cs_prepped not found — run the data prep steps above first.</em>")
} else {
  peek <- peek_cols[peek_cols %in% names(cs_prepped)]

  if (length(peek) == 0) {
    htmltools::HTML(
      "<em>None of the named columns are present — adjust peek_cols to match the profile above.</em>"
    )
  } else {
    explore_render(
      setNames(
        lapply(peek, function(col_name) explore_value_counts(cs_prepped, col_name, top = 40)),
        peek
      )
    )
  }
}
```

## From here → promote the recipe

Once the prepped profile looks right, promote each piece to its durable home so
the pipeline is reusable, not one-off:

- **Renames** → `column_map` in `district_config.yml`.
- **Join** → a config location-join spec (source file · key · columns).
- **Filter** → the Tier 1 schools-in-scope decision (config).
- **Derivations** (like `D_term` from the course number) → a derivations block —
  the thing `column_map` can't express, because these fields are computed.
- **Demographics** → profile it here when it arrives; unlocks review check 0.6.

```{r prep-export}
#| eval: true

# ---- Export the prepped data for Stage 02 ----
# This is the handoff point: Stage 02 will pick up this file and apply
# course coding + exclusion rules on top of it.
if (exists("cs_prepped")) {
  ers_write_sharepoint(
    data = cs_prepped,
    folder_path = raw_data_folder_path,
    file_name_with_extension = "/2. Processed Data/01_cs_data_post_explore.csv"
  )
} else {
  message("cs_prepped not found — skipping export. Run the prep chunks above first.")
}
```

```{r coding-export}
#| eval: false

# ---- Export a blank coding template ----
# This creates the starting point for the course coding file.
# Every course starts as "UNCODED" — a human (or AI-assisted process) fills in
# the actual subject area, format, etc. Stage 02 merges this back in.
#
# You only need to run this ONCE to generate the template; after that,
# the coding file lives on SharePoint and gets edited directly.

coding_template <- cs_prepped |>
  group_by(D_course_name, D_course_id) |>
  count() |>
  ungroup() |>
  mutate(
    C_course_time          = "UNCODED",
    C_course_subject_area  = "UNCODED",
    C_course_subject       = "UNCODED",
    C_course_credit_type   = "UNCODED",
    C_course_name          = "UNCODED",
    C_course_rigor         = "UNCODED",
    C_course_rigor_detail  = "UNCODED",
    C_course_intervention  = "UNCODED",
    C_course_time_of_day   = "UNCODED",
    C_course_format        = "UNCODED",
    C_course_pathway       = "UNCODED",
    C_teacher_load_metric  = "UNCODED",
    C_class_size_metric    = "UNCODED",
    C_time_metric          = "UNCODED",
    D_course_ell_flag      = "UNCODED",
    D_course_swd_flag      = "UNCODED"
  ) |>
    select(-n)

ers_write_sharepoint(
  data = coding_template,
  folder_path = raw_data_folder_path,
  file_name_with_extension = "/1. Coding/cs_course_coding_template.csv"
)
```
