---
title: "Course Schedule Pipeline"
date: today
params:
  district: "nashville"
  refresh_raw: false      # TRUE re-reads every source from SharePoint
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE)
config <- yaml::read_yaml(
  here::here("config", "districts", paste0("cs_", params$district, ".yaml")))
source(here::here(config$helpers_path, "00_cs_setup.R"))
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue::glue("\n## {config$district} · {config$year}\n\n"))
```

# What this file is

One document, run top to bottom, replacing `01_cs_explore.qmd` and
`02_cs_processing_and_rollups.qmd`. Those two were joined by a file —
`01_cs_data_post_explore.csv` was written to SharePoint at the end of the first
and read back at the start of the second. That seam is gone; the data stays in
memory.

The engine is unchanged. Every weight, explosion, flag, resolution and rollup
still comes from `templates/course_schedule/helpers/`.

**Two ways of checking, both reporting to one log.**

- `check_gt()` renders a table with a PASS / CHECK badge. Most checks are this:
  evidence to eyeball.
- `chk_stop()` halts. Reserved for structural failures — a join that changed
  the row count, an engine that returned nothing.

Both now write to a single check log, printed at the end, so a long render
finishes with a summary rather than a badge you had to have been watching for.

**On `tr_safe()`.** It swallows any error and returns a fallback. That is right
for an input the district may not have supplied and wrong for the engine, where
it turned a failure into `NULL` and let every later stage degrade quietly. It is
now `cs_optional()` for the former and `cs_require()` for the latter.

| # | Stage | Replaces | Status |
|---|-------|----------|--------|
| 00 | Setup and config | Part 0 of both files | built |
| 01 | Explore and prep | `01_cs_explore.qmd` | built |
| 02 | Coding and exclusions | `02_...` §02 | next |
| 03 | Weights, minutes, DB resolution | `02_...` §03 | |
| 04 | Rollups and export | `02_...` §04 | |

---

# Stage 00 — Setup and configuration

```{r cfg}
chk_reset()
chk_set_output_dir(file.path(cache_dir(config), "failed_checks"))
chk_stage("Stage 00 — Setup")

chk_info("District", config$district)
chk_info("Year", config$year)
chk_info("Time format", config$time_format)
chk_info("Enrollment type", config$intake$enrollment_type)
chk_info("Class ID recipe", paste(unlist(config$class_id$recipe), collapse = " + "))
chk_info("Cache", cache_dir(config))
```

Config values the run depends on, printed so a reviewer can see the decisions
without reading any code.

```{r cfg-decisions}
tibble::tibble(
  setting = c("Term weights", "Rotation default", "Period default",
              "Bell minutes default", "Coding join key",
              "Uncoded threshold", "Teacher DB consolidate max",
              "Self-contained threshold"),
  value = c(
    paste(names(config$weights$term), unlist(config$weights$term),
          sep = "=", collapse = ", "),
    as.character(config$weights$rotation$default),
    as.character(config$weights$period$default),
    as.character(config$bell_minutes$default),
    paste(unlist(config$coding$join_key), collapse = " + "),
    paste0(config$coding$uncoded_flag_threshold_pct, "%"),
    as.character(config$teacher_db$consolidate_max_size),
    paste0(config$self_contained$swd_ell_threshold_pct, "%")))
```

```{r cfg-guard}
required <- c("district", "year", "raw_data_folder_path", "raw_files",
              "column_map", "class_id", "time_format", "weights",
              "bell_minutes", "coding", "exclusions")
missing_keys <- required[!vapply(required, function(k) !is.null(config[[k]]), logical(1))]

chk_stop("Config has every required section",
         length(missing_keys) == 0,
         detail = if (length(missing_keys) > 0)
           paste("Missing:", paste(missing_keys, collapse = ", ")))

# Assign first, then check. An assignment inside chk_stop()'s condition would
# bind inside that function's frame, not here, and the object would vanish.
spec         <- cs_require(expr_spec(config),         "Expression spec builds")
term_expands <- cs_require(term_expands_spec(config), "Term expansion spec builds")
```

---

# Stage 01 — Explore and prep

Replaces `01_cs_explore.qmd`. Same profiling and the same prep recipe; what has
gone is the export at the end, since Stage 02 now runs in the same session.

## 1.1 Load sources

Each file is read from SharePoint once and cached as `.rds`. **The cache does
not know when the district re-sends a file** — set `refresh_raw: true` in the
header when new data lands.

The course schedule is required. Everything else is optional: a district that
has not sent demographics or HR data should still get a render, with the gap
reported rather than an error.

```{r load}
chk_stage("Stage 01 — Explore and prep")

cs_raw <- cs_require(
  read_source("cs_raw", config, refresh = params$refresh_raw),
  "Course schedule loaded")

stu_demographics <- cs_optional(
  read_source("stu_demographics", config, refresh = params$refresh_raw),
  "student demographics")

teacher_hr <- cs_optional(
  read_source("hr_raw", config, refresh = params$refresh_raw),
  "teacher HR")

location_coding <- cs_optional(
  read_source("location_coding", config, refresh = params$refresh_raw),
  "location coding")

cs_course_coding <- cs_optional(
  read_source("course_coding", config, refresh = params$refresh_raw),
  "course coding")
```

## 1.2 Raw column profile

Truly raw — original headers, nothing mapped. This is what the `column_map` is
written against.

```{r raw-profile}
explore_gt(explore_profile(cs_raw), "Course schedule · raw column profile")
```

```{r raw-profile-dem}
if (!is.null(stu_demographics))
  explore_gt(explore_profile(stu_demographics), "Student demographics · raw")
```

```{r raw-profile-hr}
if (!is.null(teacher_hr))
  explore_gt(explore_profile(teacher_hr), "Teacher HR · raw")
```

## 1.3 Time columns

The columns that might encode when a class meets. Getting this wrong makes
every weight wrong, so it is worth looking rather than assuming.

```{r raw-time-peek}
time_candidates <- intersect(
  unlist(config$column_map$course_schedule[
    c("D_period", "D_term", "D_course_section_id")]),
  names(cs_raw))

if (length(time_candidates) > 0) {
  explore_render(setNames(
    lapply(time_candidates, function(cl) explore_value_counts(cs_raw, cl, top = 25)),
    time_candidates))
}
```

## 1.4 Duplicates in the raw file

```{r raw-dupes}
n_raw   <- nrow(cs_raw)
n_uniq  <- nrow(dplyr::distinct(cs_raw))

data.frame(
  measure = c("Rows as received", "Distinct rows", "Exact duplicate rows"),
  value   = c(n_raw, n_uniq, n_raw - n_uniq)
) |>
  check_gt(
    title  = "1.4 · Exact duplicate rows in the raw file",
    note   = "Fully identical rows are dropped during prep. A large count is worth asking the district about \u2014 it usually means an export ran twice or a join upstream fanned out.",
    status = if (n_raw == n_uniq) "PASS" else "CHECK")
```

## 1.5 Prep

Rename to standard names, merge demographics and location coding, apply scope,
flag the student snapshot, and apply the derivations. Each step reports what it
cost.

### Rename

```{r prep-rename}
cs_map <- unlist(config$column_map$course_schedule)

cs_prepped <- cs_raw |>
  dplyr::rename(dplyr::any_of(cs_map)) |>
  dplyr::select(dplyr::any_of(names(cs_map))) |>
  dplyr::distinct()

mapped   <- names(cs_map)[names(cs_map) %in% names(cs_prepped)]
unmapped <- setdiff(names(cs_map), names(cs_prepped))

data.frame(
  standard_name   = names(cs_map),
  district_column = unname(cs_map),
  landed          = ifelse(names(cs_map) %in% names(cs_prepped), "yes", "MISSING")
) |>
  check_gt(
    title  = "1.5a · Did every mapped column land?",
    note   = "A MISSING row means column_map names a district column that is not in the file. Fix the map here rather than patching downstream.",
    status = if (length(unmapped) == 0) "PASS" else "CHECK")

chk_info("Rows after dropping exact duplicates",
         paste0(format(nrow(cs_prepped), big.mark = ","), " of ",
                format(nrow(cs_raw), big.mark = ",")))
```

### Student demographics

```{r prep-demographics}
if (!is.null(stu_demographics)) {
  dem_cfg <- config$stu_demographics_join

  demographics_cleaned <- stu_demographics |>
    dplyr::rename(dplyr::any_of(unlist(config$column_map$demographics))) |>
    dplyr::select(dplyr::all_of(c(dem_cfg$key, unlist(dem_cfg$brings_in)))) |>
    dplyr::distinct()

  if (!is.null(dem_cfg$recode)) {
    for (col_name in names(dem_cfg$recode)) {
      mapping <- dem_cfg$recode[[col_name]]
      demographics_cleaned <- demographics_cleaned |>
        dplyr::mutate(!!col_name := unlist(mapping)[as.character(.data[[col_name]])])
    }
  }

  # join_checked rather than left_join: a demographics file with duplicate
  # student rows would silently multiply the schedule, and that is exactly the
  # kind of thing a row-count check exists to catch.
  cs_prepped <- cs_prepped |>
    join_checked(demographics_cleaned, by = dem_cfg$key,
                 check = "Student demographics onto schedule")
} else {
  chk_info("Demographics merge skipped", "no demographics file")
}
```

### Location coding

```{r prep-location}
if (!is.null(location_coding)) {
  cs_prepped <- cs_prepped |>
    join_checked(location_coding, by = "D_location_id",
                 check = "Location coding onto schedule")
} else {
  chk_info("Location coding merge skipped", "no location file")
}
```

### Scope

```{r prep-scope}
cs_prepped <- apply_scope(cs_prepped, config$scope, id_col = "D_stu_id")
```

### Student snapshot

For longitudinal enrollment data, each row is flagged on whether the student
was enrolled on that term's snapshot date. Skipped entirely for districts whose
extract is already a snapshot.

```{r prep-snapshot}
if (!is.null(config$intake$enrollment_type) &&
    config$intake$enrollment_type != "snapshot") {

  cs_prepped <- cs_require(apply_stu_snapshot(cs_prepped, config),
                           "Student snapshot flag applied")

  cs_prepped |>
    dplyr::count(D_term, C_stu_snapshot, name = "n_records") |>
    tidyr::pivot_wider(names_from = C_stu_snapshot, values_from = n_records,
                       values_fill = 0) |>
    dplyr::mutate(snapshot_date = unlist(config$intake$snapshot_dates)[as.character(D_term)]) |>
    check_gt(
      title = "1.5b · Snapshot inclusion by term",
      note  = "Each term uses its own snapshot date. A term with almost nothing included usually means the date sits outside that term's enrolment window \u2014 check it against the enter/exit distribution before trusting it.")
} else {
  chk_info("Snapshot flagging skipped", "extract is already a snapshot")
}
```

### Derivations

Fields the pipeline computes rather than reads. These were hardcoded in the old
prep chunk, which put a district-specific assumption inside a file meant to be
reusable.

```{r prep-derivations}
null_to_zero <- intersect(unlist(config$derivations$null_to_zero %||% character()),
                          names(cs_prepped))

if (length(null_to_zero) > 0) {
  before <- vapply(null_to_zero, function(c) sum(is.na(cs_prepped[[c]])), numeric(1))
  cs_prepped <- cs_prepped |>
    dplyr::mutate(dplyr::across(dplyr::all_of(null_to_zero), ~ dplyr::coalesce(.x, 0)))
  chk_info("Null flags set to zero", "see table",
           data = tibble::tibble(column = null_to_zero, n_filled = unname(before)))
}

for (nm in names(config$derivations$constants %||% list())) {
  cs_prepped[[nm]] <- config$derivations$constants[[nm]]
  chk_info(paste0("Constant applied: ", nm),
           config$derivations$constants[[nm]],
           detail = "declared in config$derivations$constants")
}
```

## 1.6 Duplicates after prep

Confirms the merges did not fan out.

```{r prep-dupes}
n_prep  <- nrow(cs_prepped)
n_prepu <- nrow(dplyr::distinct(cs_prepped))

data.frame(
  measure = c("Rows after prep", "Distinct rows", "Exact duplicate rows"),
  value   = c(n_prep, n_prepu, n_prep - n_prepu)
) |>
  check_gt(
    title  = "1.6 · Duplicates after prep",
    note   = "Prep should not create duplicates. If this is non-zero when 1.4 was zero, a merge fanned out.",
    status = if (n_prep == n_prepu) "PASS" else "CHECK")
```

```{r prep-grain}
key_cols <- intersect(c("D_stu_id", "D_course_id", "D_course_section_id",
                        "D_term", "D_period"), names(cs_prepped))

dupe_keys <- cs_prepped |>
  dplyr::count(dplyr::across(dplyr::all_of(key_cols)), name = "n_rows") |>
  dplyr::filter(n_rows > 1)

dupe_keys |>
  dplyr::count(n_rows, name = "n_key_combinations") |>
  check_gt(
    title  = paste0("1.6b \u00B7 Rows per ", paste(key_cols, collapse = " + ")),
    note   = "More than one row per student per section per time slot means the grain is not what the pipeline assumes. Worth understanding before class sizes are computed.",
    status = if (nrow(dupe_keys) == 0) "PASS" else "CHECK")
```

## 1.7 Prepped profile

```{r prep-profile}
explore_gt(explore_profile(cs_prepped), "Course schedule · after prep")
```

```{r prep-time-peek}
time_cols_now <- intersect(config$time_cols[[config$time_format]], names(cs_prepped))

if (length(time_cols_now) > 0) {
  explore_render(setNames(
    lapply(time_cols_now, function(cl) explore_value_counts(cs_prepped, cl, top = 25)),
    time_cols_now))
}
```

## 1.8 Coding template

The one-time bootstrap for a district with no coding file yet. Off by default;
set `coding_template$write: true`, render once, then set it back.

The columns it groups by come from config. The previous version hardcoded
`D_course_subject_id`, a Detroit field absent from the Nashville data, so this
chunk would have errored if it had ever been run here.

```{r coding-template}
#| eval: !expr isTRUE(config$coding_template$write)

tpl_group <- intersect(unlist(config$coding_template$group_by), names(cs_prepped))

chk_stop("Coding template group_by columns exist",
         length(tpl_group) == length(unlist(config$coding_template$group_by)),
         detail = paste("Missing:", paste(
           setdiff(unlist(config$coding_template$group_by), names(cs_prepped)),
           collapse = ", ")))

coding_template <- cs_prepped |>
  dplyr::distinct(dplyr::across(dplyr::all_of(tpl_group)))

for (f in unlist(config$coding_template$uncoded_fields)) {
  coding_template[[f]] <- "UNCODED"
}

chk_info("Coding template rows", nrow(coding_template))

ers_write_sharepoint(
  data = coding_template,
  folder_path = config$raw_data_folder_path,
  file_name_with_extension = config$coding_template$file_name)
```

## 1.9 Handoff

`cs_data` is what Stage 02 picks up. In the old pipeline this was a CSV on
SharePoint; now it is an object in memory, and the row count below is the last
number Stage 01 is responsible for.

```{r prep-handoff}
cs_data <- cs_prepped

chk_info("Stage 01 output",
         paste0(format(nrow(cs_data), big.mark = ","), " rows x ",
                ncol(cs_data), " cols"))
chk_info("Distinct students",  dplyr::n_distinct(cs_data$D_stu_id))
chk_info("Distinct teachers",  dplyr::n_distinct(cs_data$D_employee_id))
chk_info("Distinct schools",   dplyr::n_distinct(cs_data$D_location_id))
```

---

# Stage 02 — Coding and exclusions

> Next step.

---

# Stage 03 — Weights, minutes and DB resolution

---

# Stage 04 — Rollups and export

---

# Check report

```{r final-report}
check_log <- chk_report(print_table = FALSE)
```

```{r final-table}
check_log |> dplyr::select(seq, stage, check, severity, result, value)
```
