---
title: "Course Schedule Processing — 02 · Coding, Resolution & Rollups"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

<!---
## What this file does (high level)

This is Stages 02–04 of the course schedule pipeline. It picks up the cleaned
data from Stage 01 and does three things:

1. **02 — Coding & Exclusions**: Merges course coding, applies exclusion rules,
   and runs quality checks (2.0–2.8).
2. **03 — Weights & Resolution**: Derives class weights/minutes, runs the
   double-booking resolution engine, and proves resolution worked (3.0–3.7).
3. **04 — Rollups & Export**: Builds class/teacher/student rollups for VISTA
   and exports them as .xlsx files (4.1–4.6).
--->


```{r setup-libraries}
rm(list = ls())
library(gt)
library(dplyr)
library(tidyr)
library(glue)
library(yaml)
library(Microsoft365R)
library(erstools)
library(openxlsx)
```

```{r setup-config}
# Read config first so helpers_path is available for sourcing below
config <- yaml::read_yaml(here::here("config", "districts", "cs_nashville.yaml"))
```

```{r setup-helpers}
# Source helpers using the repo-agnostic path from config
hp <- here::here(config$helpers_path)
source(file.path(hp, "01_cs_explore_helpers.R"))
source(file.path(hp, "01_cs_review_helpers.R"))
source(file.path(hp, "02_cs_db_helpers.R"))
source(file.path(hp, "02_cs_expression_helpers.R"))
source(file.path(hp, "03_cs_resolution_helpers.R"))
source(file.path(hp, "03_cs_weight_helpers.R"))
source(file.path(hp, "03_cs_explosion_helpers.R"))
source(file.path(hp, "03_cs_flagging_helpers.R"))
source(file.path(hp, "03_cs_consolidation_helpers.R"))
source(file.path(hp, "03_cs_resolution_helpers.R"))
source(file.path(hp, "03_cs_unification_helpers.R"))
source(file.path(hp, "04_cs_rollup_helpers.R"))
```

```{r setup-config-and-data}
# Expression spec and term expansion (needed for weights + resolution)
spec         <- tr_safe(expr_spec(config), NULL)
term_expands <- tr_safe(term_expands_spec(config), NULL)
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue::glue("\n## {config$district} · {config$year}\n\n"))

# Connect to SharePoint
orgfiles_site  <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles")
internal_drive <- orgfiles_site$get_drive("Internal")
raw_data_folder_path <- config$raw_data_folder_path
# Load Stage 01 output
cs_data <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = "/2. Processed Data/01_cs_data_post_explore.csv"
)

# Load course coding lookup
cs_course_coding <- tr_safe(ers_read_sharepoint(
    folder_path = source_of_truth_folder_path, # ARNOLD: using Source of Truth folder path
    file_name_with_extension = config$source_of_truth_files$course_coding # ARNOLD: using Source of Truth file path
  ),
  otherwise = NULL
)
```

```{r 02-define-exclusion-logic}
#| echo: true

apply_exclusions <- function(df, rules) {
  for (field in names(rules)) {
    should_exclude <- rep(FALSE, nrow(df))

    for (rule in rules[[field]]) {
      if (!rule$col %in% names(df)) next

      column_values <- df[[rule$col]]
      rule_values   <- rule$val

      matches_rule <- switch(
        rule$op,
        eq     = column_values == rule_values,
        ne     = column_values != rule_values,
        `in`   = column_values %in% rule_values,
        not_in = !(column_values %in% rule_values),
        rep(FALSE, length(column_values))
      )
      matches_rule[is.na(matches_rule)] <- FALSE
      should_exclude <- should_exclude | matches_rule
    }

    df[[field]] <- ifelse(should_exclude, "Exclude", "Include")
  }

  df
}
```

```{r 02-merge-course-coding}
# Merge course coding and apply exclusions
join_key   <- if (!is.null(config$coding$join_key)) unlist(config$coding$join_key) else "D_course_name"
merge_diag <- NULL

if (!is.null(cs_course_coding)) {
  cs_data <- tr_safe({

    coded_cols <- setdiff(names(cs_course_coding), join_key)
    coding_lookup <- cs_course_coding
    coding_lookup$.coded <- TRUE

    n_before   <- nrow(cs_data)
    key_unique <- !any(duplicated(coding_lookup[, join_key]))

    joined <- cs_data |>
      select(-any_of(coded_cols)) |>
      left_join(coding_lookup, by = join_key)

    unmatched <- joined |>
      filter(is.na(.coded)) |>
      distinct(across(all_of(join_key)))

    merge_diag <<- list(
      key          = paste(join_key, collapse = " + "),
      key_unique   = key_unique,
      n_before     = n_before,
      n_after      = nrow(joined),
      row_stable   = identical(n_before, nrow(joined)),
      n_unmatched  = nrow(joined |> filter(is.na(.coded))),
      unmatched_courses = as.data.frame(unmatched) |>
        arrange(across(all_of(join_key)))
    )

    joined |>
      select(-.coded) |>
      apply_exclusions(config$exclusions)

  }, otherwise = cs_data)
}

# Build class ID
cs_data$C_class_id <- build_class_id(cs_data, config$class_id$recipe)

# Quick helper
has_cols <- function(...) all(c(...) %in% names(cs_data))
```

## How to read this

The first formal review — `01` is investigation-only. Every check either
**computes** a status from the data or is **evidence-only** (a table to eyeball,
no badge). Nothing here is hand-maintained prose wearing a status. Checks detect
what's present and degrade gracefully: a not-yet-loaded input shows "data not
loaded," never a crash.

Golden rule: **processing flags, it never deletes** — exclusions are columns a
later metric can honor or ignore.

## 02 — Coding & exclusions review

```{r}
#| label: chk-2.0
expected_fields <- names(config$column_map$course_schedule)
data.frame(
  field   = expected_fields,
  present = ifelse(expected_fields %in% names(cs_data), "yes", "MISSING")
) |>
  check_gt(
    title  = "2.0 \u00B7 Did standardization from 01 land?",
    note   = "Every field in column_map should be present with its D_ name. If one shows MISSING, fix the rename map in 01 \u2014 don't patch it downstream.",
    status = if (length(setdiff(expected_fields, names(cs_data))) == 0) "PASS" else "CHECK"
  )
```

```{r}
#| label: chk-2.1
cs_data |>
  count(D_term, name = "n_records") |>
  arrange(D_term) |>
  mutate(weight = sapply(as.character(D_term), \(t) {
    w <- config$weights$term[[t]]; if (is.null(w)) "MISSING from config" else as.character(w) })) |>
  check_gt(
    title  = "2.1 \u00B7 Term structure \u2014 is every term value weightable?",
    note   = "Every term in the data needs a configured weight. If any shows MISSING from config, add it to weights.term.",
    status = if (all(as.character(unique(na.omit(cs_data$D_term))) %in% names(config$weights$term))) "PASS" else "CHECK"
  )
```

```{r}
#| label: chk-2.2
included_data <- cs_data
if (has_cols("C_course_time_exclude", "C_class_size_and_teacher_load_exclude"))
  included_data <- included_data |>
    filter(C_course_time_exclude == "Include", C_class_size_and_teacher_load_exclude == "Include")

records_per_student <- included_data |> count(D_location_name, D_stu_id, name = "records_per_student")
time_cols <- config$time_cols[[config$time_format]]

included_data |>
  group_by(D_location_name) |>
  summarise(n_students = n_distinct(D_stu_id), n_teachers = n_distinct(D_employee_id),
            terms = paste(sort(unique(na.omit(D_term))), collapse = ", "),
            n_time_slots = n_distinct(!!!syms(time_cols)), .groups = "drop") |>
  left_join(records_per_student |> group_by(D_location_name) |>
              summarise(avg_records_per_student = round(mean(records_per_student), 1), .groups = "drop"),
            by = "D_location_name") |>
  arrange(D_location_name) |>
  check_gt(
    title = "2.2 \u00B7 Records per student, by school",
    note  = "Per-school record counts and time-slot variety should look plausible for HS schedules. A school far off the others is worth a look."
  )
```

```{r}
#| label: chk-2.3
demographic_fields <- unlist(config$stu_demographics_join$brings_in)
data.frame(
  field      = demographic_fields,
  present    = ifelse(demographic_fields %in% names(cs_data), "yes", "no"),
  pct_filled = vapply(demographic_fields, function(f)
    if (f %in% names(cs_data)) round(100 * mean(!is.na(cs_data[[f]])), 1) else NA_real_, numeric(1))
) |>
  check_gt(
    title  = "2.3 \u00B7 Student demographics \u2014 present and populated?",
    note   = "The demographic fields from the 01 merge should exist and be well-filled. Low pct_filled may be real (a district that doesn't track a flag) or a merge miss.",
    status = if (all(demographic_fields %in% names(cs_data))) "PASS" else "CHECK"
  )
```

```{r}
#| label: chk-2.4
merge_summary <- data.frame(
  check  = c("key unique (no multiplication)", "rows stable pre/post", "unmatched course rows"),
  result = c(ifelse(merge_diag$key_unique, "yes", "NO"),
             paste0(format(merge_diag$n_before, big.mark = ","), " -> ", format(merge_diag$n_after, big.mark = ",")),
             format(merge_diag$n_unmatched, big.mark = ",")))
merge_summary |>
  check_gt(
    title  = "2.4 \u00B7 Did the course-coding merge land cleanly?",
    note   = "The coding join should have a unique key (no row multiplication), stable row count, and zero unmatched courses. Unmatched rows below need coding added.",
    status = if (isTRUE(merge_diag$key_unique) && isTRUE(merge_diag$row_stable) && merge_diag$n_unmatched == 0) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-2.4b
merge_diag$unmatched_courses |>
  check_gt(title = "2.4 \u00B7 Course names that didn't get coded")
```

```{r}
#| label: chk-2.5
cs_data |>
  count(C_course_subject_area, name = "n_records") |>
  mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
  arrange(C_course_subject_area) |>
  check_gt(
    title  = glue("2.5 \u00B7 Coding output \u2014 nothing left uncoded? ({round(100*mean(cs_data$C_course_subject_area=='UNCODED', na.rm=TRUE),1)}% uncoded)"),
    note   = glue("Uncoded should be \u2264 {config$coding$uncoded_flag_threshold_pct}% of records, and distributions should look plausible. High UNCODED means courses need coding in 01."),
    status = tr_threshold_status(100 * mean(cs_data$C_course_subject_area == "UNCODED", na.rm = TRUE),
                                 config$coding$uncoded_flag_threshold_pct, "below") |>
      (\(s) if (s == "ok") "PASS" else "CHECK")()
  )
```

```{r}
#| label: chk-2.5b
cs_data |>
  count(C_course_subject_area, C_course_subject, name = "n_records") |>
  mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
  arrange(C_course_subject_area) |>
  check_gt(title = "2.5 \u00B7 Subject \u2014 % of records")
```

```{r}
#| label: chk-2.6
cs_data |>
  filter(C_course_time_exclude == "Exclude" | C_class_size_and_teacher_load_exclude == "Exclude") |>
  count(C_course_subject, D_course_name, C_course_time_exclude,
        C_class_size_and_teacher_load_exclude, name = "n_courses", sort = TRUE) |>
  arrange(C_course_subject) |>
  check_gt(
    title = "2.6 \u00B7 Exclusions \u2014 right thing dropping, for the right reason?",
    note  = "Every excluded course should trace to a coding value that justifies it (lunch, study hall, non-instructional). Anything here that looks like a real class means a coding fix in 01."
  )
```

```{r}
#| label: chk-2.7
time_cols <- config$time_cols[[config$time_format]]
non_standard <- cs_data |>
  count(across(all_of(c("D_location_name", time_cols))), C_course_format, name = "n_records") |>
  pivot_wider(names_from = C_course_format, values_from = n_records, values_fill = 0) |>
  filter(`Standard` == 0) |>
  arrange(across(all_of(c("D_location_name", time_cols))))
non_standard |>
  check_gt(
    title = "2.7 \u00B7 Which time blocks don't carry real instruction?",
    note  = "Time-slots with no standard-format records (below), or very few distinct courses (next table), are identified so they fall out of instructional analysis."
  )
```

```{r}
#| label: chk-2.7b
time_cols <- config$time_cols[[config$time_format]]
max_courses <- if (!is.null(config$coding$max_unique_courses)) config$coding$max_unique_courses else 5
cs_data |>
  group_by(across(all_of(c("D_location_name", time_cols)))) |>
  mutate(n_courses = n_distinct(D_course_name)) |> ungroup() |>
  filter(n_courses <= max_courses, !C_course_subject_area %in% c("Core", "Non-Core")) |>
  count(across(all_of(c("D_location_name", time_cols))), C_course_subject, D_course_name, n_courses, name = "n_records") |>
  arrange(D_location_name, n_courses, desc(n_records)) |>
  check_gt(title = glue("2.7 \u00B7 Low-diversity time-slots (\u2264 {max_courses} courses, excl. Core and Non-Core)"))
```

```{r}
#| label: chk-2.8
cols <- config$class_id$recipe[config$class_id$recipe %in% names(cs_data)]
cs_data |>
  filter(!is.na(C_class_id)) |>
  group_by(across(all_of(c("C_class_id", cols)))) |>
  summarise(course = dplyr::first(D_course_name), n_students = n_distinct(D_stu_id), .groups = "drop") |>
  arrange(desc(n_students)) |>
  (\(d) slice_sample(d, n = min(8, nrow(d))))() |>
  check_gt(
    title = glue("2.8 \u00B7 Class ID \u2014 what defines a class here? ({paste(config$class_id$recipe, collapse = ' + ')})"),
    note  = "The recipe should mark a place-in-time without over-splitting real classes. Use SECTION, never the teacher \u2014 a teacher in the id hides teacher double-bookings."
  )
```


*Next -> stage 03: weights, double-booking resolution, and proof.*

## 03 — Weights, minutes & DB resolution

```{r 03-derive-weights-and-minutes}
# Derive class weights and minutes
cs_data <- tr_safe({
  derive_class_weight_minutes(cs_data, config, spec)
}, otherwise = cs_data)

# Initialize student and teacher weight views
cs_data$M_stu_class_weight <- tr_safe(cs_data$M_cls_class_weight, NA)
cs_data$M_tch_class_weight <- tr_safe(cs_data$M_cls_class_weight, NA)
```

```{r 03-run-resolution-engine}
# Row ID for unification matching
cs_data$H_row_id_before_explosions <- seq_len(nrow(cs_data))

# Resolution engine pipeline
exploded    <- tr_safe(explode_for_resolution(cs_data, config, spec), NULL)
flagged_all <- tr_safe(flag_double_bookings(exploded, config), NULL)
flagged     <- tr_safe(flag_double_bookings(exploded, config, filter_to_db = TRUE), NULL)
res         <- tr_safe(resolve_double_bookings(flagged, config), NULL)
resolved    <- tr_safe(res$resolved, NULL)
mapping     <- tr_safe(res$mapping, NULL)
loadoff     <- tr_safe(teacher_loadoff(resolved, config), NULL)
unified     <- tr_safe(unify_resolution(resolved, cs_data, loadoff), NULL)

# DB sets for checks
flagged_students <- tr_safe(unique(flagged_all$D_stu_id[flagged_all$H_db_student_row == 1]), character(0))
flagged_teachers <- tr_safe(unique(flagged_all$D_employee_id[flagged_all$H_db_teacher_row == 1]), character(0))

pct_of <- function(a, b) { r <- round(100 * a / b, 1); r[is.na(b) | b == 0] <- NA_real_; r }

# Precompute check values
multi_teacher <- tr_safe(resolved |> filter(startsWith(C_class_id_post_teachers, "CONS_")) |>
  group_by(C_class_id_post_teachers) |> summarise(n_tch = n_distinct(D_employee_id), .groups="drop") |>
  filter(n_tch > 1), data.frame())
consol_sum <- tr_safe(consolidation_summary(mapping), list())
teach_db_before <- length(flagged_teachers)
teach_db_after  <- tr_safe({
  rf <- resolved |> mutate(C_class_id = C_class_id_post_teachers) |>
    select(-any_of(c("H_db_teacher_row","H_db_student_row"))); rf <- flag_teacher_db(rf)
  n_distinct(rf$D_employee_id[rf$H_db_teacher_row == 1]) }, NA)
sec_before <- tr_safe(n_distinct(mapping$C_class_id_original), NA)
sec_after  <- tr_safe(n_distinct(mapping$C_class_id_resolved), NA)

# Monotonic shrink
mono <- tr_safe(with(resolved,
  all(M_class_weight_adj_same_course >= M_class_weight_post_class    - 1e-9, na.rm=TRUE) &&
  all(M_class_weight_post_class      >= M_class_weight_post_teachers - 1e-9, na.rm=TRUE) &&
  all(M_class_weight_post_teachers   >= M_class_weight_post_student  - 1e-9, na.rm=TRUE)), FALSE)

# Unification accounting
n_matched <- tr_safe(cs_data |> semi_join(distinct(resolved, H_row_id_before_explosions),
                     by="H_row_id_before_explosions") |> nrow(), NA)
unif_ok <- tr_safe(isTRUE(nrow(unified) == nrow(resolved) + (nrow(cs_data) - n_matched)), FALSE)
stu_before <- tr_safe(exploded |> group_by(D_stu_id) |> summarise(b = sum(M_class_weight_exploded, na.rm=TRUE), .groups="drop"), NULL)
stu_after  <- tr_safe(resolved |> group_by(D_stu_id) |> summarise(a = sum(M_class_weight_post_student, na.rm=TRUE), .groups="drop"), NULL)
stu_ba <- tr_safe(stu_before |> left_join(stu_after, by="D_stu_id") |> mutate(a = coalesce(a, b)), NULL)

# Finale example pickers
ex_student <- tr_safe(resolved |> filter(H_db_student_row == 1) |> slice_head(n=1) |> pull(D_stu_id), NA)
ex_cons    <- tr_safe(consol_sum[["Example consolidations (largest first)"]]$C_class_id_consolidated[1], NA) %||% NA
trace_cols <- function(d) d |> select(any_of(c(
  "D_stu_id","D_course_name","C_period_exploded","C_rotation_exploded","C_class_id_post_teachers",
  "M_class_weight_exploded","M_class_weight_post_class","M_class_weight_post_teachers",
  "M_class_weight_post_student","M_student_count_adjusted")))
```

```{r}
#| label: chk-3.0
tk   <- config$time_cols[[config$time_format]]
keep <- intersect(c("C_class_id", "D_location_name", tk, "D_term",
                    "D_course_name", "M_cls_class_weight", "M_cls_class_minutes"), names(cs_data))
base <- cs_data |> filter(!is.na(C_class_id)) |> distinct(across(all_of(keep)))
uniq <- base |> distinct(M_cls_class_weight, .keep_all = TRUE)
d <- if (nrow(uniq) < 10) bind_rows(uniq, base |> anti_join(uniq, by = "C_class_id") |> slice_sample(n = 10 - nrow(uniq))) else uniq
d <- d |> arrange(desc(M_cls_class_weight))
d$time   <- apply(d[intersect(tk, names(d))], 1, function(r) paste(r[!is.na(r)], collapse = " "))
d$course <- d$D_course_name
d |> select(any_of(c("D_location_name", "time", "D_term", "course", "M_cls_class_weight", "M_cls_class_minutes"))) |>
  check_gt(
    title = "3.0 \u00B7 Class weight & minutes \u2014 do a few real classes look right?",
    note  = "term \u00D7 rotation \u00D7 period should reconcile by hand; minutes come from the bell schedule. They're separate on purpose. Shows every distinct weight value."
  )
```

```{r}
#| label: chk-3.1
student_weight_summary <- cs_data |>
  filter(C_course_time_exclude == "Include") |>
  distinct(D_location_name, D_stu_id, C_class_id, M_stu_class_weight) |>
  group_by(D_location_name, D_stu_id) |>
  summarise(total_weight = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |>
  group_by(D_location_name) |>
  summarise(n_students = n_distinct(D_stu_id), avg_weight_per_student = round(mean(total_weight), 2), .groups = "drop")
student_weight_summary |>
  check_gt(
    title = "3.1 \u00B7 Weight per student, by school",
    note  = "Sums should reflect the schedule structure (blocks vs. periods) \u2014 NOT forced to a single target. Teacher version in the next table."
  )
```

```{r}
#| label: chk-3.1b
cs_data |>
  filter(C_class_size_and_teacher_load_exclude == "Include") |>
  distinct(D_location_name, D_employee_id, C_class_id, M_tch_class_weight) |>
  group_by(D_location_name, D_employee_id) |>
  summarise(total_weight = sum(M_tch_class_weight, na.rm = TRUE), .groups = "drop") |>
  group_by(D_location_name) |>
  summarise(n_teachers = n_distinct(D_employee_id), avg_weight_per_teacher = round(mean(total_weight), 2), .groups = "drop") |>
  check_gt(title = "3.1 \u00B7 Weight per teacher, by school")
```

```{r}
#| label: chk-3.2
cs_data |>
  filter(C_course_time_exclude == "Include") |>
  distinct(D_location_name, D_stu_id, C_class_id, M_stu_class_weight, M_cls_class_minutes) |>
  group_by(D_location_name, D_stu_id) |>
  summarise(total_weight = sum(M_stu_class_weight, na.rm = TRUE),
            total_minutes_per_cycle = sum(M_cls_class_minutes, na.rm = TRUE), .groups = "drop") |>
  group_by(D_location_name) |>
  summarise(avg_weight = round(mean(total_weight), 1),
            avg_minutes_per_cycle = round(mean(total_minutes_per_cycle), 0), .groups = "drop") |>
  check_gt(
    title = "3.2 \u00B7 Minutes layer (separate) \u2014 instructional time, by school",
    note  = "Avg weight and avg minutes side by side. weight \u00D7 minutes should be comparable across schools even when period lengths differ."
  )
```

```{r}
#| label: chk-3.3
n_inc_stu <- n_distinct(cs_data$D_stu_id[cs_data$C_course_time_exclude == "Include"])
n_inc_tch <- n_distinct(cs_data$D_employee_id[cs_data$C_class_size_and_teacher_load_exclude == "Include"])
avg_class_size <- cs_data |> filter(C_class_size_and_teacher_load_exclude == "Include") |>
  distinct(C_class_id, D_stu_id) |> count(C_class_id, name = "size") |> summarise(v = round(mean(size), 1)) |> pull(v)
avg_stu_w <- cs_data |> filter(C_course_time_exclude == "Include") |>
  distinct(D_stu_id, C_class_id, M_stu_class_weight) |> group_by(D_stu_id) |>
  summarise(t = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |> summarise(v = round(mean(t), 1)) |> pull(v)
avg_tch_w <- cs_data |> filter(C_class_size_and_teacher_load_exclude == "Include") |>
  distinct(D_employee_id, C_class_id, M_tch_class_weight) |> group_by(D_employee_id) |>
  summarise(t = sum(M_tch_class_weight, na.rm = TRUE), .groups = "drop") |> summarise(v = round(mean(t), 1)) |> pull(v)
data.frame(
  metric = c("Students double-booked","Teachers double-booked","Avg class size","Avg weight / student","Avg weight / teacher"),
  before = c(paste0(length(flagged_students), " (", pct_of(length(flagged_students), n_inc_stu), "%)"),
             paste0(length(flagged_teachers), " (", pct_of(length(flagged_teachers), n_inc_tch), "%)"),
             avg_class_size, avg_stu_w, avg_tch_w),
  after  = rep("pending resolution", 5)) |>
  check_gt(
    title = "3.3 \u00B7 Before-resolution snapshot (the XYZ of XYZ \u2192 ABC)",
    note  = "The starting picture \u2014 double-bookings by type, avg class size, avg weights, all unresolved. The 'after' fills in once resolution runs (3.6)."
  )
```

```{r}
#| label: chk-3.4
multi_teacher <- resolved |> filter(startsWith(C_class_id_post_teachers, "CONS_")) |>
  group_by(C_class_id_post_teachers) |> summarise(n_tch = n_distinct(D_employee_id), .groups = "drop") |> filter(n_tch > 1)
consol_sum[["Component-size distribution"]] |>
  check_gt(
    title  = glue("3.4 \u00B7 consolidate_sections() \u2014 overlapping sections merge into one class (teacher DBs {teach_db_before} \u2192 {teach_db_after} \u00B7 sections {sec_before} \u2192 {sec_after})"),
    note   = "Every CONS_ class must be exactly one teacher's merged sections. If any spans >1 teacher, the CONS_ id is colliding \u2014 flag it, don't edit the helper.",
    status = if (nrow(multi_teacher) == 0) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-3.5
resolved |> filter(C_course_time_exclude == "Include") |>
  group_by(D_location_name, D_stu_id) |>
  summarise(exploded = sum(M_class_weight_exploded, na.rm = TRUE),
            post_class = sum(M_class_weight_post_class, na.rm = TRUE),
            post_teachers = sum(M_class_weight_post_teachers, na.rm = TRUE),
            post_student = sum(M_class_weight_post_student, na.rm = TRUE), .groups = "drop") |>
  group_by(D_location_name) |>
  summarise(n_db_students = n_distinct(D_stu_id),
            across(c(exploded, post_class, post_teachers, post_student), ~ round(mean(.x), 2)), .groups = "drop") |>
  check_gt(
    title  = "3.5 \u00B7 resolve_stage() \u2014 weight splits at conflicts, only ever shrinking",
    note   = "adj \u2265 post_class \u2265 post_teachers \u2265 post_student everywhere. The columns show a double-booked student's weight moving down through each stage.",
    status = if (mono) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-3.6
data.frame(
  metric = c("Students double-booked","Teachers double-booked","Distinct sections","Avg weight / student"),
  before = c(length(flagged_students), teach_db_before, sec_before, round(mean(stu_ba$b), 2)),
  after  = c(0, teach_db_after, sec_after, round(mean(stu_ba$a), 2))) |>
  check_gt(
    title  = "3.6 \u00B7 unify_resolution() \u2014 before \u2192 after",
    note   = "Double-bookings resolve to zero and sections consolidate to their true count. Row accounting in the next table confirms nothing was lost or duplicated.",
    status = if (unif_ok) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-3.6b
unification_summary(unified, resolved, cs_data) |>
  check_gt(title = "3.6 \u00B7 Row accounting (unified = resolved + non-DB)")
```

```{r}
#| label: chk-3.7
trace_cols(trace_record(resolved, stu_id = ex_student)) |>
  check_gt(
    title = glue("3.7 \u00B7 A student double-booking, traced end to end (stu {ex_student})"),
    note  = "Follow one student's weight across the stage columns \u2014 it should sum to one class-worth, no time lost or double-counted."
  )
```

```{r}
#| label: chk-3.7b
trace_cols(trace_record(resolved, class_id = ex_cons)) |>
  check_gt(
    title = glue("3.7 \u00B7 A teacher consolidation, traced ({ex_cons})"),
    note  = "The teacher's sections merge into one CONS_ class, but each student keeps only their own meetings \u2014 the union never inflates anyone's time."
  )
```

*Next -> stage 04: rollups and export.*

## 04 — Rollups & export-readiness

```{r 04-rollup-layer}
# Core rollup layer (Tier 1 + Tier 2), in dependency order
unified <- add_minutes(unified, config, spec)
unified <- add_class_size(unified, config)
unified <- add_class_primaries(unified)
unified <- add_class_fields(unified)
unified <- add_class_weight_times_size(unified)
unified <- add_teacher_metrics(unified, config)
unified <- add_teacher_primary_fields(unified)
unified <- add_teacher_flags(unified)
```

```{r 04-add-ons}
# Optional add-ons (Tier 3), only if enabled in config$add_ons
hr_on     <- isTRUE(config$add_ons$teacher_hr$enabled)
grades_on <- isTRUE(config$add_ons$course_grades$enabled)
grades_rollup <- NULL

if (hr_on) {
  hr_data <- tr_safe(ers_read_sharepoint(config$raw_data_folder_path,
               config$add_ons$teacher_hr$file), NULL)
  if (!is.null(hr_data)) unified <- apply_teacher_hr(unified, hr_data, config)
}

if (grades_on) {
  grades_raw    <- tr_safe(ers_read_sharepoint(config$raw_data_folder_path,
                     config$add_ons$course_grades$file), NULL)
  course_coding <- tr_safe(ers_read_sharepoint(config$raw_data_folder_path,
                     config$raw_files$course_coding), NULL)
  if (!is.null(grades_raw) && !is.null(course_coding)) {
    gr <- process_course_grades(grades_raw, course_coding, config)
    unified <- apply_course_grades(unified, gr$student_proficiency)
    grades_rollup <- gr$course_grades_rollup
  }
}
```

```{r 04-build-rollups}
teacher_rollup <- build_teacher_rollup(unified, config)
class_rollup   <- build_class_rollup(unified, teacher_rollup, config)
student_rollup <- build_student_rollup(unified, config)
```

```{r 04-precompute-checks}
# 4.2 class-size hook: adjusted size never exceeds raw headcount
size_invariant <- tr_safe(all(class_rollup$M_cls_class_size <= class_rollup$M_cls_num_stu + 1e-9, na.rm=TRUE), FALSE)

# 4.3 buckets populated for INCLUDED classes / teachers
cls_bucket_na <- tr_safe(class_rollup |> filter(C_class_size_exclude == "Include") |>
  summarise(n = sum(is.na(C_cls_class_size_bucket))) |> pull(n), NA)
tch_bucket_na <- tr_safe(teacher_rollup |> summarise(n = sum(is.na(M_tch_load_bucket))) |> pull(n), NA)
bucket_levels <- tr_safe(levels(.bucketize(0, config$buckets$class_size)), character(0))

# 4.4 teacher utilization populated
util_na <- tr_safe(teacher_rollup |> summarise(n = sum(is.na(M_tch_utilization))) |> pull(n), NA)

# 4.5 column contract
want_class <- c("C_class_id","C_cls_course_subject","C_cls_grade","C_cls_course_name",
                "M_cls_class_size","C_cls_class_size_bucket","M_cls_class_weight","M_meeting_minutes",
                "M_tch_utilization","M_tch_load")
want_tch   <- c("D_employee_id","C_tch_course_subject","C_tch_location_name",
                "C_teacher_load_exclude",
                "M_tch_num_periods","M_tch_utilization","M_tch_load","M_tch_load_bucket","M_tch_num_preps")
want_stu   <- c("D_stu_id","C_class_id","C_course_subject","M_class_weight",
                "M_cls_class_size","C_proficiency")
miss_class <- setdiff(want_class, names(class_rollup))
miss_tch   <- setdiff(want_tch,   names(teacher_rollup))
miss_stu   <- setdiff(want_stu,   names(student_rollup))
contract_ok <- length(miss_class)==0 && length(miss_tch)==0 && length(miss_stu)==0

bucket_levels <- tr_safe(levels(.bucketize(0, config$buckets$class_size)), character(0))

# 4.2 class size distribution by school
dist_by_school <- tr_safe(
  class_rollup |>
    filter(C_class_size_exclude == "Include") |>
    group_by(C_cls_location_name, C_class_id) |>
    summarise(bucket = C_cls_class_size_bucket[which.max(M_cls_class_size)], .groups = "drop") |>
    count(C_cls_location_name, bucket) |>
    group_by(C_cls_location_name) |>
    mutate(pct = round(100 * n / sum(n), 1)) |>
    ungroup() |> select(-n) |>
    pivot_wider(names_from = bucket, values_from = pct, values_fill = 0, names_expand = TRUE) |>
    select(C_cls_location_name, any_of(bucket_levels)) |>
    rename(School = C_cls_location_name),
  data.frame())
```

```{r}
#| label: chk-4.1
data.frame(
  rollup = c("class","teacher","student", if (!is.null(grades_rollup)) "course_grades"),
  rows   = c(nrow(class_rollup), nrow(teacher_rollup), nrow(student_rollup),
             if (!is.null(grades_rollup)) nrow(grades_rollup)),
  cols   = c(ncol(class_rollup), ncol(teacher_rollup), ncol(student_rollup),
             if (!is.null(grades_rollup)) ncol(grades_rollup))) |>
  check_gt(
    title = "4.1 \u00B7 Rollups built \u2014 shapes look right",
    note  = "One row per class-slot, per teacher, and per student-class-slot. Class rows > distinct classes is expected (grain is class \u00D7 term \u00D7 period \u00D7 rotation)."
  )
```

```{r}
#| label: chk-4.2
dist_by_school |>
  check_gt(
    title  = "4.2 \u00B7 Class size \u2014 distribution by school (% of classes per size band)",
    note   = "Adjusted size never exceeds raw headcount (double-booked students count fractionally). Each school's row shows the share of its classes in each size band.",
    status = if (size_invariant) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-4.2b
class_rollup |>
  summarise(classes = n_distinct(C_class_id),
            avg_raw_headcount = round(mean(M_cls_num_stu, na.rm = TRUE), 1),
            avg_adjusted_size = round(mean(M_cls_class_size, na.rm = TRUE), 1)) |>
  check_gt(title = "4.2 \u00B7 Adjusted vs raw headcount")
```

```{r}
#| label: chk-4.3
data.frame(
  check = c("class-size NA buckets", "teacher-load NA buckets"),
  n     = c(cls_bucket_na, tch_bucket_na)) |>
  check_gt(
    title  = "4.3 \u00B7 Class-size & teacher-load buckets populated",
    note   = "Every included class needs a size bucket; every teacher a load bucket. If n > 0, a value falls outside the config buckets \u2014 extend the top bucket in config$buckets.",
    status = if (isTRUE(cls_bucket_na == 0 && tch_bucket_na == 0)) "PASS" else "CHECK"
  )
```

```{r}
#| label: chk-4.4
teacher_rollup |>
  group_by(C_tch_location_name) |>
  summarise(teachers = n(), avg_util = round(mean(M_tch_utilization, na.rm = TRUE), 2),
            na = sum(is.na(M_tch_utilization)), .groups = "drop") |>
  check_gt(
    title  = "4.4 \u00B7 Teacher utilization populated",
    note   = "Every teacher needs a utilization value. NA means a school has no periods_per_school entry \u2014 confirm config$periods_per_school keys exactly match the school names above.",
    status = if (isTRUE(util_na == 0)) "PASS" else "CHECK"
  )
```

```{r}
#| label: chk-4.5
data.frame(
  rollup  = c("class","teacher","student"),
  missing = c(ifelse(length(miss_class) == 0, "none", paste(miss_class, collapse = ", ")),
              ifelse(length(miss_tch) == 0, "none", paste(miss_tch, collapse = ", ")),
              ifelse(length(miss_stu) == 0, "none", paste(miss_stu, collapse = ", ")))) |>
  check_gt(
    title  = "4.5 \u00B7 VISTA column contract",
    note   = "Each rollup must have the columns VISTA expects (add-on fields present even when NA). Anything under 'missing' will break VISTA \u2014 add it to the rollup builder.",
    status = if (contract_ok) "PASS" else "FAIL"
  )
```

```{r}
#| label: chk-4.6
data.frame(
  add_on  = c("teacher_hr","course_grades"),
  enabled = c(hr_on, grades_on),
  mode    = c(tr_safe(config$add_ons$teacher_hr$mode, NA), "\u2014"),
  result  = c(if (hr_on && "C_tch_novice_indicator" %in% names(teacher_rollup)) "novice indicator populated" else "skipped / NA",
              if (!is.null(grades_rollup)) glue("{nrow(grades_rollup)} grade rows; proficiency merged") else "skipped / NA")) |>
  check_gt(
    title = "4.6 \u00B7 Add-on status",
    note  = "The optional add-ons run (or skip) based on config$add_ons. Skipped is fine \u2014 the core rollups carry those fields as NA and VISTA drops the metrics that need them."
  )
```

## Export

```{r 04-export}
#| echo: true
out_folder <- config$raw_data_folder_path
put <- function(data, name)
  ers_write_sharepoint(data = data, folder_path = out_folder,
                       file_name_with_extension = paste0("/2. Processed Data/", name, ".xlsx"))

put(class_rollup,   "02_cs_class_rollup")
put(teacher_rollup, "02_cs_teacher_rollup")
put(student_rollup, "02_cs_student_rollup")
if (!is.null(grades_rollup)) put(grades_rollup, "02_cs_course_grades_rollup")
```
