---
title: "Course Schedule Processing — 02 · Coding & Exclusions Review"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

<!---
## What this file does (high level)

This is Stage 02 of the course schedule pipeline. It picks up the cleaned
data from Stage 01 and does three things:

1. Merges in **course coding** — a hand-built lookup that classifies each
   course (subject area, format, whether to exclude it from certain metrics).
2. Applies **exclusion rules** from the config — marking rows as "Include"
   or "Exclude" for specific metrics (time analysis, class size, etc.).
3. Runs a series of **quality checks** (2.0 through 2.9) that produce
   tables and status badges for review.

At the end it exports the updated data for Stage 03.
--->


```{r setup-libraries}
# Start fresh and load the packages we need.
rm(list = ls())
library(gt)
library(dplyr)
library(tidyr)
library(glue)
library(yaml)
library(Microsoft365R)
library(erstools)
```

```{r setup-helpers}
# Load helper functions that are shared across stages.
# Each file provides a set of utility functions:
#   - explore_helpers: general data inspection tools
#   - review_helpers:  the tr_check / tr_stage framework for building the report
#   - expression_helpers: parsing schedule expressions (period/rotation/day)
#   - db_helpers: double-booking detection logic
source(here::here("templates", "course_schedule", "helpers", "01_cs_explore_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "01_cs_review_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "02_cs_db_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "02_cs_expression_helpers.R"))
```

```{r setup-config-and-data}
# Read the district configuration file — THIS DRIVES NEARLY EVERYTHING!
# column mappings, exclusion rules, term weights, class ID recipes, etc.
config <- yaml::read_yaml(here::here("config", "districts", "cs_nashville.yaml"))
raw_data_folder_path <- config$raw_data_folder_pathz
source_of_truth_folder_path <- config$source_of_truth_folder_path
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue::glue("\n## {config$district} · {config$year}\n\n"))

# Connect to the SharePoint site where data lives.
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
orgfiles_site     <- get_sharepoint_site(site_url = orgfiles_site_url)
internal_drive    <- orgfiles_site$get_drive("Internal")

# Load the output of Stage 01 — already cleaned and column-renamed.
cs_data <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = "/2. Processed Data/01_cs_data_post_explore.csv"
)

# Load the Claude-and-hand-coded course coding file.
# This is a lookup table that tells us each course's subject area, format, etc.
# Wrapped in tr_safe() so the document still renders even if the file doesn't exist yet.
cs_course_coding <- tr_safe(ers_read_sharepoint(
    folder_path = source_of_truth_folder_path,
    file_name_with_extension = config$source_of_truth_files$course_coding
  ),
  otherwise = NULL
)
```

```{r define-exclusion-logic}
# ---- How exclusions work ----
# The config file lists rules like:
#   "If C_course_format equals 'Non-Standard', exclude from time analysis."
#
# This function reads those rules and adds columns like
# C_course_time_exclude = "Include" or "Exclude" to each row.
# A row is "Exclude" if ANY rule for that metric matches it.
# Rules referencing columns that don't exist yet are simply skipped.

apply_exclusions <- function(df, rules) {
  for (field in names(rules)) {
    should_exclude <- rep(FALSE, nrow(df))

    for (rule in rules[[field]]) {
      # Skip rules that reference a column we don't have yet
      if (!rule$col %in% names(df)) next

      column_values <- df[[rule$col]]
      rule_values   <- rule$val

      matches_rule <- switch(
        rule$op,
        eq     = column_values == rule_values,
        ne     = column_values != rule_values,
        `in`   = column_values %in% rule_values,
        not_in = !(column_values %in% rule_values),
        rep(FALSE, length(column_values))
      )
      matches_rule[is.na(matches_rule)] <- FALSE
      should_exclude <- should_exclude | matches_rule
    }

    # Write the result as a new column: "Include" or "Exclude"
    df[[field]] <- ifelse(should_exclude, "Exclude", "Include")
  }

  df
}
```

```{r define-sizing-filters}
# ---- Sizing filters (for double-booking analysis only) ----
# These are optional "what-if" filters that drop rows from the double-booking
# sizing step ONLY — they don't affect the main data. For example, you might
# want to exclude a specific school from the DB analysis to see what happens.

.sizing_match <- function(df, rule) {
  if (!rule$col %in% names(df)) return(rep(FALSE, nrow(df)))
  column_values <- df[[rule$col]]
  rule_values   <- rule$val
  matches_rule <- switch(
    rule$op,
    equals     = column_values == rule_values,
    not_equals = column_values != rule_values,
    `in`       = column_values %in% rule_values,
    not_in     = !(column_values %in% rule_values),
    rep(FALSE, length(column_values))
  )
  matches_rule[is.na(matches_rule)] <- FALSE
  matches_rule
}

apply_sizing_filters <- function(df, filters) {
  if (is.null(filters) || length(filters) == 0) return(df)
  drop <- rep(FALSE, nrow(df))
  for (rule in filters) drop <- drop | .sizing_match(df, rule)
  df[!drop, , drop = FALSE]
}

sizing_filter_summary <- function(df, filters) {
  if (is.null(filters) || length(filters) == 0) {
    return(data.frame(rule = "(none — no sizing filters set)", rows_removed = 0L))
  }
  do.call(rbind, lapply(filters, function(rule) {
    data.frame(
      rule = paste(rule$col, rule$op, paste(rule$val, collapse = " | ")),
      rows_removed = sum(.sizing_match(df, rule))
    )
  }))
}
```

```{r merge-course-coding}
# ---- Merge the course coding into the main data ----
# The course coding file is a lookup table keyed on course name (or another
# field set in config). We join it onto cs_data so every row gets its
# subject area, format, and other coded fields.
#
# We also capture "merge diagnostics" — did the join go cleanly?
# - Was the key unique (no accidental row multiplication)?
# - Did every course in the data match something in the coding file?

join_key   <- if (!is.null(config$coding$join_key)) unlist(config$coding$join_key) else "D_course_name"
merge_diag <- NULL

if (!is.null(cs_course_coding)) {
  cs_data <- tr_safe({

    # Identify which columns the coding file brings in (everything except the join keys)
    coded_cols <- setdiff(names(cs_course_coding), join_key)
    coding_lookup <- cs_course_coding
    coding_lookup$.coded <- TRUE

    # Record state before the join
    n_before   <- nrow(cs_data)
    key_unique <- !any(duplicated(coding_lookup[, join_key]))

    # Perform the join: drop any old coded columns first, then bring in fresh ones
    joined <- cs_data |>
      select(-any_of(coded_cols)) |>
      left_join(coding_lookup, by = join_key)

    # Find rows that didn't match anything in the coding file
    unmatched <- joined |>
      filter(is.na(.coded)) |>
      distinct(across(all_of(join_key)))

    # Save diagnostics for the check report later
    merge_diag <<- list(
      key          = paste(join_key, collapse = " + "),
      key_unique   = key_unique,
      n_before     = n_before,
      n_after      = nrow(joined),
      row_stable   = identical(n_before, nrow(joined)),
      n_unmatched  = nrow(joined |> filter(is.na(.coded))),
      unmatched_courses = as.data.frame(unmatched) |>
        arrange(across(all_of(join_key)))
    )

    # Clean up and apply the exclusion rules from config
    joined |>
      select(-.coded) |>
      apply_exclusions(config$exclusions)

  }, otherwise = cs_data)
}

# Quick helper: check if certain columns exist in the data
has_cols <- function(...) all(c(...) %in% names(cs_data))
```

```{r setup-double-booking}
# ---- Double-booking detection setup ----
# Double-booking = a student (or teacher) assigned to 2+ classes at the same time.
#
# To detect this, we need to:
# 1. Parse schedule expressions into individual time-slots ("explode" them)
# 2. Look for students/teachers who appear in multiple classes in the same slot
#
# "spec" defines how expressions are structured (days, periods, rotations).
# "term_expands" defines how terms map to time-slots.

spec         <- tr_safe(expr_spec(config), NULL)
term_expands <- tr_safe(term_expands_spec(config), NULL)

# Build a class ID for each row — this identifies a unique "class" using
# the recipe defined in config (e.g., course + section + term + expression).
cs_data$C_class_id <- build_class_id(cs_data, config$class_id$recipe)

# Apply any sizing filters before doing the double-booking analysis
cs_size <- apply_sizing_filters(as.data.frame(cs_data), config$db_sizing_filters)

# "Explode" each row into one row per time-slot it covers.
# A class that meets Mon/Wed periods 1-2 becomes 4 rows (Mon P1, Mon P2, Wed P1, Wed P2).
# This lets us find overlaps by looking for duplicate student+timeslot combos.
cs_long <- tr_safe(explode_conflicts(cs_size, spec, term_expands, config), NULL)

# Flag students who appear in 2+ classes in the same time-slot
# (only among rows that are "Include" for time analysis).
flagged_students <- tr_safe(

  unique(
    db_flagged(
      filter(cs_long, C_course_time_exclude == "Include"),
      "D_stu_id",
      "C_class_id"
    )$D_stu_id
  ),
  character(0)
)

# Flag teachers who appear in 2+ classes in the same time-slot
# (only among rows that are "Include" for teacher load analysis).
flagged_teachers <- tr_safe(
  unique(
    db_flagged(
      filter(cs_long, C_class_size_and_teacher_load_exclude == "Include"),
      "D_employee_id",
      "C_class_id"
    )$D_employee_id
  ),
  character(0)
)

# Helper: calculate a percentage safely (returns NA if denominator is 0 or missing)
pct_of <- function(a, b) {
  r <- round(100 * a / b, 1)
  r[is.na(b) | b == 0] <- NA_real_
  r
}

# Count how many students/teachers are in the "Include" pool (denominators for DB rates)
n_stu_time <- tr_safe(
  n_distinct(cs_data$D_stu_id[cs_data$C_course_time_exclude == "Include"]),
  NA
)
n_tch_load <- tr_safe(
  n_distinct(cs_data$D_employee_id[cs_data$C_class_size_and_teacher_load_exclude == "Include"]),
  NA
)
```

## How to read this

The first formal review — `01` is investigation-only. Every check either
**computes** a status from the data or is **evidence-only** (a table to eyeball,
no badge). Nothing here is hand-maintained prose wearing a status. Checks detect
what's present and degrade gracefully: a not-yet-loaded input shows "data not
loaded," never a crash.

Golden rule: **processing flags, it never deletes** — exclusions are columns a
later metric can honor or ignore.

## 02 — Coding & exclusions review

```{r stage02}
#| results: 'asis'

# ---- Build the stage report ----
# Each tr_check() defines one numbered check with:
#   - A question it answers
#   - "good": what the ideal outcome looks like
#   - "value": a summary number or phrase shown in the badge
#   - "status": ok / flag / none (drives the badge color)
#   - "evidence": a function that returns a table (or list of tables) to display

s02 <- tr_stage(
  number = "02",
  title  = "Coding & Exclusions Review",
  question = "Did standardization land, do we know what each course is, and is the right thing excluded?",
  checks = list(

    # ================================================================
    # CHECK 2.0: Did all expected columns arrive from Stage 01?
    # ================================================================
    tr_check("2.0", "Did standardization from 01 land — every mapped field present?",
      good   = "Every field in column_map is present with its D_ name.",
      value  = tr_safe(
        length(setdiff(names(config$column_map$course_schedule), names(cs_data))),
        NA
      ),
      status = tr_safe(
        if (length(setdiff(names(config$column_map$course_schedule), names(cs_data))) == 0) {
          "ok"
        } else {
          "flag"
        },
        "none"
      ),
      evidence = function() {
        expected_fields <- names(config$column_map$course_schedule)
        data.frame(
          field   = expected_fields,
          present = ifelse(expected_fields %in% names(cs_data), "yes", "MISSING")
        )
      }
    ),

    # ================================================================
    # CHECK 2.1: Can every term value be weighted?
    # The config defines weights for terms (e.g., "S1" = 0.5, "FY" = 1.0).
    # If a term appears in the data but not in the config, we can't weight it.
    # ================================================================
    tr_check("2.1", "Term structure — is every term value weightable?",
      good   = "Every term in the data has a configured weight.",
      status = tr_safe(
        if (all(as.character(unique(na.omit(cs_data$D_term))) %in% names(config$weights$term))) {
          "ok"
        } else {
          "flag"
        },
        "none"
      ),
      evidence = function() {
        cs_data |>
          count(D_term, name = "n_records") |>
          arrange(D_term) |>
          mutate(
            weight = sapply(as.character(D_term), \(term_value) {
              configured_weight <- config$weights$term[[term_value]]
              if (is.null(configured_weight)) "MISSING from config" else as.character(configured_weight)
            })
          )
      }
    ),

    # ================================================================
    # CHECK 2.2: How many records per student at each school?
    # A sanity check — if a school has 2 records/student on average,
    # something is probably wrong (should be ~6-8 for a typical HS schedule).
    # ================================================================
    tr_check("2.2", "Records per student, by school",
      good = "Per-school record distributions and expression variety look plausible for HS schedules.",
      evidence = function() {
        included_data <- cs_data

        # If exclusion columns exist, filter to included records only
        if (has_cols("C_course_time_exclude", "C_class_size_and_teacher_load_exclude")) {
          included_data <- included_data |>
            filter(
              C_course_time_exclude == "Include",
              C_class_size_and_teacher_load_exclude == "Include"
            )
        }

        records_per_student <- included_data |>
          count(D_location_name, D_stu_id, name = "records_per_student")

        time_cols <- time_key_cols(config)

        included_data |>
          group_by(D_location_name) |>
          summarise(
            n_students   = n_distinct(D_stu_id),
            n_teachers   = n_distinct(D_employee_id),
            terms        = paste(sort(unique(na.omit(D_term))), collapse = ", "),
            n_time_slots = n_distinct(!!!syms(time_cols)),
            .groups = "drop"
          ) |>
          left_join(
            records_per_student |>
              group_by(D_location_name) |>
              summarise(
                avg_records_per_student = round(mean(records_per_student), 1),
                .groups = "drop"
              ),
            by = "D_location_name"
          ) |>
          arrange(D_location_name)
      }
    ),

    # ================================================================
    # CHECK 2.3: Did demographic columns come through from Stage 01?
    # Stage 01 merges student demographics (race, grade, etc.) onto the
    # schedule data. This checks they arrived and are mostly filled in.
    # ================================================================
    tr_check("2.3", "Student demographics — present and populated?",
      good   = "The demographic fields from the 01 merge exist and are well-filled.",
      status = tr_safe(
        if (all(unlist(config$stu_demographics_join$brings_in) %in% names(cs_data))) {
          "ok"
        } else {
          "none"
        },
        "none"
      ),
      evidence = function() {
        demographic_fields <- unlist(config$stu_demographics_join$brings_in)
        is_present <- demographic_fields %in% names(cs_data)
        pct_filled <- vapply(demographic_fields, function(field_name) {
          if (field_name %in% names(cs_data)) {
            round(100 * mean(!is.na(cs_data[[field_name]])), 1)
          } else {
            NA_real_
          }
        }, numeric(1))

        data.frame(
          field      = demographic_fields,
          present    = ifelse(is_present, "yes", "no"),
          pct_filled = pct_filled
        )
      }
    ),

    # ================================================================
    # CHECK 2.4: Did the course coding merge go cleanly?
    # We want: no row multiplication, row count stays the same, and
    # every course in the data matched something in the coding file.
    # ================================================================
    tr_check("2.4", "Did the course-coding merge land cleanly?",
      good   = "Unique key (no row multiplication), row count stable pre/post, and every course matched.",
      value  = tr_safe(glue("{merge_diag$n_unmatched} unmatched course rows"), NA),
      status = tr_safe(
        if (isTRUE(merge_diag$key_unique) &&
            isTRUE(merge_diag$row_stable) &&
            merge_diag$n_unmatched == 0) {
          "ok"
        } else {
          "flag"
        },
        "none"
      ),
      evidence = function() {
        if (is.null(merge_diag)) return(NULL)

        summ <- data.frame(
          check = c(
            "key unique (no multiplication)",
            "rows stable pre/post",
            "unmatched course rows"
          ),
          result = c(
            ifelse(merge_diag$key_unique, "yes", "NO"),
            paste0(
              format(merge_diag$n_before, big.mark = ","), " -> ",
              format(merge_diag$n_after, big.mark = ",")
            ),
            format(merge_diag$n_unmatched, big.mark = ",")
          )
        )

        list(
          "Merge summary" = summ,
          "Course names that didn't get coded" = merge_diag$unmatched_courses
        )
      }
    ),

    # ================================================================
    # CHECK 2.5: Are coding results sensible?
    # After the merge, every course should have a subject area.
    # "UNCODED" means the coding file didn't assign it — too many is a problem.
    # ================================================================
    tr_check("2.5", "Coding output — sane distributions, nothing left uncoded?",
      good   = glue("Uncoded <= {config$coding$uncoded_flag_threshold_pct}% of records; distributions plausible."),
      value  = tr_safe(
        paste0(
          round(100 * mean(cs_data$C_course_subject_area == "UNCODED", na.rm = TRUE), 1),
          "% uncoded"
        ),
        NA
      ),
      status = tr_safe(
        tr_threshold_status(
          100 * mean(cs_data$C_course_subject_area == "UNCODED", na.rm = TRUE),
          config$coding$uncoded_flag_threshold_pct,
          "below"
        ),
        "none"
      ),
      evidence = function() {
        # How many records fall into each subject area?
        by_subject_area <- cs_data |>
          count(C_course_subject_area, name = "n_records") |>
          mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
          arrange(C_course_subject_area)

        # More granular: subject within subject area
        by_subject <- cs_data |>
          count(C_course_subject_area, C_course_subject, name = "n_records") |>
          mutate(pct_records = round(100 * n_records / sum(n_records), 1)) |>
          arrange(C_course_subject_area)

        list(
          "Subject Area — % of records" = by_subject_area,
          "Subject — % of records" = by_subject
        )
      }
    ),

    # ================================================================
    # CHECK 2.6: What's being excluded, and does it make sense?
    # Shows every course that's been marked "Exclude" for any metric,
    # so a human can verify it's the right stuff (e.g., lunch, study hall).
    # ================================================================
    tr_check("2.6", "Exclusions — right thing dropping, for the right reason?",
      good = "Every excluded course traces to a coding value that justifies it.",
      evidence = function() {
        cs_data |>
          filter(
            C_course_time_exclude == "Exclude" |
              C_class_size_and_teacher_load_exclude == "Exclude"
          ) |>
          count(
            C_course_subject, D_course_name,
            C_course_time_exclude, C_class_size_and_teacher_load_exclude,
            name = "n_courses",
            sort = TRUE
          ) |>
          arrange(C_course_subject)
      }
    ),

    # ================================================================
    # CHECK 2.7: Which time-slots have no real instruction?
    # Some time blocks (like "Lunch" or "Advisory") contain only
    # non-standard-format courses or very few distinct courses.
    # These should probably be excluded from time analysis.
    # ================================================================
    tr_check("2.7", "Which time blocks don't carry real instruction?",
      good = "Time-slots that are entirely non-standard formats, or hold very few distinct courses, are identified so they fall out.",
      evidence = function() {
        # time_key_cols picks the right columns based on the schedule format:
        # either "D_expression" alone, or c("D_period", "D_rotation")
        time_cols <- time_key_cols(config)

        # Find time-slots where ZERO records are "Standard" format
        non_standard <- cs_data |>
          count(
            across(all_of(c("D_location_name", time_cols))),
            C_course_format,
            name = "n_records"
          ) |>
          pivot_wider(
            names_from = C_course_format,
            values_from = n_records,
            values_fill = 0
          ) |>
          filter(`Standard` == 0) |>
          arrange(across(all_of(c("D_location_name", time_cols))))

        # Find time-slots with very few distinct courses (likely not real periods)
        max_courses <- if (!is.null(config$coding$max_unique_courses)) {
          config$coding$max_unique_courses
        } else {
          5
        }

        low_diversity <- cs_data |>
          group_by(across(all_of(c("D_location_name", time_cols)))) |>
          mutate(n_courses = n_distinct(D_course_name)) |>
          ungroup() |>
          filter(n_courses <= max_courses, !C_course_subject_area %in% c("Core", "Non-Core")) |>
          count(
            across(all_of(c("D_location_name", time_cols))),
            C_course_subject, D_course_name, n_courses,
            name = "n_records"
          ) |>
          arrange(D_location_name, n_courses, desc(n_records))

        out <- list(non_standard, low_diversity)
        names(out) <- c(
          "No standard-format records",
          glue("Low-diversity time-slots (<= {max_courses} courses, excl. Core and Non-Core)")
        )
        out
      }
    ),

    # ================================================================
    # CHECK 2.8: What defines a "class" in this data?
    # The class ID is built from a recipe (e.g., course + section + term +
    # expression OR period + rotation). This shows a sample so you can 
    # verify classes aren't being over-split or under-split.
    # ================================================================
    tr_check("2.8", "Class ID — what defines a class here?",
      good  = "The recipe marks a place-in-time without over-splitting real classes.",
      value = tr_safe(paste(config$class_id$recipe, collapse = " + "), NA),
      evidence = function() {
        cols <- config$class_id$recipe[config$class_id$recipe %in% names(cs_data)]

        cs_data |>
          filter(!is.na(C_class_id)) |>
          group_by(across(all_of(c("C_class_id", cols)))) |>
          summarise(
            course     = dplyr::first(D_course_name),
            n_students = n_distinct(D_stu_id),
            .groups = "drop"
          ) |>
          arrange(desc(n_students)) |>
          (\(d) slice_sample(d, n = min(8, nrow(d))))()
      }
    ),

    # ================================================================
    # CHECK 2.9: How widespread is double-booking?
    # This is the big one. It tells us:
    #   - What % of students/teachers are double-booked
    #   - Which schools have the most
    #   - Sample rows showing actual collisions
    #   - Classes where 100% of students are double-booked (likely data artifacts)
    # ================================================================
    tr_check("2.9", "Double-booking sizing — how big, and of what kind?",
      good   = "<2-3% may skip; concentrated in a few schools/classes -> hand-resolve; widespread -> full.",
      value  = tr_safe(
        glue(
          "students {pct_of(length(flagged_students), n_stu_time)}% \u00B7 ",
          "teachers {pct_of(length(flagged_teachers), n_tch_load)}%"
        ),
        NA
      ),
      status = tr_safe({
        r <- c(
          pct_of(length(flagged_students), n_stu_time),
          pct_of(length(flagged_teachers), n_tch_load)
        )
        if (all(is.na(r))) "none" else if (max(r, na.rm = TRUE) < 3) "ok" else "flag"
      }, "none"),
      evidence = function() {

        # -- Subset to included records for student and teacher analyses --
        included_student_rows <- cs_long |>
          filter(C_course_time_exclude == "Include")
        included_teacher_rows <- cs_long |>
          filter(C_class_size_and_teacher_load_exclude == "Include")

        # -- Student double-booking rates by school --
        student_db_rates <- cs_data |>
          filter(C_course_time_exclude == "Include") |>
          group_by(D_location_name) |>
          summarise(
            n_students = n_distinct(D_stu_id),
            n_db = n_distinct(D_stu_id[D_stu_id %in% flagged_students]),
            .groups = "drop"
          ) |>
          mutate(pct_db = pct_of(n_db, n_students)) |>
          arrange(desc(pct_db))

        # -- Teacher double-booking rates by school --
        teacher_db_rates <- cs_data |>
          filter(C_class_size_and_teacher_load_exclude == "Include") |>
          group_by(D_location_name) |>
          summarise(
            n_teachers = n_distinct(D_employee_id),
            n_db = n_distinct(D_employee_id[D_employee_id %in% flagged_teachers]),
            .groups = "drop"
          ) |>
          mutate(pct_db = pct_of(n_db, n_teachers)) |>
          arrange(desc(pct_db))

        time_cols <- time_key_cols(config)

        # Helper: collapse time key cols into one readable string per row
        time_slot_str <- function(df) {
          apply(df[, time_cols, drop = FALSE], 1, function(r) paste(na.omit(r), collapse = "/"))
        }

        # -- Sample student collisions (show actual conflicting rows) --
        # Find up to 10 student+timeslot combos that have >1 class
        student_collision_groups <- included_student_rows |>
          group_by(D_stu_id, M_conflict_key) |>
          summarise(n_classes = n_distinct(C_class_id), .groups = "drop") |>
          filter(n_classes > 1) |>
          (\(d) slice_sample(d, n = min(10, nrow(d))))()

        # Pull the actual rows for those collisions
        student_collision_rows <- included_student_rows |>
          semi_join(student_collision_groups, by = c("D_stu_id", "M_conflict_key")) |>
          distinct(
            D_location_name, D_stu_id, M_conflict_key, C_class_id,
            D_term, across(all_of(time_cols)), D_course_section_id, D_course_name
          ) |>
          mutate(time_slot = time_slot_str(pick(all_of(time_cols)))) |>
          arrange(D_location_name, D_stu_id, M_conflict_key) |>
          select(
            D_location_name, D_stu_id, D_term, time_slot,
            D_course_section_id, D_course_name, M_conflict_key
          )

        # -- Sample teacher collisions (largest combined rosters) --
        teacher_collision_rows <- included_teacher_rows |>
          distinct(
            D_location_name, D_employee_id, M_conflict_key,
            C_class_id, D_stu_id, D_term, across(all_of(time_cols)),
            D_course_section_id, D_course_name
          )

        teacher_collision_groups <- teacher_collision_rows |>
          group_by(D_location_name, D_employee_id, M_conflict_key) |>
          summarise(
            n_classes         = n_distinct(C_class_id),
            n_student_rows    = n(),
            n_unique_students = n_distinct(D_stu_id),
            .groups = "drop"
          ) |>
          filter(n_classes > 1) |>
          slice_max(n_student_rows, n = 10, with_ties = FALSE)

        teacher_collision_detail <- teacher_collision_rows |>
          semi_join(
            teacher_collision_groups,
            by = c("D_location_name", "D_employee_id", "M_conflict_key")
          ) |>
          mutate(time_slot = time_slot_str(pick(all_of(time_cols)))) |>
          group_by(D_location_name, D_employee_id, M_conflict_key, C_class_id) |>
          summarise(
            D_term              = paste(sort(unique(na.omit(D_term))), collapse = ", "),
            time_slot           = paste(sort(unique(na.omit(time_slot))), collapse = ", "),
            D_course_section_id = paste(sort(unique(na.omit(D_course_section_id))), collapse = ", "),
            D_course_name       = paste(sort(unique(na.omit(D_course_name))), collapse = " | "),
            n_students_in_class = n_distinct(D_stu_id),
            .groups = "drop"
          ) |>
          left_join(
            teacher_collision_groups |>
              select(D_location_name, D_employee_id, M_conflict_key,
                     n_student_rows, n_unique_students),
            by = c("D_location_name", "D_employee_id", "M_conflict_key")
          ) |>
          arrange(
            desc(n_student_rows), D_location_name,
            D_employee_id, M_conflict_key, desc(n_students_in_class)
          ) |>
          select(
            D_location_name, D_employee_id, D_term, time_slot,
            D_course_section_id, D_course_name, M_conflict_key,
            n_students_in_class, n_student_rows, n_unique_students
          )

        # -- Classes where 100% of students are double-booked --
        # These are often data artifacts (co-taught classes, linked sections).
        full_db_classes <- cs_data |>
          filter(C_course_time_exclude == "Include") |>
          distinct(C_class_id, D_course_name, D_location_name, D_stu_id) |>
          group_by(C_class_id, D_course_name, D_location_name) |>
          summarise(
            n_students = n_distinct(D_stu_id),
            n_db = n_distinct(D_stu_id[D_stu_id %in% flagged_students]),
            .groups = "drop"
          ) |>
          mutate(pct_db = pct_of(n_db, n_students)) |>
          filter(pct_db == 100) |>
          arrange(desc(n_students)) |>
          head(20)

        # For each 100%-DB class, find what class(es) its students collide WITH
        student_timeslot_classes <- included_student_rows |>
          distinct(D_stu_id, M_conflict_key, C_class_id)

        target_students <- student_timeslot_classes |>
          filter(C_class_id %in% full_db_classes$C_class_id) |>
          distinct(D_stu_id, M_conflict_key)

        target_rows <- student_timeslot_classes |>
          semi_join(target_students, by = c("D_stu_id", "M_conflict_key"))

        class_name_lookup <- cs_data |>
          distinct(C_class_id, D_course_name)

        collision_partners <- target_rows |>
          filter(C_class_id %in% full_db_classes$C_class_id) |>
          rename(focus = C_class_id) |>
          inner_join(
            rename(target_rows, partner = C_class_id),
            by = c("D_stu_id", "M_conflict_key"),
            relationship = "many-to-many"
          ) |>
          filter(focus != partner) |>
          left_join(
            rename(class_name_lookup, partner = C_class_id, partner_course = D_course_name),
            by = "partner"
          ) |>
          distinct(focus, M_conflict_key, partner, partner_course) |>
          group_by(focus) |>
          summarise(
            collides_with = paste(
              head(unique(paste0(partner_course, " @", M_conflict_key)), 3),
              collapse = "; "
            ),
            .groups = "drop"
          )

        full_db_classes <- full_db_classes |>
          left_join(rename(collision_partners, C_class_id = focus), by = "C_class_id")

        # -- Class-level DB rates by school --
        class_db_rates <- cs_data |>
          filter(C_course_time_exclude == "Include") |>
          group_by(D_location_name) |>
          summarise(
            n_classes = n_distinct(C_class_id),
            n_db = n_distinct(C_class_id[C_class_id %in% full_db_classes$C_class_id]),
            .groups = "drop"
          ) |>
          mutate(pct_db = pct_of(n_db, n_classes)) |>
          arrange(desc(pct_db))

        # Return all evidence tables as a named list
        list(
          "Sizing filters applied" =
            sizing_filter_summary(as.data.frame(cs_data), config$db_sizing_filters),
          "Student double-bookings by school" = student_db_rates,
          "Student collisions \u2014 sample rows (max 20)" = student_collision_rows,
          "Teacher double-bookings by school" = teacher_db_rates,
          "Teacher collisions \u2014 sample rows (max 20)" = teacher_collision_detail,
          "Class double-bookings by school" = class_db_rates,
          "Class-level (100%) double-bookings (max 20)" = full_db_classes
        )
      }
    )
  )
)

htmltools::browsable(render_stage(s02))
```

*Next -> stage 03: TPR weights, class weight, and double-booking resolution.*

## Re-export for the next stage

```{r export}
#| eval: true

# Save the updated data (now with coding + exclusion columns) for Stage 03.
ers_write_sharepoint(
  data = cs_data,
  folder_path = raw_data_folder_path,
  file_name_with_extension = "/2. Processed Data/02_cs_data_post_code_and_exclude.csv"
)
```

```{r unmatched export}
#| eval: false

# Save the updated data (now with coding + exclusion columns) for Stage 03.
ers_write_sharepoint(
  data = unmatched,
  folder_path = raw_data_folder_path,
  file_name_with_extension = "/1. Coding/cs_unmatched_courses.csv"
)
```