---
title: "Course Schedule Processing — 03 · Weights, Minutes & DB Resolution"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

<!---
## What this file does (high level)

This is Stage 03 of the course schedule pipeline. It picks up the coded +
excluded data from Stage 02 and does two things:

1. Derives **class weights** and **class minutes** for each record.
2. Produces a "before resolution" snapshot of double-booking so we can
   see the starting point before any adjustments are made.

Key concepts:
- **Weight** = a dimensionless number representing what fraction of a full
  schedule this class occupies (based on term length × rotation × periods).
  Think of it as "how big a slice of the pie is this class?"
- **Minutes** = actual clock time from the bell schedule (e.g., 43 min or 90 min).
- **Weight × Minutes** = real instructional time for a class.
- **Double-booking resolution** (done in the NEXT pass) = adjusting weights
  when a student or teacher is scheduled in 2+ classes at the same time.
--->


```{r setup-libraries}
# Start fresh and load packages
rm(list = ls())
library(gt)
library(dplyr)
library(tidyr)
library(glue)
library(yaml)
library(Microsoft365R)
library(erstools)
```

```{r setup-helpers}
# Load helper functions shared across stages
source(here::here("templates", "course_schedule", "helpers", "01_cs_explore_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "01_cs_review_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "02_cs_db_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "02_cs_expression_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "03_cs_resolution_helpers.R"))
source(here::here("templates", "course_schedule", "helpers", "03_cs_weight_helpers.R"))
```

```{r setup-config-and-data}
# Read the district config (drives everything: weights, bell schedule, etc.)
config <- yaml::read_yaml(here::here("config", "districts", "nashville.yaml"))

# "spec" tells us how schedule expressions are structured (days, periods, rotations)
# "term_expands" tells us how terms map to time-slots for conflict detection
spec         <- expr_spec(config)
term_expands <- term_expands_spec(config)

raw_data_folder_path <- config$raw_data_folder_path
```

```{r subtitle}
#| results: 'asis'
#| echo: false
cat(glue::glue("\n## {config$district} · {config$year}\n\n"))

# Connect to SharePoint
orgfiles_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles")
internal_drive <- orgfiles_site$get_drive("Internal")

# Load Stage 02 output (coded + exclusions applied).
# Wrapped in tr_safe() so this file renders even if 02 hasn't been exported yet.
cs_data <- tr_safe(ers_read_sharepoint(
    folder_path = raw_data_folder_path,
    file_name_with_extension = "/2. Processed Data/02_cs_data_post_code_and_exclude.csv"
  ),
  otherwise = NULL
)
```

```{r derive-weights-and-minutes}
# ---- Build class weights and minutes ----
# For each record, we calculate:
#   M_cls_class_weight  = term_weight × rotation_weight × period_weight
#                         (e.g., a full-year class meeting every day for 1 period = 1.0)
#   M_cls_class_minutes = clock time from the bell schedule
#
# These are properties of the CLASS (not the student), so they're the same
# for every student in a given class.

cs_data <- tr_safe({
  cs_data$C_class_id <- build_class_id(cs_data, config$class_id$recipe)
  derive_class_weight_minutes(cs_data, config, spec)
}, otherwise = cs_data)

# ---- Initialize student and teacher weight "views" ----
# Before resolution, every student/teacher gets the full class weight.
# After resolution (next pass), double-booked students/teachers will get
# ADJUSTED weights (split between their overlapping classes).
#
# We keep three separate weight columns:
#   M_cls_class_weight  = the class's intrinsic weight (never changes)
#   M_stu_class_weight  = the weight from the STUDENT's perspective (adjusted during resolution)
#   M_tch_class_weight  = the weight from the TEACHER's perspective (adjusted during resolution)

cs_data$M_stu_class_weight <- tr_safe(cs_data$M_cls_class_weight, NA)
cs_data$M_tch_class_weight <- tr_safe(cs_data$M_cls_class_weight, NA)
```

```{r detect-double-bookings-before-resolution}
# ---- "Before" double-booking snapshot ----
# This measures how much double-booking exists BEFORE any resolution.
# We'll compare this to the "after" picture once resolution runs.

# Explode each row into individual time-slots (same logic as Stage 02)
cs_long <- tr_safe(
  explode_conflicts(cs_data, spec, term_expands, config),
  NULL
)

# resolution: student weights (meeting-level 1/N split)
resolved_stu <- tr_safe(resolve_student_weights(cs_long, config), NULL)
if (!is.null(resolved_stu)) {
  cs_data <- cs_data |>
    select(-any_of("M_stu_class_weight")) |>          # drop the pass-1 placeholder
    left_join(resolved_stu |> select(D_stu_id, C_class_id, M_stu_class_weight),
              by = c("D_stu_id", "C_class_id"))
}
# fall back to base class weight where resolution didn't apply (excluded / no conflict)
cs_data$M_stu_class_weight <- ifelse(is.na(cs_data$M_stu_class_weight),
                                     cs_data$M_cls_class_weight, cs_data$M_stu_class_weight)

# Find students scheduled in 2+ classes at the same time
flagged_students <- tr_safe(
  unique(
    db_flagged(
      filter(cs_long, C_course_time_exclude == "Include"),
      "D_stu_id",
      "C_class_id"
    )$D_stu_id
  ),
  character(0)
)

# Find teachers scheduled in 2+ classes at the same time
flagged_teachers <- tr_safe(
  unique(
    db_flagged(
      filter(cs_long, C_class_size_and_teacher_load_exclude == "Include"),
      "D_employee_id",
      "C_class_id"
    )$D_employee_id
  ),
  character(0)
)

# Helper: calculate a percentage safely (returns NA if denominator is 0 or missing)
pct_of <- function(a, b) {
  result <- round(100 * a / b, 1)
  result[is.na(b) | b == 0] <- NA_real_
  result
}
```

## How to read this

This stage derives the two things double-booking resolution needs and keeps
them **rigorously separate**:

- **Weight** — a *dimensionless share* (`term x rotation x period`). It drives
  resolution (how to split shared time) and only has meaning once multiplied by
  minutes. It does **not** measure course count or full load.
- **Minutes** — *real clock time* from the bell schedule. The units.

`weight x minutes` is where real instructional time lands. This pass builds and
validates both, plus the "before" DB picture. **Resolution comes next** — here
you're confirming the foundation is sound.

## 03 — Weights, minutes & the pre-resolution picture

```{r stage03}
#| results: 'asis'

# ---- Build the stage report ----
# Each check validates a piece of the weight/minutes/DB foundation.

s03 <- tr_stage(
  number = "03",
  title  = "Weights, Minutes & DB Resolution",
  question = "Are the class weights and minutes right, and how big is the problem before we resolve it?",
  checks = list(

    # ================================================================
    # CHECK 3.0: Do the computed weights and minutes look right?
    # Shows a handful of real classes so you can verify by hand:
    #   - Does the weight make sense for that term/expression combo?
    #   - Does the minutes value match the bell schedule?
    # ================================================================
    tr_check("3.0", "Class weight & minutes — do a few real classes look right?",
      good  = "term x rotation x period reconciles by hand; minutes come from the bell schedule; the two are separate.",
      evidence = function() {
        # Pick 10 classes with distinct weight values (to show variety)
        sample_classes <- cs_data |>
          filter(!is.na(C_class_id)) |>
          distinct(
            C_class_id, D_location_name, D_expression, D_term,
            D_course_name, M_cls_class_weight, M_cls_class_minutes
          ) |>
          distinct(M_cls_class_weight, .keep_all = TRUE) |>
          slice_sample(n = 10) |>
          arrange(desc(M_cls_class_weight)) |>
          # "serialized" shows the parsed expression in a readable form
          # (e.g., "Mon P1, Mon P2, Wed P1, Wed P2")
          mutate(
            serialized = vapply(
              D_expression,
              function(expr) serialize_meetings(parse_expression(expr, spec)),
              character(1)
            )
          ) |>
          transmute(
            D_location_name,
            expression = D_expression,
            serialized,
            D_term,
            course = D_course_name,
            M_cls_class_weight,
            M_cls_class_minutes
          )

        sample_classes
      }
    ),

    # ================================================================
    # CHECK 3.1: Are per-person weight totals sensible?
    # Each student's total weight should reflect how many periods they
    # have in a day. E.g., an 8-period school ≈ 8.0, a 4-block school ≈ 4.0.
    # Same logic for teachers.
    # ================================================================
    tr_check("3.1", "Weight per student & per teacher — sane by school?",
      good  = "Sums reflect the schedule structure (e.g. Sci-Tech ~4 blocks, Putnam ~8 periods) — NOT forced to a single target.",
      evidence = function() {

        # Sum each student's class weights across their schedule
        student_weight_summary <- cs_data |>
          filter(C_course_time_exclude == "Include") |>
          distinct(D_location_name, D_stu_id, C_class_id, M_stu_class_weight) |>
          group_by(D_location_name, D_stu_id) |>
          summarise(
            total_weight = sum(M_stu_class_weight, na.rm = TRUE),
            .groups = "drop"
          ) |>
          group_by(D_location_name) |>
          summarise(
            n_students = n_distinct(D_stu_id),
            avg_weight_per_student = round(mean(total_weight), 2),
            .groups = "drop"
          )

        # Sum each teacher's class weights across their schedule
        teacher_weight_summary <- cs_data |>
          filter(C_class_size_and_teacher_load_exclude == "Include") |>
          distinct(D_location_name, D_employee_id, C_class_id, M_tch_class_weight) |>
          group_by(D_location_name, D_employee_id) |>
          summarise(
            total_weight = sum(M_tch_class_weight, na.rm = TRUE),
            .groups = "drop"
          ) |>
          group_by(D_location_name) |>
          summarise(
            n_teachers = n_distinct(D_employee_id),
            avg_weight_per_teacher = round(mean(total_weight), 2),
            .groups = "drop"
          )

        list(
          "Avg weight per student, by school" = student_weight_summary,
          "Avg weight per teacher, by school" = teacher_weight_summary
        )
      }
    ),

    # ================================================================
    # CHECK 3.2: Does weight × minutes give comparable real time across schools?
    # A 4-block school with 90-min blocks and an 8-period school with
    # 43-min periods should produce similar total instructional minutes.
    # ================================================================
    tr_check("3.2", "Minutes layer (separate) — instructional time, by school",
      good  = "Avg weight and avg minutes per cycle shown side by side — weight x minutes should be comparable across schools (e.g., Sci-Tech 4.2 x 90 ~ Putnam 8.0 x 43).",
      evidence = function() {
        cs_data |>
          filter(C_course_time_exclude == "Include") |>
          distinct(
            D_location_name, D_stu_id, C_class_id,
            M_stu_class_weight, M_cls_class_minutes
          ) |>
          group_by(D_location_name, D_stu_id) |>
          summarise(
            total_weight = sum(M_stu_class_weight, na.rm = TRUE),
            total_minutes_per_cycle = sum(M_cls_class_minutes, na.rm = TRUE),
            .groups = "drop"
          ) |>
          group_by(D_location_name) |>
          summarise(
            avg_weight = round(mean(total_weight), 1),
            avg_minutes_per_cycle = round(mean(total_minutes_per_cycle), 0),
            .groups = "drop"
          )
      }
    ),

    # ================================================================
    # CHECK 3.3: The "before resolution" snapshot
    # This captures the starting state so we can compare to "after" once
    # resolution runs. It shows:
    #   - How many students/teachers are double-booked
    #   - Average class size
    #   - Average weight per student/teacher
    # The "after" column will be filled in by the resolution pass.
    # ================================================================
    tr_check("3.3", "Before-resolution snapshot (the 'XYZ' of XYZ -> ABC)",
      good  = "The starting picture: DBs by type + avg class size + avg weights, unresolved. 'After' fills in once resolution runs.",
      value = tr_safe(
        glue(
          "students {pct_of(length(flagged_students), n_distinct(cs_data$D_stu_id[cs_data$C_course_time_exclude=='Include']))}% \u00B7 ",
          "teachers {pct_of(length(flagged_teachers), n_distinct(cs_data$D_employee_id[cs_data$C_class_size_and_teacher_load_exclude=='Include']))}%"
        ),
        NA
      ),
      evidence = function() {

        # Calculate average class size (how many students per class)
        avg_class_size <- cs_data |>
          filter(C_class_size_and_teacher_load_exclude == "Include") |>
          distinct(C_class_id, D_stu_id) |>
          count(C_class_id, name = "size") |>
          summarise(value = round(mean(size), 1)) |>
          pull(value)

        # Calculate average total weight per student
        avg_student_weight <- cs_data |>
          filter(C_course_time_exclude == "Include") |>
          distinct(D_stu_id, C_class_id, M_stu_class_weight) |>
          group_by(D_stu_id) |>
          summarise(total_weight = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |>
          summarise(value = round(mean(total_weight), 2)) |>
          pull(value)

        # Calculate average total weight per teacher
        avg_teacher_weight <- cs_data |>
          filter(C_class_size_and_teacher_load_exclude == "Include") |>
          distinct(D_employee_id, C_class_id, M_tch_class_weight) |>
          group_by(D_employee_id) |>
          summarise(total_weight = sum(M_tch_class_weight, na.rm = TRUE), .groups = "drop") |>
          summarise(value = round(mean(total_weight), 2)) |>
          pull(value)

        # Count how many students/teachers are in the "included" pool
        n_included_students <- n_distinct(
          cs_data$D_stu_id[cs_data$C_course_time_exclude == "Include"]
        )
        n_included_teachers <- n_distinct(
          cs_data$D_employee_id[cs_data$C_class_size_and_teacher_load_exclude == "Include"]
        )

        data.frame(
          metric = c(
            "Students double-booked",
            "Teachers double-booked",
            "Avg class size",
            "Avg weight / student",
            "Avg weight / teacher"
          ),
          before = c(
            paste0(length(flagged_students), " (", pct_of(length(flagged_students), n_included_students), "%)"),
            paste0(length(flagged_teachers), " (", pct_of(length(flagged_teachers), n_included_teachers), "%)"),
            avg_class_size,
            avg_student_weight,
            avg_teacher_weight
          ),
          after = rep("pending resolution", 5),
          stringsAsFactors = FALSE
        )
      }
    ),
        tr_check("3.4", "Student resolution — weight reclaimed from double-bookings",
      good = "After the meeting-level split, avg student weight drops only by the double-booked share (small if DBs are rare).",
      value = tr_safe({
        ba <- cs_data |> filter(C_course_time_exclude == "Include") |>
          group_by(D_stu_id) |>
          summarise(b = sum(M_cls_class_weight, na.rm = TRUE),
                    a = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop")
        glue("avg {round(mean(ba$b),2)} -> {round(mean(ba$a),2)} per student")
      }, NA),
      evidence = function() {
        cs_data |> filter(C_course_time_exclude == "Include") |>
          group_by(D_location_name, D_stu_id) |>
          summarise(before = sum(M_cls_class_weight, na.rm = TRUE),
                    after  = sum(M_stu_class_weight, na.rm = TRUE), .groups = "drop") |>
          group_by(D_location_name) |>
          summarise(n_students        = n_distinct(D_stu_id),
                    avg_weight_before = round(mean(before), 2),
                    avg_weight_after  = round(mean(after), 2),
                    weight_reclaimed  = round(mean(before) - mean(after), 3), .groups = "drop")
      })
  )
)

htmltools::browsable(render_stage(s03))
```

*Next (pass 2) -> resolution: class 100%-DB weight adjustment, teacher consolidate/load-off (union + serialized expression), student-weight adjustment, then the "after" column fills in and we re-export.*
