---
title: "1_raw_data_HR_validation"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: "July 31 2024"
review:
last update: "March 12 2025"
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

=======
##### JK comments Oveall:

-   *As we talked about - I think we can get rid of saving each check to a sep df object as it can just create more clutter - maybe worth it if coming back to a check a lot*

-   *Also can get rid of print() since the output is shown below the code chunk regardless*

-   *Also a lot of the checks is specific to 2023 - meta question around whether we want to keep that or group by year?*

    -   *If its more helpful to check one year at a time, then maybe we should create a year variable at the beginning of 2.7 (where the code get;s into specific years) and the user can just update that variable one time and does not need to worry about changing it in each check......*

# Part 0: Read Libraries and connect to Sharepoint


```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```


# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. If your data is spread out over multiple files - i.e., **one file per year, go to step 1.1**

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning = F, message = F}

raw_data_folder_path <- "**Set Folder Path**"

#bring in file
hr_raw_data <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "**file_name_with_extension**")

# Clean column names to remove spaces and special characters
hr_raw_data <- hr_raw_data %>% clean_names()
```

## 1.1 Bring in data for Multiple Files

-   **What:** The code below grabs all csv files in a folder and stacks the data into one data frame. Use this if your data is spread out across multiple files (e.g., student_dem_2019, student_dem_2020, student_dem_2021). Otherwise, Jump to Part II

-   **Project Team Action:**
    -   Make sure that all relevant files are in one folder in SharePoint.

    -   Replace the file path below with your file path (in the code chunk below the first one)
    
-   **NOTES:** *This long function will most likely be replaced by an ers_tools package by end of March 2025. Check back for details!*

```{r}
#1. Run the function
ers_read_sharepoint_folder <- function(folder_path,
                                file_name_with_extension = NULL,
                                sheet_name = NULL,
                                drive_name = "client_work_drive",
                                skip_rows = 0,
                                na_values = c(""))
{
  # Clean the folder path and extract the drive name from it if not explicitly provided
  cleaned <- erstools::ers_sharepoint_path_clean(folder_path)
  folder_path <- cleaned$path_clean  # Cleaned folder path for SharePoint
  if (missing(drive_name)) {
    drive_name <- cleaned$drive_name
  }
  # Assign Drive
  orgfiles_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles") %>% suppressMessages
  external_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/external") %>% suppressMessages
  data_hub_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/ers-data") %>% suppressMessages
  client_work_drive <- orgfiles_site$get_drive("Client Work")
  internal_drive <- orgfiles_site$get_drive("Internal")
  external_drive <- external_site$get_drive("Documents")
  data_hub_drive <- data_hub_site$get_drive("Documents")
  drive <- switch(drive_name,
                  client_work_drive = client_work_drive,
                  internal_drive = internal_drive,
                  external_drive = external_drive,
                  data_hub_drive = data_hub_drive,
                  stop("Invalid drive name. Use 'client_work_drive', 'internal_drive', 'external_drive', or 'data_hub_drive'."))
  # Error message if folder path doesn't exist
  folder_item <- try(drive$get_item(folder_path), silent = TRUE)
  if (inherits(folder_item, "try-error")) {
    stop("Folder path does not exist: '", folder_path, "'")
  }
  # STEPS FOR READING IN WHOLE FOLDER OF FILES
  if (is.null(file_name_with_extension)) {
    files <- drive$list_items(folder_path)
    valid_files <- files[grepl("\\.(csv|xls|xlsx)$", files$name, ignore.case = TRUE), ]
    # Error message if no files in folder
    if (nrow(valid_files) == 0) {
      stop("No valid files (.csv, .xls, .xlsx) found in the folder.")
    }
    # Create a data list and use fread or read_excel to read in the files
    data_list <- list()
    for (file in valid_files$name) {
      file_path <- file.path(folder_path, file)
      temp_file <- tempfile(fileext = paste0(".", tools::file_ext(file)))
      drive$download_file(file_path, dest = temp_file)
      file_data <- if (tools::file_ext(file) == "csv") {
        fread(temp_file, skip = skip_rows, na.strings = na_values)
      } else {
        read_excel(temp_file, na = na_values, guess_max = 100000)
      }
      data_list[[tools::file_path_sans_ext(file)]] <- file_data
    }
    #assign dataframes to environment
    return(invisible(data_list))
  }
  #PROCESS SINGLE FILE ONLY
  file_name_with_extension <- gsub("%20", " ", file_name_with_extension)
  data_path <- file.path(folder_path, file_name_with_extension)
  #Error message if file doesn't exist
  file_item <- try(drive$get_item(data_path), silent = TRUE)
  if (inherits(file_item, "try-error")) {
    stop("File '", file_name_with_extension, "' does not exist in folder: '", folder_path, "'")
  }
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  temp_file <- tempfile(fileext = file_extension)
  drive$download_file(data_path, dest = temp_file)
  hr_raw_data <-
    if (tools::file_ext(file_name_with_extension) == "csv") {
      fread(temp_file, skip = skip_rows, na.strings = na_values)
    } else if (tools::file_ext(file_name_with_extension) %in% c("xls", "xlsx")) {
      read_excel(temp_file, sheet = sheet_name, skip = skip_rows, na = na_values, guess_max = 100000)
    } else if (tools::file_ext(file_name_with_extension) == "shp") {
      # Create a temporary directory
      temp_dir <- tempdir()
      # Define possible extensions
      shp_components <- c("shp", "shx", "dbf", "prj", "cpg")
      # Download each component
      for (ext in shp_components) {
        file_path <- file.path(temp_dir,
                               paste0(tools::file_path_sans_ext(file_name_with_extension), ".", ext))
        drive$download_file(file.path(folder_path,
                                      paste0(tools::file_path_sans_ext(file_name_with_extension),
                                             ".", ext)), dest = file_path)
      }
    } else {
      stop("Unsupported file format. Supported formats are xls, xlsx, csv, shp: '", file_name_with_extension, "'")
    }
  return(hr_raw_data)
}
```


```{r}
#get the list of files
data_list <- ers_read_sharepoint_folder("https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/District%20Partners/Denver/2024%20Transcript%20Analysis/1.%20Data%20&%20Analysis%20-%20Secure/1.0%20Raw%20Data/Student%20Dem%20Files")


# Use list2env to assign to your environment
list2env(data_list, envir = .GlobalEnv)

# Bind all tables into a single dataframe
hr_raw_data <- bind_rows(data_list, .id = "source_file")  # Adds a column indicating source file (optional)

# Clean column names to remove spaces and special characters
hr_raw_data<- hr_raw_data %>% clean_names()

# View combined data
glimpse(hr_raw_data)
```


## Part 2: Validate Data

### 2.1 Run ers_column_info on your data

```{r}
ers_column_info(hr_raw_data)
```

### 2.2 Manual Checks

-   **What:** a handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input the result of the check in the Data Validation Checks Table.

-   Do we have right year(s) of data?

-   **Check:** Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

##### JK Comments:

-   *are the term checks relevant for HR?*

-   *Re: Race/Ethnicity - may be helpful to note why this is important/matters. Is this specific to transcript? What are implications for sep columns vs singular columns?*

-   *Are there other manual checks that would be helpful for an HR file?*

-   Are course ids organized by term? And does term signify years, semesters, or quarters? Do they match the notation of course IDs in the transcript file?

    -   Check: look at the term column to determine how term is defined as that will have implications for the number of rows you would expect to see in your data, by teacher

-   Does the district report race/ethnicity as one field or separate fields?

    -   Check: look at the term column to determine how race/ethnicity is reported. Some districts report race and ethnicity as separate fields

**Project Team Action:** Gut check/track manual checks

### 2.3 Understand File Structure

##### JK comments

-   *Should we update the check language to reflect HR files that are just one row per teacher?*

#### a. Number of Total Rows

-   **What:** Understand the structure of your file by loading the number of rows

-   **Check:** The number of rows depends on how many sections you expect teachers to teach per year which is explained by how many terms are covered in the data. For example, if teachers teach 6 sections per semester, and rows are organized by semester, you expect to see the number of rows be equal to 6 \* 2 \* count of high school teachers. If there is only one row per teacher or you don't have complete term or course ID information, you would expect to see the number of rows equal to the number of high school teachers.

-   **Project Team Action:** Gut check/track the result of your check. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY high school teachers who teach X sections per semester""

```{r}
#Understand the total number of rows in your file

hr_raw_data %>% nrow()

```

#### b. Resolution Remove Rows

##### JK comments

-   *I know there is a world where there may be "obvious" rows to remove like "Alternative Schools", "Blank Rows", etc... but it would be hard to know which rows based on the previous check without doing deeper looks into data. So if there are more rows then we would expect, is it because of duplicates? And what are driving those duplicates, etc. etc. (which the code gets into below). So I think we either just remove this OR its renaming the step to be "Remove Rows for obvious things"... or something more elegant then that lol*

-   **What:** Transform your data if necessary to ensure that you have just teachers/the personnel needed for analysis. You can also leave a note about district follow-up required if the team did not receive the full data set

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before

-   **Project Team Action:** Maybe change (see guidance in code block)


```{r}
#remove rows
# Remove rows where a specific condition is met
# For example, remove rows where age is less than 18
hr_raw_data <- hr_raw_data %>%
  filter(age >= 18)

# Print the result
print(head(hr_raw_data))
```

### 2.4 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. See resolution of critical/important fields that are missing in step 2.2b. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present. This also creates a new data-frame just for teachers, so if you have any non-teaching staff at this point you will want to make sure they are removed from the file.

-   **Project Team Action:** Definitely change (see guidance in code block)

##### JK Comments

-   *Let's make sure these column names match the human capital practice area naming conventions... (see tettra page linked above)*

-   *And maybe we move this right after manual checks so steps 2.3 don't need to be updated?*

```{r}
#if the ERS Standard column is present in your data then type your district's column on the right hand side (it is okay if the names are identical)
#if the ERS Standard column is not present in your data then type "MISSING"
#you can also exclude the columns you do not need for your analysis by simply not renaming those columns in this step (the mutate function with the .keep = none argument only retains columns that follow the ERS column tagging convention)

teachers_combined_file <- hr_raw_data %>%
  mutate(
#critical    
         D_employee_id = `teacher_id`, 
         D_location_id = `school_code`,
         D_location_desc = `school_name`,
         D_employee_gender = `gender_code`,
         D_employee_race = `race_ethnicity_code`,
         D_years_exp = `years_of_experience`,
         D_course_id = NA,
         D_course_section = NA,
         D_year= "2023",
         D_term = NA,

#nice-to-have
         D_hire_date = `date_of_hire`,
         D_as_of_date = `as_of_date`,         
         D_position_control_num = `position_control_number`,
         .keep = "none")
```

#### b. Review columns

-   **What:** Manually check that you have the same number of columns + the number of missing columns. Review the columns you are missing

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them for later. Update the Data Validation Checks table with the result of your manual checks

### 2.5 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting


```{r}
#View duplicates
teachers_combined_file_dups <- teachers_combined_file %>%
  group_by_all() %>%
  mutate(n = n()) %>%
  filter(n>1) %>%
  print()
```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates

-   **Check:** That the number of duplicates is now 0

```{r}
#removes duplicates
# teachers_combined_file <- 
#      teachers_combined_file[!duplicated(teachers_combined_file), ]
```

### 2.6 Missing & Unique values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action**: None

```{r}
# Update blanks or NULL to NA
teachers_combined_file <- teachers_combined_file %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates. Update Data Validation Checks table

```{r}
#Create new data frame with column information
col_info <-  data.frame(
  Column_Name = names(teachers_combined_file),
  
  Count_Rows = sapply(teachers_combined_file, function(x) length(na.omit(x))),
  NA_Rows = sapply(teachers_combined_file, function(x) sum(is.na(x))),
  Unique_Rows = sapply(teachers_combined_file, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)

#Calc % missing
col_info <- col_info %>%
            mutate(Percent_Missing = NA_Rows / Count_Rows * 100)

#Print new data frame
col_info
```

### 2.7 Count of locations assigned to each teacher

##### JK Comments:

-   *This would get annoying to do for each year of data separately, so* would consider instead of filtering to a specific year grouping by year and then pivoting so the years are in columns

-   *Also, not sure we need print() or to save these checks to a different data frame..could create more clutter - (should apply this lens throughout!)*

-   **What:** Check how many teachers are assigned to more than one location

-   **Check:** Make sure code block runs without errors

-   **Project Team Action:** Update year as necessary

```{r}

#count of locations assigned to each teacher
teachers_combined_file %>% 
  group_by(D_employee_id) %>%
  summarize(School_Count = n_distinct(D_location_id)) %>% 
  count(School_Count) %>% 
  print()

#how many total teachers in the file?
teachers_combined_file %>%
 filter(D_year == 2023) %>% 
  group_by(D_employee_id) %>% 
  summarise(Teacher_Count = n_distinct(D_employee_id)) %>%
  count(Teacher_Count) %>% 
  print()

```

#### a. Add flag for teachers who are assigned to multiple locations

##### JK cmments

-   *probably don't need the to save these as dfs and can get rid of print()*

-   **What:** Check how many teachers are assigned to more than one location

-   **Check:** Make sure code block runs without errors

-   **Project Team Action:** Update year as necessary

```{r}
#group by D_year and D_employee_id, then count the distinct D_location_id values
teachers_combined_file %>%
  group_by(D_year, D_employee_id) %>%
  mutate(C_employee_multiple_locations = if_else(n_distinct(D_location_id) > 1, "Y", "N")) 

#count teachers assigned to multiple locations 
teachers_combined_file %>%
  filter(C_employee_multiple_locations == "Y") %>%
  summarise(distinct_employee_count = n_distinct(D_employee_id))

```

#### b. What are the total number of rows by teacher? (How big of an issue is this)

##### JK comments:

-   *probably don't need the to save these as dfs and can get rid of print()* 

<!-- -->

-   **What:** How many rows do most teachers have? Should be about the number of sections X locations that they teach

-   **Check:** Make sure code block runs without errors

-   **Project Team Action:** Update year as necessary, Gut check/track, follow up as needed with irregularities (teachers with no rows, teachers with many more than expected rows), Check the total rows against the multiple locations flag. If they have multiple rows, but are at multiple locations

```{r}
total_rows_by_teacher_2023 <- teachers_combined_file %>%
 # filter(D_year == "2023") %>%
  group_by(D_year, D_employee_id, C_employee_multiple_locations) %>%
  summarise(total_rows = n()) %>% 
  arrange(desc(total_rows)) %>%
  print()
```

#### c. Investigate individual teachers

##### JK comments:

-   *probably don't need the to save these as dfs and can get rid of print()*

<!-- -->

-   **What:** Check random teacher IDs with many rows

-   **Check:** Make sure code block runs without errors

-   **Project Team Action:** Update year as necessary, Gut check/track, follow up as needed with irregularities

-   **Note:** Large numbers of sections might often be attributed to a PE teacher (teach many sections of a sport, etc. Could also be attirubted to substitute teachers.

```{r}
#Investigate teacher ids with the most/fewest number of rows
teacher_id_dive_1 <- teachers_combined_file %>% 
  filter(D_year == 2023, D_employee_id == XXX) %>% #replace with an ID from your own file
  print()

teacher_id_dive_2 <- teachers_combined_file %>% 
  filter(D_year == 2023, D_employee_id == XXX) %>% #replace with an ID from your own file
  print()

teacher_id_dive_3 <- teachers_combined_file %>% 
  filter(D_year == 2023, D_employee_id == XXX) %>% #replace with an ID from your own file
  print()
```

### 2.8 Distinct Count of Course ID by Teacher ID (schools of interest)

#### a. Return the count

##### JK comments:

-   *probably don't need the to save these as dfs and can get rid of print()*

<!-- -->

-   **What:** Count of Teachers by distinct course count to understand how many **distinct** courses teachers are teaching

-   **Check:** Review the counts Teachers by course count. We would expect majority of teachers to have a low number of distinct course IDs, since most would teach multiple sections of the same course

-   **Project Action:** Gut check/track, follow up as needed with irregularities

```{r echo=TRUE}
#filter to high schools / summer schools of interest
filtered_high_schools <- 
     teachers_combined_file[teachers_combined_file$D_location_id %in% c(20, 21, 22, 23, 24),] #replace with your schools of interest

#creates a table that is the employee ID by number of courses they are teaching
course_counts_2023_filtered <- filtered_high_schools %>% 
  filter(D_year == 2023) %>%
  group_by(D_employee_id) %>%
  summarize(Course_Count = n_distinct(D_course_id)) %>% 
  mutate(Course_Count_Bucket = case_when(Course_Count == 0 ~ "0",
          Course_Count <= 2 ~ "01-02",
          Course_Count <= 4 ~ "03-04",
          Course_Count <= 6 ~ "05-06",
          Course_Count <= 8 ~ "07-08",
          Course_Count <= 10 ~ "09-10",
          Course_Count <= 12 ~ "11-12",
          Course_Count <= 16 ~ "13-16",
          Course_Count <= 20 ~ "17-20",
          Course_Count <= 24 ~ "21-24",
          Course_Count <= 28 ~ "25-28",
          Course_Count <= 32 ~ "29-32",
          Course_Count >32 ~ "33+"
          ))

#returns the percent of teachers that fall into each course bucket
course_count_buckets_2023_filtered <- course_counts_2023_filtered %>%
  group_by(Course_Count_Bucket) %>%
  summarize(total_employees = n_distinct(D_employee_id)) %>%
  ungroup() %>%
  mutate(total_employees_all_buckets = sum(total_employees)) %>%
  group_by(Course_Count_Bucket) %>%
  summarize(n = total_employees,
            pct_of_total = round(100 * (n / total_employees_all_buckets), 2)) %>% 
  print()
```

### 2.9 Count of Sections per teacher

#### a. Return the count (note this is not a distinct count)

-   **What:** Returns the \# of sections per teacher (this is not a distinct count since section numbers repeat for different courses)

-   **Check:** You would expect to see most teachers teaching \~6 sections if operating on an 8-period schedule where classes repeat daily. You are looking for wonky things - teachers teaching a ton of sections, teachers teaching 1 section, etc.

-   **Project Action:** Gut check/track, follow up as needed with irregularities

```{r}
#returns a count of sections
section_counts_2023_filtered <- filtered_high_schools %>% 
  filter(D_year == 2023) %>%
  group_by(D_employee_id) %>%
  summarize(Section_Count = n()) %>% 
  mutate(Section_Count_Bucket = case_when(Section_Count == 0 ~ "0",
          Section_Count <= 2 ~ "01-02",
          Section_Count <= 4 ~ "03-04",
          Section_Count <= 6 ~ "05-06",
          Section_Count <= 8 ~ "07-08",
          Section_Count <= 10 ~ "09-10",
          Section_Count <= 12 ~ "11-12",
          Section_Count <= 16 ~ "13-16",
          Section_Count <= 20 ~ "17-20",
          Section_Count <= 24 ~ "21-24",
          Section_Count <= 28 ~ "25-28",
          Section_Count <= 32 ~ "29-32",
          Section_Count >32 ~ "33+"
          ))

#returns the percent of teachers that fall into each course bucket
section_count_buckets_2023_filtered <- section_counts_2023_filtered %>%
  group_by(Section_Count_Bucket) %>%
  summarize(total_employees = n_distinct(D_employee_id)) %>%
  ungroup() %>%
  mutate(total_employees_all_buckets = sum(total_employees)) %>%
  group_by(Section_Count_Bucket) %>%
  summarize(n = total_employees,
            pct_of_total = round(100 * (n / total_employees_all_buckets), 2)) %>% 
  print()
```

#### b. Investigate individual teachers

-   **What:** View your data file, and filter to the year and employee id to a teacher that showed up with the most/fewest rows

-   **Check:** Look up the course IDs in the district catalog. Are these self-study courses or another type of non-traditional course?

-   **Project Action:** Gut check/track, follow up as needed with irregularities

```{r}
#Investigate teacher ids with the most/fewest number of rows
teacher_id_dive_1 <- filtered_high_schools %>% 
  filter(D_year == 2023, D_employee_id == XXX) %>% #replace with an ID from your own file
  print()

teacher_id_dive_2 <- teachers_combined_file %>% 
  filter(D_year == 2023, filtered_high_schools == XXX) %>% #replace with an ID from your own file
  print()

teacher_id_dive_3 <- teachers_combined_file %>% 
  filter(D_year == 2023, filtered_high_schools == XXX) %>% #replace with an ID from your own file
  print()
```

# Part 3: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None


```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/District%20Partners/Columbus%20City/22-23%20Financial%20Equity%20&%20Funding%20System%20Design/1.%20Data%20&%20Analysis%20-%20Secure/Processed%20Data%20for%20R"

ers_write_sharepoint(data = XXX,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "XXX.csv")
```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 6: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
