---
title: "1_raw_data_student_validation"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: July 23 2024
review:
last update: March 12 2025
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. If your data is spread out over multiple files - i.e., one file per year, go to step 1.1

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning = F, message = F}

raw_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

#bring in file
student_dem_raw <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "HPS_Student_Data.csv",
  drive_name = "internal_drive")

# Clean column names to remove spaces and special characters
student_dem_raw <- student_dem_raw %>% clean_names()
```

## 1.2 Bring in data for Multiple Files

##### *JK Comments*

-   *Given those a lot of this code assumed multiple years of data and multiple years often comes in sep files, i added code below to pull in multiple years and stack them alll. This is a new function that Zach made!*
-   **What:** The code below grabs all csv files in a folder and stacks the data into one data frame. Use this if your data is spread out across multiple files (e.g., student_dem_2019, student_dem_2020, student_dem_2021)
-   **Project Team Action:**
    -   Make sure that all relevant files are in one folder in SharePoint.

    -   Replace the file path below with your file path (in the code chunk below the first one)
-   **NOTES:** *This long function will most likely be replaced by an ers_tools package by end of March 2025. Check back for details!*

```{r}
#1. Run the function
ers_read_sharepoint_folder <- function(folder_path,
                                file_name_with_extension = NULL,
                                sheet_name = NULL,
                                drive_name = "client_work_drive",
                                skip_rows = 0,
                                na_values = c(""))
{
  # Clean the folder path and extract the drive name from it if not explicitly provided
  cleaned <- erstools::ers_sharepoint_path_clean(folder_path)
  folder_path <- cleaned$path_clean  # Cleaned folder path for SharePoint
  if (missing(drive_name)) {
    drive_name <- cleaned$drive_name
  }
  # Assign Drive
  orgfiles_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles") %>% suppressMessages
  external_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/external") %>% suppressMessages
  data_hub_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/ers-data") %>% suppressMessages
  client_work_drive <- orgfiles_site$get_drive("Client Work")
  internal_drive <- orgfiles_site$get_drive("Internal")
  external_drive <- external_site$get_drive("Documents")
  data_hub_drive <- data_hub_site$get_drive("Documents")
  drive <- switch(drive_name,
                  client_work_drive = client_work_drive,
                  internal_drive = internal_drive,
                  external_drive = external_drive,
                  data_hub_drive = data_hub_drive,
                  stop("Invalid drive name. Use 'client_work_drive', 'internal_drive', 'external_drive', or 'data_hub_drive'."))
  # Error message if folder path doesn't exist
  folder_item <- try(drive$get_item(folder_path), silent = TRUE)
  if (inherits(folder_item, "try-error")) {
    stop("Folder path does not exist: '", folder_path, "'")
  }
  # STEPS FOR READING IN WHOLE FOLDER OF FILES
  if (is.null(file_name_with_extension)) {
    files <- drive$list_items(folder_path)
    valid_files <- files[grepl("\\.(csv|xls|xlsx)$", files$name, ignore.case = TRUE), ]
    # Error message if no files in folder
    if (nrow(valid_files) == 0) {
      stop("No valid files (.csv, .xls, .xlsx) found in the folder.")
    }
    # Create a data list and use fread or read_excel to read in the files
    data_list <- list()
    for (file in valid_files$name) {
      file_path <- file.path(folder_path, file)
      temp_file <- tempfile(fileext = paste0(".", tools::file_ext(file)))
      drive$download_file(file_path, dest = temp_file)
      file_data <- if (tools::file_ext(file) == "csv") {
        fread(temp_file, skip = skip_rows, na.strings = na_values)
      } else {
        read_excel(temp_file, na = na_values, guess_max = 100000)
      }
      data_list[[tools::file_path_sans_ext(file)]] <- file_data
    }
    #assign dataframes to environment
    return(invisible(data_list))
  }
  #PROCESS SINGLE FILE ONLY
  file_name_with_extension <- gsub("%20", " ", file_name_with_extension)
  data_path <- file.path(folder_path, file_name_with_extension)
  #Error message if file doesn't exist
  file_item <- try(drive$get_item(data_path), silent = TRUE)
  if (inherits(file_item, "try-error")) {
    stop("File '", file_name_with_extension, "' does not exist in folder: '", folder_path, "'")
  }
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  temp_file <- tempfile(fileext = file_extension)
  drive$download_file(data_path, dest = temp_file)
  raw_data <-
    if (tools::file_ext(file_name_with_extension) == "csv") {
      fread(temp_file, skip = skip_rows, na.strings = na_values)
    } else if (tools::file_ext(file_name_with_extension) %in% c("xls", "xlsx")) {
      read_excel(temp_file, sheet = sheet_name, skip = skip_rows, na = na_values, guess_max = 100000)
    } else if (tools::file_ext(file_name_with_extension) == "shp") {
      # Create a temporary directory
      temp_dir <- tempdir()
      # Define possible extensions
      shp_components <- c("shp", "shx", "dbf", "prj", "cpg")
      # Download each component
      for (ext in shp_components) {
        file_path <- file.path(temp_dir,
                               paste0(tools::file_path_sans_ext(file_name_with_extension), ".", ext))
        drive$download_file(file.path(folder_path,
                                      paste0(tools::file_path_sans_ext(file_name_with_extension),
                                             ".", ext)), dest = file_path)
      }
    } else {
      stop("Unsupported file format. Supported formats are xls, xlsx, csv, shp: '", file_name_with_extension, "'")
    }
  return(raw_data)
}
```

```{r}
#get the list of files
data_list <- ers_read_sharepoint_folder("https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/District%20Partners/Denver/2024%20Transcript%20Analysis/1.%20Data%20&%20Analysis%20-%20Secure/1.0%20Raw%20Data/Student%20Dem%20Files")


# Use list2env to assign to your environment
list2env(data_list, envir = .GlobalEnv)

# Bind all tables into a single dataframe
student_dem_raw <- bind_rows(data_list, .id = "source_file")  # Adds a column indicating source file (optional)

# Clean column names to remove spaces and special characters
student_dem_raw <- student_dem_raw %>% clean_names()

# View combined data
ers_column_info(student_dem_raw)
```

## Part 2: Validate Data

### 2.1 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. See resolution of critical/important fields that are missing in step. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present

-   **Project Team Action:** Definitely change (see guidance in code block)

```{r}
#if the ERS Standard column is present in your data then type your district's column on the right hand side (it is okay if the names are identical)
#if the ERS Standard column is not present in your data then type "MISSING"
#you can also exclude the columns you do not need for your analysis by simply not renaming those columns in this step (the mutate function with the .keep = none argument only retains columns that follow the ERS column tagging convention)

student_dem_raw <- student_dem_raw %>%
  mutate(
#critical    
         D_stu_id = as.character(`masked_student_id`),
         D_location_name = as.character(`school_name_x`),
         D_stu_enter_date = as.Date(`stu_entry_date`, format = "%Y/%m/%d"),
         D_stu_exit_date = as.Date(`stu_exit_date`, format = "%Y/%m/%d"),,
         D_year = as.character(`school_year`),
         D_location_id = as.character(`stu_school_id`),
         D_stu_grade = as.numeric(`stu_grade`),
         D_stu_race = as.character(stu_race),
         D_stu_gender = NA,
         D_stu_swd = as.character(`stu_swd_flag`),
         D_stu_ell = as.character(`stu_ell_flag`),
         D_stu_frl = as.character(`stu_frl_flag`),,
      
#nice-to-have
         D_stu_enrollment_status = as.character(`stu_enrollment_status`),
        D_stu_exit_code = as.character(`stu_exit_code`),
         D_stu_exit_code_desc = as.character(`stu_exit_code_description`),
         .keep = "none") #"Keep-None" here will ONLY keep the columns listed above, so if you do not list a column, it will be deleted. Make sure to keep any district specific columns or other columns you need for your analysis.

#district-specific- add your own!
#XXX = XXX

```

#### b. Review columns

-   **What:** Manually check that you have the columns you need/would expect after renaming, and have not lost any critical data. Review the columns you are missing (as necessary).

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them. Update the Data Validation Checks table with the result of your manual checks

### 2.2 Manual Checks

##### JK comments:

-   *Updated for readability (bolded/bullets)*

-   *Re: Race/Ethnicity - may be helpful to note why this is important/matters. Is this specific to transcript? What are implications for sep columns vs singular columns?*

-   **What:** A handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input them into a checks table in OneNote, Word, or another tool.

    1.  **Do we have right year(s) of data?** Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

    2.  **Does the district report race/ethnicity as one field or separate fields?** Look at the term column to determine how race/ethnicity is reported. Some districts report race and ethnicity as separate fields, (ie. Race: Caucasian, Ethnicity: Middle Eastern) in which case its helpful to create a rollup column to accurately answer district questions

-   **Project Team Action:** Update the Data Validation Checks table with the result of your manual checks

#### a. Number of Total Rows

-   **What:** Understand the structure of your file (or files) by loading the number of rows, check the number of unique students and counts by year. If your files were in miltiple years, you should make sure they are now stacked with the code in step 1.1 above!

-   **Check:** Number of rows should be roughly the number of high school students in the district

-   **Project Team Action:** Update the Data Validation Checks table with the result of your check. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY high school students in the district"

##### JK Comments:

-   *How often do files contain all years of data vs one year at a time? For Denver, each file is separate... so below code would not apply...* **JK 4/7 Comments:** *Hmm maybe we just remove it then?? Seems like calculating the year was pretty specific to Lubbock and that most files come in separate years file so calculating the year would be easy? Maybe we can something like "this loooks at the count by year.. if you need to calculate a year consider using student/enter dates, etc. See Lubbock script for example.. " and then link to it?*

-   *Or is it that we should be recommending that files should be stacked at this point? I added a function to the beginning that does this!*

-   *Maybe also helpful to mention that "if your file did not come with a school year column then you can create one with enrollment dates, etc."*

```{r}
#student demographics
student_dem_raw %>% nrow() #View output of check to know what to write for your status column

# Calculate the number of days between EntryDate and ExitDate
student_dem_raw <- student_dem_raw %>%
  mutate(
    M_days_enrolled = as.numeric(difftime(`D_stu_exit_date`, `D_stu_enter_date`, units = "days"))
  )

# write table of the count of unique Student_number by school_year
student_dem_raw %>%
  group_by(`D_year`) %>%
  summarise(n_distinct(`D_stu_id`))

## Number of unique values of student number in student_dem_raw
student_dem_raw %>%
  summarise(n_distinct(`D_stu_id`))

```

#### b. Resolution Remove Rows

-   **What:** Transform your data if necessary. You can also leave a note about district follow-up required if the team did not receive the full data set

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before

-   **Project Team Action:** Update code as needed to filter/remove data

```{r}
#remove rows
# Remove rows where a specific condition is met
# For example, remove rows where age is less than 18
student_dem_raw <- student_dem_raw %>%
  filter(age >= 18)

# Print the result
print(head(student_dem_raw))

```

### 2.3 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting

-   **Project Action:** Update Data Validation Checks table

```{r}
#view duplicates in student enrollment file
student_enrollment_combined_dups <- student_dem_raw %>%
  group_by_all() %>%
 mutate(n = n()) %>%
    filter(n>1)
view(student_enrollment_combined_dups)
nrow(student_enrollment_combined_dups)
```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates

-   **Check:** That the number of duplicates is now 0

-   **Project Action:** Update Data Validation Checks table

```{r}
#removes duplicates
student_dem_raw <- student_dem_raw[!duplicated(student_dem_raw), ]
#count to make sure the right number of rows were removed
nrow(student_dem_raw)

```

### 2.4 Missing & Unique Values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action**: None

```{r}
# Update blanks or NULL to NA
student_dem_raw <- student_dem_raw %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates. Update Data Validation Checks table

```{r}
#create new data frame with column information
col_info_enrollment <-  data.frame(
  Column_Name = names(student_dem_raw),
  Data_Type = sapply(student_dem_raw, class),
  Count_Rows = sapply(student_dem_raw, function(x) length(na.omit(x))),
  NA_Rows = sapply(student_dem_raw, function(x) sum(is.na(x))),
  Unique_Rows = sapply(student_dem_raw, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)
# #calc % missing
 col_info_enrollment <- col_info_enrollment %>%
            mutate(Percent_Missing = NA_Rows / nrow(student_dem_raw) * 100)

#print new data frame
col_info_enrollment

## Now do the same thing, but instead check for values of 0 
#create new data frame with column information
col_info_enrollment <-  data.frame(
  Column_Name = names(student_dem_raw),
  Data_Type = sapply(student_dem_raw, class),
  Count_Rows = sapply(student_dem_raw, function(x) length(na.omit(x))), 
  Zero_Rows = sapply(student_dem_raw, function(x) sum(x == 0)),
  Unique_Rows = sapply(student_dem_raw, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)
# #calc % missing
 col_info_enrollment <- col_info_enrollment %>%
            mutate(Percent_Zero = Zero_Rows / nrow(student_dem_raw) * 100)

#print new data frame
col_info_enrollment
```

### 2.5 Distinct Counts of Students

#### a. Return the counts of student id

-   **What:** Count of student ids

-   **Check:** Review the counts of student id. Compare with public data

-   **Project Action:** Update Data Validation Checks table

```{r}
#student enrollment
student_enrollment_count <- student_dem_raw %>% 
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

```

#### b. Return Distinct Counts of Students by Year

-   **What:** Count of student ids by year

-   **Check:** Review the counts of student id by year. Are students evenly distributed across years as expected?

-   **Project Action:** Update Data Validation Checks table

```{r}
#count of unique student ids for each year of data
student_enrollment_count_by_year <- student_dem_raw %>%
  group_by(D_year) %>%
  summarise(count_students = n_distinct(`D_stu_id`)) %>% 
  print()

#count of unique student ids for each year of data in the high schools or schools of interest
student_enrollment_count_by_year <- student_dem_raw %>%
  #filter(D_location_name %in% c("XXX High School", #add schools of interest here
                                #"XXX High School", 
                                                #)) %>% 
  group_by(D_year) %>%
  summarise(count_students = n_distinct(`D_stu_id`)) %>% 
  print()
```

### 2.6 Number of years of data we have for each student

-   **What:** Creates a variable which assigns new value to student based on number of years of data we have for each unique student ID. This will be a valuable flag later when choosing cohort of students to analyze

-   **Check:** Choose at least three student ids to check. Filter to the student id in the combined file to identify how many years of data we have for them. Check the count in "num_years_data" column to make sure it is the same. You can also just visually inspect the file to make sure the "num_years_data" for each student id matches the number of rows for each student id

-   **Project Action:** None (run as is)

```{r}
#counts how many yrs of data we have for each student
student_enrollment_yrs_of_data <- student_dem_raw %>%
  group_by(D_stu_id) %>%
  summarise(C_years_of_data_enrollment = n_distinct(D_year)) 

#check the year counts
print(student_enrollment_yrs_of_data)

#then adds that variable  as a column to the original combined data file
student_enrollment_combined <- merge(student_dem_raw, student_enrollment_yrs_of_data, by = "D_stu_id", all.x = TRUE)

# write a table with the count of unique student IDs with each number of years of data
student_enrollment_combined %>%
  group_by(C_years_of_data_enrollment) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

```

### 2.7 Demographic Checks in district

#### a. % FRL, % ELL, % SWD

-   **What:** Check the percentage of students who qualify for free or reduced lunch by the year to compare to publicly available data

-   **Check:** (1) Check to make sure code worked - do total number of rows each year add up to what you would expect? (2) Compare % FRL to publicly available data

-   **Project Team Action:**

    -   Update the values of the columns to your district's values. E.G., ELL="Yes" or ELL = "True"

    -   Add to Data Validation Checks table

#### b. % FRL

##### JK COmments

##### JK Comments

-   *I updated these code chunks to pivot the data to make it easier to read.*

```{r}
#calculate percent FRL in district by year
 percent_frl_check <- student_dem_raw %>%
       mutate(criteria = ifelse(`D_stu_frl` == "Y",
                                "Yes",
                                "No")) %>% 
       group_by(criteria, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
       select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()

```

#### c. % ELL

```{r}
#calculate percent ELL in district by year
percent_ell_check <- student_dem_raw %>%
       mutate(criteria = ifelse(`D_stu_ell` == "ELL",
                                "Yes",
                                "No")) %>% 
       group_by(criteria, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
       select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()
```

#### d. % SWD

```{r}
#calculate percent SWD in district by year
 percent_swd_check <- student_dem_raw %>%
       mutate(criteria = ifelse(`D_stu_swd` == "Y",
                                "Yes",
                                "No")) %>% 
       group_by(criteria, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
     select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()
```

#### e. % Race/ethnicity

-   **What:** Check the percentage of students who identify as each race/ethnicity category to compare to publicly available data. If you created a rollup column earlier, make sure to use the correct column.

-   **Check:** Compare to publicly available data

-   **Project Team Action:** Gut check/track

```{r}
#group by D_year and D_stu_race, then calculate the percentage of each race category for each year
race_percent_year <- student_dem_raw %>%
  group_by(D_year, D_stu_race) %>%
  summarise(row_count = n()) %>%
  mutate(pct_rows = row_count / sum(row_count) * 100) %>%
  select(-row_count) %>%  # Remove row_count if you only need percentages
  pivot_wider(names_from = D_year, values_from = pct_rows) %>%
  print()
```

## Part 3: File Cleaning for export & Checks

-   **What:** Look over the data to ensure it is clean and ready for the next phase of analysis, add any final checks or cleaning

-   **Check:** Gut check/check against publicly available data (as needed)

-   **Project Team Action:** None

### 3.0 Check Random Student IDs

```{r}
# Create a function to pick a random student ID and create a new data frame
pick_random_student <- function(data) {
  # Pick a random student ID
  random_student_id <- sample(data$D_stu_id, 1)
  
  # Filter the data frame to create a new data frame for the selected student
  student_Check <- data %>%
    filter(D_stu_id == random_student_id)

  
  # Return the new data frame
  return(student_Check)
}

#student check 1
student_check_1 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_1)

#student check 2
student_check_2 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_2)

#student check 3
student_check_3 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_3)

```

### 3.1 Rearrange columns into desired order if needed

-   **What:** Creates new data-frame with changed order of appearance of your columns in your data-frame for an ease of viewing

-   **Check:** None

-   **Project Team Action:** None

```{r}
#rearrange columns by listing existing columns in the order you want them to appear
stu_roll_up <- student_dem_raw %>% 
  select(D_stu_id,
           D_location_id,
           D_location_name, 
           D_year, 
           D_stu_enter_date, 
           D_stu_exit_date, 
           D_stu_enrollment_status,  
           D_stu_exit_code, 
           D_stu_exit_code_desc, 
           D_stu_grade, 
           D_stu_gender, 
           D_stu_race, 
           D_stu_frl, 
           D_stu_ell, 
           D_stu_swd, 
           everything()) #Keeps all columns (if not listed above)
```

# Part 4: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** None

```{r}
processed_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)/Dummy Data"

ers_write_sharepoint(data = stu_roll_up,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "HPS_stu_rollup.csv",
                    drive_name = "internal_drive")
```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 6: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
<<<<<<< HEAD:1_raw_data_student_validation_NEW.Rmd
```---
title: "RMarkdown Template"
author: "Zach Friedman"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
=======
>>>>>>> a0949010c2b085e9a2bc38a97cfa4d327c784160:0. Archive Files/1_raw_data_student_validation.Rmd
```
