---
title: "2_cross_file_validation"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: "July 30 2024"
review:
last update: "March 12 2025"
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### JK comments:

-   *Can you please replace - through 1.4 this with the newest template format :)*

-   *Also apply the "do we really need to store this check" as a data frame object and remove when you think we don't really need it?*

    -   *Same with print()*

    -   

## Part 0: Read Libraries and connect to Sharepoint

### 0.1 (OPTIONAL) Install the latest version of the erstools library!

**ACTION: If you aren't sure if you have the most up to date release of erstools, follow these instructions: <https://app.tettra.co/teams/ersknowledge/pages/updating-erstools-functions-markdown-template>**

### 0.2 Read All Libraries

**ACTION: Run the code and read the comments for different functions you will have access to**

```{r, warning = F, message = F}
#ERS-specific functions!!
library(erstools) #includes functions like ers_read_sharepoint, ers_write_sharepoint, ers_table, and ers_column_info

#load data
library(Microsoft365R) # Use Microsoft365R to connect to SharePoint and load files
library(readxl) # Functions for reading Excel files: read_excel
library(writexl) # Functions for writing Excel files: write_xlsx
library(openxlsx) # Functions for writing Excel files: write.xlsx

#manipulate data
library(tidyverse) # Includes ggplot2, dplyr, tidyr, readr, etc.
library(data.table) # fread(): load large CSVs quickly
library(janitor) # Functions for cleaning and examining data: clean_names() instantly formats all column names
library(safejoin) # Provides safer join operations for 1:many matching - safe_left_join()
library(stringr) # str_detect(): pattern matching using regular expressions (see regex101.com)

#format and display data
library(scales) # Functions for scaling and formatting axes in plots: comma(), percent()
library(gt) # Creates and formats tables in a publication-ready format: gt()
library(flextable) # Creates and formats tables in a publication-ready format: flextable()
library(extrafont) # Manages and uses custom fonts in plots - loadfonts(): Loads fonts available on the system

#share output with supervisor
library(rmarkdown) # needed to create knitted files in output folder
library(downloadthis) # create download buttons in your knitted file so that a supervisor can download data in Excel!
```

## Part 1: Bring in your data

### 1.0 Authenticate to sharepoint to be able to read in files

**ACTION: After running the code, you may need to look at your console code window below for instructions to log in. A window may open in your browser to ask you to log into Microsoft to confirm your identity**

**NOTE: If you ever have authentication issues, first run this in your console then try running the code chunk below again - AzureAuth::clean_token_directory()**

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

### 1.1 Set your folder path(s)

**ACTION: Change the raw_data_folder_path to the path to your data. The folder path must start with a folder that is below the "Client Work" level. For example, it should start with "District Partners", "Internal", "State Partners", etc.**

**NOTE: If you don't want to type the name of your folder path, you can copy it directly from Sharepoint: <https://app.tettra.co/teams/ersknowledge/pages/copying-a-folder-path-from-sharepoint-to-use-in-r>**

```{r}
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/District%20Partners/Denver/2024%20Transcript%20Analysis/1.%20Data%20&%20Analysis%20-%20Secure/1.2%20Data%20Validated%20in%20R"
```

### 1.2 See guidance on using the "ers_read_sharepoint" function

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_read_sharepoint
```

### 1.3 Read in your data

**ACTION: Replace "raw_data" with the desired name of your data frame and replace "XXX.csv" with the file name. We suggest doing this for all instances of raw data by using the find and replace tool (Ctrl + F and then replace all instances of "raw_data" with the desired name.**

```{r}
#transcript data
transcript_data_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "1_transcript_data_validated.xlsx")

#student data
student_data_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "1_student_data_validated.xlsx")

#course catalog data
course_catalog_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "1_course_catalog_validated.xlsx")

```

### 1.4 View your data

**ACTION: Replace "raw_data" with the same name of your data frame you used above, and run the code to inspect your data.**

```{r, eval = interactive()}
View(raw_data)
```

## Part 2: Validate Data

### 2.0 Check out the ers_column_info function as a starting place in your data validation (credits to Mauro!)

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_column_info
```

### 2.1 Run ers_column_info on your data

**ACTION: Replace "raw_data" with the name of your data frame**

```{r}
ers_column_info(student_data_validated)
ers_column_info(transcript_data_validated)
ers_column_info(course_catalog_validated)

transcript_data_validated %>%
  pull(D_year) %>%
  unique()
```

### 2.2 Student counts

#### JK Comments:

*I asked chat GPT to see if there is a more effifient way to get the student counts and compare across years, etc. I added the code below this chunk. This would also make the year dynamic so would show for all years and a team wouldnt have to update this based on the year of data that they get. Take a spin through! If you think more clear, okay to work with then you can replace the original code with the new function...*

#### a. Return distinct student counts in student, transcript, and school files

-   **What:** Comparing student counts across files to check that they are comparable

-   **Check:** Are the student counts the same?

-   **Project Team Action:** Gut check/track

```{r}


#calculate the distinct count of D_stu_id for student file
# Define a function to calculate distinct counts by dataset and year
calculate_distinct_counts <- function(data, year = NULL) {
  result <- data
  
  if (!is.null(year)) {
    result <- filter(result, D_year == "2023") #change the year!
  }
  
  result %>%
    summarise(distinct_count = n_distinct(D_stu_id)) %>%
    pull(distinct_count)
}

# Calculate distinct counts for different datasets and years
stu_count_stu_data <- calculate_distinct_counts(student_data_validated)
stu_count_transcript_data <- calculate_distinct_counts(transcript_data_validated)
stu_count_stu_data_2023 <- calculate_distinct_counts(student_data_validated)
stu_count_transcript_data_2023 <- calculate_distinct_counts(transcript_data_validated)

# Create a comparison dataframe
compare_student_counts <- data.frame(
  Distinct_IDs = c("All Years"),
  Student_data = c(stu_count_stu_data, stu_count_stu_data_2023),
  Transcript_data = c(stu_count_transcript_data, stu_count_transcript_data_2023)
)

print(compare_student_counts)
 
```

```{r}
#NEW FUNCTION To REPLACE ABOVE POSSIBLY
# Function: takes a dataset + label and returns distinct counts by year + overall
calculate_dynamic_counts <- function(data, label) {
  # Get all unique years
  years <- data %>% 
    distinct(D_year) %>% 
    pull(D_year)
  
  # Calculate per-year counts
  per_year_counts <- map_dfr(years, function(y) {
    data %>%
      filter(D_year == y) %>%
      summarise(
        Source = label,
        Year = as.character(y),
        Distinct_Student_Count = n_distinct(D_stu_id)
      )
  })

  # Add all-years summary
  all_years_count <- tibble(
    Source = label,
    Year = "All",
    Distinct_Student_Count = n_distinct(data$D_stu_id)
  )

  bind_rows(per_year_counts, all_years_count)
}


# Named list of datasets and their labels
datasets <- list(
  "Student Data" = student_data_validated,
  "Transcript Data" = transcript_data_validated
)

# Apply function to all datasets
compare_student_counts <- imap_dfr(datasets, ~ calculate_dynamic_counts(.x, .y)) %>%
   pivot_wider(
    names_from = Year,
    values_from = Distinct_Student_Count
  )
 
```

#### b. Filter to schools of interest and return student counts

#### JK Comments

*I wonder if we can apply similar logic below that we did above... Is there a way to make the location, year and grade dynamic so that future teams don't have to update and/or for each of the different chunks manually update them? Is the idea just to get a count per students by school? Can we adapt the above code?*

**What:** Filtering to schools of interest and comparing student counts across files to check that they are comparable

**Check:** Are the student counts similar?

**Project Team Action:** Gut check/track, update schools of interest to specific district schools (ie. only schools included in analysis)

```{r}
# Table of values of D_location_id and counts in student_data
student_location_counts <- student_data_validated %>%
  filter(D_location_id %in% c(1, 4, 29, 39)) %>% #update to your schools of interest
  group_by(D_location_id, D_location_name) %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()

#calculate the distinct count of D_stu_id for transcript file
stu_count_transcript_data_filtered <- transcript_data_validated %>% 
  filter(D_location_name %in% c("MANOR HIGH SCHOOL", "MANOR NEW TECH HS", "MANOR EXCEL ACADEMY")) %>% #update to your schools of interest
  summarise(distinct_count = n_distinct(D_stu_id)) %>% 
  print()

#counts for SY 2023- student
stu_count_stu_data_2023_filtered <- student_data_validated %>% 
  filter(D_year == 2023, D_location_id %in% c(1,4,29,39)) %>% #update to your schools of interest
  summarise(distinct_count = n_distinct(D_stu_id)) %>% 
  print()

#counts for SY 2023- transcript
stu_count_transcript_data_2023_filtered <- transcript_data_validated %>% 
  filter(D_year == 2023, D_location_name %in% c("MANOR HIGH SCHOOL", "MANOR NEW TECHNOLOGY HS", "MANOR EXCEL ACADEMY") ) %>% #update to your schools of interest
  summarise(distinct_count = n_distinct(D_stu_id)) %>% 
  print()

#Count the number of distinct student IDs by grade and year
grade_counts <- student_data_validated %>%
  filter(D_stu_grade %in% c(9, 10, 11, 12)) %>%
  group_by(D_stu_grade, D_year) %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()

```

#### c. Return student ids that don't match between student and transcript files

-   **What:** This is an important step to return student ids that don't match between files

-   **Check:** Count of unique student ids should match your counts from step 2.1a (see your global environment under values for these new counts and compare). After returning the differing sets of student ids, check your work by choosing a few student ids to search for in each file. For example, if XXX is returned as a student id existing in transcript but not student, search for that in in both to confirm

-   **Project Team Action:** Gut check/track. Discuss whether it makes sense to exclude these students with your team

```{r}
#get the unique D_stu_id values in student file
stu_ids_stu_data <- unique(student_data_validated$D_stu_id) 

#get the unique D_stu_id values in transcript file
stu_ids_transcript_data <- unique(transcript_data_validated$D_stu_id)

#get the ids that are different
#this line of code returns student ids that exist in student data but not transcript data
stu_ids_in_stu_not_transcript <- setdiff(stu_ids_stu_data, stu_ids_transcript_data) 

#this line of code returns student ids that exist in transcript data but not student data
stu_ids_in_transcript_not_stu <- setdiff(stu_ids_transcript_data, stu_ids_stu_data)

#SE Adding this in to get a simpler view of number of students
# Get the IDs that are the same in both files
stu_ids_in_both <- intersect(stu_ids_stu_data, stu_ids_transcript_data)

# Count the number of distinct IDs that are the same between the two files
count_same_ids <- length(stu_ids_in_both)

# Count the number of distinct IDs that are different between the two files
count_different_ids <- length(unique(c(stu_ids_in_stu_not_transcript, stu_ids_in_transcript_not_stu)))

# Print the counts
cat("Number of distinct IDs that are the same between the two files:", count_same_ids, "\n")
cat("Number of distinct IDs that are different between the two files:", count_different_ids, "\n")


```

#### d. Filter to rows of student ids that exist in transcript file but not student file, and vice versa

-   **What:** Filters to rows of student ids that exist in transcript file but not student file, and vice versa

-   **Check:** Check that distinct counts in new filtered sets of student ids match distinct counts from step 2.1c

-   **Project Team Action:** Gut check/track

```{r}
#filter to student ids that exist in student file but not transcript file
filtered_stu_ids_stu_data <- student_data_validated[student_data_validated$D_stu_id %in% stu_ids_in_stu_not_transcript, ]

#filter to student ids that exist in transcript file but not student file
filtered_stu_ids_trancript_data <- transcript_data_validated[transcript_data_validated$D_stu_id %in% stu_ids_in_transcript_not_stu, ]

#check that distinct count in filtered_diff_stu_ids data frames matches distinct count in diff_stu_ids
count_filtered_stu_ids_stu_data <- filtered_stu_ids_stu_data %>%
  distinct(D_stu_id)

#check that distinct count in filtered_diff_stu_ids2 data frames matches distinct count in diff_stu_ids2
count_filtered_stu_ids_trancript_data <- filtered_stu_ids_trancript_data %>%
  distinct(D_stu_id)

# create a table with the counts of year in filtered_stu_ids_transcript_data
year_counts_filtered_stu_ids_transcript_data <- filtered_stu_ids_trancript_data %>%
  group_by(D_year) %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()

```

#### e. Explore exit codes and enter/exit dates of mismatched student ids

#### *JK Comments:*

*is there relevant tettra pages that would be helpful to link to here? What other exclusions should a team consider? Also, should there be some sort of code for the exit/enter codes section?*

-   **What:** To determine if these mismatched student ids should be excluded and removed, explore the students' enter and exit dates and exit code descriptions (if you have them) in your filtered student data. In Lubbock, for example, these appeared to be students who left the district before they could get any grades. We removed these students (see next steps)

-   **Check:** You are checking to see if these mismatched student ids are students who left the district to go to another educational institution before they could receive any grades

-   **Project Team Action:** Discuss whether it makes sense to exclude these students with your team

#### f. Add exclude flag to the student ids that don't match & remove students

-   **What:** Add exclude flag to students who don't match between files and remove them

-   **Check:** After removing the students, check that the number of rows is equal to the original number of rows minus the number of rows in your filtered data (when you filtered to the mismatched student ids). For example, if you had 2,000 rows of students who needed to be excluded from student file, and 60,000 rows, you should have 58,000 rows after removing those students from the file

-   **Project Team Action:** None (run as is)

##### i. Student File

```{r}
#add column with exclude flag
 student_data_validated <- student_data_validated %>%
   mutate(C_stu_exclude_flag = ifelse(D_stu_id %in% filtered_stu_ids_stu_data$D_stu_id, "exclude", "include"))
 
# #check count with exclude flag to make sure it matches # rows in filtered student data set
 exclude_count_stu_data <- sum(student_data_validated$C_stu_exclude_flag == "exclude") %>% 
   print()

#check distinct count with exclude flag to make sure it matches
 distinct_exclude_count_stu_data <- student_data_validated %>%
   filter(C_stu_exclude_flag == "exclude") %>%
   summarise(distinct_count = n_distinct(D_stu_id)) %>%
   print()


 #check a student id that was a mismatch and should have exclude flag
 stu_id_dive <- student_data_validated %>%
   filter(D_stu_id == "XXX") %>% #update with student ID from your file
   print()

#removing excluded students from data
student_data_validated <- student_data_validated %>% 
   filter(C_stu_exclude_flag == "include")
 
# #remove C_stu_exclude_flag column since you don't need it anymore
student_data_validated <- select(student_data_validated, -C_stu_exclude_flag)

#so from this point onwards, student data validated should ONLY have data that includes students that we have transcript data for. Does not imply a 1:1 year match, but does imply that we have transcript data for them
```

##### ii. Transcript File

```{r}
#add column with exclude flag
 transcript_data_validated <- transcript_data_validated %>%
   mutate(C_stu_exclude_flag = ifelse(D_stu_id %in% filtered_stu_ids_trancript_data$D_stu_id, "exclude", "include"))
 
#check count with exclude flag to make sure it matches
 exclude_count_transcript_data <- sum(transcript_data_validated$C_stu_exclude_flag == "exclude") %>% 
   print()

#check distinct count with exclude flag to make sure it matches
 distinct_exclude_count_transcript_data <- transcript_data_validated %>%
   filter(C_stu_exclude_flag == "exclude") %>%
   summarise(distinct_count = n_distinct(D_stu_id)) %>%
   print()
 
# #check a student id that was a mismatch and should have exclude flag
 stu_id_dive2 <- transcript_data_validated %>%
   filter(D_stu_id == "XXX") %>% #update with student ID from your file
   print()
 
#removing excluded students from data
 transcript_data_validated <- transcript_data_validated %>% 
   filter(C_stu_exclude_flag == "include")
#remove C_stu_exclude_flag column since you don't need it anymore
 transcript_data_validated <- select(transcript_data_validated, -C_stu_exclude_flag)


```

#### g. Check to make sure all student ids match now

-   **What:** Check that count of distinct student ids in student and transcript file match

-   **Check:** Check counts on unique student ids in each file type by looking in your environment under "values". There should not be any ids that are different between files now

-   **Project Team Action:** None (run as is)

```{r}
#get the unique D_stu_id values in student file
 stu_ids_stu_data2 <- unique(student_data_validated$D_stu_id) 
 
#get the unique D_stu_id values in transcript file
 stu_ids_transcript_data2 <- unique(transcript_data_validated$D_stu_id)

#get the ids that are different
 check_diff_stu_ids <- setdiff(stu_ids_stu_data2, stu_ids_transcript_data2 )
 check_diff_transcript_ids <- setdiff(stu_ids_transcript_data2, stu_ids_stu_data2)
 
# Count the number of distinct IDs that are different between the two files
count_different_ids <- length(unique(c(check_diff_stu_ids, check_diff_transcript_ids)))

# Print the counts
cat("Number of distinct IDs that are different between the two files:", count_different_ids, "\n")
 
```

### 2.2 Compare counts of locations across files

#### a. Return counts of schools across files

-   **What:** Return counts of schools across files

-   **Check:** Compare counts across files, next step will further investigate

-   **Project Team Action:** None (run as is)

```{r}
 #student file
 school_count_stu <- student_data_validated %>%
   summarise(student_count = n_distinct(D_location_id)) %>% 
   print()
 
 #transcript file
 school_count_transcript <- transcript_data_validated %>%
   summarise(transcript_count = n_distinct(D_location_id)) %>% 
   print()
 
```

#### b. Manual check: investigate location ids in each file to make sure you have all high schools, AND that location ids match and location names match

-   **What:** Review the location ids in each file by viewing the data frames in new tabs

-   **Check:** Check that you have the location ids you need in each file, and that the location names are consistent

-   **TO RETURN TO:** Most districts won't have too many high schools - I think it's easier to just look in each file to see if you have the school codes you need, and that they match. Or you could also do a filter to high schools and then print the unique location codes in each dataframe (maybe quicker than the manual process)

-   **Project Team Action:** Update Data Validations Table. If the school codes don't match, discuss with your team and reach out to your point person in the district

#### c. Mismatched location ids

#### JK Comments:

I *think that a nice to have here would be would be "start with exact matches then move to partial matches"..since we all know that school matching can be such a headache... OR if not, maybe its just a chunk of code that says "fix school names/ids to match" (or something like that... let me know if this doesn't make sense!). Maybe Chat GPT has some suggestions around how we could do partial matches in R. to get a first pass at this.. also, per the point above, i realize a lot of these districts have very few high schools so probably not necessary to do a whole thing! if not we can table... I think this is on the list of things for DSI team to try and figure out :)*

-   **What:** Return location ids that do not match between transcript and student files

-   **Check:** Investigate \~3 student ids that attended a location that is not in the student file. Search for that student in the transcript file and review their credit/grade history. Then conduct the deep dive codes to investigate their enter/exit information into the district in the student file. It is likely okay to keep these students - they won't be a part of your cohort analysis, but will be a part of your retention/transiency analysis and snapshot. You are mostly trying to get a better understanding of the sources of the mismatched location ids

-   **Project Team Action:** Decide how to handle students who were enrolled in other location ids outside of the district, transferred in for a period of time, and then left

```{r}
 #get unique D_location_id values from each dataset
 transcript_location_ids <- unique(transcript_data_validated$D_location_id)
 student_location_ids <- unique(student_data_validated$D_location_id)
 
 #find the D_location_id values that are in transcript_location_ids but not in student_location_ids
 location_ids_not_in_student_data <- setdiff(transcript_location_ids, student_location_ids) %>% 
   print()
 
# #student deep dive
 student_deep_dive <- student_data_validated %>% 
   filter(D_stu_id == XXX) %>%  #update with student ID from your file
   print()
 
# #student deep dive2
 student_deep_dive2 <- student_data_validated %>% 
   filter(D_stu_id == XXX) %>%  #update with student ID from your file
   print()
 
# #student deep dive3
 student_deep_dive3 <- student_data_validated %>% 
   filter(D_stu_id == XXX) %>% #update with student ID from your file
   print()
```

#### d. Add school exclude flag

**What:** There are two code chunks here:

-   The first one creates a table that allows you to explore grade/school configurations. Or how many students are enrolled at each school per grade.

-   The second code chunk creates a "exclude/include" flag based on the list of schools below. The schools in that list will get a "exclude" flag. Schools that usually considered for exclusion are Alternative, special ed or schools with specialized programming where the courses and information may not be comparable or reliable. This should be discussed with your project team ahead of time.

-   **Check:** Review the list of schools by their flag

-   **Project Team Action:**

    1.  First, explore the different schools/grade configurations - this might provide some insight into alt/special ed schools and which ones to add to the exclude list

    2.  Then update the second code chunk **schools_to_drop** section one with the ids of the schools in your district that you wish to exclude

*Explore Schools Code Chunk*

```{r}
#review count of students by school across grades
school_enroll_by_grade <- transcript_data_validated %>%
  drop_na(D_stu_grade) %>%
  group_by(D_location_name, D_stu_grade) %>%
  summarise(Unique_Count = n_distinct(D_stu_id)) %>%
  pivot_wider(names_from = D_stu_grade, 
              values_from = Unique_Count, 
              values_fill = 0) %>%
  mutate(total_enrollment = rowSums(across(where(is.integer)), na.rm = TRUE),
  across(where(is.integer), ~ round(. / total_enrollment, digits=2), .names = "{.col}_pct")) %>%
   View()
```

*Create Exclude School Flag Column and Exclude Schools*

##### i. Transcript File

```{r}
#set exclusion critera
exclude_condition_transcript <- transcript_data_validated$D_location_id %in% c(0, 714, 716, 718, 721, 1010, 1714, 1718, 1721, 999999) #update to your schools of interest

#add C_school_exclude_flag based on the condition
transcript_data_validated <- transcript_data_validated %>%
  mutate(C_school_exclude_flag = if_else(exclude_condition_transcript, "exclude", "include"))

#check list of excluded schools to make sure the code worked
exclude_location_list_transcript <- transcript_data_validated %>%
  filter(C_school_exclude_flag == "exclude") %>%
  distinct(D_location_id) %>% 
  print()

#check list of included schools to make sure the code worked
include_location_list_transcript <- transcript_data_validated %>%
  filter(C_school_exclude_flag == "include") %>%
  distinct(D_location_id) %>% 
   print()
```

##### ii. Student File

```{r}
#set exclusion critera
exclude_condition_student <- student_data_validated$D_location_id %in% c(0, 714, 716, 718, 721, 1010, 1714, 1718, 1721, 999999) #update to your schools of interest

#add C_school_exclude_flag based on the condition
student_data_validated <- student_data_validated %>%
  mutate(C_school_exclude_flag = if_else(exclude_condition_student, "exclude", "include"))

#check list of excluded schools to make sure the code worked
exclude_location_list_student <- student_data_validated %>%
  filter(C_school_exclude_flag == "exclude") %>%
  distinct(D_location_id) %>% 
  print()

#check list of included schools to make sure the code worked
include_location_list_student <- student_data_validated %>%
  filter(C_school_exclude_flag == "include") %>%
  distinct(D_location_id) %>% 
  print()
```

### 2.3 Course ID Counts

#### a. Return distinct course id counts

-   **What:** Return distinct course id counts in transcript, teacher, and course catalog files

-   **Check:** Compare across files - are they similar? If not, next steps will help you investigate

-   **Project Team Action:** Gut check/track

#### JK Comments:

*there was no D_course_key in the transcript file i was working with - I think this is probably a product of me not being consistent with what files I am using, etc. But wantedt o flag in case it makes sense to update to a critical column that would be in all transcript files after validation*

```{r}
#calculate the distinct count of D_course_id for transcript file
 course_id_count_transcript_data <- transcript_data_validated %>% 
   summarise(distinct_count = n_distinct(D_course_key)) %>% 
   print()

 
#calculate the distinct count of D_course_id for course catalog file
 course_id_count_cc_data <- course_catalog_validated %>% 
   summarise(distinct_count = n_distinct(D_course_id)) %>% 
   print()
```

#### b. Mistmatched course ids

#### JK Comments:

*I wonder if it makes sense to clean up capitilizations, etc so you can get rid of some of the "noise" in the mismatches since you are joining on D_Course Name.. this would look like:* D_course_name = str_to_lower(str_trim(D_course_name))

-   **What:** Return course ids that don't match between transcript, and course catalog files in schools of interest

-   **Check:** Count of unique course ids should match your counts from step 2.3a (see your global environment under values for these new counts and compare). After returning the differing sets of course ids, check your work by choosing a few course ids to search for in each file. For example, if XXX is returned as a course id existing in transcript but not student, search for that in in both to confirm

-   **Project Team Action:** Gut check/track

```{r}
#first, get the unique D_course_id values in course catalog file
 course_ids_cc_data <- unique(course_catalog_validated$D_course_id)

transcript_data_validated <- transcript_data_validated %>%
  mutate(`course_in_catalog?` = ifelse(D_course_key %in% course_ids_cc_data, "yes", "no"))

#count distinct course Ids where the course is not in the course catalog
transcript_data_validated %>%
  filter(`course_in_catalog?` == "no") %>%
  summarise(distinct_count = n_distinct(D_course_key)) %>%
  print()

#another layer of validation that looks at course names to see how many names match between the transcript and course catalog
#first, get the unique D_course_id values in course catalog file
 course_name_cc_data <- unique(course_catalog_validated$D_course_name)

transcript_data_validated <- transcript_data_validated %>%
  mutate(`course_name_in_catalog?` = ifelse(D_course_name %in% course_name_cc_data, "yes", "no"))

#count distinct course Ids where the course is not in the course catalog
transcript_data_validated %>%
  filter(`course_name_in_catalog?` == "no") %>%
  summarise(distinct_count = n_distinct(D_course_name)) %>%
  print()

#and now look at the two columns together to look for no, no's, or courses that don't match either by ID or name
transcript_data_validated %>%
  group_by(`course_in_catalog?`, `course_name_in_catalog?`) %>%
  summarise(count = n_distinct(D_course_key))

#pause here to review missing courses and manually check

#Lets filter to just the no, no's to see what we are looking at
missing_courses <- transcript_data_validated %>%
  filter(`course_in_catalog?` == "no", `course_name_in_catalog?` == "no") %>%
  select(D_course_key, D_course_name) %>%
  distinct() %>%
  print()

#There might be small naming differences in both the course key and the course name, so check that the titles look similar enough. Option for correciton here- write a .csv of these no's to Fuzzy lookup into the course catalog in excel, and then bring back into R. Follow the guideance listed here in Tettra, and uncomment the code below to export and run- https://app.tettra.co/teams/ersknowledge/pages/matching-names-in-excel-with-fuzzy-lookup

#current_date <- format(Sys.Date(), "%m.%d.%y") # will update with the date it was last run, thats just what I did!

#print(full_file_path)

#Export Missing Courses to CSV
#write.csv(missing_courses, file = paste0(full_file_path, "missing_courses", current_date, ".csv"), row.names = FALSE)

```

### 2.4 Student Grade Level

#### a. More than one grade level

-   **What:** Return student ids with more than one grade level in the same year

-   **Check:** Check so that we can work to develop a source of truth for student grade level

-   **Project Team Action:** None (run as is)

```{r}
#filter for rows where there are multiple distinct D_stu_grade values within the same D_year within the student data frame
multiple_grades <- student_data_validated %>%
  filter(!is.na(D_year)) %>% # Filter out rows where D_year is NA
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade) > 1) %>%
  distinct()

#get unique D_stu_id values with more than one D_stu_grade value within the same D_year
unique_multiple_grades <- unique(multiple_grades$D_stu_id)

# create a new dataframe with just these students
multiple_grades_df <- student_data_validated %>%
  filter(D_stu_id %in% unique_multiple_grades) %>% 
  print()

#count distinct students in the multiple grades dataframe
count_distinct_multiple_grades <- multiple_grades_df %>%
  summarise(count = n_distinct(D_stu_id)) %>%
  print()
```

#### b. Investigate student ids with different grade levels

#### JK Comments:

*I wonder if it would be more streamlined to combine the above code block with the below code bock.. so that you are just looking at the first 3 students are something like that of the ones with more then one grade, so you are not having to manually investigate? Or if its more helpful to manually investigate one student at a time, and specific students, then ignore this suggestion below from Chat GPT :)*

```{r}
library(dplyr)

# Step 1: Identify students with multiple grades in the same year
multiple_grades_df <- student_data_validated %>%
  filter(!is.na(D_year)) %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade) > 1) %>%
  ungroup()

# Step 2: Get first 3 unique student IDs from this subset
sample_ids <- multiple_grades_df %>%
  distinct(D_stu_id) %>%
  slice_head(n = 3) %>%
  pull(D_stu_id)

# Step 3: Deep dive on those students
student_deep_dive <- student_data_validated %>%
  filter(D_stu_id %in% sample_ids) %>%
  arrange(D_stu_id, D_year)

# Optional: print for manual review
print(student_deep_dive)

```

-   **What:** Investigate student ids with different grade levels in the same year. The source of the mismatch can sometimes be when a student moves from one grade level to the next within the same school year, but in between semesters, and may be represented by a unique exit code.

-   **Check:** Use code below to check three student ids in both student and transcript file to see why grade levels might differ

-   **Project Team Action:** Decide on source of truth for student grade level(D_stu_id)

```{r}
#student deep dive 1
stu_grade_dive <- student_data_validated %>%
  filter(D_stu_id == "309094") %>% #update with student ID from your file
  arrange(D_year) %>% 
  print()

#student deep dive 2
stu_grade_dive2 <- student_data_validated %>%
  filter(D_stu_id == "309078") %>% #update with student ID from your file
  arrange(D_year) %>% 
  print()

#student deep dive 3
stu_grade_dive3 <- student_data_validated %>%
  filter(D_stu_id == "XXX") %>%  #update with student ID from your file
  arrange(D_year) %>% 
  print()

```

#### c. Pull grade level that students finished the year on

#### JK Comments

*C_summer_school and D_stu_enter/exit dates threw an error for me. Wanted to flag in case it might sense to update to a critical column*

*Also might want to add a note that is like "use this space to do additional data cleaning, such as removing summer school rows (or something)*

-   **Project:** To create one source of truth for student grade level, so we will pull the grade level that students finished the year on (max exit date each year)

-   **Check:** There are checks built in after each step including, checking that students don't have multiple rows with the same exit date by year

-   **Project Team Action:** None (run as is)

##### i. Summer School (if necessary)

-   **Comment Out if your district does not have summer school**

```{r}
# Write a new variable called C_summer_school2 that has a value of Y if any value of Summer school when grouped by student ID and year is Y 
# This variable gives students a yes flag if they ever went to summer school in that year
student_data_validated <- student_data_validated %>% 
  group_by(D_stu_id, D_year) %>% 
  mutate(C_summer_school2 = ifelse(any(D_summer_school == "Y"), "Y", "N")) %>% 
  ungroup() %>% 
  print()

# Remove rows for summer school because we do not need them and add a helper column with max and min exit date
student_data_validated <- student_data_validated %>%
 # filter(D_summer_school != "Y") %>% # Filter out rows where D_summer_school = #Y
  group_by(D_stu_id, D_year) %>%
  mutate(H_max_exit_date = max(D_stu_exit_date), 
             H_min_enter_date = min(D_stu_enter_date)) %>%
  ungroup() %>%
  print()

# check for how many IDs still have multiple rows for the same student ID and year with different values of grade after removing summer school
multiple_grades <- student_data_validated %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade) > 1) %>%
  distinct

#get unique D_stu_id values with more than one D_stu_grade value within the same D_year
unique_multiple_grades <- unique(multiple_grades$D_stu_id)

# create a new dataframe with just these students to take a look
multiple_grades_df <- student_data_validated %>%
  filter(D_stu_id %in% unique_multiple_grades) %>% 
print()

```

##### ii. Create EOY Grade Level Flag

```{r}
# Create a new variable called C_eoy_grade that groups by student ID and year and selects the grade level that the student spent the most time in (aka D_stu_grade for the student with the highest value of M_days)
student_one_grade <- student_data_validated %>%
  group_by(`D_stu_id`, `D_year`) %>%
  slice(which.max(D_days_helper)) %>%
  ungroup() %>%
  mutate(C_eoy_grade = D_stu_grade) %>%
  select(D_stu_id, D_year, C_eoy_grade)

# check for how many IDs still have multiple rows for the same student ID and year with different values of grade after removing summer school

# should be zero
multiple_grades <- student_one_grade %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(C_eoy_grade) > 1) %>%
  distinct
print(multiple_grades)

# Now do a left join with student_data_validated by D_stu_id and D_year
student_data_validated <- student_data_validated %>%
  left_join(student_one_grade, by = c("D_stu_id", "D_year")) %>%
  print()

#Extra check here of checking the same students as before to make sure they only have one consistent grade level for EOY grade, and they do! 
stu_grade_dive <- student_data_validated %>%
  filter(D_stu_id == XXX) %>% #update with student ID from your file, should be the same student ID with mismatched grade level as before, to ensure that join worked correctly
  arrange(D_year) %>% 
  print()

#student deep dive 2
stu_grade_dive2 <- student_data_validated %>%
  filter(D_stu_id == XXX) %>% #update with student ID from your file, should be the same student ID with mismatched grade level as before, to ensure that join worked correctly
  arrange(D_year) %>% 
  print()

#student deep dive 3
stu_grade_dive3 <- student_data_validated %>%
  filter(D_stu_id == XXX) %>% #update with student ID from your file, should be the same student ID with mismatched grade level as before, to ensure that join worked correctly
  arrange(D_year) %>% 
  print()

```

#### d. Check grade level distribution of C_eoy_grade

```{r}
# Create a count of unique student IDs by year by C_eoy_grade
grade_counts_eoy <- student_data_validated %>% 
  filter(C_eoy_grade %in% c(9, 10, 11, 12)) %>% 
  group_by(D_year, C_eoy_grade) %>% 
  summarize(n = n()) %>% 
  arrange(D_year, C_eoy_grade) %>%
  print()

# Now run this same table, but just for our schools of interest 
grade_counts_filtered <- student_data_validated %>% 
  filter(D_location_id %in% c(714, 716, 718, 721)) %>% #updated to your schools of interest (as necessary)
  group_by(D_year, C_eoy_grade) %>% 
  summarize(n = n()) %>% 
  arrange(D_year, C_eoy_grade) %>%
  print()

```

## Part 3: Upload File to Sharepoint

### 3.1 Set output data folder Path

**ACTION: Change the processed_data_folder_path to the path to your data. The folder path must start with a folder that is just below the "Client Work" level. For example, it should start with "District Partners", "Internal", "State Partners", etc..**

```{r}
processed_data_folder_path <- "District Partners/Yakima"
```

### 3.2 See guidance on using the "ers_write_sharepoint" function

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_write_sharepoint
```

### 3.3 Write files

**ACTION: Replace "raw_data" with your desired name of your data and replace "XXX.csv" with the desired name of your output file. You can output CSV or XLSX files**

```{r}
#course catalog
ers_write_sharepoint(data = course_catalog_validated,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "XXX.csv")

#transcript
ers_write_sharepoint(data = transcript_data_validated,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "XXX.csv")

#student
ers_write_sharepoint(data = student_data_validated,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "XXX.csv")
```

## Part 4: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 5: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
