---
title: "3_transcript_rollup"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: "July 31 2024"
review:
last update: "March 12 2025"
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### JK comments:

-   *Can you please replace - through 1.4 this with the newest template format :) and also the last part on export*

## Part 0: Read Libraries and connect to Sharepoint

### 0.1 (OPTIONAL) Install the latest version of the erstools library!

**ACTION: If you aren't sure if you have the most up to date release of erstools, follow these instructions: <https://app.tettra.co/teams/ersknowledge/pages/updating-erstools-functions-markdown-template>**

### 0.2 Read All Libraries

**ACTION: Run the code and read the comments for different functions you will have access to**

```{r, warning = F, message = F}
#ERS-specific functions!!
library(erstools) #includes functions like ers_read_sharepoint, ers_write_sharepoint, ers_table, and ers_column_info

#load data
library(Microsoft365R) # Use Microsoft365R to connect to SharePoint and load files
library(readxl) # Functions for reading Excel files: read_excel
library(writexl) # Functions for writing Excel files: write_xlsx
library(openxlsx) # Functions for writing Excel files: write.xlsx

#manipulate data
library(tidyverse) # Includes ggplot2, dplyr, tidyr, readr, etc.
library(data.table) # fread(): load large CSVs quickly
library(janitor) # Functions for cleaning and examining data: clean_names() instantly formats all column names
library(safejoin) # Provides safer join operations for 1:many matching - safe_left_join()
library(stringr) # str_detect(): pattern matching using regular expressions (see regex101.com)

#format and display data
library(scales) # Functions for scaling and formatting axes in plots: comma(), percent()
library(gt) # Creates and formats tables in a publication-ready format: gt()
library(flextable) # Creates and formats tables in a publication-ready format: flextable()
library(extrafont) # Manages and uses custom fonts in plots - loadfonts(): Loads fonts available on the system

#share output with supervisor
library(rmarkdown) # needed to create knitted files in output folder
library(downloadthis) # create download buttons in your knitted file so that a supervisor can download data in Excel!
```

## Part 1: Bring in your data

### 1.0 Authenticate to sharepoint to be able to read in files

**ACTION: After running the code, you may need to look at your console code window below for instructions to log in. A window may open in your browser to ask you to log into Microsoft to confirm your identity**

**NOTE: If you ever have authentication issues, first run this in your console then try running the code chunk below again - AzureAuth::clean_token_directory()**

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

### 1.1 Set your folder path(s)

**ACTION: Change the raw_data_folder_path to the path to your data. The folder path must start with a folder that is below the "Client Work" level. For example, it should start with "District Partners", "Internal", "State Partners", etc.**

**NOTE: If you don't want to type the name of your folder path, you can copy it directly from Sharepoint: <https://app.tettra.co/teams/ersknowledge/pages/copying-a-folder-path-from-sharepoint-to-use-in-r>**

```{r}
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Bill%20&%20Melinda%20Gates%20Foundation/Gates%20HS%20Transformation/HS%20Diagnostics/Lubbock%20ISD/02.%20Transcript%20Analysis/Lubbock%20Data/Lubbock%20Data%20-%20Final%20(November%202023)/"
```

### 1.2 See guidance on using the "ers_read_sharepoint" function

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_read_sharepoint
```

### 1.3 Read in your data

**ACTION: Replace "XXX.csv" with the file name. We suggest doing this for all instances of raw data by using the find and replace tool (Ctrl + F and then replace all instances of "raw_data" with the desired name.**

```{r}
#course catalog
course_catalog_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "2023 Lubbock Course Catalog.xlsx")
#transcript
transcript_data_validated <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = "Transcript Rollup 12.19.23.xlsx", sheet_name = "Transcript Rollup")
```

### 1.4 View your data

**ACTION: Replace "raw_data" with the same name of your data frame you used above, and run the code to inspect your data.**

```{r, eval = interactive()}
View(transcript_data_validated)
View(course_catalog_validated)
```

## Part 2: Validate Data

### 2.0 Check out the ers_column_info function as a starting place in your data validation (credits to Mauro!)

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_column_info
```

### 2.1 Run ers_column_info on your data

**ACTION: Replace "raw_data" with the name of your data frame**

```{r}
ers_column_info(transcript_data_validated) 
ers_column_info(course_catalog_validated)
```

# Part III: Roll Up

### 3.1 Add new variables to the transcript data

#### a. C_grade_pct_rollup

**What:** Add new variables that will be necessary for later analysis. In this case, we are creating a new variable that will roll up the final grade into grade buckets.

**check:** the code block should print without error

**Project Team Action:** Adjusted grade buckets based on district guidance/team guidance

#### JK COmments

*Using Lubbok's data here which did not contain D_final_grade.......is this a critical step/column we would expect going forward?*

```{r}
## Create a new variable called C_grade_pct_rollup
transcript_data <- transcript_data_validated %>% 
  mutate(C_grade_pct_rollup = case_when(
    D_final_grade == 0 ~ "0", 
    D_final_grade >= 90 ~ "90+",
    D_final_grade >= 80 ~ "80-89",
    D_final_grade >= 70 ~ "70-79",
    D_final_grade >= 60 ~ "60-69",
    TRUE ~ "Not a grade"
  ))
```

### 3.2 Remove columns that are not needed

#### JK Comments:

*The "why" on this doesn't feel as clear as some of the other steps!* *Also, the course catelog file (from Lubbok) did not have D_Course_Credit.....,*

**What:** Removes columns from the course catalog that are not needed for the roll-up

**check:** the code block should print without error

```{r}
course_catalog_validated <- course_catalog_validated %>% 
  select(-c(
            D_course_type,
            D_course_credit #this because we WANT to force transcript course credit
            ))
```

### 3.3 Validation Checks

**What:** Run checks on the course catalog to make sure that the data is clean and that the course IDs are unique.

**check:** the code block should print without error

**Project Team Action:** Gut check with team

#### JK COmments:

*This is one of those sections that have both print and assign objects - simon use your discretion in terms of whether you think it would be helpful to keep those assigned objects or not and same with print! You are right that some of them didn't automatically print for me.. I think it was the code with "length" and not summarize. Chat GPT said we could do something like this instead of lenght:*

```{r}
tibble(D_course_id = intersect(course_catalog_validated$D_course_id, transcript_data_validated$D_course_id)) %>%
  summarize(n_courses_in_both = n_distinct(D_course_id))
```

*\
*

```{r}
#how many unique course IDs in the catalog
course_catalog_validated %>% 
  summarize(count_courses = n_distinct(D_course_id)) %>% 
  print()

#how many unique course IDs in the transcript_data
transcript_data_validated %>% 
  summarize(count_courses = n_distinct(D_course_id)) %>% 
  print()

#And how many match between the two?
course_ids_in_both <- intersect(course_catalog_validated$D_course_id, transcript_data_validated$D_course_id) %>% 
  length()  #%>% 
  #print()

#and how many are in the transcript data but not in the course catalog?
course_ids_not_in_catalog <- setdiff(transcript_data_validated$D_course_id, course_catalog_validated$D_course_id) %>% 
  length() %>% 
  print()

#Should be zero, but if not, add additional code and investigation to identify which courses need to be added or do not exist between the two files
transcript_data_validated %>%
  filter(!(D_course_id %in% course_catalog_validated$D_course_id)) %>%
  distinct(D_course_id) %>%
  arrange(D_course_id)




```

### 3.4 Join data frames

#### JK Comments:

-   *Note I replaced a regular left join with safe_left_join as it will give you warnings if there are duplicates and mismatches, etc.*

-   *Got an error on the second part of the script because course_catalog_roll_up has not been created yet. I duplicated the course_catalog_validated file so I could keep going, but wanted to flag in case that logic needs updating*

-   *The last part about removing columns did not work for the file I was using. So maybe we should make this a Project Team Action: "Update as neccessary" or something like that?*

**What:** Using leftjoin, join the course catalog and course coding from excel with the transcript data to create one source of truth for student grade/credit/and course taking information

**check:** The code block should print without error

**Note:** Based on your data structure, you may see a warning that says "warning: one to many relationship". You should check this, as it would mean you have a mismatch in course ID between the transcript and course catalog files

```{r}

#Check if there are any duplicate IDs in the course catalog unique 
course_catalog_validated %>% 
  count(D_course_id) %>% 
  filter(n > 1) %>% 
  print()

#JK ADDED
course_catalog_roll_up <- course_catalog_validated
 


#merge into transcript file by course ID and course Name
transcript_roll_up <- transcript_data_validated %>% 
 safe_left_join(course_catalog_roll_up, by = c("D_course_id" = "D_course_id")) %>% 
  print()

#count course ids in rollup file
transcript_roll_up %>% 
  summarize(count_courses = n_distinct(D_course_id)) %>% 
  print()

#rename fields, lose columns you no longer need
transcript_roll_up <- transcript_roll_up %>% 
  select(-c(
            `course_in_catalog?`,
            `course_name_in_catalog?`,
            `D_course_name.y`
            ))

```

### 3.5 Student Deep Dives

#### JK Comments:

*This is cool that you dont have to manually update the student id! However, it seems like you need to re-run that first line to get a different student each time? Whenever there is repeated code its always a sign that it might be a good candidate for a function. So I think you could either a) Make it more clear to the user that they have to re-run that first line of code before each student dive OR b) Create a function.. Chat GPT came up with this, which seems to work nicely! Also this avoids creating 5 different tables too L:\_*

```{r}
# ---- Function: Check Random Students ----
check_random_students <- function(data, student_id_col = "D_stu_id", n = 5, seed = NULL) {
  # Optional seed for reproducibility
  if (!is.null(seed)) set.seed(seed)
  
  # Sample n unique student IDs
  random_ids <- sample(unique(data[[student_id_col]]), n)

  # Filter and return rows for those students, grouped for easy viewing
  data %>%
    filter(.data[[student_id_col]] %in% random_ids) %>%
    arrange(.data[[student_id_col]]) %>%
    group_by(.data[[student_id_col]])
}


# Check 5 random students from transcript_roll_up
check_random_students(transcript_roll_up, n = 5, seed = 123)

```

**What:** Now that we have a rolled up transcript and course catalog, just hand check a few students to make sure that their enrollments make sense, and that the course coding seems to be copying over correctly

**check:** Gut check that everything has copied over as you would expect

**Project Team Action:** Gut check, track, follow up with any inconsistencies

```{r}
#check a few students
random_student_id <- sample(unique(transcript_roll_up$D_stu_id), 1)

#Check Student 1
student_deep_dive_1 <- transcript_roll_up %>% 
  filter(D_stu_id == random_student_id) %>% 
  print()

#Check Student 2
student_deep_dive_2 <- transcript_roll_up %>% 
  filter(D_stu_id == random_student_id) %>% 
  print()

#Check Student 3
student_deep_dive_3 <- transcript_roll_up %>% 
  filter(D_stu_id == random_student_id) %>% 
  print()

#Check Student 4
student_deep_dive_4 <- transcript_roll_up %>% 
  filter(D_stu_id == random_student_id) %>% 
  print()

#Check Student 5
student_deep_dive_5 <- transcript_roll_up %>% 
  filter(D_stu_id == random_student_id) %>% 
  print()

#and just quickly double check the distinct count of Student Ids that we have in the transcript file
transcript_roll_up %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  print()

```

## Part 4: Upload File to Sharepoint

### 4.1 Set output data folder Path

**ACTION: Change the processed_data_folder_path to the path to your data. The folder path must start with a folder that is just below the "Client Work" level. For example, it should start with "District Partners", "Internal", "State Partners", etc..**

```{r}
processed_data_folder_path <- "District Partners/Yakima"
```

### 4.2 See guidance on using the "ers_write_sharepoint" function

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_write_sharepoint
```

### 4.3 Write files

**ACTION: Replace "raw_data" with your desired name of your data and replace "XXX.csv" with the desired name of your output file. You can output CSV or XLSX files**

```{r}
#Example
ers_write_sharepoint(data = raw_data,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "XXX.csv")
```

## Part 5: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 6: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```
