---
title: "04_snapshot_analysis-MANORISD"
author: "Simon Ellis"
date: "06/07/24"
output:
  word_document: default
  html_document: default
---

# Part I: Setup

This part of the script allows you to:

-   Load the functions that R will be using to work with your data (Excel automatically loads all its functions - if, countif, index, match, etc. - when you open it, but R requires manual loading of functions if you want full access to what it offers)

-   Specify the Sharepoint folder where your raw data lives so R can pull it in

-   Actually load in the raw data you'll be working with throughout the rest of the script

## 1.1 Load Packages

-   **What:** This code block loads in all relevant function packages (a.k.a. R functions you'll be needing in this script). Note that if you haven't installed the packages into your computer yet, R will print an error and you'll need to install them first by running install.packages(tidyverse) and install.packages(formattable), etc.

-   **Check:** No error message coming out of code block

-   **Project Team Action:** Do not change (run as is)

```{r, warning = F, message = F}
library(tidyverse)
library(data.table)
library(dplyr)
library(formattable)
library(data.table)
library(here)
library(flextable)
library(openxlsx)
library(readxl)
library(lubridate)
library(safejoin)
library(ggplot2)
```

## 1.2 Declare Variables

-   **What:** This code block establishes a variable for code-users to define a consistent file path for loading data

-   **Check:** No error message coming out of code block

-   **Project Team Action:** Do not change (run as is)

### a. User Name

```{r, warning = F, message = F}
#action: add your user name in the file path. Add more user names for more team members
user_name_team_member1 <- "C:/Users/sellis"
user_name_team_member2 <- "XXX"
user_name_team_member3 <- "XXX"
```

### b. File Path

-   **What:** This code block creates a condensed file-path to the locally synced versions of your demographic data from sharepoint. You will have to edit to make it user/project specific.

-   **Check:** No error message coming out of code block

-   **Note:** File paths in R always use "/" instead of "", so you might need to manually change the direction of your slashes when you copy your folder path from Share point

-   **Project Team Action:** Change the code below based on where your file is located. More specifically - find your path to the data by navigating to your folder in file explorer and then copying the path at the top of your file explorer. This path might be different for each team member depending on which folders they have synced to their One Drive

```{r, warning = F, message = F}
folder_path_team_member1 <- "/Education Resource Strategies/Org-wide Files - TX Network - HS Diagnostic for CCMR Pathways - 2024/Manor ISD 2024/1. Data & Analysis - Secure/1.2 Data Validated in R"
folder_path_team_member2 <- "XXX"
folder_path_team_member3 <- "XXX"
```

### c. Full Folder Path

-   **What:** This code block combines the variables you created in the two previous steps to create one continuous file path to your data.

-   **Check:** No error message coming out of code block, file path should match folder destination locally

-   **Project Team Action:** Do not change (run as is)

```{r, warning = F, message = F}
#action: update the numbers at the end of the variables with your user number. E.g. if you are number 2, then change the number to 2 at the end of user_name_team_member and folder_path_team_member
user_name <- user_name_team_member1
folder_path <- folder_path_team_member1


#action: leave unchanged (this combines your values above into a single file path)

full_file_path <- paste0(user_name, folder_path)

#check
print(full_file_path)
```

### d. File Names

-   **What:** This code block creates condensed versions of files names for ease of upload and analysis. You should update the code based on the specific names of your demographic files

-   **Check:** No error message coming out of code block

-   **Project Team Action:** Update with specific file names

-   **RT Methodological Note:** Lubbock sent three sets of student files, so need to upload all. This will not be the case for most districts

```{r, warning = F, message = F}
#update to your file name
#note that your data should be in CSV format and that your file name should always end with .csv

#pathways
pathways <- "/00_pathways_cluster_companion.xlsx"

#cluster_map
cluster_map <- "/00_pathways_cluster_companion.xlsx"
  
#Snapshot_transcript_rollup
merged_transcript_snapshot <- "/3_merged_transcript_snapshot_06.06.24.xlsx"
  
  
```

## 1.3 Load Data

**What:** Now that R knows how to access the folder path, it's time to load in your raw data

**Check:** The code block should print the names of all data files you need

**Project Team Action:** None

```{r}
#pathways
pathways <- read_xlsx(paste0(full_file_path, pathways), sheet = "Pathways", skip = 2)

#cluster_map
cluster_map <- read_xlsx(paste0(full_file_path, cluster_map), sheet = "Cluster Map")

#Snapshot_transcript_rollup
merged_transcript_snapshot <- read_xlsx(paste0(full_file_path, merged_transcript_snapshot))

#Check them and make sure they look good
```

#1.4- Pre-Analysis File Cleaning and Merging

**What:** Previously, joining in the pathways tables and cluster coding had been done in excel and added to the course catalog roll-up, but I am implementing R code to do it here instead.

**Check:**

**Project Team Action:**

```{r}
#So first we want to add the pathways information onto the snapshot_transcropt data frame using a left join by course Id to bring in the pathway fields

#First I need to rename the pathways data frame to match the naming of the transcript for course Id, and remove the columns I don't want
pathways <- pathways %>%
  rename(D_course_id = `Course ID:`)

print(pathways)

#and I am going to remove the names of the columns that I do not need
pathways <- pathways %>%
  select(-`Course Name (PDF)`, -`Course Name Match (Course Catalog)`, -`Is Pathway?`)

#add a multiple pathways flag that flags each course that has a "Y" value for more than one pathway column(this will help later)
# Create a new column to count the number of "Y" values in pathway columns
pathways <- pathways %>%
  mutate(
    pathway_count = rowSums(across(c(`Automotive`,`Animal Science`,`Applied Agriculture Engineering`,`Plant Science`,`Carpentry`,`Digital Communications`,`Design and Multi-Media Arts`,`Accounting and Financial Services`,`Early Learning`,`Healthcare Therapeutic`,`Culinary Arts`,`Web Development`,`Law Enforcement`,`Cosmetology`,`Advadnced Manufacturing and Machinery (Robotics)`,`Biomedical Science`,`Engineering`,`Cybersecurity`,	`Marketing & Sales`)) == "Y"),
    Multiple_pathways = case_when(
      pathway_count > 1 ~ "Y",
      TRUE ~ "N"
    )
  )

# Drop the temporary pathway_count column
pathways <- pathways %>%
  select(-pathway_count)

# Print the updated data frame
print(pathways)

#then I need to filter to JUST pathways courses in the transcript data-frame before I merge on the pathways data- want to keep all courses for now 06/07/24
#merged_transcript_snapshot <- merged_transcript_snapshot %>%
  #filter(D_pathways_flag == "Y")

#and before we join in lets add some checks that all the courses are present in the pathways
#unique ids in the transcript rollup
unique(merged_transcript_snapshot$D_course_id) -> unique_transcript_ids

#unique ids in the pathways
unique(pathways$D_course_id) -> unique_pathways_ids

#create a new data frame of all of the course Ids that are NOT in the pathways data frame
setdiff(unique_transcript_ids, unique_pathways_ids) -> missing_pathways_ids
print(missing_pathways_ids)


```

#a. preform the join

```{r}
#And then join it in!
left_join(merged_transcript_snapshot, pathways, by = "D_course_id") -> merged_transcript_snapshot_pathways

#don't worry it will be a many to many join
#the number of rows should NOT change
nrow(merged_transcript_snapshot)

#view the new data-frame to see that the merge worked correctly
print(merged_transcript_snapshot_pathways)


#something I am also noticing is that there are students who show identical course enrollment between two high schools, which show as unique rows. I think we should get rid of that and just use home school as the descriptor to get around that?

print(merged_transcript_snapshot_pathways)

```

# Part II: Pathway and Advandced Course Analysis

## 2.1 Create new data frame

### a. Pathways - 2023

```{r}
#time to make the table! SE need to go back in and re-name this
snapshot_pathways_table_2023 <- merged_transcript_snapshot_pathways %>% 
  select(D_stu_id, 
         D_course_id, 
         D_course_name.x,
         D_creds_earn,
         D_final_grade,
         C_grade_pct_rollup,
         D_pass_fail,
         D_stu_swd,
         D_course_rigor,
         D_dual_credit_flag,
         D_AP_flag,
         D_course_subj,
         D_pathways_flag,
         D_stu_race,
         D_stu_frl,
         C_home_school_code, 
         C_home_school_name, 
         C_eoy_grade,
         `Course Level`,
         Multiple_pathways,
         `Automotive`,
         `Animal Science`,
         `Applied Agriculture Engineering`,
         `Plant Science`,
         `Carpentry`,
         `Digital Communications`,
         `Design and Multi-Media Arts`,
         `Accounting and Financial Services`,
         `Early Learning`,
         `Healthcare Therapeutic`,
         `Culinary Arts`,
         `Web Development`,
         `Law Enforcement`,
         `Cosmetology`,
         `Advadnced Manufacturing and Machinery (Robotics)`,
         `Biomedical Science`,
         `Engineering`,
         `Cybersecurity`,
         `Marketing & Sales`
         )

#06/07 BEFORE pivoting longer and creating the data-frame that is more suited for pathways analysis, I want to create a distinct data frame that has the same information above, but actually only one row per pathways course to avoid some of the problems I am experiencing later on

base_analysis <- snapshot_pathways_table_2023

#So now this "base_analysis" table is what I will use to run the checks on things that are NOT pathways courses

#pivot longer
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  pivot_longer(cols = c(  
         `Automotive`,
         `Animal Science`,
         `Applied Agriculture Engineering`,
         `Plant Science`,
         `Carpentry`,
         `Digital Communications`,
         `Design and Multi-Media Arts`,
         `Accounting and Financial Services`,
         `Early Learning`,
         `Healthcare Therapeutic`,
         `Culinary Arts`,
         `Web Development`,
         `Law Enforcement`,
         `Cosmetology`,
         `Advadnced Manufacturing and Machinery (Robotics)`,
         `Biomedical Science`,
         `Engineering`,
         `Cybersecurity`,
         `Marketing & Sales`), 
               names_to = "C_pathway", 
               values_to = "C_pathway_Y_or_N") #Y flag represents courses that are pathway courses - so essentially a student seat in a pathway

#check that the distinct count of students has not changed-> this makes sense because now we are JUST looking at pathway takers
snapshot_pathways_table_2023 %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup() %>% 
  print()


#convert Y to 1 and N to 0 because they are easier to deal with
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  mutate(C_pathway_Y_or_N = ifelse(C_pathway_Y_or_N == "Y", "1", "0"))


#change data type
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  mutate(C_pathway_Y_or_N = as.numeric(C_pathway_Y_or_N))

#remove all Ns because it will make file easier to deal with and you already have total counts of students so you don't need the Ns
#06/07 Do not exclude for demo information
#snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>% 
  #filter(C_pathway_Y_or_N == "1")

#Do a quick check to make sure we have no totally duplicate rows
num_duplicates <- snapshot_pathways_table_2023 %>%
  duplicated() %>%
  sum()
cat("Number of duplicate rows:", num_duplicates, "\n")

#filter to just the duplicate rows to see whats going on
duplicates_to_check <-snapshot_pathways_table_2023 %>%
  filter(duplicated(.))
print(duplicates_to_check)

#okay lets remove the duplicated rows from the snapshot_pathways_table_2023 so that its only unique values
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  distinct()

#check that the distinct count of students has not changed
snapshot_pathways_table_2023 %>% 
  summarize(count_students = n_distinct(D_stu_id)) %>% 
  ungroup() %>% 
  print()

#great
```

#b. merge in cluster coding as last step before analysis!

```{r}
#left join to create a new column which is C_cluster
left_join(snapshot_pathways_table_2023, cluster_map, by = c("C_pathway" = "Pathway")) -> snapshot_pathways_table_2023

#the number of rows should not change
#If you get NAs make sure to check spelling in the cluster sheet matches the pathways sheet
#filter to NAs for cluster
unique_pathways <- snapshot_pathways_table_2023 %>%
  filter(is.na(Cluster)) %>%
  select(C_pathway) %>%
  distinct() %>% 
  print()
```

#c. pre-Analysis Clean Up (Just renaming a couple of things to be easier in the editing process)

```{r}
#rename the cluster column
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  rename(C_cluster = Cluster)

#rename the course level column
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  rename(C_pathway_course_level = `Course Level`)

```

#d. creating additional variables

```{r}
# Adding in new race variable to match what I added into the pathways table
#merged_snapshot <- merged_snapshot %>%
  #mutate(C_stu_race_rollup = case_when(
    #D_stu_race == "Black or African-American" ~ "Black",
    #D_stu_race == "Hispanic" ~ "Hispanic",
    #D_stu_race == "White" ~ "White",
    #TRUE ~ "Other"
  #))

# Create a new variable called C_stu_grade_buckets
# This should also ideally come earlier
#merged_snapshot <- merged_snapshot %>%
  #mutate(C_stu_grade_buckets = case_when(
    #D_stu_grade == "A+" | 
      #D_stu_grade == "A" | 
#       D_stu_grade == "A-" |
#       D_stu_grade == "A+1"| # This is some kind of error, but grouping here to fix it
#       D_stu_grade == "B+"| 
#       D_stu_grade == "B"|
#       D_stu_grade == "B-"~ "A-B", # Create A-B bucket
#     D_stu_grade == "C+" | 
#       D_stu_grade == "C" | 
#       D_stu_grade == "C-" | 
#       D_stu_grade == "D+"| 
#       D_stu_grade == "D"| 
#       D_stu_grade == "D-" ~ "C-D", # Create C-D bucket
#     D_stu_grade == "F" ~ "F", # Create F bucket
#     D_stu_grade == "R" |
#       D_stu_grade == "r"|
#       D_stu_grade == "FR" ~ "R", # Create R bucket
#     TRUE ~ "Other"
#   ))
# 
# # Make a table of values of D_creds_possible
# merged_snapshot %>%
#   group_by(D_creds_possible) %>%
#   summarise(count = n(), .groups = 'drop')
# 


#quick change to run LB Scripts
#merged_snapshot <- snapshot_pathways_table_2023
#clean up the old table a bit to help with space
#rm(snapshot_pathways_table_2023)

#### Pathways variables

# Convert credits possible to be numeric
base_analysis <- base_analysis %>%
  mutate(D_creds_earn = as.numeric(D_creds_earn))

#Make sure to edit these sorts of things for NA 
# Create a new variable called C_any_pathway
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_pathway = ifelse(any(D_pathways_flag == "Y"), "Y", "N")) %>%
  mutate(C_any_pathway =ifelse(is.na(C_any_pathway), "N", as.character(C_any_pathway))) %>% 
  ungroup()

#Check: quickly summarise rows to see if it worked
base_analysis %>%
  group_by(C_any_pathway) %>%
  summarise(distinct_count = n_distinct(D_stu_id))


# Create a new variable called M_total_pathways_creds_possible that sums all of the pathways credits possible for each student
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(M_total_pathways_creds_possible = sum(ifelse(D_pathways_flag == "Y", as.numeric(D_creds_earn), 0), na.rm = TRUE)) %>%
  ungroup()

#as a check, return the number of distinct students by credits possible -> come back to check on this, seems very high
base_analysis %>%
  group_by(M_total_pathways_creds_possible) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

#### Advanced coursework variables

# Create a new variable that identifies if a student took any above standard coursework
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_advanced = ifelse(any(D_course_rigor == "Above Standard"), "Y", "N")) %>%
  ungroup()

#Check that the flag worked
base_analysis %>%
  group_by(C_any_advanced) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

# Create a new variable that sums all of the above standard credits possible for each student -> found an issue with this where it will sum every row because of the pathways transformation. -> I think I essentially just need to go back and create a new data frame that doesn't have the pivoted longer pathways stuff to do these transformations on. Hold here.
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(M_total_advanced_creds_possible = sum(ifelse(D_course_rigor == "Above Standard", as.numeric(D_creds_earn), 0), na.rm = TRUE)) %>%
  ungroup()

#and check it
base_analysis %>%
  group_by(M_total_advanced_creds_possible) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

#### AP coursework variables

# Create a new variable that identifies if a student took any ap courses
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_ap = ifelse(any(D_AP_flag == "Y"), "Y", "N")) %>%
  ungroup()

#Check that the flag worked
base_analysis %>%
  group_by(C_any_ap) %>%
  summarise(distinct_count = n_distinct(D_stu_id))


# Create a new variable called M_total_AP_creds_possible that sums all of the dc credits possible for each student
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(M_total_ap_creds_possible = sum(ifelse(D_AP_flag == "Y", as.numeric(D_creds_earn), 0), na.rm = TRUE)) %>%
  ungroup()

#and check it
base_analysis %>%
  group_by(M_total_ap_creds_possible) %>%
  summarise(distinct_count = n_distinct(D_stu_id))


#### Dual credit coursework variables 

# Create a new variable called C_any_dc that identifies if a student took any dual credit
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_dc = ifelse(any(D_dual_credit_flag == "Y"), "Y", "N")) %>%
  ungroup()

#Check that the flag worked
base_analysis %>%
  group_by(C_any_dc) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

# Create a new variable called M_total_AP_creds_possible that sums all of the advanced credits possible for each student
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(M_total_dc_creds_possible = sum(ifelse(D_dual_credit_flag == "Y", as.numeric(D_creds_earn), 0), na.rm = TRUE)) %>%
  ungroup()

#and check it
base_analysis %>%
  group_by(M_total_dc_creds_possible) %>%
  summarise(distinct_count = n_distinct(D_stu_id))



### Create new variables around failures of important courses 

# Create H_fail_alg_1
alg_1_course_ids <- c("AL1PAP", "ALG1A", "ALG1B", "ALG1A", "ALG1AI", "ALG1B", "ALG1BI", "ALG1EA", "ALG1EB", 
                      "ALG1FA", "ALG1FB", "ALG1HA", "ALG1HB", "ALG1IA", "ALG1IB", "ALG1LA", "ALG1LB", 
                      "ALG1MA", "ALG1MB", "ALG1RA", "ALG1RB", "ALGALT", "ALGAMA", "ALGAMB", "ALGEMA", 
                      "ALGEMB", "ALGIMA", "ALGIMB", "ALGMIA", "ALGMIB", "PAPLGA", "PAPLGB")
  
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(H_fail_alg_1 = case_when(
    any(D_course_id %in% alg_1_course_ids & D_pass_fail == "N") ~ "Y",
    any(D_course_id %in% alg_1_course_ids & D_pass_fail == "Y") ~ "N",
    TRUE ~ NA_character_
  )) %>%
  ungroup()

#check it -> should add up to total students WATCH OUT for multiple pass/fail students, add checks as needed if distinct count does not match total
base_analysis %>%
  group_by(H_fail_alg_1) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

#and a quick check how many distinct 9th graders are in the base_analysis data-frame?
base_analysis %>%
  filter(C_eoy_grade == "9") %>%
  group_by(D_stu_id) %>%
  summarise(count = n()) %>%
  summarise(distinct_count = n_distinct(D_stu_id))


# Create H_fail_ela_1
ela_1_course_ids <- c("ENALA", "ENALB", "ENG1A", "ENG1B", "ENG1EA", "ENG1EB", "ENG1FA", "ENG1FB", "ENG1HA", 
                "ENG1HB", "ENG1IA", "ENG1IB", "ENG1LA", "ENG1LB", "ENG1RA", "ENG1RB", "ENG1WA", 
                "ENG1WB", "ESL1A", "ESOL1A", "ESOL1B")
  
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(H_fail_ela_1 = case_when(
    any(D_course_id %in% ela_1_course_ids & D_pass_fail == "N") ~ "Y",
    any(D_course_id %in% ela_1_course_ids & D_pass_fail == "Y") ~ "N",
    TRUE ~ NA_character_
  )) %>%
  ungroup()

#check it -> should add up to total students WATCH OUT for multiple pass/fail students, add checks as needed if distinct count does not match total
base_analysis %>%
  group_by(H_fail_ela_1) %>%
  summarise(distinct_count = n_distinct(D_stu_id))

#Lets filter to just 9th grade students, and then filter to just students who have an N/A for H_fail_ela_1
ela_9_nas <- base_analysis %>%
  filter(C_eoy_grade == "9") %>%
  filter(is.na(H_fail_ela_1)) %>%
  group_by(D_stu_id)

```

#2.2 High Level Context

#a. count of included students

```{r}
# Count of distinct student IDs in merged snapshot
base_analysis %>%
  select(D_stu_id) %>%
  distinct() %>%
  count()
```

#b. overall student demographic info

```{r}
# Count of students by D_stu_race
# Count of distinct students by D_stu_race
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_race) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_race = "Total", count_students = sum(.$count_students), percentage = 100)

# Count of students by SWD
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_swd) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_swd = "Total", count_students = sum(.$count_students), percentage = 100)


# Count of students by FRL
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_frl) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_frl = "Total", count_students = sum(.$count_students), percentage = 100)
```

#c. overall student grade level

```{r}
# Count of students by C_eoy_grade
summary_table <- base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  mutate(C_eoy_grade = as.character(C_eoy_grade)) %>%
  group_by(C_eoy_grade) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  bind_rows(tibble(C_eoy_grade = "Total", 
                   count_students = sum(.$count_students), 
                   percentage = 100)) %>%
  arrange(C_eoy_grade)

summary_table %>% flextable()
```

#d. student demographics by school

```{r}
#before I do these for display purposes, I think I want to roll-up Manor High, Manor Early College, and Manor Senior High to one reporting field that will just be called "Manor High Comprehensive". So first lets make a new column called "location_rollup" that will take values from C_home_school_name and then replace the values of Manor High, Manor Early College, and Manor Senior High with "Manor High Comprehensive"
base_analysis <- base_analysis %>%
  mutate(location_rollup = case_when(
    C_home_school_name %in% c("MANOR HIGH SCHO", "MANOR EARLY COL", "MANOR SENIOR HI") ~ "Manor High Comprehensive",
    TRUE ~ C_home_school_name
  ))

#and now using this location-rollup for reporting purposes


# Create a table by race
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(location_rollup, D_stu_race) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(location_rollup = "Total", count_students = sum(.$count_students), percentage = 100)

# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, D_stu_race) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = D_stu_race, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(location_rollup = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()

#count of distinct students by location_rollup
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(location_rollup) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  add_row(location_rollup = "Total", count_students = sum(.$count_students))

```

#e. student performance by school

```{r}
#before I do this I need to cleanup my C_grade_pct_rollup column because its not displaying exactly how I want it to
base_analysis <- base_analysis %>%
  mutate(C_grade_pct_rollup = case_when(
    D_final_grade >= 90 ~ "90+",
    D_final_grade >= 80 & D_final_grade < 90 ~ "80-89",
    D_final_grade >= 70 & D_final_grade < 80 ~ "70-79",
    D_final_grade >= 60 & D_final_grade < 70 ~ "60-69",
    D_final_grade >= 50 & D_final_grade < 60 ~ "50-59",
    D_final_grade < 50 ~ "Lower than 50",
    is.na(D_final_grade) ~ "Not graded",
    TRUE ~ as.character(D_final_grade) # This handles any other cases, just in case
  ))

#and as a quick check before I move on lets check the unique course Ids of courses that show up as "Not graded"
unique(base_analysis$D_course_id[base_analysis$C_grade_pct_rollup == "Not graded"]) -> unique_not_graded_courses

#and print them
print(unique_not_graded_courses)

#okay there are defintley courses there that I would expect to be graded, so maybe sometimes students just didn't receive a grade for the course

# Create pivot table with row totals
pivot_table <- base_analysis %>%
  group_by(location_rollup, C_grade_pct_rollup) %>%
  summarise(count_courses = n(), .groups = 'drop') %>%
  pivot_wider(names_from = C_grade_pct_rollup, values_from = count_courses, values_fill = 0) %>%
  mutate(row_total = rowSums(across(-location_rollup)))

# Print the pivot table
pivot_table %>% flextable()

# Function to create pivot table with percentages
create_percentage_table <- function(data) {
  # Create pivot table with counts
  pivot_table <- data %>%
    group_by(location_rollup, C_grade_pct_rollup) %>%
    summarise(count_courses = n(), .groups = 'drop') %>%
    pivot_wider(names_from = C_grade_pct_rollup, values_from = count_courses, values_fill = 0)

  # Calculate percentages for each row and round to one decimal place
  percentage_table <- pivot_table %>%
    mutate(across(-location_rollup, ~ round((. / rowSums(across(-location_rollup))) * 100, 1)))

  # Return the percentage table
  percentage_table
}

# Create the percentage table
percentage_table <- create_percentage_table(base_analysis)

# Print the percentage table
percentage_table %>% flextable()


#After talking with Ryan, before we move on, we actually want to exclude any rows that have N/A Final grade values from the analysis, so at this point in our base analysis frame thats filter out any row that has a value of N/A in D_final_grade
base_analysis <- base_analysis %>%
  filter(!is.na(D_final_grade))

#So now there should be no N/A Rows moving forward
```

#f. student performance in Alg 1 & ELA 1

```{r}
#before checking this I want to make sure that no student is counted as both passing and failing the same course ID
duplicate_pass_fail <- base_analysis %>%
  group_by(D_stu_id, D_course_id) %>%
  filter(any(D_pass_fail == "Y") & any(D_pass_fail == "N")) %>%
  ungroup()

# View the result
print(duplicate_pass_fail)

#how many distinct students with duplicates?
duplicate_pass_fail %>%
  summarise(count = n_distinct(D_stu_id))

#Lets create a flag for these students to easily filter out their "yes" rows. First I want to tag every "Y" row for the "D_pass_fail" variable where a unique student also has an "N" for row for the same course. This new column should be called "credit_recovery_duplicate", and the "yes" flags should only be rows marked "Y" where a student also has a "N" row for the same course.
base_analysis <- base_analysis %>%
  group_by(D_stu_id, D_course_id) %>%
  mutate(credit_recovery_duplicate = ifelse(any(D_pass_fail == "Y") & any(D_pass_fail == "N") & D_pass_fail == "Y", "Y", "N")) %>%
  ungroup()

#and check that the flag worked by summing the number of rows marked "Y" for credit_recovery_duplicate
base_analysis %>%
  filter(credit_recovery_duplicate == "Y") %>%
  summarise(count = n())

#It should be the same number as "duplicates"Y" duplicate rows, which it is so we are good!
#NOTE- the new column did not "display" in the viewer for me very easily because it was at the end of the data frame, so I had to scroll to the right to see it. R can only display 50 columns at a time

#06/14 update after talking with Ryan we are actually just going to exclude these "Y" rows from the analysis, so we will filter them out now
base_analysis <- base_analysis %>%
  filter(credit_recovery_duplicate == "N")

#and quickly double check from your duplicates frame that a couple of the student IDs with duplicate pass/fail rows were filtered out of the data-frame correctly

## Algebra 1 overall 
# Create a table of distinct counts of students who fail algebra 1
base_analysis %>%
  group_by(H_fail_alg_1) %>%
  distinct(D_stu_id) %>%
  summarise(count = n_distinct(D_stu_id), .groups = 'drop')

# Create a table of distinct students H_fail_alg_1 with counts and percents
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_alg_1)) %>%
  # Find distinct combinations of D_stu_id
  distinct(D_stu_id, H_fail_alg_1) %>%
  # Summarize data by H_fail_alg_1
  group_by(H_fail_alg_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id
    count = n_distinct(D_stu_id),
    # Calculate the percentage of each group out of the total distinct students
    percent = (n_distinct(D_stu_id) / n_distinct(.$D_stu_id)) * 100,
    .groups = 'drop'
  )

result

## Algebra 1 failures by school
# Create a table of distinct students H_fail_alg_1 with counts and percents within each school
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_alg_1)) %>%
  # Find distinct combinations of D_stu_id, H_home_school_name, and H_fail_alg_1
  distinct(D_stu_id, location_rollup, H_fail_alg_1) %>%
  # Group by H_home_school_name and H_fail_alg_1
  group_by(location_rollup, H_fail_alg_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id within each school and H_fail_alg_1
    count = n_distinct(D_stu_id),
    .groups = 'drop_last'
  ) %>%
  # Calculate the total distinct students within each school
  mutate(total_school = sum(count)) %>%
  # Calculate the percentage of each group out of the total distinct students within the same school
  mutate(percent = (count / total_school) * 100) %>%
  # Drop the grouping structure
  ungroup() %>%
  # Drop the total_school column
  select(-total_school)

result

## Algebra 1 failures by race
# Create a table of distinct students H_fail_alg_1 with counts and percents within each school
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_alg_1)) %>%
  # Find distinct combinations of D_stu_id, H_home_school_name, and H_fail_alg_1
  distinct(D_stu_id, D_stu_race, H_fail_alg_1) %>%
  # Group by H_home_school_name and H_fail_alg_1
  group_by(D_stu_race, H_fail_alg_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id within each school and H_fail_alg_1
    count = n_distinct(D_stu_id),
    .groups = 'drop_last'
  ) %>%
  # Calculate the total distinct students within each school
  mutate(total_race = sum(count)) %>%
  # Calculate the percentage of each group out of the total distinct students within the same school
  mutate(percent = (count / total_race) * 100) %>%
  # Drop the grouping structure
  ungroup() %>%
  # Drop the total_school column
  select(-total_race)

result

## ELA 1 overall
# Create a table of distinct counts of students who fail ELA 1
base_analysis %>%
  group_by(H_fail_ela_1) %>%
  distinct(D_stu_id) %>%
  summarise(count = n_distinct(D_stu_id), .groups = 'drop')

# Create a table of distinct students H_fail_ela_1 with counts and percents
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_ela_1)) %>%
  # Find distinct combinations of D_stu_id
  distinct(D_stu_id, H_fail_ela_1) %>%
  # Summarize data by H_fail_alg_1
  group_by(H_fail_ela_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id
    count = n_distinct(D_stu_id),
    # Calculate the percentage of each group out of the total distinct students
    percent = (n_distinct(D_stu_id) / n_distinct(.$D_stu_id)) * 100,
    .groups = 'drop'
  )

result

## ELA 1 failures by school
# Create a table of distinct students H_fail_alg_1 with counts and percents within each school
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_ela_1)) %>%
  # Find distinct combinations of D_stu_id, H_home_school_name, and H_fail_alg_1
  distinct(D_stu_id,location_rollup, H_fail_ela_1) %>%
  # Group by H_home_school_name and H_fail_alg_1
  group_by(location_rollup, H_fail_ela_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id within each school and H_fail_alg_1
    count = n_distinct(D_stu_id),
    .groups = 'drop_last'
  ) %>%
  # Calculate the total distinct students within each school
  mutate(total_school = sum(count)) %>%
  # Calculate the percentage of each group out of the total distinct students within the same school
  mutate(percent = (count / total_school) * 100) %>%
  # Drop the grouping structure
  ungroup() %>%
  # Drop the total_school column
  select(-total_school)

result


## ELA 1 failures by race
# Create a table of distinct students H_fail_alg_1 with counts and percents within each school
result <- base_analysis %>%
  # Filter out rows where H_fail_alg_1 is NA
  filter(!is.na(H_fail_ela_1)) %>%
  # Find distinct combinations of D_stu_id, H_home_school_name, and H_fail_alg_1
  distinct(D_stu_id, D_stu_race, H_fail_ela_1) %>%
  # Group by H_home_school_name and H_fail_alg_1
  group_by(D_stu_race, H_fail_ela_1) %>%
  summarise(
    # Calculate the count of distinct D_stu_id within each school and H_fail_ela_1
    count = n_distinct(D_stu_id),
    .groups = 'drop_last'
  ) %>%
  # Calculate the total distinct students within each school
  mutate(total_race = sum(count)) %>%
  # Calculate the percentage of each group out of the total distinct students within the same school
  mutate(percent = (count / total_race) * 100) %>%
  # Drop the grouping structure
  ungroup() %>%
  # Drop the total_school column
  select(-total_race)

result
```

##2.3 Pathways Summary

#a. pathways credits

```{r}
# Create a table to look at how many student were enrolled in ANY pathways 
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(C_any_pathway) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(C_any_pathway = "Total", count_students = sum(.$count_students), percentage = 100)

# Make sure M_total_pathways_creds_possible is numeric
base_analysis <- base_analysis %>%
  mutate(M_total_pathways_creds_possible = as.character(M_total_pathways_creds_possible))

# Create a summary table of pathways credits possible
summary_table <- base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(M_total_pathways_creds_possible) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1))

# Calculate the grand total
grand_total <- base_analysis %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(M_total_pathways_creds_possible = "Total")

# Combine the summary table with the grand total row
summary_table <- bind_rows(summary_table, grand_total)

# Print the summary table
summary_table %>% flextable()
```

#b. pathways enrollment by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, C_any_pathway) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_pathway, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(location_rollup = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()
```

#c. pathways credits possible by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, M_total_pathways_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_pathways_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(location_rollup = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()


```

#d. pathways credits by grade level

```{r}
# Convert C_eoy_grade to a leveled factor initially
base_analysis <- base_analysis %>%
  mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, C_any_pathway) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_pathway, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, M_total_pathways_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_pathways_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables with pathways
tables_with_pathways <- create_tables_with_pathways(base_analysis)

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#e. pathways enrollment by race / ethnicity

```{r}
# Function to create the count and percentage tables
create_tables <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], C_any_pathway) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_pathway, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and C_any_pathway
tables_any_pathway <- create_tables(base_analysis, "D_stu_race")

# Print the final tables
tables_any_pathway$final_count_data %>% flextable()
tables_any_pathway$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], M_total_pathways_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_pathways_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and M_total_pathways_creds_possible
tables_with_pathways <- create_tables_with_pathways(base_analysis, "D_stu_race")

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

##2.4 Pathways Analysis

#a. enrollment by pathway (no cross-pathway courses)

```{r}
#first lets make the snapshot_pathways_table 2023 a little easier to navigate and read by filtering to only rows with a value of 1 in the column C_pathway_y_or_n
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  filter(C_pathway_Y_or_N == 1)

#so now the data-frame should be filtered to one pathway course per student PER pathway (for multipless). For example, student 101098 shows 3 rows for one course because it could be in 3 pathways

# Count of students by pathway 
pathways_student_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "N") %>%
  group_by(C_pathway) %>%
  summarise(distinct_student_count = n_distinct(D_stu_id), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_student_count <- pathways_student_count %>%
  summarise(C_pathway = "Total", distinct_student_count = sum(distinct_student_count))

final_student_count <- bind_rows(pathways_student_count, total_student_count)

# Print the table
final_student_count %>% flextable()

# Count of course seats by pathway
# Count of D_course_id by C_pathway with filter for C_multiple_cats == "N" and total row
pathways_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "N") %>%
  group_by(C_pathway) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_course_count <- pathways_course_count %>%
  summarise(C_pathway = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(pathways_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()
```

#b. enrollment by pathway (cross-pathway courses)

```{r}
# Count of students by pathway 
pathways_student_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "Y") %>%
  group_by(C_pathway) %>%
  summarise(distinct_student_count = n_distinct(D_stu_id), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_student_count <- pathways_student_count %>%
  summarise(C_pathway = "Total", distinct_student_count = sum(distinct_student_count))

final_student_count <- bind_rows(pathways_student_count, total_student_count)

# Print the table
final_student_count %>% flextable()

# Count of course seats by pathway
# Count of D_course_id by C_pathway with filter for C_multiple_cats == "Y" and total row
pathways_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "Y") %>%
  group_by(C_pathway) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_course_count <- pathways_course_count %>%
  summarise(C_pathway = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(pathways_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()

```

#c. enrollment by cluster

```{r}
#First I have to look at the number of courses that could be cross-cluster courses, so I want to look by course ID to see if any courses have more than one cluster tagged to them
snapshot_pathways_table_2023 %>%
  group_by(D_course_id) %>%
  summarise(cluster_count = n_distinct(C_cluster), .groups = 'drop') %>%
  filter(cluster_count > 1)

#so its two courses, MDTEMA and PRNHUA. Medical Terminology and Principles of human services. Similar to the multiple pathways flag, lets create a multiple cluster flag to flag any rows for those two course IDs as "Y" for multiple clusters, all other rows get an N flag
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  mutate(Multiple_clusters = ifelse(D_course_id %in% c("MDTEMA", "PRNHUA"), "Y", "N"))

#and make sure to apply that flag in filters


#count of students by cluster- NOT multiple clusters
# Count of students by pathway 
cluster_student_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_clusters == "N") %>%
  group_by(C_cluster) %>%
  summarise(distinct_student_count = n_distinct(D_stu_id), .groups = 'drop') %>%
  arrange(C_cluster)

# Add total row
total_student_count <- cluster_student_count %>%
  summarise(C_cluster = "Total", distinct_student_count = sum(distinct_student_count))

final_student_count <- bind_rows(cluster_student_count, total_student_count)

# Print the table
final_student_count %>% flextable()


#course seats by cluster-NOT multiple clusters
cluster_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_clusters == "N") %>%
  group_by(C_cluster) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_cluster)

# Add total row
total_course_count <- cluster_course_count %>%
  summarise(C_cluster = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(cluster_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()

#and then you can do the same for courses that span multiple clusters
#count of students by cluster

# Count of students by pathway - YES multiple clusters
cluster_student_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_clusters == "N") %>%
  group_by(C_cluster) %>%
  summarise(distinct_student_count = n_distinct(D_stu_id), .groups = 'drop') %>%
  arrange(C_cluster)

# Add total row
total_student_count <- cluster_student_count %>%
  summarise(C_cluster = "Total", distinct_student_count = sum(distinct_student_count))

final_student_count <- bind_rows(cluster_student_count, total_student_count)

# Print the table
final_student_count %>% flextable()


#course seats by cluster  - YES multiple clusters
cluster_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_clusters == "Y") %>%
  group_by(C_cluster) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_cluster)

# Add total row
total_course_count <- cluster_course_count %>%
  summarise(C_cluster = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(cluster_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()


```

##2.5 Success by pathway

#a. sucess by pathway by enrollments

```{r}
#Two quick methodology framing notes here-
#1, 06/14 before this I want to actually exclude any rows that have an N/A value for "D_final_grade" because it will mess up our success data, so lets filter those out first
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  filter(!is.na(D_final_grade))

#2, 06/14 we actually don't need to filter out multiple catagories for this analysis because we want to count the failure as a potential failure in both pathways, just note that they will be double counted

# went from 3048 rows to 2987, which is right. NOTE- this WILL change N-size of "pathway takers", which is okay!

# Count of D_course_id by C_pathway and C_stu_grade_buckets with filter for C_multiple_cats == "N"
pathways_student_count <- snapshot_pathways_table_2023 %>%
  group_by(C_pathway, C_grade_pct_rollup) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_pathway = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

## Also look at success by courses that span pathways 
# Count of D_course_id by C_pathway and C_stu_grade_buckets with filter for C_multiple_cats == "N"
pathways_student_count <- snapshot_pathways_table_2023 %>%
  group_by(C_pathway, C_grade_pct_rollup) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_pathway = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()


#I am going to try to simply this and instead just get a count of the % of students that passed courses (by pathway) and then the % of students that failed

# Count of D_course_id by C_pathway and C_stu_grade_buckets with filter for C_multiple_cats == "N"
pathways_student_count <- snapshot_pathways_table_2023 %>%
  group_by(C_pathway, D_pass_fail) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_pathway = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

## Also look at success by courses that span pathways 
# Count of D_course_id by C_pathway and C_stu_grade_buckets with filter for C_multiple_cats == "N"
pathways_student_count <- snapshot_pathways_table_2023 %>%
  group_by(C_pathway, D_pass_fail) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_pathway = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

```

#b. success by course level

```{r}
# Count of D_course_id by D_course_level and C_stu_grade_buckets with filter for C_multiple_cats == "N"
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(Multiple_pathways == "N") %>% # Note that we could leave in all here because we are not looking by pathway
  group_by(C_pathway_course_level, D_pass_fail) %>%  # Changed to D_course_level and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from =D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_course_level = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

#quickly count the number of courses that do not have a grade associated with them- NOTE should be zero now that we fixed this from before
snapshot_pathways_table_2023 %>%
  filter(is.na(D_final_grade)) %>%
  #count distinct course Ids where this is the case
count(D_course_id)

#And the number of courses that count for multiple pathways because I think its small-> but fine because they're counted on top and bottom so not necessary
snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "Y") %>%
  #count distinct course Ids where this is the case
count(D_course_id)

```

#c. success by school

```{r}
#First I am needing to go back in again and fix C_grade_pct_rollup because I only edited it for the "base-analysis" document, so I have to implement the fix here as well
snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  mutate(C_grade_pct_rollup = case_when(
    D_final_grade >= 90 ~ "90+",
    D_final_grade >= 80 ~ "80-89",
    D_final_grade >= 70 ~ "70-79",
    D_final_grade >= 60 ~ "60-69",
    D_final_grade >= 50 ~ "50-59",
    D_final_grade < 50 ~ "Below 50",
    TRUE ~ "Not a grade"
  ))

#and quickly summarize the number of rows that fall into each bucket now
snapshot_pathways_table_2023 %>%
  group_by(C_grade_pct_rollup) %>%
  summarise(count = n())


#same as above, checking for credit-recovery doubles
#before checking this I want to make sure that no student is counted as both passing and failing the same course ID
duplicate_pass_fail <- snapshot_pathways_table_2023 %>%
  group_by(D_stu_id, D_course_id) %>%
  filter(any(D_pass_fail == "Y") & any(D_pass_fail == "N")) %>%
  ungroup()

# View the result
print(duplicate_pass_fail)

#there are zero so we actually don't need to do this any further here!

#how many distinct students with duplicates?
#duplicate_pass_fail %>%
  #summarise(count = n_distinct(D_stu_id))

#Lets create a flag for these students to easily filter out their "yes" rows. First I want to tag every "Y" row for the "D_pass_fail" variable where a unique student also has an "N" for row for the same course. This new column should be called "credit_recovery_duplicate", and the "yes" flags should only be rows marked "Y" where a student also has a "N" row for the same course.
#snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  #group_by(D_stu_id, D_course_id) %>%
  #mutate(credit_recovery_duplicate = ifelse(any(D_pass_fail == "Y") & any(D_pass_fail == "N") & D_pass_fail == "Y", "Y", "N")) %>%
  #ungroup()

#and check that the flag worked by summing the number of rows marked "Y" for credit_recovery_duplicate
#snapshot_pathways_table_2023 %>%
  #filter(credit_recovery_duplicate == "Y") %>%
  #summarise(count = n())

#It should be the same number as "duplicates"Y" duplicate rows, which it is so we are good!


#exclude these "Y" rows from the analysis, so we will filter them out now
#snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  #filter(credit_recovery_duplicate == "N")

snapshot_pathways_table_2023 <- snapshot_pathways_table_2023 %>%
  mutate(location_rollup = case_when(
    C_home_school_name %in% c("MANOR HIGH SCHO", "MANOR EARLY COL", "MANOR SENIOR HI") ~ "Manor High Comprehensive",
    TRUE ~ C_home_school_name
  ))

#okay great, moving on

#And before checking the success of pathway courses, need to do the Manor specific check due to credit-recovery coding that no student has both passed and failed the same course


# Count of D_course_id by C_home_school_name and C_stu_grade_buckets
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(location_rollup, C_grade_pct_rollup) %>%  # Changed to C_home_school_name and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(location_rollup = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total

pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()


#same thing but this time just pass/fail because I want to have that look too

# Count of D_course_id by C_home_school_name and C_stu_grade_buckets
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(location_rollup, D_pass_fail) %>%  # Changed to C_home_school_name and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(location_rollup = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total

pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

```

#d. success by demographic groups

```{r}
# Count of D_course_id by C_stu_race_rollup and C_stu_grade_buckets
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(D_stu_race, C_grade_pct_rollup) %>%  # Changed to C_stu_race_rollup and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

#same thing but just by pass/fail
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(D_stu_race, D_pass_fail) %>%  # Changed to C_stu_race_rollup and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()


```

#e. Success broken out by pathway and race

```{r}
#-> SE Might come back to this later, but I dont know that I need this level of detail right now

# Filter the data to include only relevant entries
filtered_pathways_table <-snapshot_pathways_table_2023
  #filter(C_multiple_cats == "N")  # Don't need to filter for % success looks

# Create tables filtered by each value of C_stu_race_rollup
filtered_tables <- filtered_pathways_table %>%
  split(.$D_stu_race) %>%
  lapply(function(df) {
    # Count of D_course_id by C_pathway and C_stu_grade_buckets
    counts_table <- df %>%
      group_by(C_pathway, C_grade_pct_rollup) %>%
      summarise(course_count = n(), .groups = 'drop') %>%
      pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))
    
    # Add a row to sum the column totals
    counts_table <- counts_table %>%
      bind_rows(
        counts_table %>%
          summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
          mutate(C_pathway = "Total")
      )
    
    # Add a column for the column total
    counts_table <- counts_table %>%
      mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))
    
    # Create a percentage table
    percentage_table <- counts_table %>%
      rowwise() %>%
      mutate(across(2:(ncol(counts_table)-1), ~ round(. / sum(c_across(2:(ncol(counts_table)-1))) * 100, 1), .names = "{col}_percent")) %>%
      ungroup()
    
    # Return the combined table
    return(percentage_table)
  })

# Print each filtered table with clear identification in R Markdown
for (name in names(filtered_tables)) {
  cat("\n### Table for C_stu_race_rollup =", name, ":\n\n")
  print(filtered_tables[[name]])
}
```

##2.6 Advanced Coursework Enrollment

#Ω. Dual-Credit Code Fix - **What:** I discovered additional information related to dual credit courses, and instead of re-running the previous inputs, I am going to patch on the additional courses here

-   **Check:** Included in the check, the number of dual credit students should increase

-   **Project Action:** None

```{r}
#First I want to check the number of dual credit students before I add the additional courses
num_distinct_students_with_dual_credit <- base_analysis %>%
  filter(D_dual_credit_flag == "Y") %>%
  distinct(D_stu_id) %>%
  count()

# Print the result
print(num_distinct_students_with_dual_credit)


#alright lets add in the new courses
# Mutate D_dual_credit_flag for specified courses
base_analysis <- base_analysis %>%
  mutate(D_dual_credit_flag = ifelse(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD"), "Y", D_dual_credit_flag))

#and check that the number of students increased
num_distinct_students_with_dual_credit <- base_analysis %>%
  filter(D_dual_credit_flag == "Y") %>%
  distinct(D_stu_id) %>%
  count()

# Print the result
print(num_distinct_students_with_dual_credit)

#First add the courses to above standard rigor
base_analysis <- base_analysis %>%
  mutate(D_course_rigor = ifelse(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD"), "Above Standard", D_course_rigor))


# Re-run the C_any_advanced flag to add students to any advanced
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_advanced = ifelse(any(D_course_rigor == "Above Standard"), "Y", "N")) %>%
  ungroup()

#and add to C_any_dc
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(C_any_dc = ifelse(any(D_dual_credit_flag == "Y"), "Y", "N")) %>%
  ungroup()

```

#a. advanced coursework high-level enrollment

```{r}
#-> back to base analysis data frame!
#Looking at this code, it looks like I first need to mutate the base analysis data-frame to create a new column called "M_total_advanced_creds_possible" that groups by student ID and then sums the "D_creds_earned" column for all "Above Standard" courses under "D_course_rigor"
# Create a new column to sum the total advanced credits possible
base_analysis <- base_analysis %>%
  group_by(D_stu_id) %>%
  mutate(M_total_advanced_creds_possible = sum(D_creds_earn[D_course_rigor == "Above Standard"], na.rm = TRUE)) %>% 
ungroup()


# Create a table to look at how many student were enrolled in ANY advanced coursework
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_course_rigor) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_course_rigor = "Total", count_students = sum(.$count_students), percentage = 100)

# Make sure M_total_pathways_creds_possible is numeric
base_analysis <- base_analysis %>%
  mutate(M_total_advanced_creds_possible = as.character(M_total_advanced_creds_possible))

# Create a summary table of pathways credits possible
summary_table <- base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(M_total_advanced_creds_possible) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1))

# Sort the combined table by M_total_advanced_creds_possible
summary_table <- summary_table %>%
  arrange(as.numeric(M_total_advanced_creds_possible))

# Calculate the grand total
grand_total <- summary_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(M_total_advanced_creds_possible = "Total")

# Combine the summary table with the grand total row
summary_table <- bind_rows(summary_table, grand_total)

# Print the summary table
summary_table %>% flextable()

#Not that many students, will have to keep that in mind... would expect far more than 27 out of 2822... I think that first table is not correctly working because the later table summarized more accuratley how I would expect to see advadnced enrollment 
```

#b.advanced coursework enrollment by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, C_any_advanced) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_advanced, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_home_school_name = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()
```

#c. advanced coursework credits possible by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, M_total_advanced_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_advanced_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_home_school_name = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()
```

#d. advanced credits by grade level

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, C_any_advanced) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_advanced, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, M_total_advanced_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_advanced_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables with pathways
tables_with_pathways <- create_tables_with_pathways(base_analysis)

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#e. advanced coursework enrollment by race / ethnicity

```{r}
# Function to create the count and percentage tables
create_tables <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], C_any_advanced) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_advanced, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and C_any_pathway
tables_any_pathway <- create_tables(base_analysis, "D_stu_race")

# Print the final tables
tables_any_pathway$final_count_data %>% flextable()
tables_any_pathway$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], M_total_advanced_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_advanced_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and M_total_pathways_creds_possible
tables_with_pathways <- create_tables_with_pathways(base_analysis, "D_stu_race")

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#f. advanced coursework enrollment by grade level

```{r}
#Count the number of students with any advanced course-work by grade level
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(C_eoy_grade, C_any_advanced) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(C_eoy_grade = "Total", count_students = sum(.$count_students), percentage = 100)







```

##2.7 AP Enrollment

#a. AP high-level enrollment

```{r}
## Advanced coursework enrollment
# Create a table to look at how many students were enrolled in ANY advanced coursework
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(C_any_ap) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(C_any_ap = "Total", count_students = sum(.$count_students), percentage = 100)

# Make sure M_total_ap_creds_possible is numeric
base_analysis <- base_analysis %>%
  mutate(M_total_ap_creds_possible = as.character(M_total_ap_creds_possible))

# Create a summary table of pathways credits possible
summary_table <- base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(M_total_ap_creds_possible) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1))

# Sort the combined table by M_total_ap_creds_possible
summary_table <- summary_table %>%
  arrange(as.numeric(M_total_ap_creds_possible))

# Calculate the grand total
grand_total <- summary_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(M_total_ap_creds_possible = "Total")

# Combine the summary table with the grand total row
summary_table <- bind_rows(summary_table, grand_total)

# Print the summary table
summary_table %>% flextable()
```

#b. AP enrollment by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, C_any_ap) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_ap, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_home_school_name = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()
```

#c. AP enrollment by grade level

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, C_any_ap) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_ap, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_ap_creds_possible
create_tables_with_pathways <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, M_total_ap_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_ap_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables with pathways
tables_with_pathways <- create_tables_with_pathways(base_analysis)

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#d. AP Enrollment by race / ethnicity

```{r}
# Function to create the count and percentage tables
create_tables <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], C_any_ap) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_ap, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and C_any_pathway
tables_any_pathway <- create_tables(base_analysis, "D_stu_race")

# Print the final tables
tables_any_pathway$final_count_data %>% flextable()
tables_any_pathway$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], M_total_ap_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_ap_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and M_total_pathways_creds_possible
tables_with_pathways <- create_tables_with_pathways(base_analysis, "D_stu_race")

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

##2.8 Dual Credit Enrollment

#a. DC high-level enrollment

```{r}
# Function to create the count and percentage tables
create_tables <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], C_any_dc) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_dc, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and C_any_pathway
tables_any_pathway <- create_tables(base_analysis, "D_stu_race")

# Print the final tables
tables_any_pathway$final_count_data %>% flextable()
tables_any_pathway$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], M_total_dc_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_dc_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and M_total_pathways_creds_possible
tables_with_pathways <- create_tables_with_pathways(base_analysis, "D_stu_race")

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#b. DC enrollment by school

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(location_rollup, C_any_dc) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_dc, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(location_rollup = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()
```

#c. DC enrollment by grade level

```{r}
# Function to create the count and percentage tables
create_tables <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, C_any_dc) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_dc, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables
tables <- create_tables(base_analysis)

# Print the final tables
tables$final_count_data %>% flextable()
tables$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_dc_creds_possible
create_tables_with_pathways <- function(data) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(C_eoy_grade, M_total_dc_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_dc_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Ensure C_eoy_grade maintains its factor levels
  pivoted_data <- pivoted_data %>%
    mutate(C_eoy_grade = factor(C_eoy_grade, levels = c("9", "10", "11", "12"), ordered = TRUE))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(C_eoy_grade = "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the tables with pathways
tables_with_pathways <- create_tables_with_pathways(base_analysis)

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#d. DC enrollment by race / ethnicity

```{r}
# Function to create the count and percentage tables
create_tables <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], C_any_dc) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = C_any_dc, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and C_any_pathway
tables_any_pathway <- create_tables(base_analysis, "D_stu_race")

# Print the final tables
tables_any_pathway$final_count_data %>% flextable()
tables_any_pathway$final_percent_data %>% flextable()

# Function to create the count and percentage tables with M_total_pathways_creds_possible
create_tables_with_pathways <- function(data, group_var) {
  # Calculate distinct count of D_stu_id
  summary_data <- data %>%
    distinct(D_stu_id, .keep_all = TRUE) %>%
    group_by(.data[[group_var]], M_total_dc_creds_possible) %>%
    summarise(count_students = n(), .groups = 'drop')

  # Pivot the data to wide format
  pivoted_data <- summary_data %>%
    pivot_wider(names_from = M_total_dc_creds_possible, values_from = count_students, values_fill = list(count_students = 0))

  # Add Grand Total row at the bottom for counts
  grand_total <- pivoted_data %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
    mutate(!!group_var := "Grand Total")

  # Combine the pivoted data with the Grand Total row
  final_count_data <- bind_rows(pivoted_data, grand_total)

  # Calculate percentages for each row and round to one decimal place
  percent_data <- pivoted_data %>%
    rowwise() %>%
    mutate(across(where(is.numeric), ~ round((. / sum(c_across(where(is.numeric)))) * 100, 1))) %>%
    ungroup()

  # Return the final count and percentage tables
  list(final_count_data = final_count_data, final_percent_data = percent_data)
}

# Create the table with C_stu_race_rollup and M_total_pathways_creds_possible
tables_with_pathways <- create_tables_with_pathways(base_analysis, "D_stu_race")

# Print the final tables
tables_with_pathways$final_count_data %>% flextable()
tables_with_pathways$final_percent_data %>% flextable()
```

#2.9 Success by course rigor (Advanced coursework)

```{r}
#something about the percentage calculations here is not working, so double-check that and calculate on your own. Come back to this 06/24
# Count of D_course_id by C_rigor and C_stu_grade_buckets with filter for C_multiple_cats == "N"
student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(location_rollup, C_grade_pct_rollup) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
student_count <- student_count %>%
  bind_rows(
    student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_rigor = "Total")
  )

# Add a column for the column total
student_count <- student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
print(student_count)

# Create a percentage table
percentage_student_count <- student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

#same thing for race/ethnicity
# Count of D_course_id by C_rigor and C_stu_grade_buckets with filter for C_multiple_cats == "N"
student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(D_stu_race, C_grade_pct_rollup) %>%  # Changed to C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
student_count <- student_count %>%
  bind_rows(
    student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_rigor = "Total")
  )

# Add a column for the column total
student_count <- student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
print(student_count)

# Create a percentage table
percentage_student_count <- student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()
```

#a. success for FRL Students- pathways

```{r}
# Count of D_course_id by FRL status and C_stu_grade_buckets
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(D_stu_frl, C_grade_pct_rollup) %>%  # Changed to C_stu_race_rollup and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()

#same thing but just by pass/fail
pathways_student_count <- snapshot_pathways_table_2023 %>%
  #filter(C_multiple_cats == "N") %>%  # Note that we could leave in all here because we are not looking by pathway
  group_by(D_stu_frl, D_pass_fail) %>%  # Changed to C_stu_race_rollup and C_stu_grade_buckets
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()


```

#b. success for FRL Students- advadnced coursework

```{r}
# Count of D_course_id by D_stu_frl and C_stu_grade_buckets
advanced_student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>%
  group_by(D_stu_frl, D_pass_fail) %>%  
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
advanced_student_count <- advanced_student_count %>%
  bind_rows(
    advanced_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_frl = "Total")
  )

# Add a column for the column total
advanced_student_count <- advanced_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
advanced_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- advanced_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()
```

#Part III: Deeper Dive Investigations

#3.1 Re-Checking Figures

```{r}
#First I really quickly need to double check the race/ethnicity numbers because they weren't really adding up for me
# Count of students by D_stu_race
# Count of distinct students by D_stu_race
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_race) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_race = "Total", count_students = sum(.$count_students), percentage = 100)

# Count of students by SWD
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_swd) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_swd = "Total", count_students = sum(.$count_students), percentage = 100)


# Count of students by FRL
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_frl) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(D_stu_frl = "Total", count_students = sum(.$count_students), percentage = 100)

#count of students by location rollup
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(location_rollup) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(location_rollup = "Total", count_students = sum(.$count_students), percentage = 100)

#count of students by C_eoy_grade
base_analysis %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(C_eoy_grade) %>%
  summarise(count_students = n(), .groups = 'drop') %>%
  mutate(percentage = round((count_students / sum(count_students)) * 100, 1)) %>%
  add_row(C_eoy_grade = "Total", count_students = sum(.$count_students), percentage = 100)
```

#3.2. Pathways deep dives #a. Healthcare Theraputics

```{r}
#first lets take a look at healthcare theraputics, who is taking healthcare theraputics (demos), and where are they taking them?

#summarize by location
healthcare_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Healthcare Therapeutic") %>% 
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by courses
healthcare_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Healthcare Therapeutic") %>% 
group_by(D_course_id) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by race/ethnicity
healthcare_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Healthcare Therapeutic") %>% 
distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_race) %>%
  summarize(number_of_students = n_distinct(D_stu_id)) %>%
  arrange(desc(number_of_students)) %>%
  print()

#what else could I look at here, grade buckets?
healthcare_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Healthcare Therapeutic") %>% 
  group_by(C_grade_pct_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

```

#b. Accounting and Finance

```{r}
#first lets take a look at healthcare theraputics, who is taking healthcare theraputics (demos), and where are they taking them?

#summarize by location
Accounting_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Accounting and Financial Services") %>% 
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by courses
Accounting_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Accounting and Financial Services") %>% 
group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by race/ethnicity
Accounting_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Accounting and Financial Services") %>% 
distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_race) %>%
  summarize(number_of_students = n_distinct(D_stu_id)) %>%
  arrange(desc(number_of_students)) %>%
  print()

#151/240= 62.9% of students in accounting and finance are Hispanic
#72/240=30% of students in accounting and finance are Black
#6/240=2.5% of students in accounting and finance are White
#11/240=4.6% of students in accounting and finance are other


```

#c. Students who failed Engineering

```{r}
#first lets take a look at healthcare theraputics, who is taking healthcare theraputics (demos), and where are they taking them?

#summarize by location
Engineering_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Engineering") %>% 
  filter(D_pass_fail == "N") %>%
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by courses
Engineering_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Engineering") %>% 
   filter(D_pass_fail == "N") %>%
group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#summarize by race/ethnicity
Engineering_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Engineering") %>% 
   filter(D_pass_fail == "N") %>%
distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(D_stu_race) %>%
  summarize(number_of_students = n_distinct(D_stu_id)) %>%
  arrange(desc(number_of_students)) %>%
  print()

```

#d. AP Enrollments

```{r}
#summarize by courses
AP_deep_dive <- base_analysis %>%
  filter(D_AP_flag == "Y") %>% 
group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

```

#d. Advadnced Course-Taking by Type, Distinct Students

```{r}
#First lets create an "honors" flag that will be true if the course is honors by looking at the D_course_rigor and Dual Credit and AP flags to flag any course that might be honors
base_analysis <- base_analysis %>%
  mutate(D_honors_flag = ifelse(D_course_rigor == "Above Standard" & D_dual_credit_flag == "N" & D_AP_flag == "N", "Y", "N"))

#and lets check that it worked by looking at a couple of the courses
honors_flag_check <- base_analysis %>%
  filter(D_honors_flag == "Y") %>%
  select(D_course_name.x, D_course_rigor, D_dual_credit_flag, D_AP_flag, D_honors_flag)
print(honors_flag_check)
  
#looks right to me, its all these Pre-AP classes. Now lets look at the breakdown of what percent of overall course-taking at each school is one of these advadnce course types

# Calculate the breakdown of students taking advanced courses by campus
advanced_student_breakdown <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>%
  mutate(
    advanced_type = case_when(
      D_dual_credit_flag == "Y" ~ "Dual Credit",
      D_AP_flag == "Y" ~ "AP",
      D_honors_flag == "Y" ~ "Honors",
      TRUE ~ "Other"  # Just in case there's an unflagged advanced course
    )
  ) %>%
  group_by(location_rollup, advanced_type) %>%
  summarise(distinct_students = n_distinct(D_stu_id), .groups = 'drop') # Count distinct students

# Calculate the total number of unique students per campus
total_students_per_campus <- base_analysis %>%
  group_by(location_rollup) %>%
  summarise(total_students = n_distinct(D_stu_id), .groups = 'drop') # Count total unique students

# Join the breakdown with the total students and calculate percentages
advanced_student_percentage <- advanced_student_breakdown %>%
  left_join(total_students_per_campus, by = "location_rollup") %>%
  mutate(percentage = round((distinct_students / total_students) * 100, 1)) %>%
  arrange(location_rollup, desc(percentage))

# Print the summary table
print(advanced_student_percentage %>% flextable())



```

#e. count of advanced course enrollments by advanced course type

```{r}
# Create an "honors" flag
base_analysis <- base_analysis %>%
  mutate(D_honors_flag = ifelse(D_course_rigor == "Above Standard" & D_dual_credit_flag == "N" & D_AP_flag == "N", "Y", "N"))

# Check that the honors flag worked
honors_flag_check <- base_analysis %>%
  filter(D_honors_flag == "Y") %>%
  select(D_course_name.x, D_course_rigor, D_dual_credit_flag, D_AP_flag, D_honors_flag)
print(honors_flag_check)

# Calculate the breakdown of advanced course enrollments by campus
advanced_course_breakdown <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>%
  mutate(
    advanced_type = case_when(
      D_dual_credit_flag == "Y" ~ "Dual Credit",
      D_AP_flag == "Y" ~ "AP",
      D_honors_flag == "Y" ~ "Honors",
      TRUE ~ "Other"  # Just in case there's an unflagged advanced course
    )
  ) %>%
  group_by(location_rollup, advanced_type) %>%
  summarise(course_enrollments = n(), .groups = 'drop') # Count course enrollments

# Calculate the total number of course enrollments per campus
total_enrollments_per_campus <- base_analysis %>%
  group_by(location_rollup) %>%
  summarise(total_enrollments = n(), .groups = 'drop') # Count total course enrollments

# Join the breakdown with the total enrollments and calculate percentages
advanced_course_percentage <- advanced_course_breakdown %>%
  left_join(total_enrollments_per_campus, by = "location_rollup") %>%
  mutate(percentage = round((course_enrollments / total_enrollments) * 100, 1)) %>%
  arrange(location_rollup, desc(percentage))

# Print the summary table
print(advanced_course_percentage %>% flextable())


```

#f. Overall FRL of snapshot students here

```{r}
# Count of students by FRL status
FRL_percentage <- base_analysis %>%
  group_by(location_rollup, D_stu_frl) %>%
  summarise(count_students = n_distinct(D_stu_id), .groups = 'drop') %>%
  pivot_wider(names_from = D_stu_frl, values_from = count_students, values_fill = list(count_students = 0)) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  mutate(FRL_percentage = round((Y / Total) * 100, 1)) %>%
  select(location_rollup, Y, N, Total, FRL_percentage) %>%
  arrange(desc(FRL_percentage))

print(FRL_percentage)



```

#g. Advanced Course Success ###i. by school

```{r}
#grade bucket distribution
advandced_student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(location_rollup, C_grade_pct_rollup) %>%  
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
advandced_student_count <- advandced_student_count %>%
  bind_rows(
    advandced_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(location_rollup = "Total")
  )

# Add a column for the column total
advandced_student_count <- advandced_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
advandced_student_count %>% flextable()

# Create a percentage table
advandced_student_count <- advandced_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
advandced_student_count %>% flextable()

#same thing but just by pass/fail

advandced_student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(location_rollup, D_pass_fail) %>%  
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
advandced_student_count <- advandced_student_count %>%
  bind_rows(
    advandced_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(location_rollup = "Total")
  )

# Add a column for the column total
advandced_student_count <- advandced_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
advandced_student_count %>% flextable()

# Create a percentage table
advandced_student_count <- advandced_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
advandced_student_count %>% flextable()

```

###ii. by advadnced course type

```{r}

# Create an "honors" flag
base_analysis <- base_analysis %>%
  mutate(D_honors_flag = ifelse(D_course_rigor == "Above Standard" & D_dual_credit_flag == "N" & D_AP_flag == "N", "Y", "N"))

# Check that the honors flag worked
honors_flag_check <- base_analysis %>%
  filter(D_honors_flag == "Y") %>%
  select(D_course_name.x, D_course_rigor, D_dual_credit_flag, D_AP_flag, D_honors_flag)
print(honors_flag_check)

# Calculate the breakdown of advanced course enrollments by campus and pass/fail status
advanced_course_breakdown <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>%
  mutate(
    advanced_type = case_when(
      D_dual_credit_flag == "Y" ~ "Dual Credit",
      D_AP_flag == "Y" ~ "AP",
      D_honors_flag == "Y" ~ "Honors",
      TRUE ~ "Other"  # Just in case there's an unflagged advanced course
    )
  ) %>%
  group_by(location_rollup, advanced_type, D_pass_fail) %>%
  summarise(course_enrollments = n(), .groups = 'drop') # Count course enrollments

# Calculate the total number of course enrollments per campus
total_enrollments_per_campus <- base_analysis %>%
  group_by(location_rollup) %>%
  summarise(total_enrollments = n(), .groups = 'drop') # Count total course enrollments

# Join the breakdown with the total enrollments and calculate percentages
advanced_course_percentage <- advanced_course_breakdown %>%
  left_join(total_enrollments_per_campus, by = "location_rollup") %>%
  mutate(percentage = round((course_enrollments / total_enrollments) * 100, 1)) %>%
  arrange(location_rollup, desc(percentage))

# Print the summary table
print(advanced_course_percentage %>% flextable())

```

###iii. by race/ethnicity

```{r}
#same thing but just by pass/fail

# Count of D_course_id by FRL status and C_stu_grade_buckets
advandced_student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(D_stu_race, D_pass_fail) %>%  
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
advandced_student_count <- advandced_student_count %>%
  bind_rows(
    advandced_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
advandced_student_count <- advandced_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
advandced_student_count %>% flextable()

# Create a percentage table
advandced_student_count <- advandced_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
advandced_student_count %>% flextable()

#grade bucket distribution
advandced_student_count <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>% 
  group_by(D_stu_race, C_grade_pct_rollup) %>%  
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = C_grade_pct_rollup, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
advandced_student_count <- advandced_student_count %>%
  bind_rows(
    advandced_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(D_stu_race = "Total")
  )

# Add a column for the column total
advandced_student_count <- advandced_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric) & !contains("Total"))))

# Print the table with column totals and row total
advandced_student_count %>% flextable()

# Create a percentage table
advandced_student_count <- advandced_student_count %>%
  rowwise() %>%
  mutate(across(2:5, ~ round(. / sum(c_across(2:5)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
advandced_student_count %>% flextable()

```

#h. Dual Credit enrollment at New-Tech

```{r}
#The dual credit enrollment at new-tech is highrer than Ryan would expect, so want to understand what courses those students are taking/why its so much higher!
# Filter to students at New Tech and dual credit courses, then count enrollments by unique course names
new_tech_dual_enrollment <- base_analysis %>%
  filter(location_rollup == "MANOR NEW TECHN", D_dual_credit_flag == "Y") %>%
  group_by(D_course_name.x) %>%
  summarise(enrollments = n(), .groups = 'drop') %>%
  arrange(desc(enrollments))

# Print the result
print(new_tech_dual_enrollment)


#same thing for manor high
manor_high_dual_enrollment <- base_analysis %>%
  filter(location_rollup == "Manor High Comprehensive", D_dual_credit_flag == "Y") %>%
  group_by(D_course_name.x) %>%
  summarise(enrollments = n(), .groups = 'drop') %>%
  arrange(desc(enrollments))

# Print the result
print(manor_high_dual_enrollment)

```

#i. Advanced course type enrollment by year

```{r}
#looks right to me, its all these Pre-AP classes. Now lets look at the breakdown of each year by the percent of above standard courses that were taken broken out by Honors, Dual Credit, and APs
# Filter for advanced courses and group by year and type of advanced course
advanced_course_breakdown <- base_analysis %>%
  filter(D_course_rigor == "Above Standard") %>%
  mutate(
    advanced_type = case_when(
      D_dual_credit_flag == "Y" ~ "Dual Credit",
      D_AP_flag == "Y" ~ "AP",
      D_honors_flag == "Y" ~ "Honors",
      TRUE ~ "Other"  # Just in case there's an unflagged advanced course
    )
  ) %>%
  group_by(C_eoy_grade, advanced_type) %>%
  summarise(count_students = n_distinct(D_stu_id), .groups = 'drop') %>%
  arrange(C_eoy_grade, desc(count_students))

# Print the summary table
print(advanced_course_breakdown %>% flextable())



```

### Part IV: Follow on Analyses

-   **What:** Conducting a set of follow on analyses for Manor ISD based on district interest

-   **Check:** When writing new code, double check that everything is methodologically sound and that the code is clean and easy to read

-   **Project Action:** None

#4.1 Course seats by pathway w/out principles courses

```{r}
# Count of course seats by pathway
# Count of D_course_id by C_pathway with filter for C_multiple_cats == "N" and total row
pathways_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "N") %>%
  filter(C_pathway_course_level != "Level 1") %>% #filters out level 1 courses
  group_by(C_pathway) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_course_count <- pathways_course_count %>%
  summarise(C_pathway = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(pathways_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()


#same thing but for coruses that go across multiple pathways
# Count of D_course_id by C_pathway with filter for C_multiple_cats == "N" and total row
pathways_course_count <- snapshot_pathways_table_2023 %>%
  filter(Multiple_pathways == "Y") %>%
   filter(C_pathway_course_level != "Level 1") %>% #filters out level 1 courses
  group_by(C_pathway) %>%
  summarise(course_count = n(), .groups = 'drop') %>%
  arrange(C_pathway)

# Add total row
total_course_count <- pathways_course_count %>%
  summarise(C_pathway = "Total", course_count = sum(course_count))

final_course_count <- bind_rows(pathways_course_count, total_course_count)

# Print the table
final_course_count %>% flextable()

```

#4.2 Pathway Failure Rate w/out Principles Courses

```{r}

pathways_student_count <- snapshot_pathways_table_2023 %>%
  group_by(C_pathway, D_pass_fail) %>% 
  filter(C_pathway_course_level != "Level 1") %>% #filters out level 1 courses
  summarise(course_count = n(), .groups = 'drop') %>%  # Changed to count D_course_id
  pivot_wider(names_from = D_pass_fail, values_from = course_count, values_fill = list(course_count = 0))

# Add a row to sum the column totals
pathways_student_count <- pathways_student_count %>%
  bind_rows(
    pathways_student_count %>%
      summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
      mutate(C_pathway = "Total")
  )

# Add a column for the column total
pathways_student_count <- pathways_student_count %>%
  mutate(Total = rowSums(across(where(is.numeric))))

# Print the table with column totals and row total
pathways_student_count %>% flextable()

# Create a percentage table
percentage_student_count <- pathways_student_count %>%
  rowwise() %>%
  mutate(across(2:4, ~ round(. / sum(c_across(2:4)) * 100, 1), .names = "percent_{col}"))

# Print the percentage table sorted by the total column
percentage_student_count %>% flextable()


```

#4.3 Engineering enrollment

```{r}
#what courses were students enrolled in engineering taking?
engineering_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Engineering") %>% 
  group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()
```

#4.4 Web Development Enrollment

```{r}
#what courses were students enrolled in web development taking?
web_dev_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Web Development") %>% 
  group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#and where are those enrollments
web_dev_deep_dive <- snapshot_pathways_table_2023 %>%
  filter(C_pathway == "Web Development") %>% 
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()
```

#4.5 Dual Credit Investiagation

```{r}
#Lets reminder ourselves what overall dual credit enrollment looks like
dual_credit_deep_dive <- base_analysis %>%
  filter(D_dual_credit_flag == "Y") %>% 
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#so 133 total enrollments

#and how many unique students
dual_credit_deep_dive <- base_analysis %>%
  filter(D_dual_credit_flag == "Y") %>% 
  distinct(D_stu_id, .keep_all = TRUE) %>%
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#For 93 students, and it sounds like Dr.Harvey was expecting the number at Manor High to be much higher.

#what courses are they enrolling in?
dual_credit_deep_dive <- base_analysis %>%
  filter(D_dual_credit_flag == "Y") %>% 
  group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#Okay, I am going to go back and check the course catalog for some of these similar course names to see if we were missing anything. Looks like we missed 7 courses across a number of course IDs, I am going to check the enrollment size of those courses to see how large the discrepency is
missing_dual_credits <- base_analysis %>%
  filter(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD")) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#so this captures 149 additional enrollments, across how many students?
missing_dual_credits <- base_analysis %>%
  filter(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD")) %>%
  distinct(D_stu_id, .keep_all = TRUE) %>%
  summarize(unique_students = n()) %>%
  print()

#for 134 students, lets see where they were
missing_dual_credits <- base_analysis %>%
  filter(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD")) %>%
  group_by(location_rollup) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()

#okay mostly Manor High, which makes sense for what Dr.Harvey was saying on the call.

#lets make a list of all of the distinct IDs in the missing_dual_credits data-frame
# Step 1: Get a string of all unique student IDs in missing_dual_credits
missing_dual_credits <- base_analysis %>%
  filter(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD")) %>%
  distinct(D_stu_id, .keep_all = TRUE)

# Extract unique student IDs as a string
unique_student_ids <- missing_dual_credits %>%
  pull(D_stu_id) %>%
  unique()

# Step 2: Filter a new dataframe to just dual credit IDs and see how many of the student IDs from step 1 are present
dual_credit_students <- base_analysis %>%
  filter(D_dual_credit_flag == "Y")

# Find the intersection of student IDs from missing_dual_credits and dual_credit_students
students_with_dual_credit <- dual_credit_students %>%
  filter(D_stu_id %in% unique_student_ids)

# Count the number of unique student IDs
num_students_with_dual_credit <- students_with_dual_credit %>%
  distinct(D_stu_id) %>%
  nrow()

# Print the result
print(num_students_with_dual_credit)

#so 20 already dual credit enrollment tagged, which means we were missing 114 students who were taking dual credit courses (seems about right!). Now what I will probably do is go back, add in these courses as flag "Y" for dual credit in the script, re-run the script, and re-run the charts with dual credit enrollments.

#just out of curosity, what was the enrollment in each of those classes that I missed
missing_dual_credits <- base_analysis %>%
  filter(D_course_id %in% c("DRAM 1310", "EDUC 1300", "GOVT 2036", "HUMA 1301", "HUMA 1302", "CMTHEA", "CMTHEB", "CMTHIA", "CMTHIB", "COALGA", "COALGB", "COALGM", "PSYC", "PSYCD")) %>%
  group_by(D_course_name.x) %>%
  summarize(enrollments = n()) %>%
  arrange(desc(enrollments)) %>%
  print()


```

#Part V: Export File - **What:** The below exports all your check tables to an excel file in the folder path you set above

-   **Check:** look at your data in the environment after running the code block below and be sure that all your variables have the follow the standard naming convention (D tag, lower case, underscores)

-   **Project Action:** Optional - change the file name

```{r}
#save merged files for next phase
# Get today's date
current_date <- format(Sys.Date(), "%m.%d.%y") # will update with the date it was last run

# Write dataframe to Excel file with today's date included in the filename
write.xlsx(cohort_pathways_table,
           file = paste0(full_file_path, "XXX", current_date, ".xlsx"),
           rowNames = FALSE)

# Write dataframe to Excel file with today's date included in the filename
write.xlsx(merged_transcript_cohort,
           file = paste0(full_file_path, "XXX", current_date, ".xlsx"),
           rowNames = FALSE)

```
