---
title: "ADMIN- Fake Transcript Data Creation"
author: "Zach Friedman"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 0: Read Libraries and connect to Sharepoint

### 0.1 (OPTIONAL) Install the latest version of the erstools library!

**ACTION: If you aren't sure if you have the most up to date release of erstools, follow these instructions: https://app.tettra.co/teams/ersknowledge/pages/updating-erstools-functions-markdown-template**

### 0.2 Read All Libraries

**ACTION: Run the code and read the comments for different functions you will have access to**

```{r, warning = F, message = F}
#ERS-specific functions!!
library(erstools) #includes functions like ers_read_sharepoint, ers_write_sharepoint, ers_table, and ers_column_info

#load data
library(Microsoft365R) # Use Microsoft365R to connect to SharePoint and load files
library(readxl) # Functions for reading Excel files: read_excel
library(writexl) # Functions for writing Excel files: write_xlsx
library(openxlsx) # Functions for writing Excel files: write.xlsx

#manipulate data
library(tidyverse) # Includes ggplot2, dplyr, tidyr, readr, etc.
library(data.table) # fread(): load large CSVs quickly
library(janitor) # Functions for cleaning and examining data: clean_names() instantly formats all column names
library(safejoin) # Provides safer join operations for 1:many matching - safe_left_join()
library(stringr) # str_detect(): pattern matching using regular expressions (see regex101.com)

#format and display data
library(scales) # Functions for scaling and formatting axes in plots: comma(), percent()
library(gt) # Creates and formats tables in a publication-ready format: gt()
library(flextable) # Creates and formats tables in a publication-ready format: flextable()
library(extrafont) # Manages and uses custom fonts in plots - loadfonts(): Loads fonts available on the system

#share output with supervisor
library(rmarkdown) # needed to create knitted files in output folder
library(downloadthis) # create download buttons in your knitted file so that a supervisor can download data in Excel!
```

## Part 1: Create Fake Data

### 1.0 Authenticate to sharepoint to be able to read in files

**ACTION: After running the code, you may need to look at your console code window below for instructions to log in. A window may open in your browser to ask you to log into Microsoft to confirm your identity**

**NOTE: If you ever have authentication issues, first run this in your console then try running the code chunk below again - AzureAuth::clean_token_directory()**

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

### 1.1 Create Fake Student Data
```{r}
library(tidyverse)
library(lubridate)

set.seed(42)  # for reproducibility

# Generate 100 unique 8-digit student IDs
n_students <- 100
student_ids <- sample(1e7:1e8 - 1, n_students)

# Possible values
school_years <- c(2024, 2023, 2022)
grades <- c(9, 10, 11, 12, 13)
statuses <- c("Active", "Graduated", "Transferred", "Dropped Out")
school_ids <- 1:5
exit_codes <- c("Graduated", "Transferred", "Dropped Out")
exit_descriptions <- c("Graduated", "Transferred or Dropped")
races <- c("Asian", "Native Hawaiian", "Black", "White", "Hispanic", "Two or More")

# Probabilities for race
race_probs <- c(0.5, 0.2, 0.1, 0.1, 0.05, 0.05)

# Create rows for each student (some will have multiple rows)
rows_per_student <- sample(1:3, n_students, replace = TRUE)

fake_data <- map2_dfr(student_ids, rows_per_student, function(id, n_rows) {
  map_dfr(1:n_rows, function(i) {
    school_year <- sample(school_years, 1, prob = ifelse(school_years == 2024, 0.7, 0.15))
    grade <- sample(grades, 1, prob = c(rep(0.225, 4), 0.1))  # favor 9–12
    status <- sample(statuses, 1)
    school_id <- sample(school_ids, 1)
    
    # Entry/exit dates
    entry_date <- if (runif(1) > 0.1) ymd("2023-08-31") else sample(seq(ymd("2023-09-01"), ymd("2024-01-31"), by = "day"), 1)
    exit_date <- if (runif(1) > 0.1) ymd("2024-06-21") else sample(seq(ymd("2024-01-01"), ymd("2024-06-01"), by = "day"), 1)
    
    exit_code <- sample(exit_codes, 1)
    exit_description <- ifelse(exit_code == "Graduated", "Graduated", "Transferred or Dropped")
    
    tibble(
      Masked_Student_ID = id,
      School_Year = school_year,
      Stu_Grade = grade,
      Stu_Enrollment_Status = status,
      Stu_School_ID = school_id,
      Stu_Entry_Date = entry_date,
      Stu_Exit_Date = exit_date,
      Stu_Exit_Code = exit_code,
      Stu_Exit_Code_Description = exit_description,
      Stu_SWD_flag = sample(c("Y", "N"), 1),
      Stu_ELL_flag = sample(c("Y", "N"), 1),
      Stu_FRL_flag = sample(c("Y", "N"), 1),
      Stu_Race = sample(races, 1, prob = race_probs)
    )
  })
})

# View first few rows
print(head(fake_data))


#create mapping for shcool IDs to names 1-5
school_id_mapping <- tibble(
  Stu_School_ID = 1:5,
  School_Name = c("Waikea High School", "Hilo High School", "Konawaena HS", "Kamehameha High", "Hawaii Preparatory Academy")
)

#add school names to data
student_data <- fake_data %>%
  left_join(school_id_mapping, by = "Stu_School_ID")



```

###1.2 Create Fake Transcript Data

```{r}
# Assume you already have `student_data` from the prior code with columns:
# Masked_Student_ID, Stu_School_ID, School_Year, Stu_Grade

set.seed(42)

# Define course catalog
course_catalog <- tribble(
  ~Crs_ID,   ~Crs_Name,                     ~Crs_Subject,
  "ENG101", "English 9",                    "English",
  "ENG201", "English 10",                   "English",
  "MAT101", "Algebra I",                    "Math",
  "MAT201", "Geometry",                     "Math",
  "SCI101", "Biology",                      "Science",
  "SCI201", "Chemistry",                    "Science",
  "SOC101", "World History",                "Social Studies",
  "SOC201", "US History",                   "Social Studies",
  "PE101",  "Physical Education",           "PE",
  "ART101", "Intro to Art",                 "Arts",
  "MUS101", "Ukulele Ensemble",             "Music",
  "HAW101", "Hawaiian Language I",          "World Language",
  "HAW201", "Hawaiian Language II",         "World Language",
  "BUS101", "Intro to Business",            "CTE",
  "CUL101", "Culinary Arts",                "CTE",
  "AG101",  "Agricultural Science",         "CTE",
  "MAR101", "Marine Biology",               "Science",
  "TECH101","Digital Media",                "Technology",
  "LIT101", "Pacific Islander Literature",  "English",
  "ENV101", "Environmental Science",        "Science",
  "DAN101", "Hula Dance",                   "Arts",
  "ADV100", "Advisory",                     "Advisory"
)

# Location lookup
location_lookup <- tribble(
  ~Detailed_Location, ~Detailed_Location_Desc,
  "1001", "Waikea High School",
  "1002", "Hilo High School",
  "1003", "Konawaena HS",
  "1004", "Kamehameha High",
  "1005", "Hawaii Preparatory Academy"
)

# Function to create transcript rows for a student
generate_courses_for_student <- function(stu_id, grade, loc_id, year) {
  # Select 5 random courses + 1 advisory
  core_courses <- course_catalog %>% 
    filter(Crs_Name != "Advisory") %>% 
    slice_sample(n = 5)
  
  advisory <- course_catalog %>% filter(Crs_Name == "Advisory")
  
  courses <- bind_rows(core_courses, advisory)
  
  # Generate transcript rows
  courses %>%
    mutate(
      Stu_ID1 = stu_id,
      Stu_Grade_Level = grade,
      Empl_ID_Tchr1 = sample(100000:999999, n(), replace = TRUE),
      Detailed_Location = loc_id,
      Detailed_Location_Desc = location_lookup$Detailed_Location_Desc[match(loc_id, location_lookup$Detailed_Location)],
      Class_ID = paste0(Crs_ID, "_", stu_id, "_", row_number()),
      Section = sample(1:5, n(), replace = TRUE),
      `Crs_Credit Recovery` = sample(c("Y", "N"), n(), prob = c(0.1, 0.9), replace = TRUE),
      Term = sample(c("Full Year", "Fall", "Spring"), n(), replace = TRUE),
      SchoolYr = year,
      Stu_Grade_Outcome = sample(c("A", "B", "C", "D", "F", "U", "Pass"), n(), replace = TRUE),
      Nbr_Credits_Possible = ifelse(Crs_Name == "Advisory", 0, sample(c(0.5, 1.0), n(), replace = TRUE)),
      Stu_Credits_Earned = ifelse(Stu_Grade_Outcome %in% c("A", "B", "C", "Pass"), Nbr_Credits_Possible, 0),
      GPA_Points_Unweighted = ifelse(Crs_Name == "Advisory", 0, round(runif(n(), 0, 4), 2)),
      GPA_Points_Weighted = round(GPA_Points_Unweighted + ifelse(Crs_Subject %in% c("Math", "Science", "English"), 0.5, 0), 2),
      GPA_Added_Value = GPA_Points_Weighted - GPA_Points_Unweighted
    )
}

# Generate full dataset
transcript_data <- student_data %>%
  select(Masked_Student_ID, Stu_School_ID, Stu_Grade, School_Year) %>%
  pmap_dfr(~generate_courses_for_student(..1, ..3, ..2, ..4))

# View a sample
head(transcript_data)



```

### 1.3 Create Fake HR Data
```{r}
# Assuming transcript_df already exists with a column `Empl_ID_Tchr1`
set.seed(123)  # for reproducibility

# Get unique teacher IDs from transcript data
teacher_ids <- unique(transcript_data$Empl_ID_Tchr1)

# Define possible race categories (mirroring student data)
race_categories <- c("Asian", "Native Hawaiian", "White", "Black", "Hispanic", "Two or More")

# Define degree levels
edu_levels <- c("BA", "MA", "MA+", "PhD")
edu_level_desc <- c("Bachelor's", "Master's", "Master's Plus", "Doctorate")

# Define subjects for certification
subjects <- c("Math", "Science", "English", "Social Studies", "PE", "Music", "Hawaiian Language", "Special Education", "CTE", "Art")

# Create HR dataset
hr_df <- data.frame(
  Empl_ID_Tchr1 = teacher_ids,
  Empl_Race = sample(race_categories, length(teacher_ids), replace = TRUE, prob = c(0.5, 0.2, 0.1, 0.1, 0.05, 0.05)),
  Empl_FTE = sample(c(1, 0.8, 0.6, 0.5), length(teacher_ids), replace = TRUE, prob = c(0.9, 0.05, 0.03, 0.02)),
  Yrs_In_Position = sample(1:15, length(teacher_ids), replace = TRUE),
  Yrs_In_Scl = sample(1:20, length(teacher_ids), replace = TRUE),
  Yrs_Tching_Exp = sample(1:30, length(teacher_ids), replace = TRUE)
)

# Add education levels and descriptions
hr_df <- hr_df %>%
  mutate(
    Edu_Level = sample(edu_levels, n(), replace = TRUE),
    Edu_Level_Desc = case_when(
      Edu_Level == "BA" ~ "Bachelor's",
      Edu_Level == "MA" ~ "Master's",
      Edu_Level == "MA+" ~ "Master's Plus",
      Edu_Level == "PhD" ~ "Doctorate"
    )
  )

# Add certification info
hr_df <- hr_df %>%
  mutate(
    Degree_Subject = sample(subjects, n(), replace = TRUE),
    Cert_Subject_Code = str_sub(Degree_Subject, 1, 3) %>% toupper(),
    Cert_Subject_Desc = Degree_Subject,
    Cert_Type_Code = sample(c("P", "I", "T"), n(), replace = TRUE),  # Provisional, Initial, Temporary
    Cert_Type_Desc = case_when(
      Cert_Type_Code == "P" ~ "Provisional",
      Cert_Type_Code == "I" ~ "Initial",
      Cert_Type_Code == "T" ~ "Temporary"
    ),
    Addl_Cert_Subj_Code = ifelse(runif(n()) < 0.3, str_sub(sample(subjects, n(), replace = TRUE), 1, 3) %>% toupper(), ""),
    Addl_Cert_Subj_Desc = ifelse(Addl_Cert_Subj_Code != "", sample(subjects, n(), replace = TRUE), ""),
    Nat_Board_Cert_Status = sample(c("Certified", "Not Certified"), n(), replace = TRUE, prob = c(0.7, 0.3))
  )

# View a few rows
head(hr_df)




```


## Part 5: Upload File to Sharepoint

### 5.1 Set output data folder Path

**ACTION: Change the processed_data_folder_path to the path to your data. The folder path must start with a folder that is just below the "Client Work" level. For example, it should start with "District Partners", "Internal", "State Partners", etc..**

```{r}
processed_data_folder_path <- "Core Services/Strategy & Innovation/Data Strategy & Innovation/3. R Stuff/Transcript Templatization (2024)"
```


### 5.2 See guidance on using the "ers_write_sharepoint" function

**ACTION: Run the code below to see documentation on our custom function!**

```{r, eval = interactive()}
?ers_write_sharepoint
```

### 5.3 Write files

**ACTION: Replace "raw_data" with your desired name of your data and replace "XXX.csv" with the desired name of your output file. You can output CSV or XLSX files**

```{r}
#student data
ers_write_sharepoint(data = student_data,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "HPS_Student_Data.csv",
                     drive_name = "internal_drive")

#transcript data
ers_write_sharepoint(data = transcript_data,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "HPS_Transcript_Data.csv",
                     drive_name = "internal_drive")

#Hr data
ers_write_sharepoint(data = hr_df,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "HPS_HR_Data.csv",
                     drive_name = "internal_drive")
```

## Part 6: Knit the file into a nicely formatted HTML that you can share with teammates!

**ACTION: ALL YOU NEED TO DO IS Use Ctrl+Shift+K to knit your file, and it will upload an html file to your Github repo in a folder called "knitted files"!**

## OPTIONAL Part 7: Upload knitted file to Sharepoint

**ACTION: If you have made changes to your file since last knitting, make sure to knit first before running the code below. It will upload your knitted file to your "processed_data_folder_path" but you can change the destination_of_knitted_file below.**

```{r eval = FALSE}
# You can change the destination path of your knitted file here if you wish
destination_of_knitted_file <- processed_data_folder_path

# Define the path to the current R Markdown file
current_file <- rstudioapi::getActiveDocumentContext()$path

# Determine the output directory and file name for the knitted HTML
output_dir <- dirname(current_file)
output_file <- gsub(" ", "-", paste0(tools::file_path_sans_ext(basename(current_file)), ".html"))
knitted_file <- file.path(output_dir, "knitted files", output_file)


client_work_drive <- orgfiles_site$get_drive("Client Work")

client_work_drive$upload_file(knitted_file, dest = file.path(destination_of_knitted_file, output_file))
```


