---
title: "1_raw_data_student_validation_MNPS"
author: "Simon Ellis"
reviewer: "Zach Friedman"
initial write-up: July 23 2024
review:
last update: April 10 2025
output: html_document
  toc: yes
  df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Start fresh
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and Connect to Sharepoint

-   **What:** Before you start with scripts in R, its recommended to review the analysis guidance in Tettra here- <https://app.tettra.co/teams/ersknowledge/pages/transcript-analysis-start-here>, or reach out to a transcript analysis expert for more information! (Valentina, Simon, Laurel, Rebel, etc.)

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
# Connect to SharePoint where the data files live
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
orgfiles_site     <- get_sharepoint_site(site_url = orgfiles_site_url)
internal_drive    <- orgfiles_site$get_drive("Internal")

# Read the district config — tells us where to find files, how to rename columns, etc.
config <- yaml::read_yaml(here::here("config", "districts", "tr_nashville.yaml"))
raw_data_folder_path <- config$raw_data_folder_path
```

# Part 1: Set your folder path(s) and read in your data

## 1.1 Bring in data for a Singular File

-   **What:** The code below sets file path and brings in data from one file. If your data is spread out over multiple files - i.e., one file per year, go to step 1.1

-   **Project Team Action:** Replace folder path with the folder path where your files is replace **file_name_with_extension** with name of your file

```{r, warning=F, message=F}

# Load the raw student demographics data
student_dem_raw <- ers_read_sharepoint(
  folder_path = raw_data_folder_path,
  file_name_with_extension = config$raw_files$stu_demographics
)

# Clean column names to remove spaces and special characters
student_dem_raw <- student_dem_raw %>% clean_names()
```



## Part 2: Validate Data

### 2.1 Rename Columns

#### a. Map project column names to ERS standard names

-   **What:** Assign your project's column names to ERS standard column names. ERS Standard names are on the left side of the equal sign, your district's columns are on the right. See resolution of critical/important fields that are missing in step. Having standard column names is critical for this code running without error

-   **Check:** In your data, all the critical fields are present and most of the important fields are present

-   **Project Team Action:** Definitely change (see guidance in code block)

```{r}
# Rename demographic columns to standard names from tr_[district].yaml
student_dem_raw <- student_dem_raw |>
  rename(any_of(unlist(config$column_map$demographics)))
```

#### b. Review columns

-   **What:** Manually check that you have the columns you need/would expect after renaming, and have not lost any critical data. Review the columns you are missing (as necessary).

-   **Check:** Ensure that none of the missing fields are critical for your analysis AND expected to be included in a transcript data file (I.e., it is okay if you are missing student demographic fields if you got a separate student file from the district)

-   **Project Team Action:** If the missing column(s) cannot be created from available data, please talk to your team lead and discuss if the column(s) is required and if yes, reach out to the district to get this and rerun this validation script on the new data file. You can disregard fields that will be merged in later, but keep track of them. Update the Data Validation Checks table with the result of your manual checks

### 2.2 Manual Checks

-   **What:** A handful of validation checks are most readily done through quick manual steps rather than R code - for the checks outlined below, run them in the non-code way suggested and then manually input them into a checks table in OneNote, Word, or another tool.

    1.  **Do we have right year(s) of data?** Look at the name of your raw data file and see if date has been included somewhere in it. Alternatively, skim your raw data and see if any of the fields outline the date that the data was pulled by the district team. If all else fails - simply ask your district contact!

    2.  **Does the district report race/ethnicity as one field or separate fields?** Look at the term column to determine how race/ethnicity is reported. Some districts report race and ethnicity as separate fields, (ie. Race: Caucasian, Ethnicity: Middle Eastern) in which case its helpful to create a rollup column to accurately answer district questions

-   **Project Team Action:** Update the Data Validation Checks table with the result of your manual checks

#### a. Entry/Exit Dates

-   **What:** Understand the structure of your file (or files) by loading the number of rows, check the number of unique students and counts by year. If your files were in multiple years, you should make sure they are now stacked with the code in step 1.1 above!

-   **Check:** Number of rows should be roughly the number of high school students in the district

-   **Project Team Action:** Update the Data Validation Checks table with the result of your check. You might write something like "all five files have roughly XX rows each, which matches our expectation since there's YY high school students in the district". This looks at the count by year. If you need to calculate a year consider using student/enter dates, etc. See guidance in Tettra for student entry/exit dates- <https://app.tettra.co/teams/ersknowledge/pages/wip-deep-dive-snapshot-analysis>

```{r}
#student demographics
student_dem_raw %>% nrow() #View output of check to know what to write for your status column

# Force exit dates for students who don't have one- in each case of a student having no exit date, we will assume they are still enrolled for that year, and set their exit date as June 30th of the value in D_year for that row. The format of the date should be year, month, day (so value from D_year %yyyy, %mm, %dd)
student_dem_raw <- student_dem_raw %>%
  mutate(
    D_stu_exit_date = ifelse(is.na(D_stu_exit_date), 
                             as.Date(paste0(D_year, "-06-30")), 
                             D_stu_exit_date)
  )


#hmm its still not making it a date. Can we force convert that column to be a date?
student_dem_raw <- student_dem_raw %>%
  mutate(
    D_stu_exit_date  = as.Date(D_stu_exit_date, format = "%Y-%m-%d"),
    D_stu_enter_date = as.Date(D_stu_enter_date, format = "%m/%d/%Y")
  )

#there we go!

# Calculate the number of days between EntryDate and ExitDate
student_dem_raw <- student_dem_raw %>%
  mutate(
    M_days_enrolled = as.numeric(difftime(`D_stu_exit_date`, `D_stu_enter_date`, units = "days"))
  )

#count of unique Student_number by school_year
student_dem_raw %>%
  group_by(`D_year`) %>%
  summarise(n_distinct(`D_stu_id`))

## Number of unique values of student number in student_dem_raw
student_dem_raw %>%
  summarise(n_distinct(`D_stu_id`))

```

#### b. Resolution Remove Rows

-   **What:** Transform your data if necessary. You can also leave a note about district follow-up required if the team did not receive the full data set

-   **Check:** If you have indeed removed rows from your tables through code, ensure that your table now has fewer rows than it did before

-   **Project Team Action:** Update code as needed to filter/remove data

```{r}
#remove rows
# Remove rows where a specific condition is met
# For example, remove rows where age is greater than 18

#student_dem_raw <- student_dem_raw %>%
#  filter(age >= 18)

# Print the result
print(head(student_dem_raw))

```


### 2.2.1 Student with multiple schools across years
```{r}

#Create a clean location name colum that removes any numbers from the start of the location name so that we have just the high schools name
student_dem_raw <- student_dem_raw %>%
  mutate(D_location_name_clean = gsub("[0-9-]", "", D_location_name)) %>%
  mutate(D_location_name_clean = trimws(D_location_name_clean))


#Now using location name clean, Check for a student that has a different location name across school years
student_multiple_schools_clean <- student_dem_raw %>%
  group_by(D_stu_id) %>%
  filter(n_distinct(D_location_name_clean) > 1) %>%
  ungroup() %>% 
  print()

#how many distrinct students in student_multiple_schools_clean?
distinct_students_multiple_schools_clean <- student_multiple_schools_clean %>%
  summarise(n_distinct(D_stu_id)) %>% 
  print()

#how many of those students in multiple schools clean had fewer than 5 days enrolled using M_days_enrolled?
students_multiple_schools_clean_fewer_than_5_days <- student_multiple_schools_clean %>%
  filter(M_days_enrolled < 5) %>%
  summarise(n_distinct(D_stu_id)) %>% 
  print()

#Create a new data-frame of students in multiple schools clean where the student had more than 50 days at a school
students_multiple_schools_clean_more_than_50_days <- student_multiple_schools_clean %>%
  filter(M_days_enrolled > 50) %>%
  print()

#how distinct students is that?
distinct_students_multiple_schools_clean_more_than_50_days <- students_multiple_schools_clean_more_than_50_days %>%
  summarise(n_distinct(D_stu_id)) %>% 
  print()


#Create Student Check 7 with student ID 190000206
student_check_7 <- student_multiple_schools_clean %>%
  filter(D_stu_id == 190001128) %>% 
  print()

```

### 2.3 Are there any duplicates?

#### a. Review Duplicates Table

-   **What:** Understand if we have any fully duplicated rows

-   **Check:** Review the transcript_combined_file_dups table to understand the duplicates and why they are appearing before deleting

-   **Project Action:** Update Data Validation Checks table

```{r}
#view duplicates in student enrollment file
student_enrollment_combined_dups <- student_dem_raw %>%
  group_by_all() %>%
 mutate(n = n()) %>%
    filter(n>1)
view(student_enrollment_combined_dups)
nrow(student_enrollment_combined_dups)
```

#### b. Remove Duplicates

-   **What:** The below code chunk will remove the duplicate rows, so only run once you feel confident they are true duplicates

-   **Check:** That the number of duplicates is now 0

-   **Project Action:** Update Data Validation Checks table

```{r}
#removes duplicates
student_dem_raw <- student_dem_raw[!duplicated(student_dem_raw), ]
#count to make sure the right number of rows were removed
nrow(student_dem_raw)

```

### 2.4 Missing & Unique Values

#### a. Update Blanks or NULL to NA

-   **What**: This will update all blanks ("") or nulls with NA, a special value in R that denotes missing data. This will make it easier to deal with missing data throughout the code

-   **Check**: See below

-   **Project Team Action**: None

```{r}
# Update blanks or NULL to NA
student_dem_raw <- student_dem_raw %>%
  mutate(across(everything(), ~if_else(. %in% c("", NULL, "(null)"), NA, .)))
```

#### b. Review missing data

-   **What:** Understand if we have any blanks or NAs in key fields

-   **Check:** Are there any critical columns with NA/blank values? Do the Unique values for things like Student ID and D_location match expectations?

-   **Project Team Action:** Review output and note any key columns with a large % of missing values. Are they critical? Do you need to ask district about missing data? Discuss any issues with project team. (2) Do the unique count of rows align with what you would expect to see? Investigate further as needed after checking for duplicates. Update Data Validation Checks table

```{r}
#create new data frame with column information
col_info_enrollment <-  data.frame(
  Column_Name = names(student_dem_raw),
  Data_Type = sapply(student_dem_raw, class),
  Count_Rows = sapply(student_dem_raw, function(x) length(na.omit(x))),
  NA_Rows = sapply(student_dem_raw, function(x) sum(is.na(x))),
  Unique_Rows = sapply(student_dem_raw, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)
# #calc % missing
 col_info_enrollment <- col_info_enrollment %>%
            mutate(Percent_Missing = NA_Rows / nrow(student_dem_raw) * 100)

#print new data frame
col_info_enrollment

## Now do the same thing, but instead check for values of 0 
#create new data frame with column information
col_info_enrollment <-  data.frame(
  Column_Name = names(student_dem_raw),
  Data_Type = sapply(student_dem_raw, class),
  Count_Rows = sapply(student_dem_raw, function(x) length(na.omit(x))), 
  Zero_Rows = sapply(student_dem_raw, function(x) sum(x == 0)),
  Unique_Rows = sapply(student_dem_raw, function(x) length(unique(na.omit(x)))),
  stringsAsFactors = FALSE,
  row.names= NULL
)
# #calc % missing
 col_info_enrollment <- col_info_enrollment %>%
            mutate(Percent_Zero = Zero_Rows / nrow(student_dem_raw) * 100)

#print new data frame
col_info_enrollment

# Optional: Visualize missing data percentages
# library(ggplot2) # Ensure ggplot2 is loaded
ggplot(col_info_enrollment, aes(x = reorder(Column_Name, Percent_Zero), y = Percent_Zero)) +
 geom_bar(stat = "identity", fill = "steelblue") +
coord_flip() + # Makes labels easier to read
labs(title = "Percentage of Missing Values by Column",
       x = "Column Name",
        y = "Percent Missing (%)") +
  theme_minimal()
```

### 2.5 Distinct Counts of Students

#### a. Return the counts of student id

-   **What:** Count of student ids

-   **Check:** Review the counts of student id. Compare with public data

-   **Project Action:** Update Data Validation Checks table

```{r}
#student enrollment
student_enrollment_count <- student_dem_raw %>% 
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

```

#### b. Return Distinct Counts of Students by Year

-   **What:** Count of student ids by year

-   **Check:** Review the counts of student id by year. Are students evenly distributed across years as expected?

-   **Project Action:** If your analysis focuses on specific schools, this is where you'd activate and update that filter. Make sure the school names here exactly match the names in your D_location_name column.

```{r}
#count of unique student ids for each year of data
student_enrollment_count_by_year <- student_dem_raw %>%
  group_by(D_year) %>%
  summarise(count_students = n_distinct(`D_stu_id`)) %>% 
  print()

#count of unique student ids for each year of data in the high schools or schools of interest
student_enrollment_count_by_year <- student_dem_raw %>%
  #filter(D_location_name %in% c("XXX High School", #add schools of interest here
                                #"XXX High School", 
                                                #)) %>% 
  group_by(D_year) %>%
  summarise(count_students = n_distinct(`D_stu_id`)) %>% 
  print()


# student_enrollment_count_by_year <- student_dem_raw %>% ... (as in template)
# print(student_enrollment_count_by_year) # Keep the print for table output

# Optional: Visualize student counts by year
 ggplot(student_enrollment_count_by_year, aes(x = D_year, y = count_students)) +
   geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = count_students), vjust = -0.5) + # Add numbers on bars
  labs(title = "Distinct Student Counts by Year",
        x = "Year",
        y = "Number of Distinct Students") +
   theme_minimal()
 
 #lets do student counts by location for just 2025
 location_by_enrollment_25 <- student_dem_raw %>%
  filter(D_year == 2025) %>%
  group_by(D_location_name) %>%
  summarise(count_students = n_distinct(`D_stu_id`)) %>% 
  print()

```

### 2.6 Number of years of data we have for each student

-   **What:** Creates a variable which assigns new value to student based on number of years of data we have for each unique student ID. This will be a valuable flag later when choosing cohort of students to analyze

-   **Check:** Choose at least three student ids to check. Filter to the student id in the combined file to identify how many years of data we have for them. Check the count in "num_years_data" column to make sure it is the same. You can also just visually inspect the file to make sure the "num_years_data" for each student id matches the number of rows for each student id

-   **Project Action:** None (run as is)


####a. Start with 9th graders in 2022

```{r}
#create a new variable that flags students who had any rows of enrollment for 9th grade in 2022
student_dem_raw <- student_dem_raw %>%
  mutate(C_9th_in_22 = ifelse(D_stu_grade_level == 9 & D_year == 2022, 1, 0))

#how many students is that?
students_9th_in_22 <- student_dem_raw %>%
  filter(C_9th_in_22 == 1) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

#Great.



```

####b. How many years of data do we have for each student?

```{r}
#counts how many yrs of data we have for each student
student_enrollment_yrs_of_data <- student_dem_raw %>%
  group_by(D_stu_id) %>%
  summarise(C_years_of_data_enrollment = n_distinct(D_year)) 

#check the year counts
print(student_enrollment_yrs_of_data)

#then adds that variable  as a column to the original combined data file
student_enrollment_combined <- merge(student_dem_raw, student_enrollment_yrs_of_data, by = "D_stu_id", all.x = TRUE)

# write a table with the count of unique student IDs with each number of years of data
student_enrollment_combined %>%
  group_by(C_years_of_data_enrollment) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

#how many students for whom we have 4 years of data in C_years_enrolled have any rows with fewer than 5 days enrolled in any row?
students_fewer_than_5_days_enrolled <- student_enrollment_combined %>%
  filter(C_years_of_data_enrollment == 4) %>%
  filter(M_days_enrolled < 5) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()

#How many students for whom we have 4 years of data in C_years_enrolled have multiple years in D_year where M_days_enrolled in less than 5?
students_multiple_years_fewer_than_5_days_enrolled <- student_enrollment_combined %>%
  filter(C_years_of_data_enrollment == 4) %>%
  group_by(D_stu_id) %>%
  filter((M_days_enrolled < 5) > 1) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()


#test table
distribution_lt5 <- student_enrollment_combined %>%
  filter(C_years_of_data_enrollment == 4) %>%
  group_by(D_stu_id) %>%
  summarise(
    rows_lt5 = sum(M_days_enrolled > 60, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  count(rows_lt5, name = "num_students")

distribution_lt5

```

####c. 4 Year 2022 9th graders
```{r}
#putting the two together, how many 9th graders in 2022 do we have 4 years of data for?
student_enrollment_combined %>%
  filter(C_9th_in_22 == 1 & C_years_of_data_enrollment == 4) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>% 
  print()


#wow thats great

```


### 2.7 Demographic Checks in district

#### a. % FRL, % ELL, % SWD

-   **What:** Check the percentage of students who qualify for free or reduced lunch by the year to compare to publicly available data

-   **Check:** (1) Check to make sure code worked - do total number of rows each year add up to what you would expect? (2) Compare % FRL to publicly available data

-   **Project Team Action:** Update the values of the columns to your district's values. E.G., ELL="Yes" or ELL = "True". Add to Data Validation Checks table

```{r}
#Before I do the checks need to change the binary for frl, swd, and ell to Yes and No
student_dem_raw <- student_dem_raw %>%
  mutate(
    D_stu_frl = ifelse(D_stu_frl == "1", "Yes", "No"),
    D_stu_ell = ifelse(D_stu_ell == "1", "Yes", "No"),
    D_stu_swd = ifelse(D_stu_swd == "1", "Yes", "No")
  )
```

#### b. % FRL

```{r}
#calculate percent FRL in district by year
# ADJUST "Y" if your data uses a different code for FRL, e.g., "1", "True", "Eligible"."
 percent_frl_check <- student_dem_raw %>%
       group_by(D_stu_frl, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
       select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()

```

#### c. % ELL

```{r}
#calculate percent ELL in district by year
percent_ell_check <- student_dem_raw %>%
       group_by(D_stu_ell, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
       select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()
```

#### d. % SWD

```{r}
#calculate percent SWD in district by year
 percent_swd_check <- student_dem_raw %>%
       group_by(D_stu_swd, D_year) %>%
       summarise(row_count = n()) %>% 
       ungroup() %>%
       group_by(D_year) %>%
       mutate(pct_rows = row_count / sum(row_count) * 100) %>% 
       arrange(desc(D_year)) %>% 
     select(-row_count) %>%  # Remove row_count if you only need percentages
       pivot_wider(names_from = D_year, values_from = pct_rows) %>%
       print()
```

#### e. % Race/ethnicity

-   **What:** Check the percentage of students who identify as each race/ethnicity category to compare to publicly available data. If you created a rollup column earlier, make sure to use the correct column.

-   **Check:** Compare to publicly available data

-   **Project Team Action:** Gut check/track

```{r}
#group by D_year and D_stu_race, then calculate the percentage of each race category for each year
race_percent_year <- student_dem_raw %>%
  group_by(D_year, D_stu_race) %>%
  summarise(row_count = n()) %>%
  mutate(pct_rows = row_count / sum(row_count) * 100) %>%
  select(-row_count) %>%  # Remove row_count if you only need percentages
  pivot_wider(names_from = D_year, values_from = pct_rows) %>%
  print()
```

### 2.8 MNPS- Checking student entry and exit dates
```{r}
#what are the unique values for entry data and exit date, and the number of unique student IDs with each value?
entry_dates <- student_dem_raw %>%
  group_by(D_year, D_stu_enter_date) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>%
  arrange(desc(count_students)) %>% 
  print()

#and exit dates
exit_dates <- student_dem_raw %>%
  group_by(D_year, D_stu_exit_date) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>%
  arrange(desc(count_students)) %>% 
  print()


#and looking at m_days_enrolled, how many students, by year, are enrolled for less than 15 days total in the same year?
students_less_than_15_days_enrolled <- student_dem_raw %>%
  filter(M_days_enrolled < 15) %>%
  group_by(D_year) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>%
  arrange(desc(count_students)) %>% 
  print()


```

### 2.9 Grade Level Double-Check for RT
```{r}
#create a table that shows by year, the distribution of distinct students by grade level
grade_level_distribution <- student_dem_raw %>%
  group_by(D_year, D_stu_grade_level) %>%
  summarise(count_students = n_distinct(D_stu_id)) %>%
  arrange(D_year, D_stu_grade_level) %>%
  print()

```


## Part 3: File Cleaning for export & Checks

-   **What:** Look over the data to ensure it is clean and ready for the next phase of analysis, add any final checks or cleaning

-   **Check:** Gut check/check against publicly available data (as needed)

-   **Project Team Action:** None

### 3.0 Check Random Student IDs

```{r}
# Create a function to pick a random student ID and create a new data frame
pick_random_student <- function(data) {
  # Pick a random student ID
  random_student_id <- sample(data$D_stu_id, 1)
  
  # Filter the data frame to create a new data frame for the selected student
  student_Check <- data %>%
    filter(D_stu_id == random_student_id)

  
  # Return the new data frame
  return(student_Check)
}

#student check 1
student_check_1 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_1)

#student check 2
student_check_2 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_2)

#student check 3
student_check_3 <- pick_random_student(student_dem_raw)

# Print the result
print(student_check_3)


#things look mostly normal here to me, doing another check of students who might have multiple grade levels in the same year (as this has been a problem in other districts)
student_check_multiple_grades <- student_dem_raw %>%
  group_by(D_stu_id, D_year) %>%
  filter(n_distinct(D_stu_grade_level) > 1) %>%
  ungroup()

#how many distinct students across all years in student_check_multiple_grades?
distinct_students_multiple_grades <- student_check_multiple_grades %>%
  summarise(n_distinct(D_stu_id)) %>% 
  print()

#how about just in 2025
distinct_students_multiple_grades_25 <- student_check_multiple_grades %>%
  filter(D_year == 2025) %>%
  summarise(n_distinct(D_stu_id)) %>% 
  print()

#okay 406, not massive. Let's look at one of those students
student_check_4 <- student_check_multiple_grades %>%
  filter(D_stu_id == 190000620) %>% 
  print()


```

### 3.1 Rearrange columns into desired order if needed

-   **What:** Creates new data-frame with changed order of appearance of your columns in your data-frame for an ease of viewing

-   **Check:** None

-   **Project Team Action:** None

```{r}
#rearrange columns by listing existing columns in the order you want them to appear
stu_roll_up <- student_dem_raw %>% 
  select(D_stu_id,
           D_location_id,
           D_location_name, 
           D_year, 
           D_stu_enter_date, 
           D_stu_exit_date, 
           D_stu_grade_level, 
           D_stu_race, 
           D_stu_frl, 
           D_stu_ell, 
           D_stu_swd, 
           everything()) #Keeps all columns (if not listed above)



```

### 3.2 MNPS- Check for student IDs with missing courses
```{r}
#can you pull the unique location names for the following student IDs- 190363571, 190152507, 190144372, 190000478, 190136958
missing_courses <- student_dem_raw %>%
  filter(D_stu_id %in% c(190363571, 190152507, 190144372, 190000478, 190136958)) %>%
  select(D_stu_id, D_location_name) %>%
  distinct() %>%
  print()


#FIND THIS STUDENT 	190000079
student_check_5 <- student_dem_raw %>%
  filter(D_stu_id == 190000079) %>% 
  print()



student_check_6 <- student_dem_raw %>%
  filter(D_stu_id == 190000206) %>% 
  print()

```

# Part 4: Upload File to Sharepoint

-   **What:** Upload validated dataframes as .csv files to sharepoint for further analysis.

-   **Check:** None/Update names of data-frames as needed for team.

-   **Project Action:** `file_name_with_extension` will need to be updated.

```{r}
# Export to 2. Processed Data --> 2. Transcript folder on SharePoint
ers_write_sharepoint(data = stu_roll_up,
                     folder_path = raw_data_folder_path,
                     file_name_with_extension = "/2. Processed Data/01_tr_student_validated.csv",
                     drive_name = "client_work_drive")
```
